### 2016-09-26 paper.tex
Initial commit

\documentclass[journal,peerreview,onecolumn]{IEEEtran}
% \documentclass[journal]{IEEEtran}

\usepackage{cite}
\usepackage[pdftex]{graphicx}
\graphicspath{{./Fig/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{url}
\usepackage{listings}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tabularx}
\usepackage{empheq}

\newcommand\MyBox[2]{
  \fbox{\lower0.75cm
    \vbox to 1.2cm{\vfil
      \hbox to 1.2cm{\hfil\parbox{1.4cm}{#1\\#2}\hfil}
      \vfil}%
  }%
}

% for peerreview
\usepackage{setspace}
\onehalfspacing

\newtheorem{prop}{Proposition}

% correct bad hyphenation here
% \hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Operational Feature Selection in Gaussian Mixture Models}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Adrien~Lagrange,~Mathieu~Fauvel~and~Manuel~Grizonnet% <-this % stops a space
\thanks{A. Lagrange and M. Fauvel are with the Université de Toulouse,
INP-ENSAT, UMR 1201 DYNAFOR, France and with the INRA, UMR 1201
DYNAFOR, France.}
\thanks{M. Grizonnet is with Centre Nat. d'Etudes Spatiales, French Space Agency, Toulouse, France.}}

% The paper headers
\markboth{Transactions on Computational Imaging,~Special Issue on Computational Imaging for Earth Sciences, September~2016}{}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
This article presents a forward feature selection algorithm based on Gaussian mixture model (GMM) classifiers. The algorithm selects iteratively features that maximize a criterion function which can be either a classification rate or a measure of divergence. Several variations of this algorithm are explored by changing the criterion function and also by testing a floating forward variation allowing backward step to discard already selected features.

An important effort is made in exploiting GMM properties to implement a fast algorithm. In particular, update rules of the GMM model are used to compute the criterion function with various sets of features. The result is a C++ remote module for the remote sensing processing toolbox Orfeo (OTB) developed by CNES.

Finally, the method is tested and also compared to other classifiers using two different datasets, one of hyperspectral images with a lot of spectral variables and one with heterogeneous spatial features. The results validate the fact that the method performs well in terms of processing time and classification accuracy in comparison to the standard classifiers available in OTB.\\
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
% \begin{IEEEkeywords}
% remote sensing, hyperspectral imaging, feature selection, gaussian mixture model, fast computing.
% \end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:intro}

\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of data available for a given landscape becomes larger and larger. Several missions are about to produce huge amount of data or have already produced it. After 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites missioned by the German space agency will produce images with 244 spectral bands with a resolution of 30x30m and revisit every 4 days\cite{Müller09enmap}. The Hyperspectral Infrared Imager (HyspIRI) of NASA will also take images with 212 spectral bands every 5 days. Additionally to hyperspectral data, hypertemporal data are also developping. The European satellites Sentinel-2 were launched successfully recently and the hypertemporal data produced by this mission will be fully available at the end of 2016\cite{drusch2012sentinel}.

However, processing this data is more and more challenging because of statistical and computational issues. This statistical issues are often referred as the \emph{curse of dimensionality}. The Hughes phenomenon \cite{hughes1968mean} states that with a given number of samples, prediction accuracy will decays when the complexity is higher than some optimum value. The problem is the fast increase of the number of parameters to estimate in order to build a model when dimension expands \cite{bouveyron2014model}. Thus, a important number of labeled samples is needed to perform learning. For example, in the case of Gaussian Mixture Models, the number of parameters progresses quadratically with the dimension, e.g. with 200 describing variables, the model estimation requires at least around 20,000 samples by class.

The computational issues are multiple. The computing infrastructure needed to process data is more and more expensive because the processing might requires a GPU and a large amount of RAM to load images which can weight a dozen of gigabytes \cite{christophe2011remote}\cite{plaza2011high}. The processing time is also limiting and requires to use parallelize computing (GPU, multi-threading).

A possible method to solve these issues is to perform a reduction of dimension. It is possible to extract a set of relevant features to get a parsimonious representation of the data \cite{jimenez1998supervised}.

For instance, in land-cover classification, given a set of spatial, temporal and spectral features, it is possible to extract those which are the most discriminant for the purpose of classification \cite{fassnacht2014comparison}. In hyperspectral data analysis from the hundreds of available spectral channels, it is possible to reduce the number of channels to make the processing more efficient in terms of statistical complexity because of the reduction of the number of parameters to estimate and thus also in term of computational time. Moreover, dimensional reduction might improves the capacity of generalization of the classifier and avoid overfitting.

There are two ways to reduce dimension \cite{Guyon:2006:FEF:1208773}: features extraction and feature selection. Feature extraction means creating new features by combining the existing ones, for example linear combination as in Principal Component Analysis \cite{jimenez1998supervised}. To the contrary, features selection selects a subset of existing features. It has the advantage to be much more interpretable for the end-user. The selected subset of features corresponds to the most important features for the given task.

There is a large diversity of methods for feature selection. However, they usually do not scale well with the number of pixels to be processed \cite{fauvel2015fast}. Nevertheless, methods based on Gaussian Mixture Models (GMM) have several interesting properties that make them suitable for feature selection in the context of large amount of data. By taking advantage of their intrinsic properties, it is possible to increase the computational efficiency with respect to standard implementation.

Several strategy have been explored to perform feature selection and, among them, wrapper methods receive a certain interest of the community. Wrapper methods can be seen as a search method to determine the best subset of variables for a given learning model. As exhaustive searches are out of question in a practical amount of time, numerous search strategies have been designed. Some optimal are under particular hypothesis \cite{narendra1977branch} and other suboptimal but easier to set up \cite{whitney1971direct}\cite{somol1999adaptive}. Such methods have the advantage to tune the selection to make the most of a particular classifier but require the training of multiple models to test various set of variables which can make them slow if the training is not optimized.

This work proposes to develop a forward feature selection method using GMM in continuation of \cite{fauvel2015fast} and an upgraded floating forward method. The basic method selects iteratively the meaningful features. At each step, the pool of selected features is used to train a GMM which allows to compute a criterion function then used to rank features at the next iteration. The upgraded method introduces possible backward steps after each addition of a feature. An efficient implementation of the method is presented in order to handle large amount of data. Moreover, the developed algorithm is made available as a remote module of the C++ Orfeo Toolbox \cite{christophe2008orfeo}. Finally, tests are conducted to compare different variations of the method and to compare also to other classifiers (GMM, Random Forest, k-nearest-neighbor) available in the Orfeo Toolbox.

The remaining of this article is organized as follows. Section~\ref{sec:gmm-hd} presents Gaussian Mixture Model classifiers and the features selection methods used to make them suitable for high-dimension space. The work done to develop a smart implementation of the proposed selection method is presented in Section~\ref{sec:implementation}. And finally, the tests conducted to explore variations of the algorithm and to compare it to other standard classifiers are detailed in Section~\ref{sec:test}.

\section{Gaussian Mixture Models in high dimension space}
\label{sec:gmm-hd}

In the remaining of the paper, the following notations are used. $\mathcal{S} = \{\mathbf{x}_i,y_i\}_{i=1}^{n}$ denotes the training set where $\mathbf{x}_i \in \mathbb{R}^d$ is the vector of features of the $i^{th}$ sample, $y_i = 1,...,C$ the associated label, C the total number of classes, $n$ the number of samples and $n_c$ the number of samples of class $c$.

    \subsection{Gaussian Mixture Models}

    The hypothesis of mixture models is that a given sample is the realization of a random vector which distribution is a mixture (convex combination) of several class conditioned distribution:
    \begin{equation}
        p(\mathbf{x}) = \sum_{c=1}^{C} \pi_c f_c(\mathbf{x}|\theta)
    \end{equation}
    where $\pi_c$ is the prior i.e. the proportion of class $c$ and $f_c$ a probability density function parametrized by $\theta$.

    The Gaussian mixture model (GMM) assumes that each $f_c$ is , conditionally to $c$, a Gaussian distribution of parameters $\boldsymbol{\mu}_c$ and $\boldsymbol{\Sigma}_c$ and so $f_c(\mathbf{x}|\theta)$ can be written as
    \begin{equation*}
        f_c(\mathbf{x}|\theta) = \frac{1}{(2\pi)^{\frac{d}{2}} |\boldsymbol{\Sigma}_c|^{\frac{1}{2}}} \exp \left( -\frac{1}{2} (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \right).
    \end{equation*}

    In supervised learning the class parameters $\boldsymbol{\mu}_c$, $\boldsymbol{\Sigma}_c$ and $\pi_c$ are usually estimated through the conventional unbiased empirical estimators:
    \begin{align}
        \hat{\pi}_c &= \frac{n_c}{n},\\
        \hat{\boldsymbol{\mu}}_c &= \frac{1}{n_c} \sum_{\{i|y_i = c\}} \mathbf{x}_i ,\\
        \hat{\boldsymbol{\Sigma}}_c &= \frac{1}{(n_c - 1)} \sum_{\{i|y_i = c\}} (\mathbf{x}_i - \boldsymbol{\mu}_c) (\boldsymbol{}x_i - \boldsymbol{\mu}_c)^t.
    \end{align}

    To predict the class of a new sample, the maximum a posteriori (MAP) rule is used and thus, using Bayes' law, the decision rule is
    \begin{equation*}
        \mathbf{x} \text{ belongs to } c \Leftrightarrow c = \text{arg} \max_{c \in C} p(c) p(\mathbf{x}|c).
    \end{equation*}
    Identifying $p(c)$ as $\pi_c$ and $p(x|c)$ as $f_c(x|\theta)$ and by taking the log, a simplified decision rule is obtained
    \begin{align}
        Q_c(\mathbf{x})
        &= 2 \log \left( p(c) p(\mathbf{x}|c) \right) \nonumber \\
        &= - (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \nonumber \\
        &~~-\log (|\boldsymbol{\Sigma}_c|) + 2 \log (\pi_c) - d \log (2\pi).
        \label{eq:decision}
    \end{align}

    It is important to notice that the inverse and the determinant of the covariance matrix are key elements of the decision function. The estimations of these elements suffer from the curse of dimensionality\cite{bouveyron2014model}. More precisely, the number of parameters to estimate increases quadratically relatively to the number of features as illustrated by Figure~\ref{fig:nb-param}. The minimum number of samples requires to estimate the model is equal to the number of parameters and it can be an issue in some case. For example, hyperspectral data is a case of high dimensional samples and, in general, very few labeled samples are available because of the difficulty and the cost to collect ground-truth.

    \begin{figure}[!t]
        \centering
        \begin{tikzpicture}
            \begin{axis}[xmin=0,xmax=500,ymin=0,width=0.5*\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = Dimension,ylabel =Number of parameters]
                \addplot+[very thick,mark=none,smooth,domain=0:500,samples=500] (\x,{\x*(\x+3)/2+1});
            \end{axis}
        \end{tikzpicture}
        \caption{Number of parameters by class in function of dimension.\label{fig:nb-param}}
    \end{figure}

    A lack of samples induces badly-conditioned covariance matrices and so unstable inversion and computation of the determinant. There are two major solutions to this problem. First option is to use a regularization method to stabilize the inversion of the covariance matrices\cite{hoerl1970ridge}. Second option is to use a features extraction/selection method in order to reduce the dimensionality of the samples. In this study, the latter option is explored and a feature selection method named sequential forward features selection is presented in the next Section.

    \subsection{Features selection}
    \label{sec:selection}

    The objective of feature selection is to retrieve the most compact representation of the data with minimum loss of information. In deleting the redundant information, feature selection allows to reduce dimension and thus avoid statistical issue, increase algorithm speed and limit storage requirements. Moreover, it adds several beneficial side effects. It improves data understanding by identify where is the most relevant information and it is then possible to save resources when organizing a new data acquisition.

    It is important to underline that the proposed method is a \emph{selection} method and not an \emph{extraction} method. It means the subset of variables obtained is composed of actual variables of the original set and not variables built as combination of several others as it is the case for example with PCA. This choice is made to assure an easier interpretation by the user of the obtained subset of variables.

    Features selection algorithms may be divided into three types. The first type called filter method is based uniquely on data analysis. Features are ranked according to a statistical analysis of the data. For example, the Principal Component Analysis (PCA) described in \cite{jimenez1998supervised} is a typical filter method and there is numerous other methods \cite{bruzzone1995extension}, \cite{biesiada2007feature}, \cite{demir2008phase}.

    The second sort are known as wrapper methods which can be seen as search method to determine the best subset of variables for a given learning model. As exhaustive searches are out of question in a practical amount of time, numerous search strategies have been designed some optimal under particular hypothesis \cite{narendra1977branch} and other suboptimal but easier to set up \cite{whitney1971direct}, \cite{somol1999adaptive}. The advantage of such methods compare to filter methods is that they are dedicated to a particular model but on the other hand, as they require the training of multiple models to test various set of variables, they tend to be slower.

    The third type corresponds to the embedded method which do not separate the features selection process from the learning algorithm and allow interaction between the two processes. The basic example of such method is the decision tree algorithm in which a feature is selected for the creation of each node. Embedded methods also exist for other models, e.g. SVM \cite{guyon2002gene} \cite{weston2003use}.
²
    The selection method proposed in this work is a wrapper method associated to GMM models. Thus, in order to set a wrapper method, two elements are needed, a function to rank the various features which is defined in Section~\ref{sec:criterion} and a search strategy. Two search algorithms are presented: the sequential forward features selection method (Section \ref{sec:forward-presentation}) and the sequential floating forward feature selection method (\ref{sec:floating-presentation}) the second being a more complex variation of the first.

        \subsubsection{Criterion function}
        \label{sec:criterion}

        \hspace{0pt} \\

        Criterion functions aim to evaluate either a rate of correct classification or the separability/similarity of class distributions. These functions are used to estimate which sets of variables are the best to represent data, to assure the separability of the classes and to perform classification. The choice of a criterion function is the choice of a way to compare sets of variables.

            \paragraph{Measures of correct classification}
            \label{sec:criterion-rate}

            \hspace{0pt} \\

            As described in \cite{congalton2008assessing} (chapter 4), all this measures of good classification are based on an error matrix $M$ called confusion matrix which is defined so that $M_{ij}$ is the number of samples of class $i$ classified as class $j$. The confusion matrix allows the computation of several interesting values relatively to each class:
            \begin{itemize}
                \item the number of True Positive (TP) corresponding to good prediction;
                \item the number of True Negative (TN) corresponding to good classification of the other class;
                \item the number of False Negative (FN) corresponding to the samples of the class labeled as an other class;
                \item the number of False Positive (FP) corresponding to the samples wrongly classified as part of of this class.
            \end{itemize}
            Figure~\ref{fig:confusion} illustrates the definition of this values.

            \begin{figure}[!ht]
                \centering
                \begin{tabular}{c >{\bfseries}r @{\hspace{0.7em}}c @{\hspace{0.4em}}c @{\hspace{0.4em}}c @{\hspace{0.7em}}l}
                  \multirow{10}{*}{\parbox{1.1cm}{\bfseries\raggedleft Actual\\ value}} &
                    & \multicolumn{3}{c}{\bfseries Prediction outcome} \\
                  & & \bfseries $c_1$ & \bfseries $c_2$ & \bfseries $c_3$ \\
                  & $c_1'$ & \MyBox{True}{Positive} & \MyBox{False}{Negative} & \MyBox{False}{Negative} \\[2.4em]
                  & $c_2'$ & \MyBox{False}{Positive} & \MyBox{True}{Negative} & \MyBox{~}{~} \\[2.4em]
                  & $c_3'$ & \MyBox{False}{Positive} & \MyBox{~}{~} & \MyBox{True}{Negative}
                \end{tabular}
                \caption{Confusion matrix with TP, TN, FP and FN relatively to $c_1$}
                \label{fig:confusion}
            \end{figure}

            \emph{The overall accuracy} (OA) is the rate of the number of samples with the correct predicted label over the total number of samples. This metric is easy to interpret but is biased in the case of unbalanced classes.

            \emph{The Cohen's kappa} (K) is a statistic which measures the probability of agreement between predictions and ground-truth.

            \emph{The mean F1 score} (F1mean) is the average of the F1 score for each class and the F1 score is the harmonic mean of the precision (number of True Positive over True Positive plus False Positive) and the recall (number of True Positive over True Positive plus False Negative).

            \begin{align}
                OA &= \frac{\text{TP}}{n}. \\
                K &= \frac{pa - pr}{1 - pr}.  \\
                \text{F1mean} &= \frac{2 \text{TP}}{2 \text{TP} + \text{FN} + \text{FP}}.
            \end{align}
            where TP stands for True Positive, FN for False Negative, FP for False Positive, $pa$ is the probability of agreement defined by $pa = OA$  and $pr$ the probability of random agreement defined by $pr = \sum_{c=1}^{C} \frac{TP_c}{FP_c} \frac{TP_c}{FN_c}$.

            In order to estimate classification rate, a cross-validation process over training set is used\cite{opac-b1127878}. The training set is divided in k folds, then, the training is done with $(k-1)$ folds and the performances are estimated with the remaining fold.

            \hspace{0pt} \\

            \paragraph{Similarity between distributions}

            \hspace{0pt} \\
            The second type of criterion functions is a measure of similarity between two distributions. These measures are called divergence function and are defined so that, if $S$ is a space of all probability distribution with same support, it satisfies
            \begin{align*}
                &\forall (p,q) \in S, \text{Div}(p,q) \geq 0, \\
                &\text{Div}(p,q) = 0 \Leftrightarrow p = q.
            \end{align*}
            More specifically, focus is made on two particular divergences: the Kullback–Leibler divergence and the Jeffries–Matusita distance. The advantage of these divergences is that they have an explicit expression in the case of Gaussian models. The simplification allows to get rid of any integration calculus which is a major problem when dealing with high-dimensional data.

            \emph{The Kullback-Leibler divergence} (KL divergence) measures the amount of information lost when the first distribution is approximated by the second one\cite{kullback1987letter}. The formal definition is
            \begin{equation}
                \text{Div}_{KL}(c_i,c_j) = \int_\mathbf{x} p(\mathbf{x}|c_i) \ln \left(\frac{p(\mathbf{x}|c_i)}{p(\mathbf{x}|c_j)}\right) d\mathbf{x}.
            \end{equation}
            And in the case of Gaussian model, it can be rewritten as follows
            \begin{align}
                &\text{Div}_{KL}(c_i,c_j) = \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_{c_i}^{-1} \boldsymbol{\Sigma}_{c_j}) \nonumber \\
                & + (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j})^t \boldsymbol{\Sigma}_{c_i}^{-1} (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j}) - d + \log \left( \frac{|\boldsymbol{\Sigma}_{c_i}|}{|\boldsymbol{\Sigma}_{c_j}|} \right) \biggr),
            \end{align}
            where Tr is the trace operator and $d$ the dimension of the distribution.

            It can be noticed that the KL divergence is not symmetric, i.e. $\text{Div}_{KL}(c_i,c_j) \ne \text{Div}_{KL}(c_j,c_i)$, and so the symmetrized version is used to compute the criterion function. In the case of Gaussian model, the symmetrization induces the following simplification of the formula
            \begin{align}
                \text{SKL}_{ij} &=\text{Div}_{KL}(c_i,c_j) + \text{Div}_{KL}(c_j,c_i) \nonumber \\
                &= \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_{c_i}^{-1} \boldsymbol{\Sigma}_{c_j} + \boldsymbol{\Sigma}_{c_j}^{-1} \boldsymbol{\Sigma}_{c_i}) \nonumber \\
                &~~+ (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j})^t (\boldsymbol{\Sigma}_{c_i}^{-1} + \boldsymbol{\Sigma}_{c_j}^{-1}) (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j}) - 2d \biggr).
            \end{align}

            Moreover, the divergence is computed between two classes and to obtain a unique value the weighted mean of divergence measures is taken as proposed in \cite{bruzzone1995extension}:
            \begin{equation}
                \text{C}_{SKL} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_{c_i} \pi_{c_j} \text{SKL}_{ij}.
            \end{equation}

            \emph{The Bhattacharyya distance} is defined as follows
            \begin{equation}
                \text{B}_{ij} = - \ln \left( \int_\mathbf{x} \sqrt{p(\mathbf{x}|c_i) p(\mathbf{x}|c_j)} d\mathbf{x} \right).
            \end{equation}
            And in the case of Gaussian model:
            \begin{align}
                \text{B}_{ij} = &\frac{1}{8} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \left( \frac{\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j}{2} \right)^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) \nonumber \\
                &+ \frac{1}{2} \ln \left( \frac{|\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j|}{\sqrt{|\boldsymbol{\Sigma}_i| |\boldsymbol{\Sigma}_j|}} \right).
            \end{align}

            \emph{The Jeffries–Matusita distance} is a measure based on the Bhattacharyya distance. It saturates if the separability between the two distribution increases \cite{bruzzone2009novel}. The JM distance is defined according to
            \begin{equation}
                \text{JM}_{ij} = \sqrt{ \int_\mathbf{x} \left[\sqrt{p(\mathbf{x}|c_i)} - \sqrt{p(\mathbf{x}|c_j)}\right]^2 d\mathbf{x} }.
            \end{equation}
            And the Jeffries–Matusita distance can be rewritten according to the Bhattacharyya distance
            \begin{equation}
                \text{JM}_{ij} = \sqrt{ 2 \{1 - \text{exp}[-B_{ij}]\} }.
            \end{equation}

            As for the KL divergence, a weighted mean of the distance between two classes is computed to aggregate the measures in a single value:
            \begin{equation}
                \text{C}_{JM} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_{c_i} \pi_{c_j} \text{JM}_{ij}.
            \end{equation}

            \vspace{10 mm}

            According to \cite{bruzzone2009novel}, it is interesting to notice that the KL divergence increases quadratically with respect to the distance between the mean vectors of the class distributions whereas the measures of correct classification asymptotically tends to one when distributions are perfectly separable. On the contrary, the JM distance tends to saturate as the measures of correct classification. Table~\ref{tab:crit} summarized the presented criterion functions and their characteristics.

            \begin{table}[!t]
                \centering
                \caption{Summary of the various criterion functions.\label{tab:crit}}
                \begin{tabular}[b]{l|c|c||c|c}
                  \hline
                  Criterion & Divergence & Classification rate & Gaussian hypothesis & Speed \\
                  \hline
                  Overall accuracy            &  & \checkmark & 0 & - \\
                  Cohen's kappa               &  & \checkmark & 0 & - \\
                  F1 mean                     &  & \checkmark & 0 & - \\
                  Kullback-Leibler divergence & \checkmark &  & ++ & + \\
                  Jeffries-Matusita distance  & \checkmark &  & ++ & + \\
                  \hline
                \end{tabular}
            \end{table}

        \subsubsection{Selection method}

            \paragraph{Sequential forward features selection}
            \label{sec:forward-presentation}

            The Sequential Forward Selection (SFS) starts with an empty set of selected features. Then, it computes at each step for all the remaining features the value of the criterion function $J$ chosen among the ones presented in Table~\ref{tab:crit} when the feature is added to the pool of selected features. The feature that maximizes the criterion function is definitively added to the pool of selected features and it moves to the next iteration. The algorithm stops when a given number of variables \emph{maxVarNb} has been selected. The Algorithm~\ref{alg:sfs} presents the process in details.

            The advantage of this search algorithm is its reasonable processing time. The trade-off is that the result is a suboptimal solution in the sense that an untested subset of variable could lead to better classification result.

            \begin{algorithm}
            \caption{Sequential forward features selection\label{alg:sfs}}
            {\footnotesize
            \begin{algorithmic}[1]
            \REQUIRE $\Omega,J,\text{maxVarNb}$
            \STATE $\Omega=\emptyset$
            \STATE $F=\text{\{all variables\}}$
            \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
            \FORALL{$f_i \in F$}
            \STATE $R_i = J(\{\Omega + f_i\})$
            \ENDFOR
            \STATE $j=\text{arg} \max_{i} R_i$
            \STATE $\Omega = \{\Omega + f_j\}$
            \STATE $F = F \setminus f_j$
            \ENDWHILE
            \RETURN $\Omega$
            \end{algorithmic}
            }
            \end{algorithm}

            \paragraph{Sequential floating forward feature selection}
            \label{sec:floating-presentation}

            The Sequential Floating Forward Selection (SFFS) is actually based on two algorithms: the SFS described above and the Sequential Backward Selection (SBS). The SBS is the backward equivalent of SFS. The difference is that it starts with every features in the pool of selected features and tries at each step to remove the less significant one in term of the given criterion function.

            The SFFS works as the SFS but between each step of the SFS algorithm a backward selection is operated. At the end of the SBS step, the value of the criterion function is compared to the best value ever obtained with a set of features of the same size and if the new value is better the feature put into question is effectively taken away and the next step is again a SBS but if the new value is not better the SBS step is forgotten and it moves to the next SFS step. The algorithm stops when a given number of features \emph{maxVarNb} has been selected. The Algorithm~\ref{alg:sffs} sums up the method.

            This SFFS algorithm eventually tests more solutions than the SFS algorithm. The results are expected to be better but the trade-off is an increased computational time which is dependent on the complexity of the dataset.

            \begin{algorithm}
            \caption{Sequential floating forward features selection\label{alg:sffs}}
            {\footnotesize
            \begin{algorithmic}[1]
            \REQUIRE $\Omega,J,\text{maxVarNb}$
            \STATE $\Omega=\emptyset$
            \STATE $F=\text{\{all variables\}}$
            \STATE $k=0$
            \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
            \FORALL{$f_i \in F$}
            \STATE $R_i = J(\{\Omega{(k)} + f_i\})$
            \ENDFOR
            \STATE $j=\text{arg} \max_{i} R_i$
            \IF{$R_j \geq J(\Omega^{(k+1)})$}
            \STATE $k=k+1$
            \ELSE
            \STATE $\Omega^{(k+1)} = \{\Omega{(k)} + f_i\}$
            \STATE $k=k+1$
            \STATE $\text{flag}=1$
            \WHILE{$k > 2 \text{ and } \text{flag}=1$}
            \FORALL{$f_i \in \Omega^{(k)}$}
            \STATE $R_i = J(\{\Omega{(k)}\setminus f_i\})$
            \ENDFOR
            \STATE $j=\text{arg} \max_{i} R_i$
            \IF{$R_j < J(\Omega^{(k-1)})$}
            \STATE $\Omega^{(k-1)} = \{\Omega{(k)} \setminus f_i\}$
            \STATE $k=k-1$
            \ELSE
            \STATE $\text{flag}=0$
            \ENDIF
            \ENDWHILE
            \ENDIF
            \ENDWHILE
            \RETURN $\Omega$
            \end{algorithmic}
            }
            \end{algorithm}

\section{Efficient implementation}
\label{sec:implementation}

    \subsection{Statistical update rules}
    A major contribution of this work is the optimization in term of computational efficiency. More precisely, three steps of the proposed method has been upgraded in order to reduce computational time. First of all, the GMM model is learned only once using samples. When a covariance matrix or a mean vector of a reduced set of variables is required, it is obtained from the global model learned at the beginning by marginalization.

    Secondly, when a cross-validation is performed, submodels, i.e. GMM model trained with $(n-1)$ folds, are not learned using the $(n-1)$ folds but, instead, are derived from the global model and from the covariance matrices and mean vectors of the fold used for validation for the given submodel. The process is described in details in Section~\ref{sec:update-cv}.

    Finally and most importantly, when variables are tested one by one during a selection step, costly operations are made when computing criterion functions for each variable especially the computation of the inverse of covariance matrices and its determinant. Several update rules are set up which allow to compute this inverse and determinant only once to test all variables. These update rules are presented in Section~\ref{sec:update-crit}.

        \subsubsection{Update for cross validation}
        \label{sec:update-cv}

        Based on \cite{fauvel2015fast}, a method to accelerate the $n_{cv}$-fold cross-validation process in the case of criterion functions based on correct classification measures was implemented. The idea is to estimate the GMM model with the whole training set and then, instead of training a model on $(n_{cv}-1)$ folds, the complete model and the mean vector and the covariance matrix of the $k^{th}$ fold is used to derive the corresponding submodel.

        The following formulae can be obtained
        \begin{prop}
            \label{eq:update-cv1}
            (Cross-validation mean update)
            \begin{equation*}
                \boldsymbol{\mu}_c^{n_c-\nu_c} = \frac{n_c \boldsymbol{\mu}_c^{n_c} - \nu_c \boldsymbol{\mu}_c^{\nu_c}}{n_c - \nu_c} \nonumber
            \end{equation*}
        \end{prop}
        \begin{prop}
            \label{eq:update-cv2}
            (Cross-validation covariance matrix update)
            \begin{align*}
                \boldsymbol{\Sigma}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl( (n_c-1) \boldsymbol{\Sigma}_c^{n_c} - (\nu_c-1) \boldsymbol{\Sigma}_c^{\nu_c} \nonumber \\
                &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})(\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})^t \biggr) \nonumber
            \end{align*}
        \end{prop}
        where $\nu_c$ is the number of samples of class $c$ removed from the initial set, $\boldsymbol{\mu}_c^{n_c}$ the mean vector estimated with this $\nu_c$ samples and $\boldsymbol{\Sigma}_c^{\nu_c}$ the covariance matrix estimated with this $\nu_c$ samples.

        \subsubsection{Criterion function computation}
        \label{sec:update-crit}

        As explained earlier, at each iteration the SFS and SFFS algorithms compute the value of a criterion function for every possible set of features composed of the selected ones augmented by one of the remaining features. One of the main achievements of this work is to reduce the computational time needed to compute the criterion function of augmented sets.

        In the remaining of the paper, $\boldsymbol{\Sigma}^{(k-1)}$ is the covariance matrix of the $(k-1)^{th}$ iteration, i.e., the covariance matrix of the selected features and $\boldsymbol{\Sigma}^{(k)}$ is a covariance matrix at the $k^{th}$ iteration, i.e., the covariance matrix of an augmented set. Then, the inverse of the covariance matrix $(\boldsymbol{\Sigma}^{(k)})^{-1}$, the quadratical term $(\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}$ and the determinant $\log |\boldsymbol{\Sigma}^{(k)}|$ can be expressed in function of terms of the $(k-1)^{th}$ iteration. These calculi use the fact that the covariance matrix is a positive definite symmetric matrix to simplify formulae using block matrices \cite{webb2003statistical} (chapter 9.2). These update rules are summed up hereafter

        In order to easily understand which terms are available and which ones needs to be computed, the terms already computed are set in green in the update rules.

        As $\boldsymbol{\Sigma}^{(k)}$ is a positive definite symmetric matrix, the following notation can be used
        \begin{equation*}
            \boldsymbol{\Sigma}^{(k)} =
            \bigg[\begin{array}{cc}
            \boldsymbol{\Sigma}^{(k-1)} & \mathbf{u}      \\
            \mathbf{u}^t          & \sigma_{kk} \\
            \end{array}\bigg],
        \end{equation*}
        where $\mathbf{u}$ is the $k^{th}$ column of the matrix without the diagonal element i.e. $\mathbf{u}_{i} = \boldsymbol{\Sigma}^{(k)}_{i,k}$ with $i \in [1,k-1]$.

        Using the formula of the inverse of a block matrix, the following formula expressing $(\boldsymbol{\Sigma}^{(k)})^{-1}$ in function of $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is obtained
        \begin{prop}
        \label{eq:update-inv}
            (Forward update rule for inverse of covariance matrix)
            \begin{equation*}
                \fbox{$
                (\boldsymbol{\Sigma}^{(k)})^{-1} =
                \bigg[\begin{array}{cc}
                A & v \\
                v^t  & \frac{1}{\alpha} \\
                \end{array}\bigg]
                $}
            \end{equation*}
        \end{prop}
        where $A = \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} + \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}}$, $v = - \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} $.

        It means that, to select a variable among one hundred, instead of computing one hundred times $(\boldsymbol{\Sigma}^{(k)})^{-1}$, $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is computed once and then with a few matrix multiplications and sums, the hundred inverses are obtained and a lot of processing time is saved.

        In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$. The update rule becomes
        \begin{prop}
            (Backward update rule for inverse of covariance matrix)
            \begin{equation*}
                \fbox{$(\boldsymbol{\Sigma}^{(k-1)})^{-1} = \textcolor[rgb]{0,0.5,0}{\mathbf{A}} - \alpha \mathbf{v} \mathbf{v}^t$}
            \end{equation*}
        \end{prop}

        \begin{proof}
            \begin{align*}
                \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
                &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
            \end{align*}
        \end{proof}

        A formula can also be deduced for the quadratical term of the decision function. Noting $(\mathbf{x}^{(k)})^t = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x^k \end{array}\right]$, the following proposition is obtained
        \begin{prop}
        \label{eq:update-quad}
            (Update rule for quadratical term)
            \begin{empheq}[box=\fbox]{align*}
                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\textcolor[rgb]{0,0.5,0}{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}} \\
                &+ \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
            \end{empheq}
        \end{prop}

        \begin{proof}
            \begin{align*}
                (&\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}
                = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
                \left[\begin{array}{cc}
                \mathbf{A}   & \mathbf{v} \\
                \mathbf{v}^t & \frac{1}{\alpha}
                \end{array}\right]
                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
                &= \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
                &= (\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \alpha \mathbf{v} \mathbf{v}^t) \mathbf{x}^{(k-1)}\\
                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} \mathbf{v}^t \mathbf{x}^{(k-1)} \\
                &~~~+ 2 \frac{x_k}{\alpha} \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha^2}) \\
                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} )^2 \\
                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
            \end{align*}
        \end{proof}

        Finally, using the formula of the determinant of a block matrix and after taking the log, a last formula is produced
        \begin{prop}
        \label{eq:update-log}
            (Update rule for logdet)
            \begin{equation*}
                \fbox{$\log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} + \log \alpha$}
            \end{equation*}
        \end{prop}

        Now using all update rules, criterion functions can be rewritten as follows
        \begin{align}
        \label{eq:q-update}
            Q(\mathbf{x}) = &- \textcolor[rgb]{0,0.5,0}{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})} \nonumber \\
            &- \alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 \nonumber \\
            &- \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} - \log \alpha  + \textcolor[rgb]{0,0.5,0}{2 \log (\pi_c)} + \textcolor[rgb]{0,0.5,0}{k \log(2 \pi)};
        \end{align}

        {\footnotesize
        \begin{align}
        \label{eq:skl-update}
            &\text{SKL}_{ij} = \frac{1}{2} \biggl( \text{Tr} \left((\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_j}^{(k)} + (\boldsymbol{\Sigma}_{c_j}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_i}^{(k)}\right) \nonumber \\
            &+ \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})^t ((\boldsymbol{\Sigma}_{c_i}^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_{c_j}^{(k-1)})^{-1}) (\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})} \nonumber \\
            &+ \alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2 \nonumber \\
            &+ \alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2
            - \textcolor[rgb]{0,0.5,0}{2k} \biggr),
        \end{align}
        }
        with $(\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};

        \begin{align}
        \label{eq:jm-update}
            \text{B}_{ij} = &\textcolor[rgb]{0,0.5,0}{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})} \nonumber \\
            &+ \frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 \nonumber \\
            &+ \textcolor[rgb]{0,0.5,0}{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)} + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right),
        \end{align}
        where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j$.

        The Algorithm~\ref{alg:sffs-update} illustrates the optimization of the Algorithm~\ref{alg:sffs} developed using these formulae.

    \subsection{Numerical issues}

    During an iteration of the selection algorithm, it still needs to invert the covariance matrix of the previous iteration. In order to so, a decomposition in eigenvalues and eigenvectors is performed using algorithm specific for symmetric matrices. More precisely, In other words, $\boldsymbol{\Sigma}$ is decomposed s.t. $\boldsymbol{\Sigma} = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$ with $\boldsymbol{\Lambda}$ the diagonal matrix of eigenvalues and $\mathbf{Q}$ the matrix of eigenvectors.

    This decomposition has several advantages. First, it is very simple to compute the determinant once the eigenvalues are available because the determinant is equal to the product of these values. Secondly, it is possible to check the eigenvalues and so to assure a better computational stability. Due to the curse of dimensionality, the covariance matrix is sometimes badly conditioned and the result is that some eigenvalues are really small (below computational precision). The decomposition helps to check their actual values and if they are below EPS\_FLT which is the machine maximum precision for a float, these eigenvalues are set to EPS\_FLT. For the same reason, the constant $\alpha$ used in update rules is also thresholded to EPS\_FLT. Algorithm~\ref{alg:sffs-update} details when computational stability is checked.

    \begin{algorithm}
    \caption{Sequential floating forward features selection with updates\label{alg:sffs-update}}
    {\footnotesize
    \begin{algorithmic}[1]
    \REQUIRE $\Omega,J,\text{maxVarNb}$
    \STATE $\Omega=\emptyset$
    \STATE $F=\text{\{all variables\}}$
    \STATE $k=0$
    \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
    \STATE {\bfseries for all} $\lambda_i$ {\bfseries do} $\lambda_i = \min (\text{EPS\_FLT},\lambda_i)$
    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
    \FORALL{$f_i \in F$}
    \STATE Compute update constant $\alpha$
    \STATE $\alpha = \min (\text{EPS\_FLT},\alpha)$
    \STATE $R_i = J(\{\Omega{(k)} + f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \IF{$R_j \geq J(\Omega^{(k+1)})$}
    \STATE $k=k+1$
    \ELSE
    \STATE $\Omega^{(k+1)} = \{\Omega{(k)} + f_i\}$
    \STATE $k=k+1$
    \STATE $\text{flag}=1$
    \WHILE{$k > 2 \text{ and } \text{flag}=1$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
    \STATE {\bfseries for all} $\lambda_i$ {\bfseries do} $\lambda_i = \min (\text{EPS\_FLT},\lambda_i)$
    \STATE Compute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
    \FORALL{$f_i \in \Omega^{(k)}$}
    \STATE Compute update constant $\alpha$
    \STATE $\alpha = \min (\text{EPS\_FLT},\alpha)$
    \STATE $R_i = J(\{\Omega{(k)}\setminus f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \IF{$R_j < J(\Omega^{(k-1)})$}
    \STATE $\Omega^{(k-1)} = \{\Omega{(k)} \setminus f_i\}$
    \STATE $k=k-1$
    \ELSE
    \STATE $\text{flag}=0$
    \ENDIF
    \ENDWHILE
    \ENDIF
    \ENDWHILE
    \RETURN $\Omega$
    \end{algorithmic}
    }
    \end{algorithm}

    \subsection{OTB external module}
    \label{sec:otb-module}

    The Orfeo Toolbox is an open-source library for remote sensing image processing. This library allows developers to code external modules and make them available to the community. It is to note that the toolbox already implements several classifiers inherited mostly from OpenCV. Nevertheless, the GMM classifier currently in place is not robust to high dimensionality and that is why a new GMM model has been developed.

    The developed module is available on Github \footnote{\url{https://github.com/Laadr/otbExternalFastFeaturesSelection}} and is a fork of the template for remote module furnished by OTB developers.

    The GMM classifier class called \emph{GMMMachineLearningModel} inherits from the OTB class \emph{MachineLearningModel} which is a basic class used for all classifier in the OTB. The \emph{MachineLearningModel} class enables the management of a list of samples used for training and also declares the virtual methods inherited: \emph{Train()}, \emph{Predict()}, \emph{Save()}, \emph{Load()}.

    The \emph{GMMMachineLearningModel} class implements all these virtual methods. The \emph{Train()} aims to train the classifier using a list of samples set as a class member. Then, it is possible to use the method \emph{Predict()} to classify a sample and optional get the confidence in the prediction. The methods \emph{Save()} and \emph{Load()} obviously are used to save and load a trained model.

    In addition to these inherited methods, two noticeable methods are implemented. The \emph{Decomposition()} method perform a decomposition of a symmetric matrix in eigenvalues and eigenvectors using a function of the VNL library. It is to notice that this method check the value of the extracted eigenvalues and minor them to EPSILON\_FLT (or EPSILON\_DBL) which corresponds to computational precision.  The idea is to limit computational instability when parameters are badly-estimated due to an insufficient amount of training samples. The second interesting method is \emph{TrainTau()} which concerns an optional ridge regularization. Given a set of possible regularization constant $\tau$, this method selects the most efficient value in estimating classification rate with cross-validation.

    To implement the selection algorithm, a new class \emph{GMMSelectionMachineLearningModel} inheriting from \emph{GMMMachineLearningModel} is defined. This class keeps the same six methods inherited from \emph{MachineLearningModel} but reimplements \emph{Predict()} to predict with a reduced set of features and  \emph{Save()}/\emph{Load()} to use a second file to handle the results of the selection.

    The other methods available in class \emph{GMMSelectionMachineLearningModel} are used to perform the selection algorithms presented in Section \ref{sec:gmm-hd}. The method \emph{Selection()} aims to set the various parameter of the selection and then calls either the \emph{ForwardSelection()} method to use forward feature selection or the \emph{FloatingForwardSelection()} method to use SFFS. Then, 3 different methods are used to evaluate criterion functions during selection. Given a pool of variables available for selection, these functions evaluate the criterion functions for all the possible augmented set. The \emph{ComputeJM()} method is used for Jeffries-Matusita distance, \emph{ComputeDivKL()} for Kullback–Leibler divergence and \emph{ComputeClassifRate()} for the three correct classification criteria. It is to notice that the correct classification criteria are evaluated using cross-validation.

    In order to provide an easy way to use the developed algorithm, OTB applications has been developed. These applications allow the user to use and parametrize the algorithm from command line. Three applications are built.

    The application \emph{otbcli\_TrainGMMApp} creates and trains a model from class \emph{GMMMachineLearningModel} and the application \emph{otbcli\_TrainGMMSelectionApp} does the same for a model of class \emph{GMMSelectionMachineLearningModel}. These two applications use a raster image and shapefile with groundtruth as input and create a model file.

    The last application \emph{otbcli\_PredictGMMApp} takes a raster image as input and a model file and perform classification of the image. It generates a classification map in an image file and optionally a confidence map. An other application is available in the OTB to compute classification rate from the image and a groundtruth file.

\section{Experimental results}
\label{sec:test}

    \subsection{Method}
    \label{sec:method}

    The aim of the experiments is to compare the proposed method to some standard classifiers. In order to do it, OTB commandline application are used. The following classifier are tested:
    \begin{itemize}
        \item a k-nearest-neighbors classifier (KNN) with OTB default parameters (32 number of neighbors);
        \item a Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 max depth, 50 size of subset of variables);
        \item a GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by gridsearch.
    \end{itemize}
    The GMM classifier is part of the external module described in Section~\ref{sec:otb-module}.

    All these classifiers are compared with 3 configurations of the proposed GMM classifier:
    \begin{itemize}
        \item one with forward selection and JM distance as criterion (GMM SFS JM);
        \item one with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
        \item one with floating forward selection and JM distance as criterion (GMM SFFS JM).
    \end{itemize}

    The results obtained with two different datasets are presented. The first dataset is an airborne hyperspectral dataset and second is a remote sensing dataset made of heterogeneous spatial features. Training set has been created with an equal number of samples for each class and additionally a spatial stratification has been performed. Several size of training set has been tested. For Aisa dataset, experiments have been conducted using 250, 500 and 1000 samples by class and for the Potsdam dataset, 1000 and 50000 samples by class.

    For SFS and SFFS selection, the number of variables to select is set to 30 for the Aisa datset and 60 for the Potsdam dataset and then the number of variables which maximizes the criterion function is used to perform classification.

    The classification rate is presented using Cohen's kappa but it has to be noticed the results were similar when looking at the mean F1-score. Processing time has been evaluated on a desktop computer with 4Gb of RAM and Intel(R) Core(TM) i5-3570 CPU \@ 3.40GHz x 4 processor. All results has been estimated with multiple repetitions (20 trials with Aisa dataset, 5 trials with Potsdam dataset).

    \subsection{Aisa dataset}
    \label{sec:aisa}
    The Aisa dataset has been acquired by the AISA Eagle sensor during a flight campaign over Heves, Hungary. It contains 252 bands ranging from 395 to 975 nm. 16 classes have been defined for a total of 361,971 referenced pixels. Figure~\ref{fig:aisa} shows a colored composition of the multispectral image and the shapefile representing groundtruth and Table \ref{tab:aisa} presents the repartition of the various classes.

    \begin{figure}[!t]
        \centering
        \begin{tabular}{c}
            \includegraphics[width=0.7\columnwidth]{Fig/aisa.png} \\
            {\bfseries{(a)}} \\
            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt.png} \\
            {\bfseries{(b)}} \\
        \end{tabular}
        \caption{Aisa dataset: {\bfseries{(a)}} colored composition of the image, {\bfseries{(b)}} groundtruth.\label{fig:aisa}}
    \end{figure}

    \begin{table}[!t]
        \centering
        \caption{Repartition of classes in Aisa dataset.\label{tab:aisa}}
        \begin{tabular}[b]{lc}\hline
          Class & Number of samples \\ \hline
          Winter wheat & 136,524 \\
          Sunflower & 61,517 \\
          Green fallow last year treatment & 30,197 \\
          Alfalfa & 17,626 \\
          Maize & 18,278 \\
          Millet & 7,199 \\
          Broadleaved forest & 10,746 \\
          Meadow & 23,283 \\
          Winter barley & 2,799 \\
          Reed & 4,222 \\
          Water course & 4,773 \\
          Rape & 26,566 \\
          Green fallow with shrub & 9,272 \\
          Green fallow last year treated & 3,426 \\
          Pasture & 2,107 \\
          Oat & 3,436 \\ \hline
        \end{tabular}
    \end{table}

    When creating training and validation sets, special care is taken to assure that training samples are picked in distinct areas than test samples. In order to do it, the polygons of the reference are split in smaller polygons and then 50\% of the polygons are taken randomly for training and the remaining 50\% for validation. Moreover 20 trials are run each time with a different training set (different polygons). An example of training and validation set is shown in Figure~\ref{fig:set-aisa}. Additionally, it is important to understand that only a subpart of this training set is actually used to train the model, a given number of samples for each class. Table~\ref{tab:aisa-otbsimu} presents the results of the experiment with mean and standard deviation over the 20 trials and Table~\ref{tab:aisa-otbsimu-time} the corresponding processing time.

    \begin{figure}[!t]
        \centering
        \begin{tabular}{c}
            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt_train.png} \\
            {\bfseries{(a)}} \\
            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt_test.png} \\
            {\bfseries{(b)}} \\
        \end{tabular}
        \caption{Aisa dataset: {\bfseries{(a)}} training polygons of first trial, {\bfseries{(b)}} test polygons of first trial.\label{fig:set-aisa}}
    \end{figure}

    \begin{table}[!t]
        \centering
        \caption{Results of classification with sampling on separate polygons and 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
        \begin{tabular}{|l|c|c|c|}\hline
             & \multicolumn{3}{c|}{\bfseries Cohen's kappa} \\ \hline
            \# samples by class & 250 & 500 & 1000 \\ \hline

            GMM SFS kappa & 0.684 (0.025) & 0.698 (0.027) & 0.711 (0.028) \\
            GMM SFS JM &    0.680 (0.028) & 0.700 (0.028) & 0.710 (0.030) \\
            GMM SFFS JM &   0.680 (0.028) & 0.700 (0.028) & 0.710 (0.030) \\
            GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
            KNN &           0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
            Random Forest & 0.645 (0.026) & 0.673 (0.023) & 0.693 (0.023) \\

            \hline
        \end{tabular}
    \end{table}

    \begin{table}[!t]
        \centering
        \caption{Mean processing time for training and classification for results of Table \ref{tab:aisa-otbsimu}.\label{tab:aisa-otbsimu-time}}
        \begin{tabularx}{\columnwidth}{|l|*{6}{>{\centering\arraybackslash}X|}}
            \hline
             & \multicolumn{3}{c|}{\bfseries Training time (s)} & \multicolumn{3}{c|}{\bfseries Classification time (s)} \\ \hline
            \# samples by class & 250 & 500 & 1000 & 250 & 500 & 1000 \\ \hline

            GMM SFS kappa & 257 & 496 & 955 & 7.7 & 8.6 & 8.7 \\
            GMM SFS JM &    8.6 & 8.9 & 9.1 & 9.7 & 9.8 & 9.6 \\
            GMM SFFS JM &   8.8 & 9.0 & 9.3 & 9.7 & 9.8 & 9.8 \\
            GMM ridge &     71.7 & 105 & 167 & 530 & 530 & 530 \\
            KNN &           8.9 & 19.6 & 59.7 & 387 & 639 & 887 \\
            Random Forest & 24.5 & 49.3 & 105 & 33.0 & 41.7 & 45.9 \\

            \hline
        \end{tabularx}
    \end{table}

    The results show that, on this dataset, GMM classifiers with selection get the best classification rate. Especially with very few training samples, GMM classifiers with selection seems to outperform the others. In term of computational time, the GMM classifiers are as expected very fast for classification and also for training if the criterion function is not a classification rate. The computation of the kappa seems to suffer from a lack of parallelization which could be improved. Among the three variations of the selection algorithm, none appears to perform better than the others. Using kappa or Jeffries-Matusita distance as criterion is equal and using SFFS does not give any advantage. In this case, using J-M distance as criterion and SFS as search strategy is the best choice in term of time efficiency.

    Nevertheless, good performance are also obtained with Random Forest which becomes more competitive when the number of samples increases.
    The KNN classifier and the GMM classifier with ridge regularization are totally outperform both in term of classification rate and of processing time.

    \subsection{Potsdam dataset}

    This second dataset is built from a dataset of remote sensing images distributed by the International Society for Photogrammetry and Remote Sensing (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}. The dataset is composed of aerial images of the urban area of Potsdam. The area is cut into 38 patches of 6000x6000 pixels with a resolution of 5cm by pixel and 4 channels are available: Red, Blue, Green and Infrared (RGBIR). A Digital Surface Model with same resolution is also provided and a so-called normalized DSM representing the height above ground. The groundtruth for 24 tiles are provided with 6 classes: Low vegetation, High vegetation, Impervious surfaces, Buildings, Cars, Clutter. Table~\ref{tab:potsdam} summarizes the number of samples of each class in the two images used during experimentations. Figure~\ref{fig:potsdam-dataset} shows the complete area and Figure~\ref{fig:potsdam-expl} a particular patch.

    \begin{figure}[!t]
        \centering
        \includegraphics[width=0.7\columnwidth]{Fig/potsdam_full.png}
        \caption{Full Potsdam dataset.\label{fig:potsdam-dataset}}
    \end{figure}

    \begin{figure}[!t]
        \centering
        \includegraphics[width=0.8\columnwidth]{Fig/potsdam_imgexpl.png}
        \caption{Example of patch with {\bfseries{(a)}} orthophoto, {\bfseries{(b)}} DSM and {\bfseries{(c)}} groundtruth.\label{fig:potsdam-expl}}
    \end{figure}

    \begin{table}[!t]
        \centering
        \caption{Repartition of classes in Aisa dataset.\label{tab:potsdam}}
        \begin{tabular}[b]{lrr}\hline
          Class & Number of samples & Number of samples \\
                & in 5\_11          & in 5\_12 \\
          \hline
          Clutter             & 1,078,611  & 812,038 \\
          Trees               & 4,493,295  & 2,132,368 \\
          Cars                & 900,076    & 1,101,541 \\
          Buildings           & 13,469,575 & 17,501,421 \\
          Low vegetation      & 4,718,219  & 3,210,596 \\
          Impervious surfaces & 11,340,224 & 11,242,036 \\
          \hline
        \end{tabular}
    \end{table}

    In the experiment, the following features are computed using the RGBIR images similar to \cite{tuia2015multiclass}:
    \begin{itemize}
        \item Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2 (15 features) \footnote{\url{https://www.orfeo-toolbox.org//Applications/RadiometricIndices.html}};
        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)\cite{tuia2015multiclass};
        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)\cite{tuia2015multiclass};
        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)\cite{tuia2015multiclass};
        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features)\cite{tuia2015multiclass}.
    \end{itemize}
    The normalized DSM and the raw RGBIR image are added to these 191 features and then then stacked to create a new image with 196 bands. The training is made with tile 5\_11 and test with tile 5\_12. Table~\ref{tab:potsdam-otbsimu} presents the results.

    \begin{table}[!t]
        \centering
        \caption{Results of classification with 1000 samples by class and 5 trials (standard deviation in parenthesis).\label{tab:potsdam-otbsimu}}
        \resizebox{\columnwidth}{!}{%
        \begin{tabular}{|l|c|c|c|c|c|}\hline
             & {\bfseries Kappa of} & {\bfseries Kappa of} & {\bfseries Training} & {\bfseries Classif.} & {\bfseries \# features} \\
             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries time (s)} & {\bfseries time (s)} & {\bfseries to classify} \\ \hline

            GMM SFS kappa & 0.694 (0.002) & 0.669 (0.005) & 400 & 310 & 13.2 \\
            GMM SFS JM &    0.624 (0.028) & 0.631 (0.034) & 2 & 310   & 11 \\
            GMM SFFS JM &   0.624 (0.028) & 0.631 (0.034) & 2.6 & 310   & 11 \\
            GMM ridge &     0.632 (0.007) & 0.592 (0.010) & 10 & 2000 & all \\
            KNN &           0.637 (0.005) & 0.607 (0.005) &  &   & all \\
            Random Forest & 0.729 (0.004) & 0.673 (0.005) & 20 & 840  & all \\

            \hline
        \end{tabular}
        }
    \end{table}

    \begin{table}[!t]
        \centering
        \caption{Results of classification with 50000 samples by class and 5 trials (standard deviation in parenthesis).\label{tab:potsdam-otbsimu-big}}
        \resizebox{\columnwidth}{!}{%
        \begin{tabular}{|l|c|c|c|c|c|}\hline
             & {\bfseries Kappa of} & {\bfseries Kappa of} & {\bfseries Training} & {\bfseries Classif.} & {\bfseries \# features} \\
             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries time (s)} & {\bfseries time (s)} & {\bfseries to classify} \\ \hline

            GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 20000 & 340 & 29 \\
            GMM SFS JM &    0.560 (0.111) & 0.576 (0.104) & 6 & 330 & 10 \\
            GMM SFFS JM &   0.560 (0.111) & 0.576 (0.104) & 6.6 & 340 & 10 \\
            GMM ridge &     0.641 (0.015) & 0.611 (0.026) & 460 & 2000 & all \\
            Random Forest & 0.851 (0.001) & 0.715 (0.001) & 2000 & 2000 & all \\

            \hline
        \end{tabular}
        }
    \end{table}

    In this case, the classification rate of GMM classifiers with Jeffries-Matusita distance as criterion function drops clearly and has a high standard deviation. It could be due to the fact that the Gaussian hypothesis, on which this metric relies a lot, is not enough verified. When the classification rate is directly used as criterion, the classifier manages to select the relevant features. It has to be noticed that with kappa only 30 features are selected on average when with Jeffries-Matusita distance, the maximum number of features is selected almost always (set to 60).

    The KNN classifier and the GMM classifier with ridge regularization are again outperformed even if they get stable results.

    Finally, the Random Forest classifier and the GMM with kappa as criterion are from far the best classifiers. The only difference is in the processing time and also a small difference of classification rate with the training image. Random Forest is faster for the training process while the GMM classifier is faster for classifying.

\section{Conclusion}

An algorithm for the classification of high dimension data as hyperspectral image has been presented. The classifier uses GMM model to select iteratively the most relevant features. Several variation of the algorithms has been explored. Experimentations show that allowing backward step in the selection to discard already selected features do not give significant advantages.

Additionally, from the comparison between all the measures used to rank the features, it has been shown that the Jeffries-Matusita distance can be a fast and accurate solution in some case but suffer from some limitations which remain to be determine. A possible explaination could be that this distance does not perform well when data distribution is not similar to a Gaussian distribution. More investigation is needed to confirm this hypothesis. In any case, it is possible to use directly kappa as criterion function even if it implies a loss of time.

Finally, experiments show that the developed GMM classifier performs at least as best as standard classifiers without features selection in particular Random Forest and even outperforms all of them in term of classification time. The next step would be to compare to other feature selection methods.

The second achievement of this work is the development of an efficient implementation using GMM properties to derive fast update rules of the model. The resulting code is available as a remote module of the Orfeo toolbox on Github and make it possible to process huge quantity of high dimension data.

Finally, an improvement could be made to increase the stability of the selected features. For example, with hyperspectral data, a selection of continuous intervals and not band is a possible solution which has already been explored in \cite{serpico2007extraction}.

The python and C++ code are available freely for download: \url{https://github.com/Laadr/FFFS}, \url{https://github.com/Laadr/otbExternalFastFeaturesSelection}.

% \appendices
% \section{Proof of the First Zonklar Equation}
% Appendix one text goes here.

% \section*{Acknowledgment}

% The authors would like to thank...

\bibliographystyle{IEEEtran}
\bibliography{biblio}

% biography section
%
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\end{document}

%\begin{figure}[!t]

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

### 2016-09-26 paper.tex
update

@@
                   \hline
-                  Criterion & Divergence & Classification rate & Gaussian hypothesis & Speed \\
+                  Criterion & Divergence & Classification & Gaussian   & Speed \\
+                            &            & rate           & hypothesis &       \\
                   \hline
@@
 
-            The Sequential Floating Forward Selection (SFFS) is actually based on two algorithms: the SFS described above and the Sequential Backward Selection (SBS).
+            The Sequential Floating Forward Selection (SFFS)\cite{somol1999adaptive} is actually based on two algorithms: the SFS described above and the Sequential Backward Selection (SBS).
 The SBS is the backward equivalent of SFS.
@@
     \subsection{Statistical update rules}
+    In the presented method, the most costly operation is the evaluation of the criterion function which needs to be done for each possible subset of variables augmented by one of the candidate variable.
+To compute these values, the most costly operation are the inversion of the covariance matrix and the computation of its determinant.
+Both of these operations have a $O(n^3)$ computational complexity.
+Additionaly, the estimation of classification rate with cross-validation is a heavy process which take a consequent amount of time.
+
     A major contribution of this work is the optimization in term of computational efficiency.
@@
     Secondly, when a cross-validation is performed, submodels, i.e.
-GMM model trained with $(n-1)$ folds, are not learned using the $(n-1)$ folds but, instead, are derived from the global model and from the covariance matrices and mean vectors of the fold used for validation for the given submodel.
+GMM model trained with $(n_{cv}-1)$ folds, are not learned using the $(n_{cv}-1)$ folds but, instead, are derived from the global model and from the covariance matrices and mean vectors of the fold used for validation for the given submodel.
 The process is described in details in Section~\ref{sec:update-cv}.
@@
         Based on \cite{fauvel2015fast}, a method to accelerate the $n_{cv}$-fold cross-validation process in the case of criterion functions based on correct classification measures was implemented.
-The idea is to estimate the GMM model with the whole training set and then, instead of training a model on $(n_{cv}-1)$ folds, the complete model and the mean vector and the covariance matrix of the $k^{th}$ fold is used to derive the corresponding submodel.
+The idea is to estimate the GMM model with the whole training set once and then, instead of training models on $(n_{cv}-1)$ folds, parameters of the complete model are used to derive those of submodels.
 
@@
         \end{prop}
-        where $\nu_c$ is the number of samples of class $c$ removed from the initial set, $\boldsymbol{\mu}_c^{n_c}$ the mean vector estimated with this $\nu_c$ samples and $\boldsymbol{\Sigma}_c^{\nu_c}$ the covariance matrix estimated with this $\nu_c$ samples.
+        where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exposants on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\nu}_c$ denotes the set of samples used to compute them.
 
@@
 
-        As explained earlier, at each iteration the SFS and SFFS algorithms compute the value of a criterion function for every possible set of features composed of the selected ones augmented by one of the remaining features.
-One of the main achievements of this work is to reduce the computational time needed to compute the criterion function of augmented sets.
+        At each iteration the SFS and SFFS algorithms compute the value of a criterion function for every possible set of features composed of the selected ones augmented by one of the remaining features.
+One of the main achievements of this work is to reduce the computational time in factorizing the computation of the inverse and determinant of the covariance matrix..
 
@@
 These calculi use the fact that the covariance matrix is a positive definite symmetric matrix to simplify formulae using block matrices \cite{webb2003statistical} (chapter 9.2).
-These update rules are summed up hereafter
-
-        In order to easily understand which terms are available and which ones needs to be computed, the terms already computed are set in green in the update rules.
+These update rules are summed up hereafter.
 

### 2016-09-26 paper.tex
update

@@
             \begin{equation*}
-                \fbox{$
                 (\boldsymbol{\Sigma}^{(k)})^{-1} =
@@
                 \end{array}\bigg]
-                $}
             \end{equation*}
         \end{prop}
-        where $A = \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} + \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}}$, $v = - \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} $.
+        where $A = \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} + \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}}$, $v = - \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$.
 
@@
         In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$.
-The update rule becomes
+The update rule becomes (proof in Appendix~\ref{app:proof-update})
         \begin{prop}
+        \label{eq:update-inv-back}
             (Backward update rule for inverse of covariance matrix)
             \begin{equation*}
-                \fbox{$(\boldsymbol{\Sigma}^{(k-1)})^{-1} = \textcolor[rgb]{0,0.5,0}{\mathbf{A}} - \alpha \mathbf{v} \mathbf{v}^t$}
+                (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \textcolor[rgb]{0,0.5,0}{\mathbf{A}} - \alpha \mathbf{v} \mathbf{v}^t
             \end{equation*}
@@
 
-        \begin{proof}
-            \begin{align*}
-                \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
-                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
-                &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
-                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
-            \end{align*}
-        \end{proof}
-
         A formula can also be deduced for the quadratical term of the decision function.
-Noting $(\mathbf{x}^{(k)})^t = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x^k \end{array}\right]$, the following proposition is obtained
+Noting $(\mathbf{x}^{(k)})^t = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x^k \end{array}\right]$, the following proposition is obtained (proof in Appendix~\ref{app:proof-update})
         \begin{prop}
@@
             (Update rule for quadratical term)
-            \begin{empheq}[box=\fbox]{align*}
+            \begin{align*}
                 (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\textcolor[rgb]{0,0.5,0}{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}} \\
                 &+ \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
-            \end{empheq}
+            \end{align*}
         \end{prop}
-
-        \begin{proof}
-            \begin{align*}
-                (&\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}
-                = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
-                \left[\begin{array}{cc}
-                \mathbf{A}   & \mathbf{v} \\
-                \mathbf{v}^t & \frac{1}{\alpha}
-                \end{array}\right]
-                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
-                &= \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
-                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
-                &= (\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
-                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
-                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
-                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \alpha \mathbf{v} \mathbf{v}^t) \mathbf{x}^{(k-1)}\\
-                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
-                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} \mathbf{v}^t \mathbf{x}^{(k-1)} \\
-                &~~~+ 2 \frac{x_k}{\alpha} \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha^2}) \\
-                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} )^2 \\
-                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
-            \end{align*}
-        \end{proof}
 
@@
             \begin{equation*}
-                \fbox{$\log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} + \log \alpha$}
+                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} + \log \alpha
             \end{equation*}
@@
 
-% \appendices
-% \section{Proof of the First Zonklar Equation}
-% Appendix one text goes here.
+\appendices
+\section{}
+\label{app:proof-update}
+\subsection{Proof of update rules}
+    \subsubsection{Proposition~\ref{eq:update-inv-back}}
+        \begin{proof}
+            \begin{align*}
+                \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
+                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
+                &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
+                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
+            \end{align*}
+        \end{proof}
+
+    \subsubsection{Proposition~\ref{eq:update-quad}}
+        \begin{proof}
+            \begin{align*}
+                (&\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}
+                = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
+                \left[\begin{array}{cc}
+                \mathbf{A}   & \mathbf{v} \\
+                \mathbf{v}^t & \frac{1}{\alpha}
+                \end{array}\right]
+                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
+                &= \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
+                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
+                &= (\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
+                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
+                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
+                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \alpha \mathbf{v} \mathbf{v}^t) \mathbf{x}^{(k-1)}\\
+                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
+                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} \mathbf{v}^t \mathbf{x}^{(k-1)} \\
+                &~~~+ 2 \frac{x_k}{\alpha} \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha^2}) \\
+                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} )^2 \\
+                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
+            \end{align*}
+        \end{proof}
 

### 2016-09-27 paper.tex
update

@@
             \begin{algorithmic}[1]
-            \REQUIRE $\Omega,J,\text{maxVarNb}$
-            \STATE $\Omega=\emptyset$
+            \REQUIRE $J,\text{maxVarNb}$
+            \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
             \STATE $F=\text{\{all variables\}}$
             \STATE $k=0$
-            \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
+            \WHILE{$k \neq \text{maxVarNb}$}
             \FORALL{$f_i \in F$}
-            \STATE $R_i = J(\{\Omega{(k)} + f_i\})$
+            \STATE $R_i = J(\{\Omega_k + f_i\})$
             \ENDFOR
             \STATE $j=\text{arg} \max_{i} R_i$
-            \IF{$R_j \geq J(\Omega^{(k+1)})$}
             \STATE $k=k+1$
-            \ELSE
-            \STATE $\Omega^{(k+1)} = \{\Omega{(k)} + f_i\}$
-            \STATE $k=k+1$
+            \IF{$R_j \geq J(\Omega_k)$}
+            \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
             \STATE $\text{flag}=1$
             \WHILE{$k > 2 \text{ and } \text{flag}=1$}
-            \FORALL{$f_i \in \Omega^{(k)}$}
-            \STATE $R_i = J(\{\Omega{(k)}\setminus f_i\})$
+            \FORALL{$f_i \in \Omega_k$}
+            \STATE $R_i = J(\{\Omega_k \setminus f_i\})$
             \ENDFOR
             \STATE $j=\text{arg} \max_{i} R_i$
-            \IF{$R_j < J(\Omega^{(k-1)})$}
-            \STATE $\Omega^{(k-1)} = \{\Omega{(k)} \setminus f_i\}$
+            \IF{$R_j > J(\Omega^{(k-1)})$}
+            \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
             \STATE $k=k-1$
@@
             \ENDWHILE
-            \RETURN $\Omega$
+            \RETURN $\Omega_{\text{maxVarNb}}$
             \end{algorithmic}
@@
         \end{prop}
-        where $A = \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} + \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}}$, $v = - \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$.
-
-        It means that, to select a variable among one hundred, instead of computing one hundred times $(\boldsymbol{\Sigma}^{(k)})^{-1}$, $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is computed once and then with a few matrix multiplications and sums, the hundred inverses are obtained and a lot of processing time is saved.
+        where $A = (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $v = - \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$.
+Then, $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is computed only once and this update update formulae is used to obtain all the $(\boldsymbol{\Sigma}^{(k)})^{-1}$ corresponding to all the possible augmented set of a given selection iteration.
 
@@
             \begin{equation*}
-                (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \textcolor[rgb]{0,0.5,0}{\mathbf{A}} - \alpha \mathbf{v} \mathbf{v}^t
+                (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \underbrace{\mathbf{A}}_{\substack{\text{computed once}\\ \text{per selection step}}} - \underbrace{\alpha \mathbf{v} \mathbf{v}^t}_{\substack{\text{computed for} \\ \text{each augmented set}}}
             \end{equation*}
@@
             \begin{align*}
-                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\textcolor[rgb]{0,0.5,0}{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}} \\
-                &+ \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
+                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once}\\ \text{per selection step}}} \\
+                &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}}
             \end{align*}
@@
             \begin{equation*}
-                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} + \log \alpha
+                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)}_{\substack{\text{computed once}\\ \text{per selection step}}} + \underbrace{\log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}}
             \end{equation*}
@@
         \label{eq:q-update}
-            Q(\mathbf{x}) = &- \textcolor[rgb]{0,0.5,0}{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})} \nonumber \\
-            &- \alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 \nonumber \\
-            &- \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} - \log \alpha  + \textcolor[rgb]{0,0.5,0}{2 \log (\pi_c)} + \textcolor[rgb]{0,0.5,0}{k \log(2 \pi)};
+            Q(\mathbf{x}) = &- \underbrace{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
+            &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
+            &- \underbrace{\alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 - \log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}};
         \end{align}
 
-        {\footnotesize
         \begin{align}
         \label{eq:skl-update}
-            &\text{SKL}_{ij} = \frac{1}{2} \biggl( \text{Tr} \left((\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_j}^{(k)} + (\boldsymbol{\Sigma}_{c_j}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_i}^{(k)}\right) \nonumber \\
-            &+ \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})^t ((\boldsymbol{\Sigma}_{c_i}^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_{c_j}^{(k-1)})^{-1}) (\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})} \nonumber \\
-            &+ \alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2 \nonumber \\
-            &+ \alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2
-            - \textcolor[rgb]{0,0.5,0}{2k} \biggr),
+            &\text{SKL}_{ij} = \frac{1}{2} \biggl( \underbrace{\text{Tr} \left((\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_j}^{(k)} + (\boldsymbol{\Sigma}_{c_j}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_i}^{(k)}\right)}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
+            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
+            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2 - 2k}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
+            &+ \underbrace{(\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})^t ((\boldsymbol{\Sigma}_{c_i}^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_{c_j}^{(k-1)})^{-1}) (\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \biggr),
         \end{align}
-        }
         with $(\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
@@
         \label{eq:jm-update}
-            \text{B}_{ij} = &\textcolor[rgb]{0,0.5,0}{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})} \nonumber \\
-            &+ \frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 \nonumber \\
-            &+ \textcolor[rgb]{0,0.5,0}{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)} + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right),
+            \text{B}_{ij} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
+            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
+            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right)}_{\substack{\text{computed for} \\ \text{each augmented set}}},
         \end{align}
@@
     \begin{algorithmic}[1]
-    \REQUIRE $\Omega,J,\text{maxVarNb}$
-    \STATE $\Omega=\emptyset$
+    \REQUIRE $J,\text{maxVarNb}$
+    \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
     \STATE $F=\text{\{all variables\}}$
     \STATE $k=0$
-    \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
+    \WHILE{$k \neq \text{maxVarNb}$}
     \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
-    \STATE {\bfseries for all} $\lambda_i$ {\bfseries do} $\lambda_i = \min (\text{EPS\_FLT},\lambda_i)$
+    \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
     \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
@@
     \STATE Compute update constant $\alpha$
-    \STATE $\alpha = \min (\text{EPS\_FLT},\alpha)$
-    \STATE $R_i = J(\{\Omega{(k)} + f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
+    \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
+    \STATE $R_i = J(\{\Omega_k + f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
     \ENDFOR
     \STATE $j=\text{arg} \max_{i} R_i$
-    \IF{$R_j \geq J(\Omega^{(k+1)})$}
     \STATE $k=k+1$
-    \ELSE
-    \STATE $\Omega^{(k+1)} = \{\Omega{(k)} + f_i\}$
-    \STATE $k=k+1$
+    \IF{$R_j \geq J(\Omega_k)$}
+    \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
     \STATE $\text{flag}=1$
@@
     \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
-    \STATE {\bfseries for all} $\lambda_i$ {\bfseries do} $\lambda_i = \min (\text{EPS\_FLT},\lambda_i)$
+    \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
     \STATE Compute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
-    \FORALL{$f_i \in \Omega^{(k)}$}
+    \FORALL{$f_i \in \Omega_k$}
     \STATE Compute update constant $\alpha$
-    \STATE $\alpha = \min (\text{EPS\_FLT},\alpha)$
-    \STATE $R_i = J(\{\Omega{(k)}\setminus f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
+    \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
+    \STATE $R_i = J(\{\Omega_k \setminus f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
     \ENDFOR
     \STATE $j=\text{arg} \max_{i} R_i$
-    \IF{$R_j < J(\Omega^{(k-1)})$}
-    \STATE $\Omega^{(k-1)} = \{\Omega{(k)} \setminus f_i\}$
+    \IF{$R_j > J(\Omega_{k-1})$}
+    \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
     \STATE $k=k-1$
@@
     \ENDWHILE
-    \RETURN $\Omega$
+    \RETURN $\Omega_{\text{maxVarNb}}$
     \end{algorithmic}

### 2016-09-27 paper.tex
update

@@
             \STATE $j=\text{arg} \max_{i} R_i$
-            \IF{$R_j > J(\Omega^{(k-1)})$}
+            \IF{$R_j > J(\Omega_{k-1})$}
             \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
@@
 
-    During an iteration of the selection algorithm, it still needs to invert the covariance matrix of the previous iteration.
-In order to so, a decomposition in eigenvalues and eigenvectors is performed using algorithm specific for symmetric matrices.
-More precisely, In other words, $\boldsymbol{\Sigma}$ is decomposed s.t.
+    During the selection process, numerical issues needs to be handled due to the computation of the inverse and determinant of the covariance matrix.
+As pointed out before, the lack of training samples can induce a badly-conditioned matrix with very small eigenvalues.
+Even with a well-estimated covariance matrix, it is still possible to encounter very correlated features resulting also in very low eigenvalues.
+The issue with these very small eigenvalues is that they reach the computational precision allowed by the computer especially when computing the logarithm of the determinant.
+
+    In order to deal with this limitation, the choice has been made to perform an explicit decomposition in eigenvalues and eigenvectors of the covariance matrix s.t.
 $\boldsymbol{\Sigma} = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$ with $\boldsymbol{\Lambda}$ the diagonal matrix of eigenvalues and $\mathbf{Q}$ the matrix of eigenvectors.
-
-    This decomposition has several advantages.
-First, it is very simple to compute the determinant once the eigenvalues are available because the determinant is equal to the product of these values.
-Secondly, it is possible to check the eigenvalues and so to assure a better computational stability.
-Due to the curse of dimensionality, the covariance matrix is sometimes badly conditioned and the result is that some eigenvalues are really small (below computational precision).
-The decomposition helps to check their actual values and if they are below EPS\_FLT which is the machine maximum precision for a float, these eigenvalues are set to EPS\_FLT.
+The advantage of an explicit decomposition is the possibility to check the eigenvalues and, if some appears to be below computational precision, to threshold them at EPS\_FLT which is the computer precision for a floating number.
 For the same reason, the constant $\alpha$ used in update rules is also thresholded to EPS\_FLT.
@@
 
+    The developed method has been made available as a remote module of Orfeo toolbox called \emph{otbFFSforGMM} and can be download on Github\footnote{\url{https://github.com/Laadr/otbExternalFastFeaturesSelection}}.
+
     The Orfeo Toolbox is an open-source library for remote sensing image processing.
@@
 
-    The developed module is available on Github \footnote{\url{https://github.com/Laadr/otbExternalFastFeaturesSelection}} and is a fork of the template for remote module furnished by OTB developers.
-
-    The GMM classifier class called \emph{GMMMachineLearningModel} inherits from the OTB class \emph{MachineLearningModel} which is a basic class used for all classifier in the OTB.
-The \emph{MachineLearningModel} class enables the management of a list of samples used for training and also declares the virtual methods inherited: \emph{Train()}, \emph{Predict()}, \emph{Save()}, \emph{Load()}.
-
-    The \emph{GMMMachineLearningModel} class implements all these virtual methods.
-The \emph{Train()} aims to train the classifier using a list of samples set as a class member.
-Then, it is possible to use the method \emph{Predict()} to classify a sample and optional get the confidence in the prediction.
-The methods \emph{Save()} and \emph{Load()} obviously are used to save and load a trained model.
-
-    In addition to these inherited methods, two noticeable methods are implemented.
-The \emph{Decomposition()} method perform a decomposition of a symmetric matrix in eigenvalues and eigenvectors using a function of the VNL library.
-It is to notice that this method check the value of the extracted eigenvalues and minor them to EPSILON\_FLT (or EPSILON\_DBL) which corresponds to computational precision.
-The idea is to limit computational instability when parameters are badly-estimated due to an insufficient amount of training samples.
-The second interesting method is \emph{TrainTau()} which concerns an optional ridge regularization.
-Given a set of possible regularization constant $\tau$, this method selects the most efficient value in estimating classification rate with cross-validation.
-
-    To implement the selection algorithm, a new class \emph{GMMSelectionMachineLearningModel} inheriting from \emph{GMMMachineLearningModel} is defined.
-This class keeps the same six methods inherited from \emph{MachineLearningModel} but reimplements \emph{Predict()} to predict with a reduced set of features and  \emph{Save()}/\emph{Load()} to use a second file to handle the results of the selection.
-
-    The other methods available in class \emph{GMMSelectionMachineLearningModel} are used to perform the selection algorithms presented in Section \ref{sec:gmm-hd}.
-The method \emph{Selection()} aims to set the various parameter of the selection and then calls either the \emph{ForwardSelection()} method to use forward feature selection or the \emph{FloatingForwardSelection()} method to use SFFS.
-Then, 3 different methods are used to evaluate criterion functions during selection.
-Given a pool of variables available for selection, these functions evaluate the criterion functions for all the possible augmented set.
-The \emph{ComputeJM()} method is used for Jeffries-Matusita distance, \emph{ComputeDivKL()} for Kullback–Leibler divergence and \emph{ComputeClassifRate()} for the three correct classification criteria.
-It is to notice that the correct classification criteria are evaluated using cross-validation.
-
-    In order to provide an easy way to use the developed algorithm, OTB applications has been developed.
-These applications allow the user to use and parametrize the algorithm from command line.
-Three applications are built.
-
-    The application \emph{otbcli\_TrainGMMApp} creates and trains a model from class \emph{GMMMachineLearningModel} and the application \emph{otbcli\_TrainGMMSelectionApp} does the same for a model of class \emph{GMMSelectionMachineLearningModel}.
-These two applications use a raster image and shapefile with groundtruth as input and create a model file.
-
-    The last application \emph{otbcli\_PredictGMMApp} takes a raster image as input and a model file and perform classification of the image.
+    In addition of the C++ class directly available to develop C++ code, three command line application has been produced.
+These applications allow the user to use and parametrize the algorithm as desire.
+
+    The first application \emph{otbcli\_TrainGMMApp} trains a GMM classifier from multiple pairs of images and training vector data and creates a model file.
+It also allows to perform ridge regularization with an embedded selection of the regularization parameter among a list of proposed values.
+The regularization values is chosen by maximizing a criterion function (Cohen's kappa, overall accuracy or mean of F1 scores) estimated by cross-validation.
+
+    The second application \emph{otbcli\_TrainGMMSelectionApp} also trains a GMM classifier and in addition perform a feature selection algorithm (SFS, SFFS) with various possible criterion (Jeffries-Matusita dist., Kullback-Lieber div., Cohen’s kappa, overall accuracy, mean F1-score).
+
+    The third application \emph{otbcli\_PredictGMMApp} takes a raster image as input and a model file and perform the classification of the image.
 It generates a classification map in an image file and optionally a confidence map.
-An other application is available in the OTB to compute classification rate from the image and a groundtruth file.
+It also allows to precise manually the number of features to use for prediction if the number automatically set does not fit the user requirement (evolution of criterion function is available in model file).
 

### 2016-09-27 paper.tex
update

@@
 
-\section{Experimental results}
-\label{sec:test}
-
-    \subsection{Method}
-    \label{sec:method}
-
-    The aim of the experiments is to compare the proposed method to some standard classifiers. In order to do it, OTB commandline application are used. The following classifier are tested:
-    \begin{itemize}
-        \item a k-nearest-neighbors classifier (KNN) with OTB default parameters (32 number of neighbors);
-        \item a Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 max depth, 50 size of subset of variables);
-        \item a GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by gridsearch.
-    \end{itemize}
-    The GMM classifier is part of the external module described in Section~\ref{sec:otb-module}.
-
-    All these classifiers are compared with 3 configurations of the proposed GMM classifier:
-    \begin{itemize}
-        \item one with forward selection and JM distance as criterion (GMM SFS JM);
-        \item one with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
-        \item one with floating forward selection and JM distance as criterion (GMM SFFS JM).
-    \end{itemize}
-
-    The results obtained with two different datasets are presented.
-The first dataset is an airborne hyperspectral dataset and second is a remote sensing dataset made of heterogeneous spatial features.
-Training set has been created with an equal number of samples for each class and additionally a spatial stratification has been performed.
-Several size of training set has been tested.
-For Aisa dataset, experiments have been conducted using 250, 500 and 1000 samples by class and for the Potsdam dataset, 1000 and 50000 samples by class.
-
-    For SFS and SFFS selection, the number of variables to select is set to 30 for the Aisa datset and 60 for the Potsdam dataset and then the number of variables which maximizes the criterion function is used to perform classification.
-
-    The classification rate is presented using Cohen's kappa but it has to be noticed the results were similar when looking at the mean F1-score.
-Processing time has been evaluated on a desktop computer with 4Gb of RAM and Intel(R) Core(TM) i5-3570 CPU \@ 3.40GHz x 4 processor.
-All results has been estimated with multiple repetitions (20 trials with Aisa dataset, 5 trials with Potsdam dataset).
+\section{Datasets}
+\label{sec:datasets}
+Experimentation has been conducted on two different datasets.
+The first dataset, called Aisa, is an airborne hyperspectral dataset and second, called Potsdam, is a remote sensing dataset made of heterogeneous spatial features.
 
     \subsection{Aisa dataset}
-    \label{sec:aisa}
+    \label{sec:aisa-dataset}
     The Aisa dataset has been acquired by the AISA Eagle sensor during a flight campaign over Heves, Hungary.
@@
 
+    \subsection{Potsdam dataset}
+    \label{sec:pots-dataset}
+
+    This second dataset is built from a dataset of remote sensing images distributed by the International Society for Photogrammetry and Remote Sensing (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}.
+The dataset is composed of aerial images of the urban area of Potsdam.
+The area is cut into 38 patches of 6000x6000 pixels with a resolution of 5cm by pixel and 4 channels are available: Red, Blue, Green and Infrared (RGBIR).
+A Digital Surface Model with same resolution is also provided and a so-called normalized DSM representing the height above ground.
+The groundtruth for 24 tiles are provided with 6 classes: Low vegetation, High vegetation, Impervious surfaces, Buildings, Cars, Clutter.
+Table~\ref{tab:potsdam} summarizes the number of samples of each class in the three images used during experimentations and Figure~\ref{fig:potsdam-expl} the data itself.
+
+    \begin{figure}[!t]
+        \centering
+        \begin{tabular}{ccc}
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_RGB.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_RGB.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_RGB.jpg} \\
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_label.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_label.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_label.jpg} \\
+            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_11_normalized_lastools.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_12_normalized_lastools.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_03_10_normalized_lastools.jpg} \\
+            {\bfseries{(a)}} & {\bfseries{(b)}}  & {\bfseries{(c)}}\\
+        \end{tabular}
+        \caption{From top to bottom, image RGB, reference and normalized DSM of: {\bfseries{(a)}} tile 5\_11, {\bfseries{(b)}} tile 5\_12 and {\bfseries{(c)}} tile 3\_10.\label{fig:potsdam-expl}}
+    \end{figure}
+
+    \begin{table}[!t]
+        \centering
+        \caption{Repartition of classes in Aisa dataset.\label{tab:potsdam}}
+        \begin{tabular}[b]{lrrr}\hline
+          Class & Number of samples & Number of samples & Number of samples \\
+                & in 5\_11          & in 5\_12          & in 3\_10 \\
+          \hline
+          Clutter             & 1,078,611  & 812,038    & 1,890,467 \\
+          Trees               & 4,493,295  & 2,132,368  & 8,780,245 \\
+          Cars                & 900,076    & 1,101,541  & 434,615 \\
+          Buildings           & 13,469,575 & 17,501,421 & 5,128,149 \\
+          Low vegetation      & 4,718,219  & 3,210,596  & 11,428,326 \\
+          Impervious surfaces & 11,340,224 & 11,242,036 & 8,338,198 \\
+          \hline
+        \end{tabular}
+    \end{table}
+
+    In the experiment, the following features are computed using the RGBIR images similar to \cite{tuia2015multiclass}:
+    \begin{itemize}
+        \item Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2 (15 features) \footnote{\url{https://www.orfeo-toolbox.org//Applications/RadiometricIndices.html}};
+        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)\cite{tuia2015multiclass};
+        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)\cite{tuia2015multiclass};
+        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)\cite{tuia2015multiclass};
+        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features)\cite{tuia2015multiclass}.
+    \end{itemize}
+    The normalized DSM and the raw RGBIR image are added to these 191 features and then then stacked to create a new image with 196 bands.
+The training is made with tile 5\_11 and test with tile 5\_12.
+
+\section{Experimental results}
+\label{sec:test}
+
+    \subsection{Method}
+    \label{sec:method}
+
+    The aim of the experiments is to compare the proposed method to some standard classifiers. In order to do it, OTB commandline application are used. The following classifier are tested:
+    \begin{itemize}
+        \item a k-nearest-neighbors classifier (KNN) with OTB default parameters (32 number of neighbors);
+        \item a Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 max depth, 50 size of subset of variables);
+        \item a GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by gridsearch.
+    \end{itemize}
+    The GMM classifier is part of the external module described in Section~\ref{sec:otb-module}.
+
+    All these classifiers are compared with 3 configurations of the proposed GMM classifier:
+    \begin{itemize}
+        \item one with forward selection and JM distance as criterion (GMM SFS JM);
+        \item one with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
+        \item one with floating forward selection and JM distance as criterion (GMM SFFS JM).
+    \end{itemize}
+
+    Training set has been created with an equal number of samples for each class and additionally a spatial stratification has been performed.
+Several size of training set has been tested.
+For Aisa dataset, experiments have been conducted using 250, 500 and 1000 samples by class and for the Potsdam dataset, 1000 and 50000 samples by class.
+
+    For SFS and SFFS selection, the number of variables to select is set to 30 for the Aisa datset and 60 for the Potsdam dataset and then the number of variables which maximizes the criterion function is used to perform classification.
+
+    The classification rate is presented using Cohen's kappa but it has to be noticed the results were similar when looking at the mean F1-score.
+Processing time has been evaluated on a desktop computer with 4Gb of RAM and Intel(R) Core(TM) i5-3570 CPU \@ 3.40GHz x 4 processor.
+All results has been estimated with multiple repetitions (20 trials with Aisa dataset, 5 trials with Potsdam dataset).
+
+    \subsection{Aisa dataset}
+    \label{sec:aisa}
+
     When creating training and validation sets, special care is taken to assure that training samples are picked in distinct areas than test samples.
@@
 
-    This second dataset is built from a dataset of remote sensing images distributed by the International Society for Photogrammetry and Remote Sensing (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}.
-The dataset is composed of aerial images of the urban area of Potsdam.
-The area is cut into 38 patches of 6000x6000 pixels with a resolution of 5cm by pixel and 4 channels are available: Red, Blue, Green and Infrared (RGBIR).
-A Digital Surface Model with same resolution is also provided and a so-called normalized DSM representing the height above ground.
-The groundtruth for 24 tiles are provided with 6 classes: Low vegetation, High vegetation, Impervious surfaces, Buildings, Cars, Clutter.
-Table~\ref{tab:potsdam} summarizes the number of samples of each class in the two images used during experimentations.
-Figure~\ref{fig:potsdam-dataset} shows the complete area and Figure~\ref{fig:potsdam-expl} a particular patch.
-
-    \begin{figure}[!t]
-        \centering
-        \includegraphics[width=0.7\columnwidth]{Fig/potsdam_full.png}
-        \caption{Full Potsdam dataset.\label{fig:potsdam-dataset}}
-    \end{figure}
-
-    \begin{figure}[!t]
-        \centering
-        \includegraphics[width=0.8\columnwidth]{Fig/potsdam_imgexpl.png}
-        \caption{Example of patch with {\bfseries{(a)}} orthophoto, {\bfseries{(b)}} DSM and {\bfseries{(c)}} groundtruth.\label{fig:potsdam-expl}}
-    \end{figure}
-
-    \begin{table}[!t]
-        \centering
-        \caption{Repartition of classes in Aisa dataset.\label{tab:potsdam}}
-        \begin{tabular}[b]{lrr}\hline
-          Class & Number of samples & Number of samples \\
-                & in 5\_11          & in 5\_12 \\
-          \hline
-          Clutter             & 1,078,611  & 812,038 \\
-          Trees               & 4,493,295  & 2,132,368 \\
-          Cars                & 900,076    & 1,101,541 \\
-          Buildings           & 13,469,575 & 17,501,421 \\
-          Low vegetation      & 4,718,219  & 3,210,596 \\
-          Impervious surfaces & 11,340,224 & 11,242,036 \\
-          \hline
-        \end{tabular}
-    \end{table}
-
-    In the experiment, the following features are computed using the RGBIR images similar to \cite{tuia2015multiclass}:
-    \begin{itemize}
-        \item Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2 (15 features) \footnote{\url{https://www.orfeo-toolbox.org//Applications/RadiometricIndices.html}};
-        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)\cite{tuia2015multiclass};
-        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)\cite{tuia2015multiclass};
-        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)\cite{tuia2015multiclass};
-        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features)\cite{tuia2015multiclass}.
-    \end{itemize}
-    The normalized DSM and the raw RGBIR image are added to these 191 features and then then stacked to create a new image with 196 bands.
-The training is made with tile 5\_11 and test with tile 5\_12.
     Table~\ref{tab:potsdam-otbsimu} presents the results.

### 2016-09-27 paper.tex
update

@@
 
-    In the experiment, the following features are computed using the RGBIR images similar to \cite{tuia2015multiclass}:
+    In the experiment, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
     \begin{itemize}
         \item Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2 (15 features) \footnote{\url{https://www.orfeo-toolbox.org//Applications/RadiometricIndices.html}};
-        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)\cite{tuia2015multiclass};
-        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)\cite{tuia2015multiclass};
-        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)\cite{tuia2015multiclass};
-        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features)\cite{tuia2015multiclass}.
+        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features);
+        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features);
+        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features);
+        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features).
     \end{itemize}
     The normalized DSM and the raw RGBIR image are added to these 191 features and then then stacked to create a new image with 196 bands.
-The training is made with tile 5\_11 and test with tile 5\_12.
+The training is made with tile 5\_11 and test with two tiles: 5\_12 which very similar with tile 5\_11 and 3\_10 which a further area a bit less urban and more industrial.
 
@@
 
-    The aim of the experiments is to compare the proposed method to some standard classifiers. In order to do it, OTB commandline application are used. The following classifier are tested:
+    The aim of the experiments is to compare the proposed method to some standard classifiers knowing that it has been compared to other selection method in \cite{fauvel2015fast}.
+The real objective is to test the operational efficiency and it justifies to compare it with the other OTB classifiers already used operationally.
+In order to do it, OTB commandline application are used.
+The following classifiers are tested:
     \begin{itemize}
-        \item a k-nearest-neighbors classifier (KNN) with OTB default parameters (32 number of neighbors);
+        \item a k-nearest-neighbors classifier (KNN) with OTB default parameters (32 as number of neighbors);
         \item a Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 max depth, 50 size of subset of variables);

### 2016-09-27 paper.tex
Section II.A et B
+  une ref en plus.

--- 2016-09-27 paper.tex
@@
 % \documentclass[journal]{IEEEtran}
-
+\usepackage{etex}
 \usepackage{cite}
@@
 \usepackage{empheq}
+\usepackage{hyperref}
 
@@
 
+\tableofcontents
+\clearpage
 % As a general rule, do not put math, special symbols or citations
@@
 
-\section{Gaussian Mixture Models in high dimension space}
+\section{Gaussian Mixture Models in high dimensional space}
 \label{sec:gmm-hd}
 
-In the remaining of the paper, the following notations are used.
-$\mathcal{S} = \{\mathbf{x}_i,y_i\}_{i=1}^{n}$ denotes the training set where $\mathbf{x}_i \in \mathbb{R}^d$ is the vector of features of the $i^{th}$ sample, $y_i = 1,...,C$ the associated label, C the total number of classes, $n$ the number of samples and $n_c$ the number of samples of class $c$.
+The    following    notations    are   used    in    the    remaining.
+$\mathcal{S}  = \{\mathbf{x}_i,y_i\}_{i=1}^{n}$  denotes the  training
+set where $\mathbf{x}_i \in \mathbb{R}^d$ is the vector of features of
+the $i^{th}$ sample, $y_i = 1,...,C$ the associated label, C the total
+number of classes,  $n$ the number of samples and  $n_c$ the number of
+samples of class $c$.
 
@@
 
-    The hypothesis of mixture models is that a given sample is the realization of a random vector which distribution is a mixture (convex combination) of several class conditioned distribution:
+For mixture models, it is assumed  that a given sample $\mathbf{x}$ is
+the realization  of a  random vector which  distribution is  a mixture
+(convex     combination)     of      several     class     conditioned
+distribution~\cite{Fraley00model-basedclustering}:
     \begin{equation}
@@
     \end{equation}
-    where $\pi_c$ is the prior i.e. the proportion of class $c$ and $f_c$ a probability density function parametrized by $\theta$.
-
-    The Gaussian mixture model (GMM) assumes that each $f_c$ is , conditionally to $c$, a Gaussian distribution of parameters $\boldsymbol{\mu}_c$ and $\boldsymbol{\Sigma}_c$ and so $f_c(\mathbf{x}|\theta)$ can be written as
+    where $\pi_c$ is  the prior, \emph{i.e.}, the  proportion of class
+    $c$  and  $f_c$  a  parametric density  function  parametrized  by
+    $\theta$.
+
+    Among the possible parametric model,  the Gaussian one is the most
+    used~\cite{bouveyron2014model}.   It assumes  that each  $f_c$ is,
+    conditionally  to  $c$,  a  Gaussian  distribution  of  parameters
+    $\boldsymbol{\mu}_c$    and    $\boldsymbol{\Sigma}_c$:
     \begin{equation*}
-        f_c(\mathbf{x}|\theta) = \frac{1}{(2\pi)^{\frac{d}{2}} |\boldsymbol{\Sigma}_c|^{\frac{1}{2}}} \exp \left( -\frac{1}{2} (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \right).
+        f_c(\mathbf{x}|\boldsymbol{\mu}_c, \boldsymbol{\Sigma}_c) = \frac{1}{(2\pi)^{\frac{d}{2}} |\boldsymbol{\Sigma}_c|^{\frac{1}{2}}} \exp \left( -\frac{1}{2} (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \right).
     \end{equation*}
 
-    In supervised learning the class parameters $\boldsymbol{\mu}_c$, $\boldsymbol{\Sigma}_c$ and $\pi_c$ are usually estimated through the conventional unbiased empirical estimators:
+    It  is  referred  to  as  Gaussian  mixture  model  (GMM).   In  a
+    supervised    learning    framework,    the    class    parameters
+    $\boldsymbol{\mu}_c$,   $\boldsymbol{\Sigma}_c$   and  the   prior
+    $\pi_c$ are  usually estimated  through the  conventional unbiased
+    empirical estimators:
     \begin{align}
@@
 
-    To predict the class of a new sample, the maximum a posteriori (MAP) rule is used and thus, using Bayes' law, the decision rule is
+    To predict the  class of a new unseen sample,  the maximum \emph{a
+      posteriori}  rule  is  used:
     \begin{equation*}
@@
     \end{equation*}
-    Identifying $p(c)$ as $\pi_c$ and $p(x|c)$ as $f_c(x|\theta)$ and by taking the log, a simplified decision rule is obtained
-    \begin{align}
-        Q_c(\mathbf{x})
-        &= 2 \log \left( p(c) p(\mathbf{x}|c) \right) \nonumber \\
-        &= - (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \nonumber \\
-        &~~-\log (|\boldsymbol{\Sigma}_c|) + 2 \log (\pi_c) - d \log (2\pi).
-        \label{eq:decision}
-    \end{align}
-
-    It is important to notice that the inverse and the determinant of the covariance matrix are key elements of the decision function.
-The estimations of these elements suffer from the curse of dimensionality\cite{bouveyron2014model}.
-More precisely, the number of parameters to estimate increases quadratically relatively to the number of features as illustrated by Figure~\ref{fig:nb-param}.
-The minimum number of samples requires to estimate the model is equal to the number of parameters and it can be an issue in some case.
-For example, hyperspectral data is a case of high dimensional samples and, in general, very few labeled samples are available because of the difficulty and the cost to collect ground-truth.
-
-    \begin{figure}[!t]
+    Under the GMM,  and identifying $p(c)$ as $\pi_c$  and $p(x|c)$ as
+    $f_c(x|\theta)$ and by taking the log, a decision rule is obtained
+    \begin{eqnarray}\label{eq:decision}
+      Q_c(\mathbf{x}) &=& 2 \log \left( p(c) p(\mathbf{x}|c) \right) \nonumber \\
+                      &=& - (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \nonumber \\
+                      & &-\log (|\boldsymbol{\Sigma}_c|) + 2 \log (\pi_c) - d \log (2\pi).
+    \end{eqnarray}
+
+    \subsection{Curse of dimensionality in GMM}
+    \label{sec:curse:gmm}
+
+    The computation of~\ref{eq:decision} requires the inversion of the
+    covariance matrix and and the  computation of the logarithm of the
+    determinant.  The estimations of these terms suffer from the curse
+    of  dimensionality\cite{bouveyron2014model}, because  in practice,
+    the  number of  parameters  $\eta_c$ to  estimate  for each  class
+    increases quadratically with respect to the number of features, as
+    illustrated in Figure~\ref{fig:nb-param}.  Hence, if the number of
+    observation  $n_c$ is  small  compared to  the  number of  samples
+    $n_c$, the  estimated covariance  matrix is badly  conditioned and
+    thus the computation  of its inverse and its  determinant would be
+    unstable.  The  worst situation  is  $n_c<\eta_c$  which leads  to
+    singular covariance matrix.
+
+    Unfortunately, this situation happens regularly in remote sensing.
+    For  instance  in  hyperspectral image  classification,  very  few
+    labeled samples  are usually  available because of  the difficulty
+    and the cost to collect ground-truth.
+
+    \begin{figure}[!t] % TODO : log plot et donner la formule dans la légende 
         \centering
         \begin{tikzpicture}
-            \begin{axis}[xmin=0,xmax=500,ymin=0,width=0.5*\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = Dimension,ylabel =Number of parameters]
-                \addplot+[very thick,mark=none,smooth,domain=0:500,samples=500] (\x,{\x*(\x+3)/2+1});
-            \end{axis}
+          \begin{semilogyaxis}[xmin=0,xmax=200,ymin=0,width=0.5*\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = $d$,ylabel = $\eta_c$]
+            \addplot+[very thick,mark=none,smooth,domain=0:200,samples=200] (\x,{\x*(\x+3)/2+1});
+            \end{semilogyaxis}
         \end{tikzpicture}
-        \caption{Number of parameters by class in function of dimension.\label{fig:nb-param}}
+        \caption{Number of parameters $\eta_c$ by class in function of dimension $d$: $\eta_c=d(d+3)/2+1$.\label{fig:nb-param}}
     \end{figure}
 
-    A lack of samples induces badly-conditioned covariance matrices and so unstable inversion and computation of the determinant.
-There are two major solutions to this problem.
-First option is to use a regularization method to stabilize the inversion of the covariance matrices\cite{hoerl1970ridge}.
-Second option is to use a features extraction/selection method in order to reduce the dimensionality of the samples.
-In this study, the latter option is explored and a feature selection method named sequential forward features selection is presented in the next Section.
-
-    \subsection{Features selection}
+    %% TODO: il faut reprendre cette partie en détaillant plus
+    %% Il y a deux approches: une qui regularise le probleme inverse,
+    %% ridge est une solution, mais il y en a d'autres. Voir le
+    %% paragraphe qui est consacré à cela dans la thèse d'Anthony ->
+    %% elle arrive pas mail. Ensuite, il y a la réduction de dimension
+    %% données des refs (voir intro du papier JSTARS) et dans la
+    %% réduction  de dimension, il  y a  la sélection de  features. La
+    %% conclusion doit s'enchainer avec la partie suivante.
+    There are two major  solutions to this problem. First
+    option  is  to  use  a  regularization  method  to  stabilize  the
+    inversion of the  covariance matrices\cite{hoerl1970ridge}. Second
+    option is to  use a features extraction/selection  method in order
+    to reduce  the dimensionality of  the samples. In this  study, the
+    latter option  is explored  and a  feature selection  method named
+    sequential  forward features  selection is  presented in  the next
+    Section.
+
+    \section{Features selection}
     \label{sec:selection}
 
-    The objective of feature selection is to retrieve the most compact representation of the data with minimum loss of information.
-In deleting the redundant information, feature selection allows to reduce dimension and thus avoid statistical issue, increase algorithm speed and limit storage requirements.
-Moreover, it adds several beneficial side effects.
-It improves data understanding by identify where is the most relevant information and it is then possible to save resources when organizing a new data acquisition.
+    The objective of feature selection is to retrieve the most compact
+    representation of  the data with  minimum loss of  information. By
+    deleting the  redundant information,  feature selection  allows to
+    reduce  dimension  and  thus  avoid  statistical  issue,  increase
+    algorithm speed and limit  storage requirements. Moreover, it adds
+    several beneficial side effects. It improves data understanding by
+    identify where  is the  most relevant information  and it  is then
+    possible to save resources when organizing a new data
+    acquisition.% TODO: add ref. et example.
 
@@
 % \fnbelowfloat command of the stfloats package.
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2016-09-27 paper.tex

@@
 
+    \subsection{Potsdam dataset}
+    \label{sec:pots-dataset}
+
+    This second dataset is built from a dataset of remote sensing images distributed by the International Society for Photogrammetry and Remote Sensing (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}.
+The dataset is composed of aerial images of the urban area of Potsdam.
+The area is cut into 38 patches of 6000x6000 pixels with a resolution of 5cm by pixel and 4 channels are available: Red, Blue, Green and Infrared (RGBIR).
+A Digital Surface Model with same resolution is also provided and a so-called normalized DSM representing the height above ground.
+The groundtruth for 24 tiles are provided with 6 classes: Low vegetation, High vegetation, Impervious surfaces, Buildings, Cars, Clutter.
+Table~\ref{tab:potsdam} summarizes the number of samples of each class in the three images used during experimentations and Figure~\ref{fig:potsdam-expl} the data itself.
+
+    \begin{figure}[!t]
+        \centering
+        \begin{tabular}{ccc}
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_RGB.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_RGB.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_RGB.jpg} \\
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_label.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_12_label.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_3_10_label.jpg} \\
+            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_11_normalized_lastools.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_05_12_normalized_lastools.jpg} &
+            \includegraphics[width=0.3\columnwidth]{Fig/dsm_potsdam_03_10_normalized_lastools.jpg} \\
+            {\bfseries{(a)}} & {\bfseries{(b)}}  & {\bfseries{(c)}}\\
+        \end{tabular}
+        \caption{From top to bottom, image RGB, reference and normalized DSM of: {\bfseries{(a)}} tile 5\_11, {\bfseries{(b)}} tile 5\_12 and {\bfseries{(c)}} tile 3\_10.\label{fig:potsdam-expl}}
+    \end{figure}
+
+    \begin{table}[!t]
+        \centering
+        \caption{Repartition of classes in Aisa dataset.\label{tab:potsdam}}
+        \begin{tabular}[b]{lrrr}\hline
+          Class & Number of samples & Number of samples & Number of samples \\
+                & in 5\_11          & in 5\_12          & in 3\_10 \\
+          \hline
+          Clutter             & 1,078,611  & 812,038    & 1,890,467 \\
+          Trees               & 4,493,295  & 2,132,368  & 8,780,245 \\
+          Cars                & 900,076    & 1,101,541  & 434,615 \\
+          Buildings           & 13,469,575 & 17,501,421 & 5,128,149 \\
+          Low vegetation      & 4,718,219  & 3,210,596  & 11,428,326 \\
+          Impervious surfaces & 11,340,224 & 11,242,036 & 8,338,198 \\
+          \hline
+        \end{tabular}
+    \end{table}
+
+    In the experiment, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
+    \begin{itemize}
+        \item Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2 (15 features) \footnote{\url{https://www.orfeo-toolbox.org//Applications/RadiometricIndices.html}};
+        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features);
+        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features);
+        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features);
+        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features).
+    \end{itemize}
+    The normalized DSM and the raw RGBIR image are added to these 191 features and then then stacked to create a new image with 196 bands.
+The training is made with tile 5\_11 and test with two tiles: 5\_12 which very similar with tile 5\_11 and 3\_10 which a further area a bit less urban and more industrial.
+
+\section{Experimental results}
+\label{sec:test}
+
+    \subsection{Method}
+    \label{sec:method}
+
+    The aim of the experiments is to compare the proposed method to some standard classifiers knowing that it has been compared to other selection method in \cite{fauvel2015fast}.
+The real objective is to test the operational efficiency and it justifies to compare it with the other OTB classifiers already used operationally.
+In order to do it, OTB commandline application are used.
+The following classifiers are tested:
+    \begin{itemize}
+        \item a k-nearest-neighbors classifier (KNN) with OTB default parameters (32 as number of neighbors);
+        \item a Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 max depth, 50 size of subset of variables);
+        \item a GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by gridsearch.
+    \end{itemize}
+    The GMM classifier is part of the external module described in Section~\ref{sec:otb-module}.
+
+    All these classifiers are compared with 3 configurations of the proposed GMM classifier:
+    \begin{itemize}
+        \item one with forward selection and JM distance as criterion (GMM SFS JM);
+        \item one with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
+        \item one with floating forward selection and JM distance as criterion (GMM SFFS JM).
+    \end{itemize}
+
+    Training set has been created with an equal number of samples for each class and additionally a spatial stratification has been performed.
+Several size of training set has been tested.
+For Aisa dataset, experiments have been conducted using 250, 500 and 1000 samples by class and for the Potsdam dataset, 1000 and 50000 samples by class.
+
+    For SFS and SFFS selection, the number of variables to select is set to 30 for the Aisa datset and 60 for the Potsdam dataset and then the number of variables which maximizes the criterion function is used to perform classification.
+
+    The classification rate is presented using Cohen's kappa but it has to be noticed the results were similar when looking at the mean F1-score.
+Processing time has been evaluated on a desktop computer with 4Gb of RAM and Intel(R) Core(TM) i5-3570 CPU \@ 3.40GHz x 4 processor.
+All results has been estimated with multiple repetitions (20 trials with Aisa dataset, 5 trials with Potsdam dataset).
+
+    \subsection{Aisa dataset}
+    \label{sec:aisa}
+
     When creating training and validation sets, special care is taken to assure that training samples are picked in distinct areas than test samples.

### 2016-09-28 paper.tex
Data sets
- rajouté de la bilbio
- changé quelque tableaux pour les passer au format booktabs
- modifié la section data set

@@
 \usepackage{hyperref}
+\usepackage{booktabs}
 
@@
     unstable.   The worst  situation  is $n_c<\eta_c$  which leads  to
-    singular covariance matrix.
-
-    Unfortunately, this situation happens regularly in remote sensing.
-    For  instance  in  hyperspectral image  classification,  very  few
-    labeled samples  are usually  available because of  the difficulty
-    and the cost to collect ground-truth.
-
-    \begin{figure}[!t] % TODO : log plot et donner la formule dans la légende 
+    singular covariance matrix.  Unfortunately, this situation happens
+    regularly in remote sensing.   For instance in hyperspectral image
+    classification,  very few  labeled samples  are usually  available
+    because of the difficulty and the cost to collect ground-truth.
+
+    \begin{figure}[!t]
         \centering
@@
     possible to save resources when organizing a new data
-    acquisition.% TODO: add ref. et example.
-
-    It is important to underline that the proposed method is a \emph{selection} method and not an \emph{extraction} method.
+    acquisition.% TODO: add ref. et example
+
+    % It is important to underline that the proposed method is a \emph{selection} method and not an \emph{extraction} method.
 It means the subset of variables obtained is composed of actual variables of the original set and not variables built as combination of several others as it is the case for example with PCA.
@@
                 \centering
-                \caption{Summary of the various criterion functions.\label{tab:crit}}
-                \begin{tabular}[b]{l|c|c||c|c}
-                  \hline
-                  Criterion & Divergence & Classification & Gaussian   & Speed \\
-                            &            & rate           & hypothesis &       \\
-                  \hline
-                  Overall accuracy            &  & \checkmark & 0 & - \\
-                  Cohen's kappa               &  & \checkmark & 0 & - \\
-                  F1 mean                     &  & \checkmark & 0 & - \\
-                  Kullback-Leibler divergence & \checkmark &  & ++ & + \\
-                  Jeffries-Matusita distance  & \checkmark &  & ++ & + \\
-                  \hline
+                \caption{Summary of the different criterion functions.\label{tab:crit}}
+                \begin{tabular}[b]{lccc}
+                  \toprule
+                  Criterion & Type Divergence & Complexity \\
+                  \midrule
+                  Overall accuracy            & Accuracy   & High \\
+                  Cohen's kappa               & Accuracy   & High\\
+                  F1 mean                     & Accuracy   & High\\
+                  \midrule
+                  Kullback-Leibler divergence & Divergence & Low \\
+                  Jeffries-Matusita distance  & Divergence & Low \\
+                  \bottomrule
                 \end{tabular}
@@
 \label{sec:datasets}
-Experimentation has been conducted on two different datasets.
-The first dataset, called Aisa, is an airborne hyperspectral dataset and second, called Potsdam, is a remote sensing dataset made of heterogeneous spatial features.
+Numerical experiments  have been conducted on  two different datasets.
+The first  dataset, called  \emph{Aisa}, is an  airborne hyperspectral
+dataset and  the second,  called \emph{Potsdam},  is an  airborne very
+high resolution multispectral image.
 
@@
     \label{sec:aisa-dataset}
-    The Aisa dataset has been acquired by the AISA Eagle sensor during a flight campaign over Heves, Hungary.
-It contains 252 bands ranging from 395 to 975 nm.
-16 classes have been defined for a total of 361,971 referenced pixels.
-Figure~\ref{fig:aisa} shows a colored composition of the multispectral image and the shapefile representing groundtruth and Table \ref{tab:aisa} presents the repartition of the various classes.
-
-    \begin{figure}[!t]
+    The Aisa dataset has been acquired by the AISA Eagle sensor during
+    a  flight campaign  over Heves,  Hungary.  It  contains 252  bands
+    ranging from  395 to 975  nm. 16 classes  have been defined  for a
+    total of 361,971  referenced pixels, Table~\ref{tab:aisa} presents
+    the number of pixel per  class.  The Figure~\ref{fig:aisa} shows a
+    colored composition of the image and the groundtruth.
+
+    \begin{figure}[!t]  %TODO: elle  est  pourrie cette  colorisation!
+      % Prend trois bandes dans le visible (RGB) et
+      % étire  correctement l'histogramme. De  plus, on ne sait  pas à
+      %  quelle  couleur correspond  la  couleur.  Tu devrais  pouvoir
+      % récupérer  cela depuis QGIS.  Il faudrait le  rajouter quelque
+      % part. J'ai du code TIKZ pour le faire ci-besoin ...
         \centering
@@
         \caption{Repartition of classes in Aisa dataset.\label{tab:aisa}}
-        \begin{tabular}[b]{lc}\hline
-          Class & Number of samples \\ \hline
+        \begin{tabular}[b]{lc}\toprule
+          Class & Number of samples \\ \midrule
           Winter wheat & 136,524 \\
@@
           Pasture & 2,107 \\
-          Oat & 3,436 \\ \hline
+          Oat & 3,436 \\ \bottomrule
         \end{tabular}
@@
 
-    This second dataset is built from a dataset of remote sensing images distributed by the International Society for Photogrammetry and Remote Sensing (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}.
-The dataset is composed of aerial images of the urban area of Potsdam.
-The area is cut into 38 patches of 6000x6000 pixels with a resolution of 5cm by pixel and 4 channels are available: Red, Blue, Green and Infrared (RGBIR).
-A Digital Surface Model with same resolution is also provided and a so-called normalized DSM representing the height above ground.
-The groundtruth for 24 tiles are provided with 6 classes: Low vegetation, High vegetation, Impervious surfaces, Buildings, Cars, Clutter.
-Table~\ref{tab:potsdam} summarizes the number of samples of each class in the three images used during experimentations and Figure~\ref{fig:potsdam-expl} the data itself.
+    This  second dataset  is built  from a  dataset of  remote sensing
+    images distributed by the International Society for Photogrammetry
+    and                         Remote                         Sensing
+    (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}. The
+    dataset is composed of aerial images of the urban area of Potsdam.
+    The area is cut into 38  patches of 6000$\times$6000 pixels with a
+    resolution  of 5cm  by pixel  and 4  channels are  available: Red,
+    Blue, Green  and Infrared (RGBIR).   A Digital Surface  Model with
+    the same  resolution is also  provided and a  so-called normalized
+    DSM representing the height above  ground. The ground-truth for 24
+    tiles  are   provided  with   6  classes:  Low   vegetation,  High
+    vegetation, Impervious surfaces,  Buildings, Cars, Clutter.  Three
+    tiles  have  been  used  in  this  work,  they  are  displayed  in
+    Figure~\ref{fig:potsdam-expl}.  Table~\ref{tab:potsdam} summarizes
+    the  number of  samples of  each class.
 
@@
         \end{tabular}
-        \caption{From top to bottom, image RGB, reference and normalized DSM of: {\bfseries{(a)}} tile 5\_11, {\bfseries{(b)}} tile 5\_12 and {\bfseries{(c)}} tile 3\_10.\label{fig:potsdam-expl}}
+        \caption{From top to bottom, true color composition, ground-truth and normalized DSM of: {\bfseries{(a)}} tile 5\_11, {\bfseries{(b)}} tile 5\_12 and {\bfseries{(c)}} tile 3\_10.\label{fig:potsdam-expl}}
     \end{figure}
@@
         \caption{Repartition of classes in Aisa dataset.\label{tab:potsdam}}
-        \begin{tabular}[b]{lrrr}\hline
-          Class & Number of samples & Number of samples & Number of samples \\
-                & in 5\_11          & in 5\_12          & in 3\_10 \\
-          \hline
+        \begin{tabular}[b]{lrrr}\toprule
+          Class & \# of samples in 5\_11  & \#  of samples in 5\_12  & \#  of samples in 3\_10 \\
+          \midrule
           Clutter             & 1,078,611  & 812,038    & 1,890,467 \\
@@
           Impervious surfaces & 11,340,224 & 11,242,036 & 8,338,198 \\
-          \hline
+          \bottomrule
         \end{tabular}
@@
 
-    In the experiment, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
+    In order to increase the dimensionality of the data, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
     \begin{itemize}
-        \item Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2 (15 features) \footnote{\url{https://www.orfeo-toolbox.org//Applications/RadiometricIndices.html}};
-        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features);
-        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features);
-        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features);
-        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features).
+        \item 15 Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2~\cite{otb}.
+        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)~\cite{fauvel2013advances};
+        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)~\cite{dalla2010morphological}.
+        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)~\cite{dalla2010morphological}.
+        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features)~\cite{otb}.
     \end{itemize}
-    The normalized DSM and the raw RGBIR image are added to these 191 features and then then stacked to create a new image with 196 bands.
-The training is made with tile 5\_11 and test with two tiles: 5\_12 which very similar with tile 5\_11 and 3\_10 which a further area a bit less urban and more industrial.
+    The normalized DSM and the raw  RGBIR image are added to these 191
+    features and  then then  stacked to  create a  new image  with 196
+    bands. %  The  training is  made with  tile 5\_11  and test  with two
+    % tiles: 5\_12 which very similar with  tile 5\_11 and 3\_10 which a
+    % further area a bit less urban and more industrial.
 

### 2016-09-28 paper.tex
update

@@
 
-    The objective of feature selection is to retrieve the most compact
-    representation of  the data with  minimum loss of  information. By
-    deleting the  redundant information,  feature selection  allows to
-    reduce  dimension  and  thus  avoid  statistical  issue,  increase
-    algorithm speed and limit  storage requirements. Moreover, it adds
-    several beneficial side effects. It improves data understanding by
-    identify where  is the  most relevant information  and it  is then
-    possible to save resources when organizing a new data
-    acquisition.% TODO: add ref. et example
+    The objective of feature selection is to retrieve the most compact representation of the data with minimum loss of information \cite{Guyon:2006:FEF:1208773}.
+By deleting the redundant information, feature selection allows to reduce dimension and thus avoid statistical issue, increase algorithm speed and limit storage requirements.
+Moreover, it adds several beneficial side effects.
+It improves data understanding by identify where is the most relevant information and it is then possible to save resources when organizing a new data acquisition.% TODO: add ref.
+et example
 
@@
 
-    \begin{figure}[!t]  %TODO: elle  est  pourrie cette  colorisation!
-      % Prend trois bandes dans le visible (RGB) et
-      % étire  correctement l'histogramme. De  plus, on ne sait  pas à
-      %  quelle  couleur correspond  la  couleur.  Tu devrais  pouvoir
-      % récupérer  cela depuis QGIS.  Il faudrait le  rajouter quelque
-      % part. J'ai du code TIKZ pour le faire ci-besoin ...
+    \begin{figure}[!t]
         \centering
@@
         \end{tabular}
-        \caption{Aisa dataset: {\bfseries{(a)}} colored composition of the image, {\bfseries{(b)}} groundtruth.\label{fig:aisa}}
+        \caption{Aisa dataset: {\bfseries{(a)}} colored composition of the image (R: 634nm, G: 519nm, B: 477nm), {\bfseries{(b)}} groundtruth.\label{fig:aisa}}
     \end{figure}
@@
         \resizebox{\columnwidth}{!}{%
-        \begin{tabular}{|l|c|c|c|c|c|}\hline
-             & {\bfseries Kappa of} & {\bfseries Kappa of} & {\bfseries Training} & {\bfseries Classif.} & {\bfseries \# features} \\
-             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries time (s)} & {\bfseries time (s)} & {\bfseries to classify} \\ \hline
-
-            GMM SFS kappa & 0.694 (0.002) & 0.669 (0.005) & 400 & 310 & 13.2 \\
-            GMM SFS JM &    0.624 (0.028) & 0.631 (0.034) & 2 & 310   & 11 \\
-            GMM SFFS JM &   0.624 (0.028) & 0.631 (0.034) & 2.6 & 310   & 11 \\
-            GMM ridge &     0.632 (0.007) & 0.592 (0.010) & 10 & 2000 & all \\
-            KNN &           0.637 (0.005) & 0.607 (0.005) &  &   & all \\
-            Random Forest & 0.729 (0.004) & 0.673 (0.005) & 20 & 840  & all \\
+        \begin{tabular}{|l|c|c|c||c|c|c|}\hline
+             & {\bfseries Kappa of}      & {\bfseries Kappa of}     & {\bfseries Kappa of}     & {\bfseries Training} & {\bfseries Classif.} & {\bfseries \# features} \\
+             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries time (s)} & {\bfseries time (s)} & {\bfseries to classify} \\
+            \hline
+
+            GMM SFS kappa & 0.694 (0.002)             & {\bfseries 0.669 (0.005)} & {\bfseries 0.533 (0.008)} & 400 & 310  & 13.2 \\
+            GMM SFS JM &    0.624 (0.028)             & 0.631 (0.034)             & 0.461 (0.027)             & 2   & 310  & 11 \\
+            GMM SFFS JM &   0.624 (0.028)             & 0.631 (0.034)             & 0.461 (0.027)             & 2.6 & 310  & 11 \\
+            GMM ridge &     0.632 (0.007)             & 0.592 (0.010)             & 0.433 (0.008)             & 10  & 2000 & all \\
+            KNN &           0.637 (0.005)             & 0.607 (0.005)             & 0.478 (0.002)             &     &      & all \\
+            Random Forest & {\bfseries 0.729 (0.004)} & {\bfseries 0.673 (0.005)} & {\bfseries 0.529 (0.009)} & 20  & 840  & all \\
 
@@
         \resizebox{\columnwidth}{!}{%
-        \begin{tabular}{|l|c|c|c|c|c|}\hline
-             & {\bfseries Kappa of} & {\bfseries Kappa of} & {\bfseries Training} & {\bfseries Classif.} & {\bfseries \# features} \\
-             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries time (s)} & {\bfseries time (s)} & {\bfseries to classify} \\ \hline
-
-            GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 20000 & 340 & 29 \\
-            GMM SFS JM &    0.560 (0.111) & 0.576 (0.104) & 6 & 330 & 10 \\
-            GMM SFFS JM &   0.560 (0.111) & 0.576 (0.104) & 6.6 & 340 & 10 \\
-            GMM ridge &     0.641 (0.015) & 0.611 (0.026) & 460 & 2000 & all \\
-            Random Forest & 0.851 (0.001) & 0.715 (0.001) & 2000 & 2000 & all \\
+        \begin{tabular}{|l|c|c|c||c|c|c|}\hline
+             & {\bfseries Kappa of}      & {\bfseries Kappa of}     & {\bfseries Kappa of}     & {\bfseries Training} & {\bfseries Classif.} & {\bfseries \# features} \\
+             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries time (s)} & {\bfseries time (s)} & {\bfseries to classify} \\ \hline
+
+            GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 0.531 (0.005) & 20000 & 340 & 29 \\
+            GMM SFS JM &    0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6 & 330 & 10 \\
+            GMM SFFS JM &   0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6.6 & 340 & 10 \\
+            GMM ridge &     0.641 (0.015) & 0.611 (0.026) & 0.440 (0.015) & 460 & 2000 & all \\
+            Random Forest & {\bfseries 0.851 (0.001)} & {\bfseries 0.715 (0.001)} & {\bfseries 0.573 (0.002)} & 2000 & 2000 & all \\
 

### 2016-09-28 paper.tex
Section Exp. results
- repris le texte et les tableaux
- rajouté une biblio
- rajouté des commentaires à Adrien

--- 2016-09-28 paper.tex
@@
 
-    The aim of the experiments is to compare the proposed method to some standard classifiers knowing that it has been compared to other selection method in \cite{fauvel2015fast}.
-The real objective is to test the operational efficiency and it justifies to compare it with the other OTB classifiers already used operationally.
-In order to do it, OTB commandline application are used.
+    The aim  of the experiments is  to compare the proposed  method to
+    some   standard  classifiers   used   in   operational  land   map
+    production~\cite{rs70912356}. A non-optimized  previous version of
+    the method has been already compared to other selection methods in
+    \cite{fauvel2015fast}.  Hence,  the primary  objective is  to test
+    the  operational  efficiency  and  it is  compared  to  other  OTB
+    classifiers   used  operationally   through  their   command  line
+    applications\footnote{\url{http://otbcb.readthedocs.io/en/latest/OTB-Applications.html}}.
+
     The following classifiers are tested:
     \begin{itemize}
-        \item a k-nearest-neighbors classifier (KNN) with OTB default parameters (32 as number of neighbors);
-        \item a Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 max depth, 50 size of subset of variables);
-        \item a GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by gridsearch.
+        \item A k-nearest-neighbors classifier (KNN) with OTB default parameters (32 as number of neighbors).
+        \item A Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 max depth, 50 size of subset of variables)%%TODO : pas clair ce qu'est "50 size of subset of variables"
+        \item A GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by gridsearch.
     \end{itemize}
@@
 
-    Training set has been created with an equal number of samples for each class and additionally a spatial stratification has been performed.
-Several size of training set has been tested.
-For Aisa dataset, experiments have been conducted using 250, 500 and 1000 samples by class and for the Potsdam dataset, 1000 and 50000 samples by class.
-
-    For SFS and SFFS selection, the number of variables to select is set to 30 for the Aisa datset and 60 for the Potsdam dataset and then the number of variables which maximizes the criterion function is used to perform classification.
-
-    The classification rate is presented using Cohen's kappa but it has to be noticed the results were similar when looking at the mean F1-score.
-Processing time has been evaluated on a desktop computer with 4Gb of RAM and Intel(R) Core(TM) i5-3570 CPU \@ 3.40GHz x 4 processor.
-All results has been estimated with multiple repetitions (20 trials with Aisa dataset, 5 trials with Potsdam dataset).
+    The training set has been created  with an equal number of samples
+    for each class and additionally  a spatial stratification has been
+    performed, \emph{i.e.}, each training  sample belongs to a spatial
+    polygon  that  does  not  intersect  spatially  with  any  spatial
+    polygons used  for the validation.   Several size of  training set
+    have been tested investigated.   For the Aisa dataset, experiments
+    have been conducted  using 250, 500 and 1000 samples  by class and
+    for the Potsdam dataset, 1000 and 50000 samples by class.
+
+    %% TODO : c'est encore le cas ? à verifier: si on prends quand cela arréte de croitre rapidement il faut changer
+    For SFS and  SFFS selection, the number of variables  to select is
+    set to 30 for  the Aisa datset and 60 for  the Potsdam dataset and
+    then  the  number  of  variables  which  maximizes  the  criterion
+    function is used to perform classification.
+
+    The classification  rate is presented  using Cohen's kappa  but it
+    has to  be noticed the  results were  similar when looking  at the
+    mean F1-score.  Processing time  has been  evaluated on  a desktop
+    computer  with 4Gb  of RAM  and  Intel(R) Core(TM)  i5-3570 CPU  @
+    3.40GHz $\times$ 4 processors.
 
@@
 
-    When creating training and validation sets, special care is taken to assure that training samples are picked in distinct areas than test samples.
-In order to do it, the polygons of the reference are split in smaller polygons and then 50\% of the polygons are taken randomly for training and the remaining 50\% for validation.
-Moreover 20 trials are run each time with a different training set (different polygons).
-An example of training and validation set is shown in Figure~\ref{fig:set-aisa}.
-Additionally, it is important to understand that only a subpart of this training set is actually used to train the model, a given number of samples for each class.
-Table~\ref{tab:aisa-otbsimu} presents the results of the experiment with mean and standard deviation over the 20 trials and Table~\ref{tab:aisa-otbsimu-time} the corresponding processing time.
+    When creating training and validation  sets, special care is taken
+    to assure that training samples are picked out from distinct areas
+    than test  samples.  The  polygons of the  reference are  split in
+    smaller polygons and then 50\%  of the polygons are taken randomly
+    for training and the remaining 50\% for validation.  An example of
+    training      and     validation      set     is      shown     in
+    Figure~\ref{fig:set-aisa}.  From the  training  polygons, a  given
+    number of samples  were selected to build the  training set, while
+    all  the pixels  from the  validation polygons  were used  for the
+    validation.  Moreover 20  random trials were run  with a different
+    training  set (different  polygons).  Table~\ref{tab:aisa-otbsimu}
+    presents  the results  of the  experiment with  mean and  standard
+    deviation  of  the  Kappa  coefficient  over  the  20  trials  and
+    Table~\ref{tab:aisa-otbsimu-time}  the   corresponding  processing
+    time.
 
@@
         \centering
-        \caption{Results of classification with sampling on separate polygons and 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
-        \begin{tabular}{|l|c|c|c|}\hline
-             & \multicolumn{3}{c|}{\bfseries Cohen's kappa} \\ \hline
-            \# samples by class & 250 & 500 & 1000 \\ \hline
+        \caption{Average classification accuracy 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
+        \begin{tabular}{lccc}\toprule
+             & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cline{2-4}
+            \# samples by class & 250 & 500 & 1000 \\ \midrule
 
@@
             Random Forest & 0.645 (0.026) & 0.673 (0.023) & 0.693 (0.023) \\
-
-            \hline
+            \bottomrule
         \end{tabular}
@@
         \centering
-        \caption{Mean processing time for training and classification for results of Table \ref{tab:aisa-otbsimu}.\label{tab:aisa-otbsimu-time}}
-        \begin{tabularx}{\columnwidth}{|l|*{6}{>{\centering\arraybackslash}X|}}
-            \hline
-             & \multicolumn{3}{c|}{\bfseries Training time (s)} & \multicolumn{3}{c|}{\bfseries Classification time (s)} \\ \hline
-            \# samples by class & 250 & 500 & 1000 & 250 & 500 & 1000 \\ \hline
+        \caption{Mean processing time for training and classification for results in Table \ref{tab:aisa-otbsimu}.\label{tab:aisa-otbsimu-time}}
+        \begin{tabularx}{\columnwidth}{l*{6}{>{\centering\arraybackslash}X}}
+            \toprule
+             & \multicolumn{3}{c}{\bfseries Training time (s)} & \multicolumn{3}{c}{\bfseries Classification time (s)} \\ \cline{2-7}
+            \# samples by class & 250 & 500 & 1000 & 250 & 500 & 1000 \\ \midrule
 
@@
             Random Forest & 24.5 & 49.3 & 105 & 33.0 & 41.7 & 45.9 \\
-
-            \hline
+            \bottomrule
         \end{tabularx}
@@
 
-    The results show that, on this dataset, GMM classifiers with selection get the best classification rate.
-Especially with very few training samples, GMM classifiers with selection seems to outperform the others.
-In term of computational time, the GMM classifiers are as expected very fast for classification and also for training if the criterion function is not a classification rate.
-The computation of the kappa seems to suffer from a lack of parallelization which could be improved.
-Among the three variations of the selection algorithm, none appears to perform better than the others.
-Using kappa or Jeffries-Matusita distance as criterion is equal and using SFFS does not give any advantage.
-In this case, using J-M distance as criterion and SFS as search strategy is the best choice in term of time efficiency.
-
-    Nevertheless, good performance are also obtained with Random Forest which becomes more competitive when the number of samples increases.
-    The KNN classifier and the GMM classifier with ridge regularization are totally outperform both in term of classification rate and of processing time.
+    %% TODO; Reprendre cette partie: ne parle pas à la fois de la
+    %% classification et du temps de calcul. D'abors du commences à
+    %% décrire et commenter sur la classification. Ensuite tu fais la
+    %% même chose sur les temps de calculs. Et il faut en dire plus !!
+    %%   Pour   les   classifs,   il  faut   parler   des   tests   de
+    %%  significativité.  Pour les  temps  de  calculs, il  faut  plus
+    %% détailler  pourquoi cela  va plus  ou moins  vite. Avec  FS, on
+    %% réduit le nombre de variables et comme la régle de décision est
+    %%  de compléxité  (d^3)  donc  quand on  réduit  d, on  améliorer
+    %% grandement  les temps de  calcul, pour les RF,  plus il y  a de
+    %% training samples, plus lse arbres vont êtres grands ...
+    The  results show  that,  on this  dataset,  GMM classifiers  with
+    feature selection  get the  best classification  rate.  Especially
+    with  very  few training  samples,  GMM  classifiers with  feature
+    selection outperform  the others.  In term  of computational time,
+    the GMM classifiers  are as expected very  fast for classification
+    and also  for training,  except when the  criterion function  is a
+    classification rate.  Among the  three variations of the selection
+    algorithm, none appears  to perform better than  the others. Using
+    kappa  or Jeffries-Matusita  distance  as criterion  is equal  and
+    using SFFS  does not give any  advantage. In this case,  using J-M
+    distance  as criterion  and SFS  as  search strategy  is the  best
+    choice in term of time efficiency.  % The computation of the kappa
+                                       % seems  to suffer from  a lack
+                                       %   of  parallelization   which
+                                       %  could be  improved. ->  plus
+                                       % tard dans la conclusion.
+
+    Nevertheless,  good  performance  are also  obtained  with  Random
+    Forest which becomes  more competitive when the  number of samples
+    increases.  The KNN  classifier and the GMM  classifier with ridge
+    regularization   are   totally   outperform  both   in   term   of
+    classification rate and of processing time.
 
@@
 
-    Table~\ref{tab:potsdam-otbsimu} presents the results.
+    For the Potsdam  dataset, training samples were  selected from one
+    tile (5\_11)  and validation  samples were all  the pixel  of tile
+    5\_12  or  3\_10.  Table~\ref{tab:potsdam-otbsimu}  presents  the
+    results.
 
@@
         \centering
-        \caption{Results of classification with 1000 samples by class and 5 trials (standard deviation in parenthesis).\label{tab:potsdam-otbsimu}}
-        \resizebox{\columnwidth}{!}{%
-        \begin{tabular}{|l|c|c|c|c|c|}\hline
-             & {\bfseries Kappa of} & {\bfseries Kappa of} & {\bfseries Training} & {\bfseries Classif.} & {\bfseries \# features} \\
-             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries time (s)} & {\bfseries time (s)} & {\bfseries to classify} \\ \hline
-
+        \caption{Kappa coefficient and processing time for 1000 samples by class and averaged over 5 trials (standard deviation in parenthesis).
+Processing times are given in second.\label{tab:potsdam-otbsimu}}
+        \begin{tabular}{lccccc}\toprule
+          & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-6}
           GMM SFS kappa & 0.694 (0.002) & 0.669 (0.005) & 400 & 310 & 13.2 \\
@@
           Random Forest & 0.729 (0.004) & 0.673 (0.005) & 20 & 840  & all \\
-
-            \hline
+          \bottomrule
         \end{tabular}
-        }
     \end{table}
 
+    %% TODO: je te laisse modifier le tableau de la même manière (style, legende etc ...)
     \begin{table}[!t]
@@
         \caption{Results of classification with 50000 samples by class and 5 trials (standard deviation in parenthesis).\label{tab:potsdam-otbsimu-big}}
-        \resizebox{\columnwidth}{!}{%
         \begin{tabular}{|l|c|c|c|c|c|}\hline
@@
         \end{tabular}
-        }
       \end{table}
 
+      %% TODO: reprendre la discussion avec les commentaires de l'image précédente
     In this case, the classification rate of GMM classifiers with Jeffries-Matusita distance as criterion function drops clearly and has a high standard deviation.

### 2016-09-28 paper.tex

@@
             Random Forest & 0.851 (0.001) & 0.715 (0.001) & 2000 & 2000 & all \\
-
             \hline

### 2016-09-28 paper.tex
update

@@
 SVM \cite{guyon2002gene} \cite{weston2003use}.
-²
+
 The selection method proposed in this work is a wrapper method associated to GMM models.
@@
 
-        \subsubsection{Criterion function}
+    \subsection{Criterion function}
     \label{sec:criterion}
-
-        \hspace{0pt} \\
 
@@
 
-            \paragraph{Measures of correct classification}
+        \subsubsection{Measures of correct classification}
         \label{sec:criterion-rate}
@@
 
-            \paragraph{Similarity between distributions}
+        \subsubsection{Similarity between distributions}
 
@@
 
-        \subsubsection{Selection method}
-
-            \paragraph{Sequential forward features selection}
+    \subsection{Selection method}
+
+        \subsubsection{Sequential forward features selection}
         \label{sec:forward-presentation}
@@
 
-            \paragraph{Sequential floating forward feature selection}
+        \subsubsection{Sequential floating forward feature selection}
         \label{sec:floating-presentation}
@@
         \item A k-nearest-neighbors classifier (KNN) with OTB default parameters (32 as number of neighbors).
-        \item A Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 max depth, 50 size of subset of variables)%%TODO : pas clair ce qu'est "50 size of subset of variables"
+        \item A Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 as max depth, 50 as size of the randomly selected subset of features at each tree node)
         \item A GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by gridsearch.
@@
 
-    %% TODO : c'est encore le cas ? à verifier: si on prends quand cela arréte de croitre rapidement il faut changer
-    For SFS and  SFFS selection, the number of variables  to select is
-    set to 30 for  the Aisa datset and 60 for  the Potsdam dataset and
-    then  the  number  of  variables  which  maximizes  the  criterion
-    function is used to perform classification.
+    For SFS and  SFFS selection, the number of variables to select is set to 30 for the Aisa dataset and 60 for the Potsdam dataset.
+Then, a small processing is made to select the number of variables used for prediction.
+The evolution of the criterion function is retrieved from the model file and the discrete derivative of the criterion \emph{DJ} is computed and normalized by its maximum.
+The subset of variables used for classification $\Omega^{*}$ is the first subset for which this value is lower than a given threshold \emph{th}, equal  to 0.001\% in this experiment,
+    \begin{equation*}
+        \Omega^{*} = \Omega_{\min \{k|DJ_k<th\}}
+    \end{equation*}
+    with $DJ_k = \frac{J(\Omega_k) - J(\Omega_{k-1})}{\max_k (J(\Omega_k) - J(\Omega_{k-1}))}$.
 
@@
 
-    %% TODO; Reprendre cette partie: ne parle pas à la fois de la
-    %% classification et du temps de calcul. D'abors du commences à
-    %% décrire et commenter sur la classification. Ensuite tu fais la
-    %% même chose sur les temps de calculs. Et il faut en dire plus !!
-    %%   Pour   les   classifs,   il  faut   parler   des   tests   de
-    %%  significativité.  Pour les  temps  de  calculs, il  faut  plus
-    %% détailler  pourquoi cela  va plus  ou moins  vite. Avec  FS, on
-    %% réduit le nombre de variables et comme la régle de décision est
-    %%  de compléxité  (d^3)  donc  quand on  réduit  d, on  améliorer
-    %% grandement  les temps de  calcul, pour les RF,  plus il y  a de
-    %% training samples, plus lse arbres vont êtres grands ...
-    The  results show  that,  on this  dataset,  GMM classifiers  with
-    feature selection  get the  best classification  rate.  Especially
-    with  very  few training  samples,  GMM  classifiers with  feature
-    selection outperform  the others.  In term  of computational time,
-    the GMM classifiers  are as expected very  fast for classification
-    and also  for training,  except when the  criterion function  is a
-    classification rate.  Among the  three variations of the selection
-    algorithm, none appears  to perform better than  the others. Using
-    kappa  or Jeffries-Matusita  distance  as criterion  is equal  and
-    using SFFS  does not give any  advantage. In this case,  using J-M
-    distance  as criterion  and SFS  as  search strategy  is the  best
-    choice in term of time efficiency.  % The computation of the kappa
-                                       % seems  to suffer from  a lack
-                                       %   of  parallelization   which
-                                       %  could be  improved. ->  plus
-                                       % tard dans la conclusion.
-
-    Nevertheless,  good  performance  are also  obtained  with  Random
-    Forest which becomes  more competitive when the  number of samples
-    increases.  The KNN  classifier and the GMM  classifier with ridge
-    regularization   are   totally   outperform  both   in   term   of
-    classification rate and of processing time.
+    The results show that, on this  dataset, GMM classifiers with feature selection get the best classification rate.
+Among the three variations of the selection algorithm, none appears to perform better than the others.
+Using kappa or Jeffries-Matusita distance as criterion is equal and using SFFS does not give any advantage.
+
+    The difference with the second best classifier, Random Forest, appears to be significant when using 250 and 500 samples.
+Random Forest has similar performance in term of classification rate with 1000 samples.
+The GMM classifier with ridge regularization and the kNN classifier are both outperformed.
+
+    In term of computational time, the GMM classifiers are as expected very fast for classification and also for training, except when the criterion function is a classification rate.
+In this case, using JM distance as criterion and SFS as search strategy is the best choice in term of time efficiency.
+The good performance in time can explain by the dimensional reduction.
+Actually, the decision rule corresponding to Equation~\ref{eq:decision} has a complexity in $d^3$ where $d$ is the dimension.
+Thus, reducing $d$ induces a reduction of the classification time.
+
+    The processing times of the three standard classifiers suffer from the increase of training samples.
+For the GMM classifier with ridge, the selection of regularization parameter is more costly with more samples because the classification rate estimation needed.
+For the kNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file.
+Finally, for the Random Forest classifier, the trees tends to be more deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.
+    % The computation of the kappa seems  to suffer from  a lack of  parallelization   which could be  improved. ->  plus tard dans la conclusion.
 
@@
         \centering
-        \caption{Kappa coefficient and processing time for 1000 samples by class and averaged over 5 trials (standard deviation in parenthesis).
+        \caption{Kappa coefficient and processing time for 1,000 samples by class and averaged over 5 trials (standard deviation in parenthesis).
 Processing times are given in second.\label{tab:potsdam-otbsimu}}
-        \begin{tabular}{lccccc}\toprule
-          & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-6}
-          GMM SFS kappa & 0.694 (0.002) & 0.669 (0.005) & 400 & 310 & 13.2 \\
-          GMM SFS JM &    0.624 (0.028) & 0.631 (0.034) & 2 & 310   & 11 \\
-          GMM SFFS JM &   0.624 (0.028) & 0.631 (0.034) & 2.6 & 310   & 11 \\
-          GMM ridge &     0.632 (0.007) & 0.592 (0.010) & 10 & 2000 & all \\
-          KNN &           0.637 (0.005) & 0.607 (0.005) &  &   & all \\
-          Random Forest & 0.729 (0.004) & 0.673 (0.005) & 20 & 840  & all \\
+        \begin{tabular}{lcccccc}\toprule
+            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-6}
+            GMM SFS kappa & 0.694 (0.002)             & {\bfseries 0.669 (0.005)} & {\bfseries 0.533 (0.008)} & 400 & 310  & 13.2 \\
+            GMM SFS JM &    0.624 (0.028)             & 0.631 (0.034)             & 0.461 (0.027)             & 2   & 310  & 11 \\
+            GMM SFFS JM &   0.624 (0.028)             & 0.631 (0.034)             & 0.461 (0.027)             & 2.6 & 310  & 11 \\
+            GMM ridge &     0.632 (0.007)             & 0.592 (0.010)             & 0.433 (0.008)             & 10  & 2000 & all \\
+            KNN &           0.637 (0.005)             & 0.607 (0.005)             & 0.478 (0.002)             &     &      & all \\
+            Random Forest & {\bfseries 0.729 (0.004)} & {\bfseries 0.673 (0.005)} & {\bfseries 0.529 (0.009)} & 20  & 840  & all \\
             \bottomrule
@@
 
-    %% TODO: je te laisse modifier le tableau de la même manière (style, legende etc ...)
     \begin{table}[!t]
         \centering
-        \caption{Results of classification with 50000 samples by class and 5 trials (standard deviation in parenthesis).\label{tab:potsdam-otbsimu-big}}
-        \begin{tabular}{|l|c|c|c|c|c|}\hline
-             & {\bfseries Kappa of} & {\bfseries Kappa of} & {\bfseries Training} & {\bfseries Classif.} & {\bfseries \# features} \\
-             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries time (s)} & {\bfseries time (s)} & {\bfseries to classify} \\ \hline
-
-            GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 20000 & 340 & 29 \\
-            GMM SFS JM &    0.560 (0.111) & 0.576 (0.104) & 6 & 330 & 10 \\
-            GMM SFFS JM &   0.560 (0.111) & 0.576 (0.104) & 6.6 & 340 & 10 \\
-            GMM ridge &     0.641 (0.015) & 0.611 (0.026) & 460 & 2000 & all \\
-            Random Forest & 0.851 (0.001) & 0.715 (0.001) & 2000 & 2000 & all \\
-            \hline
+        \caption{Kappa coefficient and processing time for 50,000 samples by class and averaged over 5 trials (standard deviation in parenthesis).
+Processing times are given in second.\label{tab:potsdam-otbsimu-big}}
+        \begin{tabular}{|lcccccc}\toprule
+            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-6}
+            GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 0.531 (0.005) & 20000 & 340 & 29 \\
+            GMM SFS JM &    0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6 & 330 & 10 \\
+            GMM SFFS JM &   0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6.6 & 340 & 10 \\
+            GMM ridge &     0.641 (0.015) & 0.611 (0.026) & 0.440 (0.015) & 460 & 2000 & all \\
+            Random Forest & {\bfseries 0.851 (0.001)} & {\bfseries 0.715 (0.001)} & {\bfseries 0.573 (0.002)} & 2000 & 2000 & all \\
+            \bottomrule
         \end{tabular}
@@
 
-      %% TODO: reprendre la discussion avec les commentaires de l'image précédente
-    In this case, the classification rate of GMM classifiers with Jeffries-Matusita distance as criterion function drops clearly and has a high standard deviation.
-It could be due to the fact that the Gaussian hypothesis, on which this metric relies a lot, is not enough verified.
-When the classification rate is directly used as criterion, the classifier manages to select the relevant features.
-It has to be noticed that with kappa only 30 features are selected on average when with Jeffries-Matusita distance, the maximum number of features is selected almost always (set to 60).
-
-    The KNN classifier and the GMM classifier with ridge regularization are again outperformed even if they get stable results.
-
-    Finally, the Random Forest classifier and the GMM with kappa as criterion are from far the best classifiers.
-The only difference is in the processing time and also a small difference of classification rate with the training image.
-Random Forest is faster for the training process while the GMM classifier is faster for classifying.
+    With this second dataset, the Random Forest classifier and the GMM classifier with kappa as selection criterion distinguished themselves by good performance.
+When using 1000 samples, no significant difference of classification rate has been observed on test set.
+But, with 50,000 samples, the Random Forest classifier better significantly better.
+
+    It can be guessed that the fact that the Gaussian hypothesis is unlikely to be verified explains the difficulty to extract the finer information available in a larger dataset.
+The JM distance has stressed in Table~\ref{tab:crit} has a strong Gaussian hypothesis and the same hypotheses could also explain the drop of the classification rate of GMM classifiers with Jeffries-Matusita distance as criterion.
+
+    The KNN classifier and the GMM classifier with ridge regularization are again outperformed even if they get more stable results than GMM classifiers with JM as criterion.
+
+    The number of variables identified by the selection process shows that selection with JM distance as criterion identifies less relevant samples than with kappa as criterion.
+Moreover, the GMM classifier with kappa manages to get good performance with only 6.7\ of the variables with 1000 samples and 15\% with 50,000 samples.
+
+    In term of processing time, results are similar than with the Aisa dataset.
+GMM classifiers with selection are very fast for prediction.
+For example, the GMM classifier with kappa as criterion is 63\% faster than the Random Forest classifier for prediction with 1000 samples and 83\% faster with 50,000 samples.
 

### 2016-09-28 paper.tex

@@
     determinant.  The estimations of these terms suffer from the curse
-    of  dimensionality\cite{bouveyron2014model}, because  in practice,
+    of  dimensionality\cite{}, because  in practice,
     the  number of  parameters  $\eta_c$ to  estimate  for each  class
@@
     %% conclusion doit s'enchainer avec la partie suivante.
-    There are two major  solutions to this problem. First
-    option  is  to  use  a  regularization  method  to  stabilize  the
-    inversion of the  covariance matrices\cite{hoerl1970ridge}. Second
-    option is to  use a features extraction/selection  method in order
-    to reduce  the dimensionality of  the samples. In this  study, the
-    latter option  is explored  and a  feature selection  method named
-    sequential  forward features  selection is  presented in  the next
-    Section.
+    There are two major  solutions to this problem.
+First option is to stabilize the inversion of the covariance matrices work.
+Some methods investigate the use of constraints on the direct problem \cite{reynolds1995robust} \cite{hathaway1985constrained} and some work on the inverse problem by using regularization method \cite{hoerl1970ridge} \cite{jensen2008regression}.
+Second option is to reduce dimension.
+Features extraction/selection methods have been developed in order to reduce the dimension.
+In this  study, the latter option is explored and a feature selection method named sequential forward features selection is presented after a brief overview of the selection algorithms in the next Section.
+%A finir
 

### 2016-09-28 paper.tex
Intro FS

@@
     determinant.  The estimations of these terms suffer from the curse
-    of  dimensionality\cite{}, because  in practice,
+    of  dimensionality\cite{bouveyron2014model}, because  in practice,
     the  number of  parameters  $\eta_c$ to  estimate  for each  class
@@
 
-Features selection algorithms may be divided into three types.
-The first type called filter method is based uniquely on data analysis.
+Features selection algorithms can be divided into three types.
+The first type, called \emph{filter methods}, is based uniquely on data analysis.
 Features are ranked according to a statistical analysis of the data.
-For example, the Principal Component Analysis (PCA) described in \cite{jimenez1998supervised} is a typical filter method and there is numerous other methods \cite{bruzzone1995extension}, \cite{biesiada2007feature}, \cite{demir2008phase}.
-
-The second sort are known as wrapper methods which can be seen as search method to determine the best subset of variables for a given learning model.
-As exhaustive searches are out of question in a practical amount of time, numerous search strategies have been designed some optimal under particular hypothesis \cite{narendra1977branch} and other suboptimal but easier to set up \cite{whitney1971direct}, \cite{somol1999adaptive}.
-The advantage of such methods compare to filter methods is that they are dedicated to a particular model but on the other hand, as they require the training of multiple models to test various set of variables, they tend to be slower.
-
-The third type corresponds to the embedded method which do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
+For example, the Principal Component Analysis (PCA) described in \cite{jimenez1998supervised} is a typical filter method and there is numerous other methods \cite{bruzzone1995extension,biesiada2007feature,demir2008phase}.
+
+The second sort are known as \emph{wrapper methods} which can be seen as search method to determine the best subset of variables for a given learning model.
+As exhaustive searches are to expansive in terms of processing time, several search strategies have been designed.
+Some are optimal under particular hypothesis \cite{narendra1977branch} and other sub-optimal but easier to set up \cite{whitney1971direct,somol1999adaptive}.
+The advantage of such methods compare to filter methods is that they are dedicated to a particular model but on the other hand, as they require the training of multiple models to test various set of variables, they tend to be more time consuming.
+
+The third type corresponds to the \emph{embedded method} which do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
 The basic example of such method is the decision tree algorithm in which a feature is selected for the creation of each node.
 Embedded methods also exist for other models, e.g.
-SVM \cite{guyon2002gene} \cite{weston2003use}.
-
-The selection method proposed in this work is a wrapper method associated to GMM models.
-Thus, in order to set a wrapper method, two elements are needed, a function to rank the various features which is defined in Section~\ref{sec:criterion} and a search strategy.
-Two search algorithms are presented: the sequential forward features selection method (Section \ref{sec:forward-presentation}) and the sequential floating forward feature selection method (\ref{sec:floating-presentation}) the second being a more complex variation of the first.
+SVM \cite{guyon2002gene,weston2003use}.
+
+The feature selection method proposed in this work is a wrapper method associated to GMM models.
+To set up a wrapper method, two elements are needed.
+A function that  ranks the features according to some criteria and a search strategy to optimize the function.
+The criterion used are described in Section~\ref{sec:criterion} and the two search strategy are discussed in~\ref{sec:selection:method}.
 
@@
         \subsection{Selection method}
+        \label{sec:selection:method}
 
@@
 The first  dataset, called  \emph{Aisa}, is an  airborne hyperspectral
-dataset and  the second,  called \emph{Potsdam},  is an  airborne very
-high resolution multispectral image.
+dataset  and  the  second,  called \emph{Potsdam},  is  an  very  high
+resolution multispectral airborne image.
 
@@
         \begin{tabular}{lcccccc}\toprule
-            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-6}
+            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-7}
             GMM SFS kappa & 0.694 (0.002)             & {\bfseries 0.669 (0.005)} & {\bfseries 0.533 (0.008)} & 400 & 310  & 13.2 \\
@@
         \begin{tabular}{|lcccccc}\toprule
-            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-6}
+            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-7}
             GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 0.531 (0.005) & 20000 & 340 & 29 \\

### 2016-09-29 paper.tex
Section II
criterion function

@@
 
-\section{Features selection}
+\section{Sequential forward features selection}
 \label{sec:selection}
 
+%%% Debut Déplacer vers l'intro
 The objective of feature selection is to retrieve the most compact representation of the data with minimum loss of information \cite{Guyon:2006:FEF:1208773}.
@@
 
-% It is important to underline that the proposed method is a \emph{selection} method and not an \emph{extraction} method.
-It means the subset of variables obtained is composed of actual variables of the original set and not variables built as combination of several others as it is the case for example with PCA.
-This choice is made to assure an easier interpretation by the user of the obtained subset of variables.
-
+%% Cette partie, dans l'intro il faut la rallonger: mieux décrire ces trois méthodes, dire les avantages et les inconvénients de chaque approche et positionner notre approche par rapport à tout cela (je dis cela sans avoir relu l'intro).
+Tu peux t'inspirer de la thèse de Arnaud Lebris et des pages 58-66)
 Features selection algorithms can be divided into three types.
@@
 SVM \cite{guyon2002gene,weston2003use}.
-
-The feature selection method proposed in this work is a wrapper method associated to GMM models.
-To set up a wrapper method, two elements are needed.
-A function that  ranks the features according to some criteria and a search strategy to optimize the function.
-The criterion used are described in Section~\ref{sec:criterion} and the two search strategy are discussed in~\ref{sec:selection:method}.
+%%% Fin déplacer
+
+The feature selection method proposed in this work is a wrapper method associated to GMM models. Two elements are needed to set up a wrapper method:
+\begin{enumerate}
+\item A function that  ranks the features according to some good classification criteria,
+\item A search strategy to optimize the function.
+\end{enumerate}
+
+Section~\ref{sec:criterion} describes the different criterion used in this work and two search strategies are discussed in~\ref{sec:selection:method}.
 
@@
     \label{sec:criterion}
-
-    Criterion functions aim to evaluate either a rate of correct classification or the separability/similarity of class distributions.
+    The criterion evaluates how a given model built with a subset of features performs for the classification task.
+It can be an estimation of the correct classification or a measure of separability/similarity between class distribution.
+The former are in general more demanding in terms of processing time than the later.
+
+    % Criterion functions aim to evaluate either a rate of correct classification or the separability/similarity of class distributions.
 These functions are used to estimate which sets of variables are the best to represent data, to assure the separability of the classes and to perform classification.
@@
 
-        \hspace{0pt} \\
-
-        As described in \cite{congalton2008assessing} (chapter 4), all this measures of good classification are based on an error matrix $M$ called confusion matrix which is defined so that $M_{ij}$ is the number of samples of class $i$ classified as class $j$.
-The confusion matrix allows the computation of several interesting values relatively to each class:
+        A measure of good classification is based on an error matrix $M$, or \emph{confusion matrix}~\cite[Chapter 4]{congalton2008assessing}.
+An element $M_{ij}$ of the matrix represents the number of samples of class $i$ classified as class $j$.
+The confusion matrix allows the computation of several global and per-class indices related to the classification accuracy~\cite{congalton2008assessing}.
+Three global criteria were used
         \begin{itemize}
-            \item the number of True Positive (TP) corresponding to good prediction;
-            \item the number of True Negative (TN) corresponding to good classification of the other class;
-            \item the number of False Negative (FN) corresponding to the samples of the class labeled as an other class;
-            \item the number of False Positive (FP) corresponding to the samples wrongly classified as part of of this class.
+        \item  \emph{The overall  accuracy} (OA)  is the  rate of  the
+          number of samples with the  correct predicted label over the
+          total  number of  samples~\cite{congalton2008assessing}. This
+          metric is  easy to interpret  but is  biased in the  case of
+          unbalanced classes.
+
+        \item \emph{The Cohen's kappa} (K) is a statistic which measures the probability of agreement between predictions and ground-truth~\cite{congalton2008assessing}.
+
+        \item \emph{The mean F1 score}  (F1mean) is the average of the
+          F1 score  for each class  and the  F1 score is  the harmonic
+          mean of  the precision  (number of  True Positive  over True
+          Positive plus False Positive) and the recall (number of True
+          Positive over True Positive plus False Negative)~\cite{powers2011evaluation}.
       \end{itemize}
-        Figure~\ref{fig:confusion} illustrates the definition of this values.
-
-        \begin{figure}[!ht]
-            \centering
-            \begin{tabular}{c >{\bfseries}r @{\hspace{0.7em}}c @{\hspace{0.4em}}c @{\hspace{0.4em}}c @{\hspace{0.7em}}l}
-              \multirow{10}{*}{\parbox{1.1cm}{\bfseries\raggedleft Actual\\ value}} &
-                & \multicolumn{3}{c}{\bfseries Prediction outcome} \\
-              & & \bfseries $c_1$ & \bfseries $c_2$ & \bfseries $c_3$ \\
-              & $c_1'$ & \MyBox{True}{Positive} & \MyBox{False}{Negative} & \MyBox{False}{Negative} \\[2.4em]
-              & $c_2'$ & \MyBox{False}{Positive} & \MyBox{True}{Negative} & \MyBox{~}{~} \\[2.4em]
-              & $c_3'$ & \MyBox{False}{Positive} & \MyBox{~}{~} & \MyBox{True}{Negative}
-            \end{tabular}
-            \caption{Confusion matrix with TP, TN, FP and FN relatively to $c_1$}
-            \label{fig:confusion}
-        \end{figure}
-
-        \emph{The overall accuracy} (OA) is the rate of the number of samples with the correct predicted label over the total number of samples.
-This metric is easy to interpret but is biased in the case of unbalanced classes.
-
-        \emph{The Cohen's kappa} (K) is a statistic which measures the probability of agreement between predictions and ground-truth.
-
-        \emph{The mean F1 score} (F1mean) is the average of the F1 score for each class and the F1 score is the harmonic mean of the precision (number of True Positive over True Positive plus False Positive) and the recall (number of True Positive over True Positive plus False Negative).
-
-        \begin{align}
-            OA &= \frac{\text{TP}}{n}. \\
-            K &= \frac{pa - pr}{1 - pr}.  \\
-            \text{F1mean} &= \frac{2 \text{TP}}{2 \text{TP} + \text{FN} + \text{FP}}.
-        \end{align}
-        where TP stands for True Positive, FN for False Negative, FP for False Positive, $pa$ is the probability of agreement defined by $pa = OA$  and $pr$ the probability of random agreement defined by $pr = \sum_{c=1}^{C} \frac{TP_c}{FP_c} \frac{TP_c}{FN_c}$.
 

### 2016-09-29 paper.tex
fin de la SII-A-1

@@
 
-        A measure of good classification is based on an error matrix $M$, or \emph{confusion matrix}~\cite[Chapter 4]{congalton2008assessing}.
+        A measure of correct classification is based on an error matrix $M$, or \emph{confusion matrix}~\cite[Chapter 4]{congalton2008assessing}.
 An element $M_{ij}$ of the matrix represents the number of samples of class $i$ classified as class $j$.
@@
 
-        In order to estimate classification rate, a cross-validation process over training set is used\cite{opac-b1127878}.
-The training set is divided in k folds, then, the training is done with $(k-1)$ folds and the performances are estimated with the remaining fold.
+      The classification rate is  estimated by a $k$- cross-validation
+      ($k$-CV)~\cite{opac-b1127878}.  To compute  the $k$-CV, a subset
+      is removed  from $\mathcal{S}$ and  the GMM is learned  with the
+      remaining training samples.   A test error is  computed with the
+      removed  training  samples  used  as  validation  samples.   The
+      process is  iterated $k$ times and  the estimated classification
+      rate is computed as the mean  test error over the $k$ subsets of
+      $\mathcal{S}$.
 
@@
         \subsubsection{Similarity between distributions}
-
-        \hspace{0pt} \\
-        The second type of criterion functions is a measure of similarity between two distributions.
         These measures are called divergence function and are defined so that, if $S$ is a space of all probability distribution with same support, it satisfies

### 2016-09-29 paper.tex
subsection Criterion

@@
 The confusion matrix allows the computation of several global and per-class indices related to the classification accuracy~\cite{congalton2008assessing}.
-Three global criteria were used
+Three global criteria were used :
         \begin{itemize}
@@
         \end{itemize}
+        High  value  of  theses  indices  correspond  to  an  accurate
+        classification.
 
@@
         \subsubsection{Similarity between distributions}
-        These measures are called divergence function and are defined so that, if $S$ is a space of all probability distribution with same support, it satisfies
-        \begin{align*}
-            &\forall (p,q) \in S, \text{Div}(p,q) \geq 0, \\
-            &\text{Div}(p,q) = 0 \Leftrightarrow p = q.
-        \end{align*}
-        More specifically, focus is made on two particular divergences: the Kullback–Leibler divergence and the Jeffries–Matusita distance.
-The advantage of these divergences is that they have an explicit expression in the case of Gaussian models.
-The simplification allows to get rid of any integration calculus which is a major problem when dealing with high-dimensional data.
-
-        \emph{The Kullback-Leibler divergence} (KL divergence) measures the amount of information lost when the first distribution is approximated by the second one\cite{kullback1987letter}.
-The formal definition is
-        \begin{equation}
-            \text{Div}_{KL}(c_i,c_j) = \int_\mathbf{x} p(\mathbf{x}|c_i) \ln \left(\frac{p(\mathbf{x}|c_i)}{p(\mathbf{x}|c_j)}\right) d\mathbf{x}.
-        \end{equation}
-        And in the case of Gaussian model, it can be rewritten as follows
+        The similarity  between two distributions is  quantified using
+        divergence  measure~\cite{opac-b1097517}. Contrary  to measure
+        of correct  classification, divergences are  computed directly
+        on the trained model, with no need of validation samples.  Two
+        particular   divergences   are   used  in   this   work:   the
+        \emph{Kullback–Leibler}        divergence       and        the
+        \emph{Jeffries–Matusita}  distance.  The  advantage  of  these
+        divergences is  that they have  an explicit expression  in the
+        case of Gaussian models. The  simplification allows to get rid
+        of  any integration  calculus which  is a  major problem  when
+        dealing with high-dimensional data.
+
+        \emph{The   Kullback-Leibler   divergence}   (KL   divergence)
+        measures  the  amount  of  information  lost  when  the  first
+        distribution     is     approximated     by     the     second
+        one\cite{kullback1987letter}. It can be explicitly computed in
+        the case of Gaussian distributions:
         \begin{align}
-            &\text{Div}_{KL}(c_i,c_j) = \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_{c_i}^{-1} \boldsymbol{\Sigma}_{c_j}) \nonumber \\
+            &KL_{ij} = \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_{c_i}^{-1} \boldsymbol{\Sigma}_{c_j}) \nonumber \\
             & + (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j})^t \boldsymbol{\Sigma}_{c_i}^{-1} (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j}) - d + \log \left( \frac{|\boldsymbol{\Sigma}_{c_i}|}{|\boldsymbol{\Sigma}_{c_j}|} \right) \biggr),
@@
 
-        It can be noticed that the KL divergence is not symmetric, i.e.
-$\text{Div}_{KL}(c_i,c_j) \ne \text{Div}_{KL}(c_j,c_i)$, and so the symmetrized version is used to compute the criterion function.
-In the case of Gaussian model, the symmetrization induces the following simplification of the formula
+        The    KL   divergence    is   not    symmetric,   \emph{i.e.}
+        $KL_{ij} \ne KL_{i,j}$.  A the symmetrized version  is used to
+        compute the criterion function:
         \begin{align}
-            \text{SKL}_{ij} &=\text{Div}_{KL}(c_i,c_j) + \text{Div}_{KL}(c_j,c_i) \nonumber \\
+          SKL_{ij} &=KL_{ij} + KL_{i,j} \nonumber \\
             &= \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_{c_i}^{-1} \boldsymbol{\Sigma}_{c_j} + \boldsymbol{\Sigma}_{c_j}^{-1} \boldsymbol{\Sigma}_{c_i}) \nonumber \\
@@
 
-        Moreover, the divergence is computed between two classes and to obtain a unique value the weighted mean of divergence measures is taken as proposed in \cite{bruzzone1995extension}:
+        The extension  to the multiclass  situation is done  by taking
+        the weighted mean  of the KL divergences computed  on all pair
+        of classes~\cite{bruzzone1995extension}:
         \begin{equation}
-            \text{C}_{SKL} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_{c_i} \pi_{c_j} \text{SKL}_{ij}.
+            C_{SKL} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_{c_i} \pi_{c_j} SKL_{ij}.
         \end{equation}
 
-        \emph{The Bhattacharyya distance} is defined as follows
-        \begin{equation}
-            \text{B}_{ij} = - \ln \left( \int_\mathbf{x} \sqrt{p(\mathbf{x}|c_i) p(\mathbf{x}|c_j)} d\mathbf{x} \right).
-        \end{equation}
-        And in the case of Gaussian model:
+        \emph{The Bhattacharyya distance} is defined in the case of Gaussian model as
         \begin{align}
-            \text{B}_{ij} = &\frac{1}{8} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \left( \frac{\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j}{2} \right)^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) \nonumber \\
+            {B}_{ij} = &\frac{1}{8} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \left( \frac{\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j}{2} \right)^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) \nonumber \\
             &+ \frac{1}{2} \ln \left( \frac{|\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j|}{\sqrt{|\boldsymbol{\Sigma}_i| |\boldsymbol{\Sigma}_j|}} \right).
@@
 
-        \emph{The Jeffries–Matusita distance} is a measure based on the Bhattacharyya distance.
-It saturates if the separability between the two distribution increases \cite{bruzzone2009novel}.
-The JM distance is defined according to
+        The \emph{Jeffries–Matusita distance} is a measure based on the Bhattacharyya distance.
+It saturates if the separability between the two distribution increases~\cite{bruzzone2009novel}.
+The JM distance is defined as
         \begin{equation}
-            \text{JM}_{ij} = \sqrt{ \int_\mathbf{x} \left[\sqrt{p(\mathbf{x}|c_i)} - \sqrt{p(\mathbf{x}|c_j)}\right]^2 d\mathbf{x} }.
+            {JM}_{ij} = \sqrt{ 2 \{1 - \text{exp}[-B_{ij}]\} }.
         \end{equation}
-        And the Jeffries–Matusita distance can be rewritten according to the Bhattacharyya distance
+
+        Similar to the KL divergence,  a weighted mean of the distance
+        between two classes is computed to aggregate the measures in a
+        single value:
         \begin{equation}
-            \text{JM}_{ij} = \sqrt{ 2 \{1 - \text{exp}[-B_{ij}]\} }.
+            {C}_{JM} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_{c_i} \pi_{c_j} {JM}_{ij}.
         \end{equation}
 
-        As for the KL divergence, a weighted mean of the distance between two classes is computed to aggregate the measures in a single value:
-        \begin{equation}
-            \text{C}_{JM} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_{c_i} \pi_{c_j} \text{JM}_{ij}.
-        \end{equation}
-
         \vspace{10 mm}
 
-        According to \cite{bruzzone2009novel}, it is interesting to notice that the KL divergence increases quadratically with respect to the distance between the mean vectors of the class distributions whereas the measures of correct classification asymptotically tends to one when distributions are perfectly separable.
+        % According to \cite{bruzzone2009novel}, it is interesting to notice that the KL divergence increases quadratically with respect to the distance between the mean vectors of the class distributions whereas the measures of correct classification asymptotically tends to one when distributions are perfectly separable.
 On the contrary, the JM distance tends to saturate as the measures of correct classification.
+
         Table~\ref{tab:crit} summarized the presented criterion functions and their characteristics.

### 2016-09-29 paper.tex
section serch strategy
done

@@
         \label{sec:selection:method}
+        Two  sequential  search  functions were  implemented  in  this
+        work~\cite{Guyon:2006:FEF:1208773}:  the   sequential  forward
+        selection and the sequential  floating forward.  The later one
+        is  an   extension  of   the  former.  Both   select  features
+        iteratively.
 
@@
         The Sequential Forward Selection (SFS) starts with an empty set of selected features.
-Then, it computes at each step for all the remaining features the value of the criterion function $J$ chosen among the ones presented in Table~\ref{tab:crit} when the feature is added to the pool of selected features.
-The feature that maximizes the criterion function is definitively added to the pool of selected features and it moves to the next iteration.
-The algorithm stops when a given number of variables \emph{maxVarNb} has been selected.
+At each  step, the feature that most  improves one selected criterion function $J$ selected in Table~\ref{tab:crit} is added  to the set.
+This feature is definitively added to the pool of selected features and the algorithm stops when a given number of variables \emph{maxVarNb} has been selected.
 The Algorithm~\ref{alg:sfs} presents the process in details.
-
-        The advantage of this search algorithm is its reasonable processing time.
-The trade-off is that the result is a suboptimal solution in the sense that an untested subset of variable could lead to better classification result.
 
@@
 
-        The Sequential Floating Forward Selection (SFFS)\cite{somol1999adaptive} is actually based on two algorithms: the SFS described above and the Sequential Backward Selection (SBS).
+        The Sequential Floating Forward Selection (SFFS)\cite{somol1999adaptive} is based on two algorithms: the SFS described above and the Sequential Backward Selection (SBS).
 The SBS is the backward equivalent of SFS.
@@
 
-        The SFFS works as the SFS but between each step of the SFS algorithm a backward selection is operated.
-At the end of the SBS step, the value of the criterion function is compared to the best value ever obtained with a set of features of the same size and if the new value is better the feature put into question is effectively taken away and the next step is again a SBS but if the new value is not better the SBS step is forgotten and it moves to the next SFS step.
+        The SFFS works like the SFS but between each step of the SFS algorithm, a backward selection is operated to identify the less important feature.
+If the criterion function is higher than the best value value ever obtained with a set of same size, the identified feature is picked out.
+The SBS step is repeated while removing the less important feature leads to an increase of the criterion function.
+Then SFS is going back.
 The algorithm stops when a given number of features \emph{maxVarNb} has been selected.
-The Algorithm~\ref{alg:sffs} sums up the method.
+The Algorithm~\ref{alg:sffs} provides details of the process.
 

### 2016-09-30 paper.tex
update

--- 2016-09-29 paper.tex
@@
     \end{equation*}
-    Under the GMM,  and identifying $p(c)$ as $\pi_c$  and $p(x|c)$ as
-    $f_c(x|\theta)$ and by taking the log, a decision rule is obtained
+    Under the GMM,  and identifying $p(c)$ as $\pi_c$  and $p(\mathbf{x}|c)$ as
+    $f_c(\mathbf{x}|\theta)$ and by taking the log, a decision rule is obtained
     \begin{eqnarray}\label{eq:decision}
@@
     There are two major solutions to this problem.
-First option is to stabilize the inversion of the covariance matrices work.
-Some methods investigate the use of constraints on the direct problem \cite{reynolds1995robust} \cite{hathaway1985constrained} and some work on the inverse problem by using regularization method \cite{hoerl1970ridge} \cite{jensen2008regression}.
+First option is to stabilize the inversion of the covariance matrices.
+Some methods investigate the use of constraints on the direct problem.
+Reynolds \emph{et al.} \cite{reynolds1995robust} proposed to use diagonal covariance matrices.
+It is also possible to force the diagonal element to be higher than a given value by maximizing the GMM likelihood \cite{hathaway1985constrained}.
+Celeux and Govaert \cite{celeux1995gaussian} suggest to use equality constraints between coefficients of the covariance matrix in a parsimonious cluster-based GMM framework.
+Others papers propose to work on the inverse problem.
+A classical method is to use a regularization method as the well-known ridge regularization \cite{hoerl1970ridge}.
+A ridge regularization aims to stabilize the inversion by replacing the covariance matrix $\boldsymbol{\Sigma}_c$ by $\boldsymbol{\Sigma}_c + \tau \mathbf{I}$ where $\tau$ is a positive parameter and $\mathbf{I}$ the identity matrix.
+Jensen \emph{et al.} \cite{jensen2008regression} propose a different approach using a sparsity approximation to inverse the covariance matrix.
+
     Second option is to reduce dimension.
-Features extraction/selection methods have been developed in order to reduce the dimension.
-In this  study, the latter option is explored and a feature selection method named sequential forward features selection is presented after a brief overview of the selection algorithms in the next Section.
-%A finir
+Features extraction/selection methods have been developed in order to reduce the dimension with various approaches described in Section~\ref{sec:intro}.
+In this study, this latter option is explored and a feature selection method named sequential forward features selection is presented after a brief overview of the selection algorithms in the next Section.
 
@@
 
-        A measure of correct classification is based on an error matrix $M$, or \emph{confusion matrix}~\cite[Chapter 4]{congalton2008assessing}.
+        A measure of good classification is based on an error matrix $M$, or \emph{confusion matrix}~\cite[Chapter 4]{congalton2008assessing}.
 An element $M_{ij}$ of the matrix represents the number of samples of class $i$ classified as class $j$.
@@
 
-      The classification rate is  estimated by a $k$- cross-validation
-      ($k$-CV)~\cite{opac-b1127878}.  To compute  the $k$-CV, a subset
-      is removed  from $\mathcal{S}$ and  the GMM is learned  with the
-      remaining training samples.   A test error is  computed with the
-      removed  training  samples  used  as  validation  samples.   The
-      process is  iterated $k$ times and  the estimated classification
-      rate is computed as the mean  test error over the $k$ subsets of
-      $\mathcal{S}$.
+        In order to estimate classification rate, a cross-validation process over training set is used\cite{opac-b1127878}.
+The training set is divided in k folds, then, the training is done with $(k-1)$ folds and the performances are estimated with the remaining fold.
 
@@
         \subsubsection{Similarity between distributions}
+
+        \hspace{0pt} \\
+        The second type of criterion functions is a measure of similarity between two distributions.
 These measures are called divergence function and are defined so that, if $S$ is a space of all probability distribution with same support, it satisfies

### 2016-09-30 paper.tex

--- 2016-09-29 paper.tex
@@
 Three global criteria were used :
+
         \begin{itemize}
@@
 
-      The classification rate is  estimated by a $k$- cross-validation
-      ($k$-CV)~\cite{opac-b1127878}.  To compute  the $k$-CV, a subset
-      is removed  from $\mathcal{S}$ and  the GMM is learned  with the
-      remaining training samples.   A test error is  computed with the
-      removed  training  samples  used  as  validation  samples.   The
-      process is  iterated $k$ times and  the estimated classification
-      rate is computed as the mean  test error over the $k$ subsets of
-      $\mathcal{S}$.
+        The classification rate is  estimated by a $k$- cross-validation ($k$-CV)~\cite{opac-b1127878}.
+To compute  the $k$-CV, a subset is removed  from $\mathcal{S}$ and  the GMM is learned  with the remaining training samples.
+A test error is  computed with the removed  training  samples  used  as  validation  samples.
+The process is  iterated $k$ times and  the estimated classification rate is computed as the mean  test error over the $k$ subsets of $\mathcal{S}$.
 

### 2016-09-30 paper.tex
update

@@
 It is possible to extract a set of relevant features to get a parsimonious representation of the data \cite{jimenez1998supervised}.
-
 For instance, in land-cover classification, given a set of spatial, temporal and spectral features, it is possible to extract those which are the most discriminant for the purpose of classification \cite{fassnacht2014comparison}.
 In hyperspectral data analysis from the hundreds of available spectral channels, it is possible to reduce the number of channels to make the processing more efficient in terms of statistical complexity because of the reduction of the number of parameters to estimate and thus also in term of computational time.
-Moreover, dimensional reduction might improves the capacity of generalization of the classifier and avoid overfitting.
+
+The objective of feature selection is to retrieve the most compact representation of the data with minimum loss of information \cite{Guyon:2006:FEF:1208773}.
+By deleting the redundant information, feature selection allows to reduce dimension and thus avoid statistical issue, increase algorithm speed and limit storage requirements.
+Moreover, it adds several beneficial side effects.
+It improves data understanding by identify where is the most relevant information and it is then possible to save resources when organizing a new data acquisition.
 
 There are two ways to reduce dimension \cite{Guyon:2006:FEF:1208773}: features extraction and feature selection.
-Feature extraction means creating new features by combining the existing ones, for example linear combination as in Principal Component Analysis \cite{jimenez1998supervised}.
-To the contrary, features selection selects a subset of existing features.
+Feature extraction means reformulate and summarize the information by creating new features in combining the existing ones.
+Linear combinations of the initial features can be extracted as in Principal Component Analysis \cite{jimenez1998supervised} or in independent component analysis \cite{villa2011hyperspectral}.
+Supervised extraction method has also been investigate such as Fischer discriminant analysis.
+To the contrary, features selection selects a subset of existing features identified as the most relevant for a given criterion.
 It has the advantage to be much more interpretable for the end-user.
@@
 
-There is a large diversity of methods for feature selection.
-However, they usually do not scale well with the number of pixels to be processed \cite{fauvel2015fast}.
+Features selection/extraction algorithms can be divided into three types.
+The first type, called \emph{filter methods}, is based uniquely on data analysis in the sense that the method is independant of any classifier.
+Features are ranked according to a statistical analysis of the data.
+For example, the Principal Component Analysis (PCA) described in \cite{jimenez1998supervised} is a typical unsupervised filter method.
+Bruzzone \emph{et al.}\cite{bruzzone1995extension} develop a supervised filter method based on Jeffries-Matusita distance to maximize the separability of class distribution.
+A filter based on the Pearson $\chi^2$ statistical test has been used on biomedical data\cite{biesiada2007feature}.
+The use of correlation between bands has been explored with hyperspectral data\cite{demir2008phase}.
+In general, these methods have the advantage to be fast and usable with any classifier but they do not take into account the strength and weakness of the chosen classifier and do not optimize directly the classification process.
+
+The second sort are known as \emph{wrapper methods} which can be seen as search method to determine the best subset of variables for a given learning model.
+As exhaustive searches are too expensive in terms of processing time, several search strategies have been designed.
+Some are optimal under particular hypothesis as the branch and bound algorithm \cite{narendra1977branch} optimal under the condition of a monotonic criterion function and other sub-optimal but easier to set up.
+Whitney \cite{whitney1971direct} proposed a forward search adding feature one by one \cite{whitney1971direct} when Marill and Green \cite{marill1963effectiveness} use a backward search eliminating features one by one.
+Some papers propose to combine these two search method \emph{somol1999adaptive}.
+The advantage of such methods compare to filter methods is that they are dedicated to a particular model but on the other hand, as they require the training of multiple models to test various set of variables, they tend to be more time consuming.
+The proposed in this paper is actually a wrapper method which has been optimized in order to accelerate model learning and testing which is the weak point of this family.
+
+The third type corresponds to the \emph{embedded method} which do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
+The basic example of such method is the decision tree algorithm in which a feature is selected for the creation of each node.
+Embedded methods also exist for other models, e.g.
+SVM \cite{guyon2002gene,weston2003use}.
+
+However, this large diversity of methods for feature selection usually do not scale well with the number of pixels to be processed \cite{fauvel2015fast}.
 Nevertheless, methods based on Gaussian Mixture Models (GMM) have several interesting properties that make them suitable for feature selection in the context of large amount of data.
@@
 
-    %% TODO: il faut reprendre cette partie en détaillant plus
-    %% Il y a deux approches: une qui regularise le probleme inverse,
-    %% ridge est une solution, mais il y en a d'autres. Voir le
-    %% paragraphe qui est consacré à cela dans la thèse d'Anthony ->
-    %% elle arrive pas mail. Ensuite, il y a la réduction de dimension
-    %% données des refs (voir intro du papier JSTARS) et dans la
-    %% réduction  de dimension, il  y a  la sélection de  features. La
-    %% conclusion doit s'enchainer avec la partie suivante.
     There are two major solutions to this problem.
@@
 
+    %% Question: tu suggérais de reprendre ici "Ensuite, il y a la réduction de dimension données des refs (voir intro du papier JSTARS) et dans la réduction  de dimension, il  y a  la sélection de features." Le soucis que je vois, c'est que je détaille déjà dans l'intro les différentes méthodes de sélection/extraction et je trouve que ça devient pas très clair si je reprends tout en me basant sur le papier JSTARS.
+J'aurais dit avt "il y a trois types de sélection: filtre, wrapper, embedded blablabla" et ensuite ça me fait reprendre ici une autre distinction "supervisé/non-supervisé" si je me base sur jstars qui est un peu redondante il me semble.
+Qu'en dis-tu ?
     Second option is to reduce dimension.
@@
 \label{sec:selection}
-
-%%% Debut Déplacer vers l'intro
-The objective of feature selection is to retrieve the most compact representation of the data with minimum loss of information \cite{Guyon:2006:FEF:1208773}.
-By deleting the redundant information, feature selection allows to reduce dimension and thus avoid statistical issue, increase algorithm speed and limit storage requirements.
-Moreover, it adds several beneficial side effects.
-It improves data understanding by identify where is the most relevant information and it is then possible to save resources when organizing a new data acquisition.% TODO: add ref.
-et example
-
-%% Cette partie, dans l'intro il faut la rallonger: mieux décrire ces trois méthodes, dire les avantages et les inconvénients de chaque approche et positionner notre approche par rapport à tout cela (je dis cela sans avoir relu l'intro).
-Tu peux t'inspirer de la thèse de Arnaud Lebris et des pages 58-66)
-Features selection algorithms can be divided into three types.
-The first type, called \emph{filter methods}, is based uniquely on data analysis.
-Features are ranked according to a statistical analysis of the data.
-For example, the Principal Component Analysis (PCA) described in \cite{jimenez1998supervised} is a typical filter method and there is numerous other methods \cite{bruzzone1995extension,biesiada2007feature,demir2008phase}.
-
-The second sort are known as \emph{wrapper methods} which can be seen as search method to determine the best subset of variables for a given learning model.
-As exhaustive searches are to expansive in terms of processing time, several search strategies have been designed.
-Some are optimal under particular hypothesis \cite{narendra1977branch} and other sub-optimal but easier to set up \cite{whitney1971direct,somol1999adaptive}.
-The advantage of such methods compare to filter methods is that they are dedicated to a particular model but on the other hand, as they require the training of multiple models to test various set of variables, they tend to be more time consuming.
-
-The third type corresponds to the \emph{embedded method} which do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
-The basic example of such method is the decision tree algorithm in which a feature is selected for the creation of each node.
-Embedded methods also exist for other models, e.g.
-SVM \cite{guyon2002gene,weston2003use}.
-%%% Fin déplacer
 

### 2016-10-03 paper.tex
update

@@
 % Do not put math or special symbols in the title.
-\title{Operational Feature Selection in Gaussian Mixture Models}
+\title{Operational Feature Selection with Gaussian Mixture Models for remote sensing images}
 %
@@
 \author{Adrien~Lagrange,~Mathieu~Fauvel~and~Manuel~Grizonnet% <-this % stops a space
-\thanks{A. Lagrange and M. Fauvel are with the Université de Toulouse,
+\thanks{A. Lagrange and M. Fauvel are with the Universit\'{e} de Toulouse,
 INP-ENSAT, UMR 1201 DYNAFOR, France and with the INRA, UMR 1201
 DYNAFOR, France.}
-\thanks{M. Grizonnet is with Centre Nat. d'Etudes Spatiales, French Space Agency, Toulouse, France.}}
+\thanks{M. Grizonnet is with Centre National d'\'{E}tudes Spatiales, French Space Agency, Toulouse, France.}}
 
@@
 
-Finally, the method is tested and also compared to other classifiers using two different datasets, one of hyperspectral images with a lot of spectral variables and one with heterogeneous spatial features.
+Finally, the method is compared to standard classifiers using two different datasets, one of hyperspectral images with a lot of spectral variables and one with heterogeneous spatial features.
 The results validate the fact that the method performs well in terms of processing time and classification accuracy in comparison to the standard classifiers available in OTB.\\
@@
 
-However, processing this data is more and more challenging because of statistical and computational issues.
+However, processing such data is more and more challenging because of statistical and computational issues.
 This statistical issues are often referred as the \emph{curse of dimensionality}.
-The Hughes phenomenon \cite{hughes1968mean} states that with a given number of samples, prediction accuracy will decays when the complexity is higher than some optimum value.
+The Hughes phenomenon \cite{hughes1968mean} states that with a given number of samples, prediction accuracy will decays when the complexity, \emph{i.e.} the dimension, is higher than some optimum value.
 The problem is the fast increase of the number of parameters to estimate in order to build a model when dimension expands \cite{bouveyron2014model}.
-Thus, a important number of labeled samples is needed to perform learning.
+Thus, an important number of labeled samples is needed to perform learning.
 For example, in the case of Gaussian Mixture Models, the number of parameters progresses quadratically with the dimension, e.g.
@@
 
-The computational issues are multiple.
+The computational issues are also multiple.
 The computing infrastructure needed to process data is more and more expensive because the processing might requires a GPU and a large amount of RAM to load images which can weight a dozen of gigabytes \cite{christophe2011remote}\cite{plaza2011high}.
@@
 For instance, in land-cover classification, given a set of spatial, temporal and spectral features, it is possible to extract those which are the most discriminant for the purpose of classification \cite{fassnacht2014comparison}.
-In hyperspectral data analysis from the hundreds of available spectral channels, it is possible to reduce the number of channels to make the processing more efficient in terms of statistical complexity because of the reduction of the number of parameters to estimate and thus also in term of computational time.
+In hyperspectral data, from the hundreds of available spectral channels, it is possible to reduce the number of channels to make the processing more efficient in terms of statistical complexity because of the reduction of the number of parameters to estimate and thus also in term of computational time.
 
@@
 Feature extraction means reformulate and summarize the information by creating new features in combining the existing ones.
-Linear combinations of the initial features can be extracted as in Principal Component Analysis \cite{jimenez1998supervised} or in independent component analysis \cite{villa2011hyperspectral}.
+Linear combinations of the initial features can be extracted using for example in Principal Component Analysis \cite{jimenez1998supervised} or Independent Component Analysis \cite{villa2011hyperspectral}.
 Supervised extraction method has also been investigate such as Fischer discriminant analysis.
@@
 As exhaustive searches are too expensive in terms of processing time, several search strategies have been designed.
-Some are optimal under particular hypothesis as the branch and bound algorithm \cite{narendra1977branch} optimal under the condition of a monotonic criterion function and other sub-optimal but easier to set up.
+Some are optimal under particular hypothesis as the branch and bound algorithm \cite{narendra1977branch} optimal under the condition of a monotonic criterion function.
+Others are sub-optimal but easier to set up.
 Whitney \cite{whitney1971direct} proposed a forward search adding feature one by one \cite{whitney1971direct} when Marill and Green \cite{marill1963effectiveness} use a backward search eliminating features one by one.
-Some papers propose to combine these two search method \emph{somol1999adaptive}.
+Some papers propose to combine these two search method \cite{somol1999adaptive}.
 The advantage of such methods compare to filter methods is that they are dedicated to a particular model but on the other hand, as they require the training of multiple models to test various set of variables, they tend to be more time consuming.
-The proposed in this paper is actually a wrapper method which has been optimized in order to accelerate model learning and testing which is the weak point of this family.
+The method proposed in this paper is actually a wrapper method which has been optimized in order to accelerate model learning and testing which is the weak point of this family.
 
@@
 
-However, this large diversity of methods for feature selection usually do not scale well with the number of pixels to be processed \cite{fauvel2015fast}.
+Despite a large diversity of methods, feature selection algorithms usually do not scale well with the number of pixels to be processed \cite{fauvel2015fast}.
 Nevertheless, methods based on Gaussian Mixture Models (GMM) have several interesting properties that make them suitable for feature selection in the context of large amount of data.
@@
 
-Several strategy have been explored to perform feature selection and, among them, wrapper methods receive a certain interest of the community.
+% Several strategy have been explored to perform feature selection and, among them, wrapper methods receive a certain interest of the community.
 Wrapper methods can be seen as a search method to determine the best subset of variables for a given learning model.
@@
 The remaining of this article is organized as follows.
-Section~\ref{sec:gmm-hd} presents Gaussian Mixture Model classifiers and the features selection methods used to make them suitable for high-dimension space.
-The work done to develop a smart implementation of the proposed selection method is presented in Section~\ref{sec:implementation}.
-And finally, the tests conducted to explore variations of the algorithm and to compare it to other standard classifiers are detailed in Section~\ref{sec:test}.
+Section~\ref{sec:gmm-hd} presents Gaussian Mixture Model classifiers and the problems link to high-dimensionality.
+The features selection methods used to make them suitable for high-dimension space is detailed in Section~\ref{sec:selection}.
+The work done to develop an efficient implementation of the proposed selection method is presented in Section~\ref{sec:implementation}.
+And finally, after introducing the two datasets in Section~\ref{sec:datasets}, the tests conducted to explore variations of the algorithm and to compare it to other standard classifiers are shown and commented in Section~\ref{sec:test}.
 
@@
     \end{align}
-
     To predict the  class of a new unseen sample,  the maximum \emph{a
@@
     The computation of~\ref{eq:decision} requires the inversion of the
-    covariance matrix and and the  computation of the logarithm of the
-    determinant.  The estimations of these terms suffer from the curse
+    covariance matrix and the  computation of the logarithm of the
+    determinant.  The estimation of these terms suffers from the curse
     of  dimensionality\cite{bouveyron2014model}, because  in practice,
@@
     illustrated in Figure~\ref{fig:nb-param}.  Hence, if the number of
-    observation  $n_c$ is  small  compared to  the  number of  samples
-    $n_c$, the  estimated covariance  matrix is badly  conditioned and
+    observation  $n_c$ is  small  compared to  the  number of parameters
+    $\eta_c$, the  estimated covariance  matrix is badly  conditioned and
     thus the computation  of its inverse and its  determinant would be
     unstable.   The worst  situation  is $n_c<\eta_c$  which leads  to
-    singular covariance matrix.  Unfortunately, this situation happens
+    a singular covariance matrix.  Unfortunately, this situation happens
     regularly in remote sensing.   For instance in hyperspectral image
     classification,  very few  labeled samples  are usually  available
-    because of the difficulty and the cost to collect ground-truth.
+    because of the difficulty and the high cost to collect ground-truth.
 
@@
 
-        The classification rate is  estimated by a $k$- cross-validation ($k$-CV)~\cite{opac-b1127878}.
-To compute  the $k$-CV, a subset is removed  from $\mathcal{S}$ and  the GMM is learned  with the remaining training samples.
+        The classification rate is  estimated by a $n_{cv}$- cross-validation ($n_{cv}$-CV)~\cite{opac-b1127878}.
+To compute the $n_{cv}$-CV, a subset is removed  from $\mathcal{S}$ and  the GMM is learned with the remaining training samples.
 A test error is computed with the removed  training  samples  used  as  validation  samples.
-The process is  iterated $k$ times and  the estimated classification rate is computed as the mean  test error over the $k$ subsets of $\mathcal{S}$.
+The process is  iterated $n_{cv}$ times and  the estimated classification rate is computed as the mean  test error over the $n_{cv}$ subsets of $\mathcal{S}$.
 
@@
         particular   divergences   are   used  in   this   work:   the
-        \emph{Kullback–Leibler}        divergence       and        the
-        \emph{Jeffries–Matusita}  distance.  The  advantage  of  these
+        \emph{Kullback-Leibler}        divergence       and        the
+        \emph{Jeffries-Matusita}  distance.  The  advantage  of  these
         divergences is  that they have  an explicit expression  in the
@@
         \begin{align}
-            &KL_{ij} = \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_{c_i}^{-1} \boldsymbol{\Sigma}_{c_j}) \nonumber \\
-            & + (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j})^t \boldsymbol{\Sigma}_{c_i}^{-1} (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j}) - d + \log \left( \frac{|\boldsymbol{\Sigma}_{c_i}|}{|\boldsymbol{\Sigma}_{c_j}|} \right) \biggr),
+            &KL_{ij} = \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j) \nonumber \\
+            & + (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \boldsymbol{\Sigma}_i^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - d + \log \left( \frac{|\boldsymbol{\Sigma}_i|}{|\boldsymbol{\Sigma}_j|} \right) \biggr),
         \end{align}
@@
         The    KL   divergence    is   not    symmetric,   \emph{i.e.}
-        $KL_{ij} \ne KL_{i,j}$.  A the symmetrized version  is used to
+        $KL_{ij} \ne KL_{ji}$.  A symmetrized version  is used to
         compute the criterion function:
@@
           SKL_{ij} &=KL_{ij} + KL_{i,j} \nonumber \\
-            &= \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_{c_i}^{-1} \boldsymbol{\Sigma}_{c_j} + \boldsymbol{\Sigma}_{c_j}^{-1} \boldsymbol{\Sigma}_{c_i}) \nonumber \\
-            &~~+ (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j})^t (\boldsymbol{\Sigma}_{c_i}^{-1} + \boldsymbol{\Sigma}_{c_j}^{-1}) (\boldsymbol{\mu}_{c_i} - \boldsymbol{\mu}_{c_j}) - 2d \biggr).
+            &= \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j + \boldsymbol{\Sigma}_j^{-1} \boldsymbol{\Sigma}_i) \nonumber \\
+            &~~+ (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t (\boldsymbol{\Sigma}_i^{-1} + \boldsymbol{\Sigma}_j^{-1}) (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - 2d \biggr).
         \end{align}
@@
         \begin{equation}
-            C_{SKL} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_{c_i} \pi_{c_j} SKL_{ij}.
+            C_{SKL} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_i \pi_j SKL_{ij}.
         \end{equation}
@@
         \begin{equation}
-            {C}_{JM} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_{c_i} \pi_{c_j} {JM}_{ij}.
+            {C}_{JM} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_i \pi_j {JM}_{ij}.
         \end{equation}
@@
             \caption{Summary of the different criterion functions.\label{tab:crit}}
-            \begin{tabular}[b]{lccc}
+            \begin{tabular}[b]{lcc}
               \toprule
-              Criterion & Type Divergence & Complexity \\
+              Criterion & Type & Complexity \\
               \midrule
@@
         \label{sec:selection:method}
-        Two  sequential  search  functions were  implemented  in  this
+        Two  sequential  search  algorithms were  implemented  in  this
         work~\cite{Guyon:2006:FEF:1208773}:  the   sequential  forward
@@
         The Sequential Forward Selection (SFS) starts with an empty set of selected features.
-At each  step, the feature that most  improves one selected criterion function $J$ selected in Table~\ref{tab:crit} is added  to the set.
+At each  step, the feature that most  improves the criterion function $J$ selected in Table~\ref{tab:crit} is added  to the set.
 This feature is definitively added to the pool of selected features and the algorithm stops when a given number of variables \emph{maxVarNb} has been selected.
@@
         The SFFS works like the SFS but between each step of the SFS algorithm, a backward selection is operated to identify the less important feature.
-If the criterion function is higher than the best value value ever obtained with a set of same size, the identified feature is picked out.
-The SBS step is repeated while removing the less important feature leads to an increase of the criterion function.
-Then SFS is going back.
+If the criterion value is higher than the best value ever obtained with a set of same size, the identified feature is picked out.
+The SBS step is repeated while removing the less important feature leads to an increase of the criterion value.
+Then SFS is called again.
 The algorithm stops when a given number of features \emph{maxVarNb} has been selected.
@@
     \subsection{Statistical update rules}
-    In the presented method, the most costly operation is the evaluation of the criterion function which needs to be done for each possible subset of variables augmented by one of the candidate variable.
-To compute these values, the most costly operation are the inversion of the covariance matrix and the computation of its determinant.
+    In the presented method, the most time-consuming operation is the evaluation of the criterion function which needs to be done for each possible subset of variables composed of the selected set augmented by one of the candidate variable.
+To compute these values, the most costly operations are the inversion of the covariance matrix and the computation of its determinant.
 Both of these operations have a $O(n^3)$ computational complexity.
-Additionaly, the estimation of classification rate with cross-validation is a heavy process which take a consequent amount of time.
+Additionally, the estimation of classification rate with cross-validation is a heavy process which takes a substantial amount of time.
 
     A major contribution of this work is the optimization in term of computational efficiency.
-More precisely, three steps of the proposed method has been upgraded in order to reduce computational time.
+More precisely, three steps of the proposed method have been upgraded in order to reduce computational time.
 First of all, the GMM model is learned only once using samples.
@@
     Secondly, when a cross-validation is performed, submodels, i.e.
-GMM model trained with $(n_{cv}-1)$ folds, are not learned using the $(n_{cv}-1)$ folds but, instead, are derived from the global model and from the covariance matrices and mean vectors of the fold used for validation for the given submodel.
+GMM trained with $(n_{cv}-1)$ folds, are not learned using the $(n_{cv}-1)$ folds but, instead, are derived from the global model and from the covariance matrices and mean vectors of the fold used for validation for the given submodel.
 The process is described in details in Section~\ref{sec:update-cv}.
 
-    Finally and most importantly, when variables are tested one by one during a selection step, costly operations are made when computing criterion functions for each variable especially the computation of the inverse of covariance matrices and its determinant.
-Several update rules are set up which allow to compute this inverse and determinant only once to test all variables.
+    Finally and most importantly, when variables are tested one by one during a selection step, costly operations are made when computing criterion functions for each candidate set especially the computation of the inverse of covariance matrices and its determinant.
+Several update rules are set up which allow to compute this inverse and determinant only once to test all sets.
 These update rules are presented in Section~\ref{sec:update-crit}.
@@
         Based on \cite{fauvel2015fast}, a method to accelerate the $n_{cv}$-fold cross-validation process in the case of criterion functions based on correct classification measures was implemented.
-The idea is to estimate the GMM model with the whole training set once and then, instead of training models on $(n_{cv}-1)$ folds, parameters of the complete model are used to derive those of submodels.
+The idea is to estimate the GMM with the whole training set once and then, instead of training models on $(n_{cv}-1)$ folds, parameters of the complete model are used to derive those of submodels.
 
@@
         \end{prop}
-        where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exposants on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\nu}_c$ denotes the set of samples used to compute them.
+        \noindent where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exposants on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\nu}_c$ denotes the set of samples used to compute them.
 
@@
                 \bigg[\begin{array}{cc}
-                A & v \\
-                v^t  & \frac{1}{\alpha} \\
+                \mathbf{A} & \mathbf{v} \\
+                \mathbf{v}^t  & \frac{1}{\alpha} \\
                 \end{array}\bigg]
@@
         \end{prop}
-        where $A = (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $v = - \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$.
-Then, $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is computed only once and this update update formulae is used to obtain all the $(\boldsymbol{\Sigma}^{(k)})^{-1}$ corresponding to all the possible augmented set of a given selection iteration.
+        \noindent where $\mathbf{A} = (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $\mathbf{v} = - \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$.
+Then, $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is computed only once and this update formulae is used to obtain all the $(\boldsymbol{\Sigma}^{(k)})^{-1}$ corresponding to all the possible augmented set of a given selection iteration.
 
@@
         \label{eq:skl-update}
-            &\text{SKL}_{ij} = \frac{1}{2} \biggl( \underbrace{\text{Tr} \left((\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_j}^{(k)} + (\boldsymbol{\Sigma}_{c_j}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_i}^{(k)}\right)}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
-            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
-            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2 - 2k}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
-            &+ \underbrace{(\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})^t ((\boldsymbol{\Sigma}_{c_i}^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_{c_j}^{(k-1)})^{-1}) (\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \biggr),
+            &\text{SKL}_{ij} = \frac{1}{2} \biggl( \underbrace{\text{Tr} \left((\boldsymbol{\Sigma}_i^{(k)})^{-1} \boldsymbol{\Sigma}_j^{(k)} + (\boldsymbol{\Sigma}_j^{(k)})^{-1} \boldsymbol{\Sigma}_i^{(k)}\right)}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
+            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
+            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2 - 2k}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
+            &+ \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ((\boldsymbol{\Sigma}_i^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_j^{(k-1)})^{-1}) (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \biggr),
         \end{align}
-        with $(\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
+        with $(\boldsymbol{\Sigma}_i^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
 
@@
 
-    The developed method has been made available as a remote module of Orfeo toolbox called \emph{otbFFSforGMM} and can be download on Github\footnote{\url{https://github.com/Laadr/otbExternalFastFeaturesSelection}}.
+    The developed method has been made available as a remote module of Orfeo toolbox called \emph{otbFFSforGMM} and can be download on Github\footnote{\url{https://github.com/Laadr/otbFFSforGMM}}.
 
@@
 It is to note that the toolbox already implements several classifiers inherited mostly from OpenCV.
-Nevertheless, the GMM classifier currently in place is not robust to high dimensionality and that is why a new GMM model has been developed.
-
-    In addition of the C++ class directly available to develop C++ code, three command line application has been produced.
-These applications allow the user to use and parametrize the algorithm as desire.
+Nevertheless, the GMM classifier currently in place is not robust to high dimensionality and that is why a new GMM model has been developed in the module.
+
+    In addition of the C++ class directly available to develop C++ code, three commandline applications have been produced.
+These applications allow the user to use and parametrize the algorithm as desired.
 
     The first application \emph{otbcli\_TrainGMMApp} trains a GMM classifier from multiple pairs of images and training vector data and creates a model file.
-It also allows to perform ridge regularization with an embedded selection of the regularization parameter among a list of proposed values.
-The regularization values is chosen by maximizing a criterion function (Cohen's kappa, overall accuracy or mean of F1 scores) estimated by cross-validation.
-
-    The second application \emph{otbcli\_TrainGMMSelectionApp} also trains a GMM classifier and in addition perform a feature selection algorithm (SFS, SFFS) with various possible criterion (Jeffries-Matusita dist., Kullback-Lieber div., Cohen’s kappa, overall accuracy, mean F1-score).
-
-    The third application \emph{otbcli\_PredictGMMApp} takes a raster image as input and a model file and perform the classification of the image.
+It also allows to perform ridge regularization with an embedded selection of the regularization parameter among a list of values proposed by the user.
+The regularization value is chosen by maximizing a criterion function (Cohen's kappa, overall accuracy or mean of F1 scores) estimated by cross-validation.
+
+    The second application \emph{otbcli\_TrainGMMSelectionApp} also trains a GMM classifier and in addition performs a feature selection algorithm (SFS, SFFS) with various possible criterion (Jeffries-Matusita dist., Kullback-Lieber div., Cohen’s kappa, overall accuracy, mean F1-score).
+
+    The third application \emph{otbcli\_PredictGMMApp} takes a raster image and a model file as input and performs the classification of the image.
 It generates a classification map in an image file and optionally a confidence map.
@@
     The classification  rate is presented  using Cohen's kappa  but it
-    has to  be noticed the  results were  similar when looking  at the
+    has to  be noticed that the  results were  similar when looking  at the
     mean F1-score.  Processing time  has been  evaluated on  a desktop
@@
 In this case, using JM distance as criterion and SFS as search strategy is the best choice in term of time efficiency.
-The good performance in time can explain by the dimensional reduction.
+The good performance in time can be explained by the dimension reduction.
 Actually, the decision rule corresponding to Equation~\ref{eq:decision} has a complexity in $d^3$ where $d$ is the dimension.
@@
     The processing times of the three standard classifiers suffer from the increase of training samples.
-For the GMM classifier with ridge, the selection of regularization parameter is more costly with more samples because the classification rate estimation needed.
+For the GMM classifier with ridge, the selection of regularization parameter is more costly with more samples because of the classification rate estimation needed.
 For the kNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file.
@@
 When using 1000 samples, no significant difference of classification rate has been observed on test set.
-But, with 50,000 samples, the Random Forest classifier better significantly better.
+But, with 50,000 samples, the Random Forest classifier becomes significantly better.
 
     It can be guessed that the fact that the Gaussian hypothesis is unlikely to be verified explains the difficulty to extract the finer information available in a larger dataset.
-The JM distance has stressed in Table~\ref{tab:crit} has a strong Gaussian hypothesis and the same hypotheses could also explain the drop of the classification rate of GMM classifiers with Jeffries-Matusita distance as criterion.
+The JM distance has stressed in Table~\ref{tab:crit} has a strong Gaussian hypothesis and the same hypotheses could also explain the drop of the classification rate of GMM classifiers with JM distance as criterion.
 
@@
 
-Additionally, from the comparison between all the measures used to rank the features, it has been shown that the Jeffries-Matusita distance can be a fast and accurate solution in some case but suffer from some limitations which remain to be determine.
-A possible explaination could be that this distance does not perform well when data distribution is not similar to a Gaussian distribution.
+Additionally, from the comparison between all the measures used to rank the features, it has been shown that the Jeffries-Matusita distance can be a fast and accurate solution in some case but suffer from some limitations which remain to be determine more precisely.
+A possible explanation could be that this distance does not perform well when data distribution is not similar to a Gaussian distribution.
 More investigation is needed to confirm this hypothesis.
@@
 
-Finally, experiments show that the developed GMM classifier performs at least as best as standard classifiers without features selection in particular Random Forest and even outperforms all of them in term of classification time.
-The next step would be to compare to other feature selection methods.
+Finally, experiments show that the developed GMM classifier performs, in most cases, at least as best as standard classifiers without features selection in particular Random Forest and even outperforms all of them in term of classification time.
 
 The second achievement of this work is the development of an efficient implementation using GMM properties to derive fast update rules of the model.
-The resulting code is available as a remote module of the Orfeo toolbox on Github and make it possible to process huge quantity of high dimension data.
+The resulting code is available as a remote module of the Orfeo toolbox on Github and makes it possible to process huge quantity of high dimension data.
 

### 2016-10-05 paper.tex
Section IV
- modification de la section IV.A

@@
     Under the GMM,  and identifying $p(c)$ as $\pi_c$  and $p(\mathbf{x}|c)$ as
-    $f_c(\mathbf{x}|\theta)$ and by taking the log, a decision rule is obtained
+    $f_c(\mathbf{x}|\theta)$ and by taking the log, a decision function is obtained
     \begin{eqnarray}\label{eq:decision}
@@
         compute the criterion function:
-        \begin{align}
+        \begin{align}\label{eq:skl}
           SKL_{ij} &=KL_{ij} + KL_{i,j} \nonumber \\
@@
 The JM distance is defined as
-        \begin{equation}
+        \begin{equation}\label{eq:jm}
             {JM}_{ij} = \sqrt{ 2 \{1 - \text{exp}[-B_{ij}]\} }.
@@
 \label{sec:implementation}
+The most  demanding part  of the  algorithm is  the evaluation  of the
+criterion  for  all   the  remaining  variables  (see   lines  5-7  in
+Algorithm~\ref{alg:sfs}). Calculus are based on linear algebra, and the
+numerical  complexity is  on average  $O(d^3)$.  Furthermore,  for the
+\emph{accuracy}-type  criterion the  complexity  is  augmented by  the
+cross-validation procedure.
+
+An efficient implementation of the criterion optimization is proposed.
+It based on  the symmetry property of the covariance  matrix and block
+inverse  formula~\cite{IMM2012-03274}. In  the following  it is  shown
+that the criterion  can be split into  two part: one that  needs to be
+computed for each  tested variable, and one that needs  to be computed
+only once per  selection step. For the  cross-validation part, updates
+rules are given to derive sub-models  without the necessity to learn a
+GMM models for each fold.
 
 \subsection{Statistical update rules}
-    In the presented method, the most time-consuming operation is the evaluation of the criterion function which needs to be done for each possible subset of variables composed of the selected set augmented by one of the candidate variable.
-To compute these values, the most costly operations are the inversion of the covariance matrix and the computation of its determinant.
-Both of these operations have a $O(n^3)$ computational complexity.
-Additionally, the estimation of classification rate with cross-validation is a heavy process which takes a substantial amount of time.
-
-    A major contribution of this work is the optimization in term of computational efficiency.
-More precisely, three steps of the proposed method have been upgraded in order to reduce computational time.
-First of all, the GMM model is learned only once using samples.
-When a covariance matrix or a mean vector of a reduced set of variables is required, it is obtained from the global model learned at the beginning by marginalization.
-
-    Secondly, when a cross-validation is performed, submodels, i.e.
-GMM trained with $(n_{cv}-1)$ folds, are not learned using the $(n_{cv}-1)$ folds but, instead, are derived from the global model and from the covariance matrices and mean vectors of the fold used for validation for the given submodel.
-The process is described in details in Section~\ref{sec:update-cv}.
-
-    Finally and most importantly, when variables are tested one by one during a selection step, costly operations are made when computing criterion functions for each candidate set especially the computation of the inverse of covariance matrices and its determinant.
-Several update rules are set up which allow to compute this inverse and determinant only once to test all sets.
-These update rules are presented in Section~\ref{sec:update-crit}.
-
         \subsubsection{Update for cross validation}
         \label{sec:update-cv}
-
         Based on \cite{fauvel2015fast}, a method to accelerate the $n_{cv}$-fold cross-validation process in the case of criterion functions based on correct classification measures was implemented.
@@
 
-        The following formulae can be obtained
+        A sub-GMM model can be estimated at a reduced complexity when one fold is remove from the training set.
+From~\cite{fauvel2015fast}, the following formulae for unbiased estimators can be derived: %% TODO: rajouter des \hat{} partout: ce sont des estimateurs.
         \begin{prop}
@@
         \label{sec:update-crit}
-
-        At each iteration the SFS and SFFS algorithms compute the value of a criterion function for every possible set of features composed of the selected ones augmented by one of the remaining features.
-One of the main achievements of this work is to reduce the computational time in factorizing the computation of the inverse and determinant of the covariance matrix..
-
-        In the remaining of the paper, $\boldsymbol{\Sigma}^{(k-1)}$ is the covariance matrix of the $(k-1)^{th}$ iteration, i.e., the covariance matrix of the selected features and $\boldsymbol{\Sigma}^{(k)}$ is a covariance matrix at the $k^{th}$ iteration, i.e., the covariance matrix of an augmented set.
-Then, the inverse of the covariance matrix $(\boldsymbol{\Sigma}^{(k)})^{-1}$, the quadratical term $(\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}$ and the determinant $\log |\boldsymbol{\Sigma}^{(k)}|$ can be expressed in function of terms of the $(k-1)^{th}$ iteration.
-These calculi use the fact that the covariance matrix is a positive definite symmetric matrix to simplify formulae using block matrices \cite{webb2003statistical} (chapter 9.2).
-These update rules are summed up hereafter.
-
-        As $\boldsymbol{\Sigma}^{(k)}$ is a positive definite symmetric matrix, the following notation can be used
-        \begin{equation*}
+        At iteration $k$, depending on the criterion, three terms have
+        to  be computed:  the inverse  of the  covariance matrix,  the
+        logarithm of the determinant of  the covariance matrix and the
+        quadratical term in  eq.~\ref{eq:decision}. However, all these
+        terms  have already  been computed  for iteration  $(k-1)$. By
+        using  the  fact that  the  covariance  matrix is  a  positive
+        definite     symmetric    matrix     and    block     matrices
+        formulae~\cite[Chapter   9.2]{webb2003statistical},    it   is
+        possible to factorize these terms at iteration $k$.
+
+        In the  remaining of the  paper, $\boldsymbol{\Sigma}^{(k-1)}$
+        denotes the  covariance matrix of the  $(k-1)^{th}$ iteration,
+        \emph{i.e.}, the  covariance matrix  of the  selected features
+        and $\boldsymbol{\Sigma}^{(k)}$ denotes a covariance matrix at
+        the $k^{th}$ iteration, \emph{i.e.},  the covariance matrix of
+        an augmented set. Then, since $\boldsymbol{\Sigma}^{(k)}$ is a
+        positive definite symmetric matrix,  the covariance matrix can
+        be written as
+
+        % Then, the inverse of  the covariance matrix
+        % $(\boldsymbol{\Sigma}^{(k)})^{-1}$,   the   quadratical   term
+        % $(\mathbf{x}^{(k)})^t         (\boldsymbol{\Sigma}^{(k)})^{-1}
+        % \mathbf{x}^{(k)}$        and        the        log-determinant
+        % $\log   |\boldsymbol{\Sigma}^{(k)}|$  can   be  expressed   in
+        % function of terms of the $(k-1)^{th}$ iteration.
+
+        \begin{equation}\label{eq:cov}
             \boldsymbol{\Sigma}^{(k)} =
@@
             \end{array}\bigg],
-        \end{equation*}
-        where $\mathbf{u}$ is the $k^{th}$ column of the matrix without the diagonal element i.e. $\mathbf{u}_{i} = \boldsymbol{\Sigma}^{(k)}_{i,k}$ with $i \in [1,k-1]$.
-
-        Using the formula of the inverse of a block matrix, the following formula expressing $(\boldsymbol{\Sigma}^{(k)})^{-1}$ in function of $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is obtained
+        \end{equation}
+        where  $\mathbf{u}$  is  the  $k^{th}$ column  of  the  matrix
+        without      the      diagonal      element,      \emph{i.e.},
+        $\mathbf{u}_{i}   =    \boldsymbol{\Sigma}^{(k)}_{i,k}$   with
+        $i  \in [1,k-1]$.  Using  block matrix  inverse formulae,  the
+        following proposition holds:
         \begin{prop}
@@
             (Forward update rule for inverse of covariance matrix)
-            \begin{equation*}
+            \begin{equation}\label{eq:cov:inv}
                 (\boldsymbol{\Sigma}^{(k)})^{-1} =
@@
                 \end{array}\bigg]
-            \end{equation*}
+            \end{equation}
         \end{prop}
-        \noindent where $\mathbf{A} = (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $\mathbf{v} = - \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$.
-Then, $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is computed only once and this update formulae is used to obtain all the $(\boldsymbol{\Sigma}^{(k)})^{-1}$ corresponding to all the possible augmented set of a given selection iteration.
-
-        In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$.
+        \noindent                                                where
+        $\mathbf{A}     =     (\boldsymbol{\Sigma}^{(k-1)})^{-1}     +
+        \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}
+        \mathbf{u}^t              (\boldsymbol{\Sigma}^{(k-1)})^{-1}$,
+        $\mathbf{v}           =           -           \frac{1}{\alpha}
+        (\boldsymbol{\Sigma}^{(k-1)})^{-1}       \mathbf{u}$       and
+        $      \alpha      =      \sigma_{kk}      -      \mathbf{u}^t
+        (\boldsymbol{\Sigma}^{(k-1)})^{-1}  \mathbf{u}$.  This  update
+        formulae      is     used      to      obtain     all      the
+        $(\boldsymbol{\Sigma}^{(k)})^{-1}$  corresponding  to all  the
+        possible augmented set of a given selection iteration. Similar
+        update formulae can be written for the backward step.
+
+        % In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$.
 The update rule becomes (proof in Appendix~\ref{app:proof-update})
-        \begin{prop}
-        \label{eq:update-inv-back}
-            (Backward update rule for inverse of covariance matrix)
-            \begin{equation*}
-                (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \underbrace{\mathbf{A}}_{\substack{\text{computed once}\\ \text{per selection step}}} - \underbrace{\alpha \mathbf{v} \mathbf{v}^t}_{\substack{\text{computed for} \\ \text{each augmented set}}}
-            \end{equation*}
-        \end{prop}
-
-        A formula can also be deduced for the quadratical term of the decision function.
-Noting $(\mathbf{x}^{(k)})^t = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x^k \end{array}\right]$, the following proposition is obtained (proof in Appendix~\ref{app:proof-update})
+        % \begin{prop}
+        % \label{eq:update-inv-back}
+        %     (Backward update rule for inverse of covariance matrix)
+        %     \begin{equation*}
+        %         (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \underbrace{\mathbf{A}}_{\substack{\text{computed once}\\ \text{per selection step}}} - \underbrace{\alpha \mathbf{v} \mathbf{v}^t}_{\substack{\text{computed for} \\ \text{each augmented set}}}
+        %     \end{equation*}
+        % \end{prop}
+
+        Using eq.~\ref{eq:cov} and~\ref{eq:cov:inv}, it is possible to
+        deduce  the   following  propositions  (proof  are   given  in
+        Appendix~\ref{app:proof-update}).
         \begin{prop}
@@
             (Update rule for quadratical term)
-            \begin{align*}
-                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once}\\ \text{per selection step}}} \\
-                &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}}
-            \end{align*}
+            \begin{align}
+                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
+                &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
+            \end{align}
         \end{prop}
-
-        Finally, using the formula of the determinant of a block matrix and after taking the log, a last formula is produced
         \begin{prop}
@@
             (Update rule for logdet)
-            \begin{equation*}
-                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)}_{\substack{\text{computed once}\\ \text{per selection step}}} + \underbrace{\log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}}
-            \end{equation*}
+            \begin{equation}
+                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)}_{\substack{\text{computed once}\\ \text{per selection step}}} + \underbrace{\log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
+            \end{equation}
         \end{prop}
 
-        Now using all update rules, criterion functions can be rewritten as follows
+        From these update rules, it is now possible to split each criterion into two parts: one computed once per selection step and one computed for each augmented set.
+
+        \begin{prop}[Decision function~\ref{eq:decision}]
         \begin{align}
@@
             &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
-            &- \underbrace{\alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 - \log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}};
+            &- \underbrace{\alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 - \log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
         \end{align}
-
+      \end{prop}
+
+      \begin{prop}[Kullback-Leibler divergence~\ref{eq:skl}]
         \begin{align}
         \label{eq:skl-update}
-            &\text{SKL}_{ij} = \frac{1}{2} \biggl( \underbrace{\text{Tr} \left((\boldsymbol{\Sigma}_i^{(k)})^{-1} \boldsymbol{\Sigma}_j^{(k)} + (\boldsymbol{\Sigma}_j^{(k)})^{-1} \boldsymbol{\Sigma}_i^{(k)}\right)}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
+            & SKL_{ij} = \frac{1}{2} \biggl( \underbrace{\text{Tr} \left((\boldsymbol{\Sigma}_i^{(k)})^{-1} \boldsymbol{\Sigma}_j^{(k)} + (\boldsymbol{\Sigma}_j^{(k)})^{-1} \boldsymbol{\Sigma}_i^{(k)}\right)}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
             &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
             &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2 - 2k}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
-            &+ \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ((\boldsymbol{\Sigma}_i^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_j^{(k-1)})^{-1}) (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \biggr),
+            &+ \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ((\boldsymbol{\Sigma}_i^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_j^{(k-1)})^{-1}) (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \biggr).
         \end{align}
         with $(\boldsymbol{\Sigma}_i^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
-
+      \end{prop}
+
+      \begin{prop}[Jeffries-Matusita distance~\ref{eq:jm}]
       \begin{align}
@@
         where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j$.
+      \end{prop}
 
@@
 \appendices
-\section{}
+\section{Proof of update rules}
 \label{app:proof-update}
-\subsection{Proof of update rules}
-    \subsubsection{Proposition~\ref{eq:update-inv-back}}
-        \begin{proof}
-            \begin{align*}
-                \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
-                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
-                &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
-                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
-            \end{align*}
-        \end{proof}
-
-    \subsubsection{Proposition~\ref{eq:update-quad}}
-        \begin{proof}
+    % \subsubsection{Proposition~\ref{eq:update-inv-back}}
+    %     \begin{proof}
+    %         \begin{align*}
+    %             \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
+    %             &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
+    %             &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
+    %             &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
+    %         \end{align*}
+    %     \end{proof}
+
+\begin{proof}[Proof of proposition~\ref{eq:update-quad}]
   \begin{align*}

### 2016-10-05 paper.tex
Section IV.B

@@
     \subsection{Numerical issues}
-
-    During the selection process, numerical issues needs to be handled due to the computation of the inverse and determinant of the covariance matrix.
-As pointed out before, the lack of training samples can induce a badly-conditioned matrix with very small eigenvalues.
-Even with a well-estimated covariance matrix, it is still possible to encounter very correlated features resulting also in very low eigenvalues.
-The issue with these very small eigenvalues is that they reach the computational precision allowed by the computer especially when computing the logarithm of the determinant.
-
-    In order to deal with this limitation, the choice has been made to perform an explicit decomposition in eigenvalues and eigenvectors of the covariance matrix s.t.
-$\boldsymbol{\Sigma} = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$ with $\boldsymbol{\Lambda}$ the diagonal matrix of eigenvalues and $\mathbf{Q}$ the matrix of eigenvectors.
-The advantage of an explicit decomposition is the possibility to check the eigenvalues and, if some appears to be below computational precision, to threshold them at EPS\_FLT which is the computer precision for a floating number.
-For the same reason, the constant $\alpha$ used in update rules is also thresholded to EPS\_FLT.
-Algorithm~\ref{alg:sffs-update} details when computational stability is checked.
+    For each  iteration $k$, after  the selection of  optimal features
+    w.r.t  the  selected criterion,  the  inverses  of the  covariance
+    matrices and their log-determinant  needs to be computed. However,
+    the lack of training samples or the highly correlated features may
+    induce  a  badly-conditioned  matrix  with  very  small,  or  even
+    negative, eigenvalues.   Such values will degrade  drastically the
+    estimation of the  inverse and of the log-determinant,  and so the
+    numerical stability.
+
+    To deal with this limitation, the choice has been made to perform an eigendecomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}$:
+
+    \begin{eqnarray}
+    \boldsymbol{\Sigma}^{(k)} = \mathbf{Q}^{(k)} \boldsymbol{\Lambda}^{(k)} (\mathbf{Q}^{(k)})^t\label{eq:eigendecomp}
+    \end{eqnarray}
+
+    where  $\boldsymbol{\Lambda}^{(k)}$ and  $\mathbf{Q}^{(k)}$ are  the
+    diagonal matrix  of eigenvalues of  the covariance matrix  and the
+    orthonormal matrix of corresponding eigenvectors, respectively. To
+    prevent numerical instability, the  eigenvalues are thresholded to
+    EPS\_FLT. In our  implementation, EPS\_FLT is set  to the floating
+    machine precision.
+
+    Then, the inverse of the covariance matrix can be computed as
+
+    \begin{eqnarray}
+      \boldsymbol{\Sigma}^{(k)} = \mathbf{Q}^{(k)} (\tilde{\boldsymbol{\Lambda}}^{(k)})^{-1} (\mathbf{Q}^{(k)})^t\label{eq:eigendecomp:inv}
+    \end{eqnarray}
+
+    and the log-determinant as
+
+    \begin{eqnarray}
+      \label{eq:log:det}
+      \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_i^{(k)}),
+    \end{eqnarray}
+    where the \textasciitilde{} indicated thresholded values.
+
+    Same reasoning applied  to the term $\alpha$ in  the update rules:
+    it       is       also        thresholded       to       EPS\_FLT.
+    Algorithm~\ref{alg:sffs-update}    details   when    computational
+    stability is enforced (lines 6, 10, 20 and 24).
 

### 2016-10-06 paper.tex
update

@@
         A sub-GMM model can be estimated at a reduced complexity when one fold is remove from the training set.
-From~\cite{fauvel2015fast}, the following formulae for unbiased estimators can be derived: %% TODO: rajouter des \hat{} partout: ce sont des estimateurs.
-        \begin{prop}
+From~\cite{fauvel2015fast}, the following formulae for unbiased estimators can be derived:
+        \begin{prop}[Cross-validation mean update]
             \label{eq:update-cv1}
-            (Cross-validation mean update)
             \begin{equation*}
-                \boldsymbol{\mu}_c^{n_c-\nu_c} = \frac{n_c \boldsymbol{\mu}_c^{n_c} - \nu_c \boldsymbol{\mu}_c^{\nu_c}}{n_c - \nu_c} \nonumber
+                \boldsymbol{\hat{\mu}}_c^{n_c-\nu_c} = \frac{n_c \boldsymbol{\hat{\mu}}_c^{n_c} - \nu_c \boldsymbol{\hat{\mu}}_c^{\nu_c}}{n_c - \nu_c} \nonumber
             \end{equation*}
         \end{prop}
-        \begin{prop}
+        \begin{prop}[Cross-validation covariance matrix update]
             \label{eq:update-cv2}
-            (Cross-validation covariance matrix update)
             \begin{align*}
-                \boldsymbol{\Sigma}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl( (n_c-1) \boldsymbol{\Sigma}_c^{n_c} - (\nu_c-1) \boldsymbol{\Sigma}_c^{\nu_c} \nonumber \\
-                &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})(\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})^t \biggr) \nonumber
+                \boldsymbol{\hat{\Sigma}}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl( (n_c-1) \boldsymbol{\hat{\Sigma}}_c^{n_c} - (\nu_c-1) \boldsymbol{\hat{\Sigma}}_c^{\nu_c} \nonumber \\
+                &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})(\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})^t \biggr) \nonumber
             \end{align*}
@@
         following proposition holds:
-        \begin{prop}
+        \begin{prop}[Forward update rule for inverse of covariance matrix]
         \label{eq:update-inv}
-            (Forward update rule for inverse of covariance matrix)
             \begin{equation}\label{eq:cov:inv}
@@
 The update rule becomes (proof in Appendix~\ref{app:proof-update})
-        % \begin{prop}
+        % \begin{prop}[Backward update rule for inverse of covariance matrix]
         % \label{eq:update-inv-back}
-        %     (Backward update rule for inverse of covariance matrix)
         %     \begin{equation*}
@@
         Appendix~\ref{app:proof-update}).
-        \begin{prop}
+        \begin{prop}[Update rule for quadratical term]
         \label{eq:update-quad}
-            (Update rule for quadratical term)
             \begin{align}
@@
         \end{prop}
-        \begin{prop}
+        \begin{prop}[Update rule for logdet]
         \label{eq:update-log}
-            (Update rule for logdet)
             \begin{equation}
@@
 Processing times are given in second.\label{tab:potsdam-otbsimu-big}}
-        \begin{tabular}{|lcccccc}\toprule
+        \begin{tabular}{lcccccc}\toprule
             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-7}
@@
 \begin{proof}[Proof of proposition~\ref{eq:update-quad}]
-  \begin{align*}
-    (&\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}
+    \begin{alignat*}{3}
+    (&\mathbf{x}^{(k)})^t && (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}
        = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
@@
                                                                                    \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
-     &= \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
+     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
                                                                                        \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
-     &= (\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
-     &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
-     &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
-     &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \alpha \mathbf{v} \mathbf{v}^t) \mathbf{x}^{(k-1)}\\
-     &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
-     &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} \mathbf{v}^t \mathbf{x}^{(k-1)} \\
-     &~~~+ 2 \frac{x_k}{\alpha} \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha^2}) \\
-     &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} )^2 \\
-     &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
+     &= &&(\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
+     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
+     & &&+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
+     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \alpha \mathbf{v} \mathbf{v}^t) \mathbf{x}^{(k-1)}\\
+     & &&+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
+     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} \mathbf{v}^t \mathbf{x}^{(k-1)} \\
+     & &&+ 2 \frac{x_k}{\alpha} \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha^2}) \\
+     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} )^2 \\
+     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
+    \end{alignat*}
+\end{proof}
+
+\begin{proof}[Proof of proposition~\ref{eq:update-log}]
+    \begin{align*}
+        \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) =
+        &\log \left(\left|
+            \bigg[\begin{array}{cc}
+                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{u}      \\
+                \mathbf{u}^t          & \sigma_{kk} \\
+            \end{array}\bigg]
+        \right|\right) \\
+        &\log \left(\left|
+            \bigg[\begin{array}{cc}
+                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
+                \mathbf{u}^t          & 1 \\
+            \end{array}\bigg]
+            \bigg[\begin{array}{cc}
+                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
+                \mathbf{0}_{k-1,1}^t          & \sigma_{kk} - \mathbf{u}^t(\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}\\
+            \end{array}\bigg]
+        \right|\right) \\
+        &\log \left(\left|
+            \bigg[\begin{array}{cc}
+                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
+                \mathbf{u}^t          & 1 \\
+            \end{array}\bigg]
+        \right|\right)
+        +\log \left(\left|
+            \bigg[\begin{array}{cc}
+                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
+                \mathbf{0}_{k-1,1}^t          & \alpha \\
+            \end{array}\bigg]
+        \right|\right) \\
+        &\log \left( |\boldsymbol{\Sigma}^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha| \right)\\
+        &\log \left( |\boldsymbol{\Sigma}^{(k-1)}| \right) + \log ( \alpha )
     \end{align*}

### 2016-10-06 paper.tex
Fin section IV
+ passage en double colonne

--- 2016-10-05 paper.tex
@@
-\documentclass[journal,peerreview,onecolumn]{IEEEtran}
-% \documentclass[journal]{IEEEtran}
+%\documentclass[journal,peerreview,onecolumn]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 \usepackage{etex}
@@
 \usepackage{booktabs}
-
+\usepackage{minted}\definecolor{bg}{rgb}{0.95,0.95,0.95}
 \newcommand\MyBox[2]{
@@
 % for peerreview
-\usepackage{setspace}
-\onehalfspacing
+% \usepackage{setspace}
+% \onehalfspacing
 
@@
 
-\tableofcontents
-\clearpage
+% \tableofcontents
+% \clearpage
 % As a general rule, do not put math, special symbols or citations
@@
 
-Finally, the method is compared to standard classifiers using two different datasets, one of hyperspectral images with a lot of spectral variables and one with heterogeneous spatial features.
+Finally, the method is compared to standard classifiers using two different data-sets, one of hyperspectral images with a lot of spectral variables and one with heterogeneous spatial features.
 The results validate the fact that the method performs well in terms of processing time and classification accuracy in comparison to the standard classifiers available in OTB.\\
@@
 Features selection/extraction algorithms can be divided into three types.
-The first type, called \emph{filter methods}, is based uniquely on data analysis in the sense that the method is independant of any classifier.
+The first type, called \emph{filter methods}, is based uniquely on data analysis in the sense that the method is independent of any classifier.
 Features are ranked according to a statistical analysis of the data.
@@
 Bruzzone \emph{et al.}\cite{bruzzone1995extension} develop a supervised filter method based on Jeffries-Matusita distance to maximize the separability of class distribution.
-A filter based on the Pearson $\chi^2$ statistical test has been used on biomedical data\cite{biesiada2007feature}.
+A filter based on the Pearson $\chi^2$ statistical test has been used on bio-medical data\cite{biesiada2007feature}.
 The use of correlation between bands has been explored with hyperspectral data\cite{demir2008phase}.
@@
 The work done to develop an efficient implementation of the proposed selection method is presented in Section~\ref{sec:implementation}.
-And finally, after introducing the two datasets in Section~\ref{sec:datasets}, the tests conducted to explore variations of the algorithm and to compare it to other standard classifiers are shown and commented in Section~\ref{sec:test}.
+And finally, after introducing the two data-sets in Section~\ref{sec:datasets}, the tests conducted to explore variations of the algorithm and to compare it to other standard classifiers are shown and commented in Section~\ref{sec:test}.
 
@@
         \begin{tikzpicture}
-          \begin{semilogyaxis}[xmin=0,xmax=200,ymin=0,width=0.5*\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = $d$,ylabel = $\eta_c$]
+          \begin{axis}[ymode=log,xmin=0,xmax=200,ymin=0,width=0.8\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = $d$,ylabel = $\eta_c$,small]
             \addplot+[very thick,mark=none,smooth,domain=0:200,samples=200] (\x,{\x*(\x+3)/2+1});
-            \end{semilogyaxis}
+            \end{axis}
         \end{tikzpicture}
@@
         \begin{align}
-            &KL_{ij} = \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j) \nonumber \\
-            & + (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \boldsymbol{\Sigma}_i^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - d + \log \left( \frac{|\boldsymbol{\Sigma}_i|}{|\boldsymbol{\Sigma}_j|} \right) \biggr),
+            &KL_{ij} = \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j) \nonumber \\
+            & + (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \boldsymbol{\Sigma}_i^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - d + \log \left( \frac{|\boldsymbol{\Sigma}_i|}{|\boldsymbol{\Sigma}_j|} \right) \biggr\},
         \end{align}
@@
           SKL_{ij} &=KL_{ij} + KL_{i,j} \nonumber \\
-            &= \frac{1}{2} \biggl( \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j + \boldsymbol{\Sigma}_j^{-1} \boldsymbol{\Sigma}_i) \nonumber \\
-            &~~+ (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t (\boldsymbol{\Sigma}_i^{-1} + \boldsymbol{\Sigma}_j^{-1}) (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - 2d \biggr).
+            &= \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j + \boldsymbol{\Sigma}_j^{-1} \boldsymbol{\Sigma}_i) \nonumber \\
+            &~~+ (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t (\boldsymbol{\Sigma}_i^{-1} + \boldsymbol{\Sigma}_j^{-1}) (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - 2d \biggr\}.
         \end{align}
@@
         \begin{equation}\label{eq:jm}
-            {JM}_{ij} = \sqrt{ 2 \{1 - \text{exp}[-B_{ij}]\} }.
+            {JM}_{ij} = \sqrt{ 2 \{1 - \text{exp}(-B_{ij})\} }.
         \end{equation}
@@
               \midrule
-              Kullback-Leibler divergence & Divergence & Low \\
+              Kullback-Leibler divergences & Divergence & Low \\
               Jeffries-Matusita distance  & Divergence & Low \\
@@
         Based on \cite{fauvel2015fast}, a method to accelerate the $n_{cv}$-fold cross-validation process in the case of criterion functions based on correct classification measures was implemented.
-The idea is to estimate the GMM with the whole training set once and then, instead of training models on $(n_{cv}-1)$ folds, parameters of the complete model are used to derive those of submodels.
+The idea is to estimate the GMM with the whole training set once and then, instead of training models on $(n_{cv}-1)$ folds, parameters of the complete model are used to derive those of sub-models.
 
@@
             \begin{align*}
-                \boldsymbol{\Sigma}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl( (n_c-1) \boldsymbol{\Sigma}_c^{n_c} - (\nu_c-1) \boldsymbol{\Sigma}_c^{\nu_c} \nonumber \\
-                &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})(\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})^t \biggr) \nonumber
+                \boldsymbol{\Sigma}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl\{ (n_c-1) \boldsymbol{\Sigma}_c^{n_c} - (\nu_c-1) \boldsymbol{\Sigma}_c^{\nu_c} \nonumber \\
+                &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})(\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})^t \biggr\} \nonumber
             \end{align*}
         \end{prop}
-        \noindent where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exposants on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\nu}_c$ denotes the set of samples used to compute them.
+        \noindent where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exponents on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\nu}_c$ denotes the set of samples used to compute them.
 
@@
         logarithm of the determinant of  the covariance matrix and the
-        quadratical term in  eq.~\ref{eq:decision}. However, all these
+        quadratic  term in  eq.~\ref{eq:decision}. However,  all these
         terms  have already  been computed  for iteration  $(k-1)$. By
@@
         \label{eq:update-inv}
-            (Forward update rule for inverse of covariance matrix)
+            (Forward update rule for the inverse of the covariance matrix)
             \begin{equation}\label{eq:cov:inv}
@@
         \label{eq:update-quad}
-            (Update rule for quadratical term)
+            (Update rule for the quadratic term)
             \begin{align}
-                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
-                &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
+              (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once per selection step}}} \nonumber \\
+                                                                                       &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for each augmented set}}}.
             \end{align}
@@
         \label{eq:skl-update}
-            & SKL_{ij} = \frac{1}{2} \biggl( \underbrace{\text{Tr} \left((\boldsymbol{\Sigma}_i^{(k)})^{-1} \boldsymbol{\Sigma}_j^{(k)} + (\boldsymbol{\Sigma}_j^{(k)})^{-1} \boldsymbol{\Sigma}_i^{(k)}\right)}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
-            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
-            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2 - 2k}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
-            &+ \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ((\boldsymbol{\Sigma}_i^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_j^{(k-1)})^{-1}) (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \biggr).
+            & SKL_{ij} = \frac{1}{2} \biggl\{ \underbrace{\text{\normalfont Tr} \left((\boldsymbol{\Sigma}_i^{(k)})^{-1} \boldsymbol{\Sigma}_j^{(k)} + (\boldsymbol{\Sigma}_j^{(k)})^{-1} \boldsymbol{\Sigma}_i^{(k)}\right)}_{\substack{\text{computed for each augmented set}}} \nonumber \\
+            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2}_{\substack{\text{computed for each augmented set}}} \nonumber \\
+            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2 - 2k}_{\substack{\text{computed for each augmented set}}} \nonumber \\
+            &+ \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t (\boldsymbol{\Sigma}_i^{(k-1)})^{-1}(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber\\
+          & + \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t(\boldsymbol{\Sigma}_j^{(k-1)})^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \biggr\}.
         \end{align}
@@
         \label{eq:jm-update}
-            \text{B}_{ij} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
-            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
-            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right)}_{\substack{\text{computed for} \\ \text{each augmented set}}},
+            \text{B}_{ij} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
+            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
+            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right)}_{\substack{\text{computed for each augmented set}}},
       \end{align}
@@
 
-    \subsection{OTB external module}
+    \subsection{Implementation}
     \label{sec:otb-module}
 
-    The developed method has been made available as a remote module of Orfeo toolbox called \emph{otbFFSforGMM} and can be download on Github\footnote{\url{https://github.com/Laadr/otbFFSforGMM}}.
-
-    The Orfeo Toolbox is an open-source library for remote sensing image processing.
-This library allows developers to code external modules and make them available to the community.
-It is to note that the toolbox already implements several classifiers inherited mostly from OpenCV.
-Nevertheless, the GMM classifier currently in place is not robust to high dimensionality and that is why a new GMM model has been developed in the module.
-
-    In addition of the C++ class directly available to develop C++ code, three commandline applications have been produced.
-These applications allow the user to use and parametrize the algorithm as desired.
-
-    The first application \emph{otbcli\_TrainGMMApp} trains a GMM classifier from multiple pairs of images and training vector data and creates a model file.
-It also allows to perform ridge regularization with an embedded selection of the regularization parameter among a list of values proposed by the user.
-The regularization value is chosen by maximizing a criterion function (Cohen's kappa, overall accuracy or mean of F1 scores) estimated by cross-validation.
-
-    The second application \emph{otbcli\_TrainGMMSelectionApp} also trains a GMM classifier and in addition performs a feature selection algorithm (SFS, SFFS) with various possible criterion (Jeffries-Matusita dist., Kullback-Lieber div., Cohen’s kappa, overall accuracy, mean F1-score).
-
-    The third application \emph{otbcli\_PredictGMMApp} takes a raster image and a model file as input and performs the classification of the image.
-It generates a classification map in an image file and optionally a confidence map.
-It also allows to precise manually the number of features to use for prediction if the number automatically set does not fit the user requirement (evolution of criterion function is available in model file).
-
-\section{Datasets}
+    The proposed  method has been implemented  in \texttt{C++} through
+    the  Orfeo  Toolbox (OTB)~\cite{christophe2008orfeo}.   The  Orfeo
+    Toolbox  is  an  open-source  library  for  remote  sensing  image
+    processing,  developed by  the  French Space  Agency (CNES).   The
+    feature selection algorithm  can be installed as  a remote module,
+    the           source            code           is           freely
+    available\footnote{\url{https://github.com/Laadr/otbFFSforGMM}}.
+
+    Following OTB framework, two application are available. The first,
+    called  \texttt{otbcli\_TrainGMMSelectionApp}, performs  a feature
+    selection algorithm (SFS or SFFS) with the one given criterion (to
+    choose  between  Jeffries-Matusita  dist.,  Kullback-Lieber  div.,
+    Cohen’s  kappa, overall  accuracy,  mean  F1-score).  The  second,
+    called \texttt{otbcli\_PredictGMMApp}, generates the thematic maps
+    according to the  learn model. The only  technical limitation that
+    the  training set  must  fit  in the  RAM  of  the computer.   The
+    classification step is streamed and there is no limitation in term
+    of image  size. The Figure~\ref{fig:otb:ffs} shows  a code excerpt
+    to run the application.
+
+    \begin{figure}\centering
+
+    \begin{minted}[fontsize=\tiny,bgcolor=bg]{sh}
+otbcli_TrainGMMSelectionApp -io.il hyper.tif \
+			    -io.vd reference.shp \
+			    -gmm.varnb 20 -gmm.method forward -gmm.crit jm\
+			    -gmm.best 1 -gmm.seed 0\
+			    -io.out model.txt
+
+otbcli_PredictGMMApp -in hyper.tif \
+		     -model model.txt -modeltype selection\
+		     -out ThematicMap.tif
+    \end{minted}
+    \caption{OTB Module: The feature selection is done on the image \emph{hyper.tif} using the training set from \emph{reference.shp}.
+The feature selection algorithm is the forward search used with the Jeffries-Matusita criterion, 20 features are extracted and the corresponding GMM model is saved in \emph{model.txt}.
+Then the whole image is classified using the model and the results is saved in the geotiff \emph{ThematicMap.tif}.}
+    \label{fig:otb:ffs}
+  \end{figure}
+\section{Data-sets}
 \label{sec:datasets}
-Numerical experiments  have been conducted on  two different datasets.
+Numerical experiments  have been conducted on  two different data-sets.
 The first  dataset, called  \emph{Aisa}, is an  airborne hyperspectral
@@
     the number of pixel per  class.  The Figure~\ref{fig:aisa} shows a
-    colored composition of the image and the groundtruth.
+    colored composition of the image and the ground-truth.
 
@@
         \end{tabular}
-        \caption{Aisa dataset: {\bfseries{(a)}} colored composition of the image (R: 634nm, G: 519nm, B: 477nm), {\bfseries{(b)}} groundtruth.\label{fig:aisa}}
+        \caption{Aisa dataset: {\bfseries{(a)}} colored composition of the image (R: 634nm, G: 519nm, B: 477nm), {\bfseries{(b)}} ground-truth.\label{fig:aisa}}
     \end{figure}
@@
         \centering
-        \caption{Repartition of classes in Aisa dataset.\label{tab:aisa}}
+        \caption{Information classes for the Aisa dataset.\label{tab:aisa}}
         \begin{tabular}[b]{lc}\toprule
@@
         \centering
-        \begin{tabular}{ccc}
+        \begin{tabular}{c@{~}c@{~}c}
             \includegraphics[width=0.3\columnwidth]{Fig/top_potsdam_5_11_RGB.jpg} &
@@
         \centering
-        \caption{Repartition of classes in Aisa dataset.\label{tab:potsdam}}
+        \caption{Information classes for the three tiles of the Postdam dataset.\label{tab:potsdam}}
         \begin{tabular}[b]{lrrr}\toprule
-          Class & \# of samples in 5\_11  & \#  of samples in 5\_12  & \#  of samples in 3\_10 \\
+          & \multicolumn{3}{c}{Number of samples per tile}\\
+            \cmidrule{2-4}
+            Class &   5\_11  &   5\_12  & 3\_10 \\
           \midrule
@@
         \item A k-nearest-neighbors classifier (KNN) with OTB default parameters (32 as number of neighbors).
-        \item A Random Forest classifier with parameters optimized by gridsearch (200 trees, 40 as max depth, 50 as size of the randomly selected subset of features at each tree node)
-        \item A GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by gridsearch.
+        \item A Random Forest classifier with parameters optimized by grid search (200 trees, 40 as max depth, 50 as size of the randomly selected subset of features at each tree node)
+        \item A GMM classifier with ridge regularization (GMM ridge) with regularization constant optimized by grid search.
     \end{itemize}
@@
         \begin{tabular}{lccc}\toprule
-             & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cline{2-4}
+             & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-4}
             \# samples by class & 250 & 500 & 1000 \\ \midrule
@@
             \toprule
-             & \multicolumn{3}{c}{\bfseries Training time (s)} & \multicolumn{3}{c}{\bfseries Classification time (s)} \\ \cline{2-7}
+             & \multicolumn{3}{c}{\bfseries Training time (s)} & \multicolumn{3}{c}{\bfseries Classification time (s)} \\ \cmidrule{2-7}
             \# samples by class & 250 & 500 & 1000 & 250 & 500 & 1000 \\ \midrule
@@
 Random Forest has similar performance in term of classification rate with 1000 samples.
-The GMM classifier with ridge regularization and the kNN classifier are both outperformed.
+The GMM classifier with ridge regularization and the KNN classifier are both outperformed.
 
@@
 For the GMM classifier with ridge, the selection of regularization parameter is more costly with more samples because of the classification rate estimation needed.
-For the kNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file.
+For the KNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file.
 Finally, for the Random Forest classifier, the trees tends to be more deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.
@@
 
-    \begin{table}[!t]
+    \begin{table*}[!t]
         \centering
@@
         \begin{tabular}{lcccccc}\toprule
-            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-7}
+            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train.
+Time} & {\bfseries Classif.
+time} & {\bfseries \# of selected features} \\ \cmidrule{2-7}
             GMM SFS kappa & 0.694 (0.002)             & {\bfseries 0.669 (0.005)} & {\bfseries 0.533 (0.008)} & 400 & 310  & 13.2 \\
@@
         \end{tabular}
-    \end{table}
-
-    \begin{table}[!t]
+    \end{table*}
+
+    \begin{table*}[!t]
         \centering
@@
 Processing times are given in second.\label{tab:potsdam-otbsimu-big}}
-        \begin{tabular}{|lcccccc}\toprule
-            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cline{2-7}
+        \begin{tabular}{lcccccc}\toprule
+            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train.
+Time} & {\bfseries Classif.
+time} & {\bfseries \# of selected features} \\ \cmidrule{2-7}
             GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 0.531 (0.005) & 20000 & 340 & 29 \\
@@
         \end{tabular}
-    \end{table}
+    \end{table*}
 
@@
 
-\begin{IEEEbiography}{Michael Shell}
-Biography text here.
-\end{IEEEbiography}
-
-% if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{John Doe}
-Biography text here.
-\end{IEEEbiographynophoto}
+% \begin{IEEEbiography}{Michael Shell}
+% Biography text here.
+% \end{IEEEbiography}
+
+% % if you will not have a photo at all:
+% \begin{IEEEbiographynophoto}{John Doe}
+% Biography text here.
+% \end{IEEEbiographynophoto}
 

### 2016-10-06 paper.tex

@@
         A sub-GMM model can be estimated at a reduced complexity when one fold is remove from the training set.
-From~\cite{fauvel2015fast}, the following formulae for unbiased estimators can be derived: %% TODO: rajouter des \hat{} partout: ce sont des estimateurs.
-        \begin{prop}
+From~\cite{fauvel2015fast}, the following formulae for unbiased estimators can be derived:
+        \begin{prop}[Cross-validation mean update]
             \label{eq:update-cv1}
-            (Cross-validation mean update)
             \begin{equation*}
-                \boldsymbol{\mu}_c^{n_c-\nu_c} = \frac{n_c \boldsymbol{\mu}_c^{n_c} - \nu_c \boldsymbol{\mu}_c^{\nu_c}}{n_c - \nu_c} \nonumber
+                \boldsymbol{\hat{\mu}}_c^{n_c-\nu_c} = \frac{n_c \boldsymbol{\hat{\mu}}_c^{n_c} - \nu_c \boldsymbol{\hat{\mu}}_c^{\nu_c}}{n_c - \nu_c} \nonumber
             \end{equation*}
         \end{prop}
-        \begin{prop}
+        \begin{prop}[Cross-validation covariance matrix update]
             \label{eq:update-cv2}
-            (Cross-validation covariance matrix update)
             \begin{align*}
-                \boldsymbol{\Sigma}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl\{ (n_c-1) \boldsymbol{\Sigma}_c^{n_c} - (\nu_c-1) \boldsymbol{\Sigma}_c^{\nu_c} \nonumber \\
-                &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})(\boldsymbol{\mu}_c^{\nu_c}-\boldsymbol{\mu}_c^{n_c})^t \biggr\} \nonumber
+              \boldsymbol{\hat{\Sigma}}_c^{n_c-\nu_c} = &\frac{1}{n_c-\nu_c-1} \biggl\{ (n_c-1) \boldsymbol{\hat{\Sigma}}_c^{n_c} - (\nu_c-1)\boldsymbol{\hat{\Sigma}}_c^{\nu_c} \nonumber \\
+                                                        &- \frac{n_c \nu_c}{(n_c-\nu_c)} (\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})(\boldsymbol{\hat{\mu}}_c^{\nu_c}-\boldsymbol{\hat{\mu}}_c^{n_c})^t \biggr\} \nonumber
             \end{align*}
@@
         following proposition holds:
-        \begin{prop}
+        \begin{prop}[Forward update rule for the inverse of the covariance matrix]
         \label{eq:update-inv}
-            (Forward update rule for the inverse of the covariance matrix)
         \begin{equation}\label{eq:cov:inv}
@@
         Appendix~\ref{app:proof-update}).
-        \begin{prop}
+        \begin{prop}[Update rule for the quadratical term]
         \label{eq:update-quad}
-            (Update rule for the quadratic term)
             \begin{align}
@@
 
-    The Algorithm~\ref{alg:sffs-update} illustrates the optimization of the Algorithm~\ref{alg:sffs} developed using these formulae.
+    The  Algorithm~\ref{alg:sffs-update} illustrates  the optimization
+    of the Algorithm~\ref{alg:sffs} using these formulae.
 

### 2016-10-07 paper.tex
complete tab

--- 2016-10-06 paper.tex
@@
             GMM ridge &     0.632 (0.007)             & 0.592 (0.010)             & 0.433 (0.008)             & 10  & 2000 & all \\
-            KNN &           0.637 (0.005)             & 0.607 (0.005)             & 0.478 (0.002)             &     &      & all \\
+            KNN &           0.637 (0.005)             & 0.607 (0.005)             & 0.478 (0.002)             & 0.7 & 9500 & all \\
             Random Forest & {\bfseries 0.729 (0.004)} & {\bfseries 0.673 (0.005)} & {\bfseries 0.529 (0.009)} & 20  & 840  & all \\

### 2016-10-07 paper.tex
improve indent
Petite modif

--- 2016-10-06 paper.tex
@@
             \label{eq:q-update}
-                Q(\mathbf{x}) = &- \underbrace{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
-                &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
-                &- \underbrace{\alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 - \log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
+                Q(\mathbf{x}) = &- \underbrace{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
+                &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once per selection step}}} \nonumber \\
+                &- \underbrace{\alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 - \log \alpha}_{\substack{\text{computed for each augmented set}}}.
             \end{align}
@@
 
-    \begin{minted}[fontsize=\tiny,bgcolor=bg]{sh}
+    \begin{minted}[fontsize=\scriptsize,bgcolor=bg]{sh}
 otbcli_TrainGMMSelectionApp -io.il hyper.tif \
 			    -io.vd reference.shp \
-			    -gmm.varnb 20 -gmm.method forward -gmm.crit jm\
-			    -gmm.best 1 -gmm.seed 0\
+			    -gmm.varnb 20 \
+                            -gmm.method forward \
+                            -gmm.crit jm \
+			    -gmm.best 1 -gmm.seed 0 \
 			    -io.out model.txt
@@
 otbcli_PredictGMMApp -in hyper.tif \
-		     -model model.txt -modeltype selection\
+                     -model model.txt \
+                     -modeltype selection \
 		     -out ThematicMap.tif

### 2016-10-07 paper.tex

--- 2016-10-07 paper.tex
@@
 
-    \begin{figure}\centering
-
+    \begin{figure}
+        \centering
         \begin{minted}[fontsize=\tiny,bgcolor=bg]{sh}
@@
     \end{figure}
+
 \section{Data-sets}

### 2016-10-07 paper.tex
Section VI.

@@
     \begin{figure} 
-        \centering
+        \centering %Pas touche l'indentation !
         \begin{minted}[fontsize=\tiny,bgcolor=bg]{sh} 
@@
     dataset is composed of aerial images of the urban area of Potsdam.
-    The area is cut into 38  patches of 6000$\times$6000 pixels with a
-    resolution  of 5cm  by pixel  and 4  channels are  available: Red,
-    Blue, Green  and Infrared (RGBIR).   A Digital Surface  Model with
-    the same  resolution is also  provided and a  so-called normalized
-    DSM representing the height above  ground. The ground-truth for 24
-    tiles  are   provided  with   6  classes:  Low   vegetation,  High
-    vegetation, Impervious surfaces,  Buildings, Cars, Clutter.  Three
-    tiles  have  been  used  in  this  work,  they  are  displayed  in
-    Figure~\ref{fig:potsdam-expl}.  Table~\ref{tab:potsdam} summarizes
-    the  number of  samples of  each class.
+    The area  is divided  into 38  patches of  6000$\times$6000 pixels
+    with a  resolution of 5cm by  pixel and 4 channels  are available:
+    Red, Blue,  Green and Infrared  (RGBIR).  A Digital  Surface Model
+    with  the  same  resolution  is  also  provided  and  a  so-called
+    normalized  DSM   representing  the   height  above   ground.  The
+    ground-truth  for  24  tiles  are provided  with  6  classes:  Low
+    vegetation, High vegetation, Impervious surfaces, Buildings, Cars,
+    Clutter.   Three tiles  have  been  used in  this  work, they  are
+    displayed             in            Figure~\ref{fig:potsdam-expl}.
+    Table~\ref{tab:potsdam} summarizes  the number of samples  of each
+    class.
 
@@
         \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)~\cite{dalla2010morphological}.
-        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features)~\cite{otb}.
+        \item Textural features for  each channel with neighborhood of
+          19x19 pixels:  mean, standard  deviation, range  and entropy
+          (16 features)~\cite{otb}.
     \end{itemize}
@@
     the method has been already compared to other selection methods in
-    \cite{fauvel2015fast}.  Hence,  the primary  objective is  to test
+    \cite{fauvel2015fast}.  Hence, the primary  objective is to assess
     the  operational  efficiency  and  it is  compared  to  other  OTB
@@
     \end{itemize}
+    Other configurations  have been  investigated and  performs either
+    equally     or     lower     in    terms     of     classification
+    accuracy~\cite{al:report}. For the sake of clarity, only the three
+    aforementioned configurations are discussed here.
 
@@
     polygons used  for the validation.   Several size of  training set
-    have been tested investigated.   For the Aisa dataset, experiments
-    have been conducted  using 250, 500 and 1000 samples  by class and
-    for the Potsdam dataset, 1000 and 50000 samples by class.
-
-    For SFS and  SFFS selection, the number of variables to select is set to 30 for the Aisa dataset and 60 for the Potsdam dataset.
-Then, a small processing is made to select the number of variables used for prediction.
+    have been  tested.  For  the Aisa  dataset, experiments  have been
+    conducted using  250, 500 and  1000 samples  by class and  for the
+    Potsdam dataset, 1000 and 50000 samples by class.
+
+    For SFS and  SFFS selection, the number of variables  to select is
+    set to  30 for the  Aisa dataset and  60 for the  Potsdam dataset.
+    After  the selection  procedure,  the optimal  number of  selected
+    variable is selected  as follow. Rather than  selecting the number
+    of variables corresponding to the  highest value oft he criterion,
+    the number of  retained variables is set when  the criterion stops
+    to increase significantly.  It is found by  computing the discrete
+    derivative of the criteria. See  %%TODO: faire une figure avec une
+                                    %%courbe réelle et montrer où l'on
+                                    %%s'arrete et où est le max.
+
+    % Then, a small processing is made to select the number of variables used for prediction.
 The evolution of the criterion function is retrieved from the model file and the discrete derivative of the criterion \emph{DJ} is computed and normalized by its maximum.
 The subset of variables used for classification $\Omega^{*}$ is the first subset for which this value is lower than a given threshold \emph{th}, equal  to 0.001\% in this experiment,
-    \begin{equation*}
-        \Omega^{*} = \Omega_{\min \{k|DJ_k<th\}}
-    \end{equation*}
-    with $DJ_k = \frac{J(\Omega_k) - J(\Omega_{k-1})}{\max_k (J(\Omega_k) - J(\Omega_{k-1}))}$.
-
-    The classification  rate is presented  using Cohen's kappa  but it
-    has to  be noticed that the  results were  similar when looking  at the
-    mean F1-score.  Processing time  has been  evaluated on  a desktop
-    computer  with 4Gb  of RAM  and  Intel(R) Core(TM)  i5-3570 CPU  @
-    3.40GHz $\times$ 4 processors.
+    % \begin{equation*}
+    %     \Omega^{*} = \Omega_{\min \{k|DJ_k<th\}}
+    % \end{equation*}
+    % with $DJ_k = \frac{J(\Omega_k) - J(\Omega_{k-1})}{\max_k (J(\Omega_k) - J(\Omega_{k-1}))}$.
+
+    The    classification   rate    is    presented   using    Cohen's
+    kappa. Processing  time has been  evaluated on a  desktop computer
+    with  8Gb of  RAM  and  Intel(R) Core(TM)  i5-3570  CPU @  3.40GHz
+    $\times$ 4 processors.
 
@@
 For the KNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file.
-Finally, for the Random Forest classifier, the trees tends to be more deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.
+Finally, for the Random Forest classifier, the trees tends to be deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.
     % The computation of the kappa seems  to suffer from  a lack of  parallelization   which could be  improved. ->  plus tard dans la conclusion.
@@
     5\_12  or  3\_10.  Table~\ref{tab:potsdam-otbsimu}  presents  the
-    results.
+    results in terms of classification accuracy and processing time.
 
@@
 
-    With this second dataset, the Random Forest classifier and the GMM classifier with kappa as selection criterion distinguished themselves by good performance.
-When using 1000 samples, no significant difference of classification rate has been observed on test set.
-But, with 50,000 samples, the Random Forest classifier becomes significantly better.
+    With this second dataset, the Random Forest classifier and the GMM classifier with kappa as selection criterion perform the best in terms of classification accuracy.
+When using 1000 samples per class, no significant difference of classification rate has been observed on test set.
+But, with 50,000 samples per class, the Random Forest classifier becomes significantly better.
 
@@
 %%% End:
+
+working...

### 2016-10-10 paper.tex
fix img + misc

@@
       \end{align}
-      where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j$.
+      where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha$ and $\mathbf{v}$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}$.
     \end{prop}
@@
 
-            GMM SFS kappa & 0.684 (0.025) & 0.698 (0.027) & 0.711 (0.028) \\
-            GMM SFS JM &    0.680 (0.028) & 0.700 (0.028) & 0.710 (0.030) \\
-            GMM SFFS JM &   0.680 (0.028) & 0.700 (0.028) & 0.710 (0.030) \\
+            GMM SFS kappa & {\bfseries 0.684 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.711 (0.028)} \\
+            GMM SFS JM &    {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
+            GMM SFFS JM &   {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
             GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
             KNN &           0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
-            Random Forest & 0.645 (0.026) & 0.673 (0.023) & 0.693 (0.023) \\
+            Random Forest & 0.645 (0.026) & 0.673 (0.023) & {\bfseries 0.693 (0.023)} \\
             \bottomrule
@@
 
-            GMM SFS kappa & 257 & 496 & 955 & 7.7 & 8.6 & 8.7 \\
-            GMM SFS JM &    8.6 & 8.9 & 9.1 & 9.7 & 9.8 & 9.6 \\
-            GMM SFFS JM &   8.8 & 9.0 & 9.3 & 9.7 & 9.8 & 9.8 \\
+            GMM SFS kappa & 257             & 496             & 955             & {\bfseries 7.7} & {\bfseries 8.6} & {\bfseries 8.7} \\
+            GMM SFS JM &    {\bfseries 8.6} & {\bfseries 8.9} & {\bfseries 9.1} & {\bfseries 9.7} & {\bfseries 9.8} & {\bfseries 9.6} \\
+            GMM SFFS JM &   {\bfseries 8.8} & {\bfseries 9.0} & {\bfseries 9.3} & {\bfseries 9.7} & {\bfseries 9.8} & {\bfseries 9.8} \\
             GMM ridge &     71.7            & 105             & 167             & 530 & 530 & 530 \\
-            KNN &           8.9 & 19.6 & 59.7 & 387 & 639 & 887 \\
+            KNN &           {\bfseries 8.9} & 19.6            & 59.7            & 387 & 639 & 887 \\
             Random Forest & 24.5            & 49.3            & 105             & 33.0 & 41.7 & 45.9 \\

### 2016-10-10 paper.tex
add fig for stopping crit + misc

@@
 
-    The computation of~\ref{eq:decision} requires the inversion of the
+    The computation of eq.~(\ref{eq:decision}) requires the inversion of the
     covariance matrix and the  computation of the logarithm of the
@@
         logarithm of the determinant of  the covariance matrix and the
-        quadratic  term in  eq.~\ref{eq:decision}. However,  all these
+        quadratic  term in  eq.(~\ref{eq:decision}). However,  all these
         terms  have already  been computed  for iteration  $(k-1)$. By
@@
 
-        Using eq.~\ref{eq:cov} and~\ref{eq:cov:inv}, it is possible to
+        Using eq.~(\ref{eq:cov}) and~(\ref{eq:cov:inv}), it is possible to
         deduce  the   following  propositions  (proof  are   given  in
@@
 
-        \begin{prop}[Decision function~\ref{eq:decision}]
+        \begin{prop}[Decision function~(\ref{eq:decision})]
             \begin{align}
@@
 
-      \begin{prop}[Kullback-Leibler divergence~\ref{eq:skl}]
+      \begin{prop}[Kullback-Leibler divergence~(\ref{eq:skl})]
         \begin{align}
@@
 
-      \begin{prop}[Jeffries-Matusita distance~\ref{eq:jm}]
+      \begin{prop}[Jeffries-Matusita distance~(\ref{eq:jm})]
       \begin{align}
@@
     \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
-    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
+    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
     \FORALL{$f_i \in F$}
@@
     \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
-    \STATE $R_i = J(\{\Omega_k + f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
+    \STATE $R_i = J(\{\Omega_k + f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
     \ENDFOR
@@
     \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
-    \STATE Compute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
+    \STATE Compute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
     \FORALL{$f_i \in \Omega_k$}
@@
     \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
-    \STATE $R_i = J(\{\Omega_k \setminus f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
+    \STATE $R_i = J(\{\Omega_k \setminus f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
     \ENDFOR
@@
     variable is selected  as follow. Rather than  selecting the number
-    of variables corresponding to the  highest value oft he criterion,
+    of variables corresponding to the  highest value of the criterion,
     the number of  retained variables is set when  the criterion stops
     to increase significantly.  It is found by  computing the discrete
-    derivative of the criteria. See  %%TODO: faire une figure avec une
-                                    %%courbe réelle et montrer où l'on
-                                    %%s'arrete et où est le max.
+    derivative of the criteria. See \ref{fig:crit-evol}.
+
+    \begin{figure}[!t]
+        \centering
+        \begin{tikzpicture}
+            \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
+                \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
+                \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
+                \draw[thick,black] (axis cs:47,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:47,\pgfkeysvalueof{/pgfplots/ymax});
+            \end{axis};
+            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
+            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_50000spl.txt};
+            %     \draw[thick,red] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
+            %     \draw[thick,black] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
+            % \end{axis};
+        \end{tikzpicture}
+        \caption{Criterion evolution in function of the number of selected variables.
+Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.\label{fig:crit-evol}}
+    \end{figure}
 
@@
 The evolution of the criterion function is retrieved from the model file and the discrete derivative of the criterion \emph{DJ} is computed and normalized by its maximum.
-The subset of variables used for classification $\Omega^{*}$ is the first subset for which this value is lower than a given threshold \emph{th}, equal  to 0.001\% in this experiment,
+The subset of variables used for classification $\Omega^{*}$ is the first subset for which this value is lower than a given threshold \emph{th}, equal  to 0.1\% in this experiment,
     % \begin{equation*}
@@
 The good performance in time can be explained by the dimension reduction.
-Actually, the decision rule corresponding to Equation~\ref{eq:decision} has a complexity in $d^3$ where $d$ is the dimension.
+Actually, the decision rule corresponding to Equation~(\ref{eq:decision}) has a complexity in $d^3$ where $d$ is the dimension.
 Thus, reducing $d$ induces a reduction of the classification time.
@@
 \label{app:proof-update}
-    % \subsubsection{Proposition~\ref{eq:update-inv-back}}
+    % \subsubsection{Proposition~(\ref{eq:update-inv-back})}
     %     \begin{proof}
@@
 
-\begin{proof}[Proof of proposition~\ref{eq:update-quad}]
+\begin{proof}[Proof of proposition~(\ref{eq:update-quad})]
     \begin{alignat*}{3}
@@
 
-\begin{proof}[Proof of proposition~\ref{eq:update-log}]
+\begin{proof}[Proof of proposition~(\ref{eq:update-log})]
     \begin{align*}

### 2016-10-11 paper.tex
fix appendix

@@
 
-The python and C++ code are available freely for download: \url{https://github.com/Laadr/FFFS}, \url{https://github.com/Laadr/otbExternalFastFeaturesSelection}.
+The python and C++ code are available freely for download: \url{https://github.com/Laadr/FFFS}, \url{https://github.com/Laadr/otbFFSforGMM}.
 
@@
 \begin{proof}[Proof of proposition~(\ref{eq:update-quad})]
+    {\footnotesize
     \begin{alignat*}{3}
@@
         \end{array}\right]
-                                                                                   \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
+        \left[\begin{array}{c} \mathbf{x}^{(k-1)} \\ x_k \end{array}\right] \\
      &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
-                                                                                       \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
+            \left[\begin{array}{c} \mathbf{x}^{(k-1)} \\ x_k \end{array}\right] \\
      &= &&(\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
@@
     \end{alignat*}
+    }
 \end{proof}
@@
 \begin{proof}[Proof of proposition~(\ref{eq:update-log})]
+    {\footnotesize
     \begin{align*}
-        \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) =
-        &\log \left(\left|
+        \log &\left(|\boldsymbol{\Sigma}^{(k)}|\right) =
+        \log \left(\left|
             \bigg[\begin{array}{cc}
@@
         \right|\right) \\
-        &\log \left(\left|
+        =&\log \left(\left|
             \bigg[\begin{array}{cc}
@@
         \right|\right) \\
-        &\log \left( |\boldsymbol{\Sigma}^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha| \right)\\
-        &\log \left( |\boldsymbol{\Sigma}^{(k-1)}| \right) + \log ( \alpha )
+        =&\log \left( |\boldsymbol{\Sigma}^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha| \right)\\
+        =&\log \left( |\boldsymbol{\Sigma}^{(k-1)}| \right) + \log ( \alpha )
     \end{align*}
+    }
+    \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.
+
 \end{proof}

### 2016-10-11 paper.tex
Intro et bio

--- 2016-10-10 paper.tex
@@
 INP-ENSAT, UMR 1201 DYNAFOR, France and with the INRA, UMR 1201
-DYNAFOR, France.}
-\thanks{M. Grizonnet is with Centre National d'\'{E}tudes Spatiales, French Space Agency, Toulouse, France.}}
+DYNAFOR, France.}%
+\thanks{M. Grizonnet is with Centre National d'\'{E}tudes Spatiales, French Space Agency, Toulouse, France.}%
+\thanks{This  work was  supported  by the  French National  Research Agency  (ANR)  under  Project Grant  ANR-13-JS02-0005-01  (Asterix project).}}
 
@@
 
-\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of data available for a given landscape becomes larger and larger.
-Several missions are about to produce huge amount of data or have already produced it.
-After 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites missioned by the German space agency will produce images with 244 spectral bands with a resolution of 30x30m and revisit every 4 days\cite{Müller09enmap}.
-The Hyperspectral Infrared Imager (HyspIRI) of NASA will also take images with 212 spectral bands every 5 days.
-Additionally to hyperspectral data, hypertemporal data are also developping.
-The European satellites Sentinel-2 were launched successfully recently and the hypertemporal data produced by this mission will be fully available at the end of 2016\cite{drusch2012sentinel}.
-
+\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available data for a given landscape becomes larger and larger.
+Satellite missions produce a huge amount of data on a regularly (daily) basis.
+From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites missioned by the German space agency will produce images with 244 spectral bands with a resolution of 30x30m with a frequency of revisit of 4 days\cite{Müller09enmap}.
+The Hyperspectral Infrared Imager (HyspIRI) of NASA will also deliver images with 212 spectral bands.
+Additionally to hyperspectral data, the amount of available hypertemporal data increases a lot.
+For instance, the European satellites Sentinel-2 were launched successfully recently and 2 Terabytes of data will be released every days~\cite{drusch2012sentinel}.
+This high volume of earth observation data provides an accurate information of land state and functioning, and help to improve the understanding of the earth~\cite{rs1010001}.
 However, processing such data is more and more challenging because of statistical and computational issues.
-This statistical issues are often referred as the \emph{curse of dimensionality}.
-The Hughes phenomenon \cite{hughes1968mean} states that with a given number of samples, prediction accuracy will decays when the complexity, \emph{i.e.} the dimension, is higher than some optimum value.
-The problem is the fast increase of the number of parameters to estimate in order to build a model when dimension expands \cite{bouveyron2014model}.
-Thus, an important number of labeled samples is needed to perform learning.
-For example, in the case of Gaussian Mixture Models, the number of parameters progresses quadratically with the dimension, e.g.
-with 200 describing variables, the model estimation requires at least around 20,000 samples by class.
-
-The computational issues are also multiple.
-The computing infrastructure needed to process data is more and more expensive because the processing might requires a GPU and a large amount of RAM to load images which can weight a dozen of gigabytes \cite{christophe2011remote}\cite{plaza2011high}.
-The processing time is also limiting and requires to use parallelize computing (GPU, multi-threading).
-
-A possible method to solve these issues is to perform a reduction of dimension.
-It is possible to extract a set of relevant features to get a parsimonious representation of the data \cite{jimenez1998supervised}.
-For instance, in land-cover classification, given a set of spatial, temporal and spectral features, it is possible to extract those which are the most discriminant for the purpose of classification \cite{fassnacht2014comparison}.
-In hyperspectral data, from the hundreds of available spectral channels, it is possible to reduce the number of channels to make the processing more efficient in terms of statistical complexity because of the reduction of the number of parameters to estimate and thus also in term of computational time.
-
-The objective of feature selection is to retrieve the most compact representation of the data with minimum loss of information \cite{Guyon:2006:FEF:1208773}.
+
+In the spectral or temporal domain, a pixel is represented by a vector
+for   wich  each   component   corresponds   to  a   spectral/temporal
+measurement.  The  size of  the vector  is the  number of  spectral or
+temporal  measurements.   For  hyperspectral images,  this  number  is
+typically   about   several   hundreds  while   for   the   Sentinel-2
+multitemporal images, the number of spectro-temporal measurement for a
+given  year  is approximately  one  thousand.   When working  in  high
+dimensional  space,  statistical  methods  made for  low  or  moderate
+dimensional  space do  not  adapt  well.  For  instance,  the rate  of
+convergence of the statistical estimation decreases when the dimension
+grows while conjointly the number of parameters to estimate increases,
+making    the    estimation    of   the    model    parameters    very
+difficult~\cite{donoho}.  Consequently,  with a limited  training set,
+beyond a certain limit, the classification accuracy actually decreases
+as the number of features increases~\cite{hughes}.  For the purpose of
+classification,  these  problems are  related  to  the \emph{curse  of
+  dimensionality}. This is a major drawback in remote sensing since it
+is  difficult to  collect a  large and  an accurate  ground-truth.  An
+intensive work has  been performed in the remote  sensing community to
+build  accurate classifiers  for  high  dimensional images.   Bayesian
+models~\cite{book:landgrebe}, feature extraction and feature reduction
+techniques~\cite{book:landgrebe,DR:guided:tour},                random
+forest~\cite{1396322},   neural  networks~\cite{5411821}   and  kernel
+methods~\cite{kernel:methods:rs}  have   been  investigated   for  the
+classification of such images.
+
+The  size of  the data  is increasing  with respect  to the  number of
+measurement per  pixel. The data  volume of an hyperspectral  image is
+typically    several    Gigabytes     per    acquisition    ($\approx$
+300km$^2$). Multitemporal data are  now available freely from internet
+stream    (see    for    instance     the    Copernicus    data    hub
+\url{https://cophub.copernicus.eu/}).  This very  large volume of data
+requires   specific   computing  infrastructure.    High   performance
+computing   is   actually   investigated   by   the   remote   sensing
+community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
+related to the use of  parallel approach (multicore, CPU, clusters) to
+improve the  processing time, and  to the use of  streaming techniques
+when data  does not fit in  memory.  One popular open  source software
+solution  is  the Orfeo  Toolbox,  developed  by the  French  National
+Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
+are  convenientely  proposed  to  the  user  through  several  modules
+``\emph{ready to use}''.
+
+A method  to reduce  both statistical and  computational issues  is to
+perform a reduction of the dimension. In fact, with the \emph{curse of
+  dimensionality}     comes     the      \emph{blessing     of     the
+  dimensionality}~\cite{bouveyron2014model}:  high   dimensional  data
+space   exhibits  interesting   properties   for  the   classification
+purpose. In  particular, it is possible  to extract a set  of relevant
+features    to   get    a   parsimonious    representation   of    the
+data~\cite{jimenez1998supervised}.    For   instance,  in   land-cover
+classification,  given  a  set   of  spatial,  temporal  and  spectral
+features,  it  is  possible  to  extract  those  which  are  the  most
+discriminant      for      the     purpose      of      classification
+\cite{fassnacht2014comparison}.   In  hyperspectral   data,  from  the
+hundreds of available spectral channels,  it is possible to reduce the
+number of channels  to make the processing more efficient  in terms of
+statistical complexity  and computational load. In  short, by reducing
+the dimension better classification results is expected with a reduced
+computational load.
+
+% Among the various dimension reduction techniques, feature selection demonstrates several appealing for practical situation.
+The objective of feature selection is to find a subset of the original features with a minimum loss according the objective.
+Thus,
+
+% retrieve the most compact representation of the data with minimum loss of information \cite{Guyon:2006:FEF:1208773}.
 By deleting the redundant information, feature selection allows to reduce dimension and thus avoid statistical issue, increase algorithm speed and limit storage requirements.
@@
 
-There are two ways to reduce dimension \cite{Guyon:2006:FEF:1208773}: features extraction and feature selection.
-Feature extraction means reformulate and summarize the information by creating new features in combining the existing ones.
-Linear combinations of the initial features can be extracted using for example in Principal Component Analysis \cite{jimenez1998supervised} or Independent Component Analysis \cite{villa2011hyperspectral}.
-Supervised extraction method has also been investigate such as Fischer discriminant analysis.
+There are two ways to reduce dimension~\cite[Chapter 1]{Guyon:2006:FEF:1208773}: features extraction and feature selection.
+Feature extraction means reformulate and summarize the information by creating new features in combining the existing ones, it is sometimes referred to \emph{feature construction}.
+Linear combination of the initial features can be extracted using Principal Component Analysis (PCA)~\cite{jimenez1998supervised} or Independent Component Analysis \cite{villa2011hyperspectral}.
+Supervised extraction method has also been investigated such as Fischer discriminant analysis and decision boundary feature extraction~\cite{book:landgrebe}.
 To the contrary, features selection selects a subset of existing features identified as the most relevant for a given criterion.
-It has the advantage to be much more interpretable for the end-user.
-The selected subset of features corresponds to the most important features for the given task.
-
-Features selection/extraction algorithms can be divided into three types.
-The first type, called \emph{filter methods}, is based uniquely on data analysis in the sense that the method is independent of any classifier.
-Features are ranked according to a statistical analysis of the data.
-For example, the Principal Component Analysis (PCA) described in \cite{jimenez1998supervised} is a typical unsupervised filter method.
+It has the additional advantage to be much more interpretable for the end-user.
+
+Features selection/extraction algorithms can be divided into three classes.
+The first class is called \emph{filter methods}.
+They select features independently of the classifier.
+Features are ranked according to some statistical measures (correlation, independence).
+For example, the PCA is a typical unsupervised filter method.
 Bruzzone \emph{et al.}\cite{bruzzone1995extension} develop a supervised filter method based on Jeffries-Matusita distance to maximize the separability of class distribution.
-A filter based on the Pearson $\chi^2$ statistical test has been used on bio-medical data\cite{biesiada2007feature}.
 The use of correlation between bands has been explored with hyperspectral data\cite{demir2008phase}.
@@
 
-The second sort are known as \emph{wrapper methods} which can be seen as search method to determine the best subset of variables for a given learning model.
-As exhaustive searches are too expensive in terms of processing time, several search strategies have been designed.
-Some are optimal under particular hypothesis as the branch and bound algorithm \cite{narendra1977branch} optimal under the condition of a monotonic criterion function.
-Others are sub-optimal but easier to set up.
-Whitney \cite{whitney1971direct} proposed a forward search adding feature one by one \cite{whitney1971direct} when Marill and Green \cite{marill1963effectiveness} use a backward search eliminating features one by one.
-Some papers propose to combine these two search method \cite{somol1999adaptive}.
-The advantage of such methods compare to filter methods is that they are dedicated to a particular model but on the other hand, as they require the training of multiple models to test various set of variables, they tend to be more time consuming.
-The method proposed in this paper is actually a wrapper method which has been optimized in order to accelerate model learning and testing which is the weak point of this family.
-
-The third type corresponds to the \emph{embedded method} which do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
+The second class are known as \emph{wrapper methods}.
+They search for the best subset of variables for a given learning model.
+Since exhaustive searches are too expensive in terms of processing time, several sub-optimal search strategies have been designed, mainly iterative forward or backward search~\cite{whitney1971direct,marill1963effectiveness} or a combination of both~\cite{somol1999adaptive}.
+The advantage of such methods compare to filter methods is that they are dedicated to a particular model and to a particular learning problem.
+On the other hand, as they require the training of multiple models to test various set of variables, they tend to be more time consuming.
+
+% The method proposed in this paper is actually a wrapper method which has been optimized in order to accelerate model learning and testing which is the weak point of this family.
+
+The third class corresponds to the \emph{embedded method}.
+They do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
 The basic example of such method is the decision tree algorithm in which a feature is selected for the creation of each node.
 Embedded methods also exist for other models, e.g.
-SVM \cite{guyon2002gene,weston2003use}.
-
-Despite a large diversity of methods, feature selection algorithms usually do not scale well with the number of pixels to be processed \cite{fauvel2015fast}.
-Nevertheless, methods based on Gaussian Mixture Models (GMM) have several interesting properties that make them suitable for feature selection in the context of large amount of data.
+SVM \cite{guyon2002gene,weston2003use,tuia2015multiclass}.
+
+Despite a large diversity of methods, feature selection algorithms usually do not scale well with the number of pixels to be processed~\cite{fauvel2015fast}.
+The training computational load is too important to compensate the reduced prediction computational load.
+Hence, feature selection is not so used in operational situation.
+However, methods based on Gaussian Mixture Models (GMM) have several interesting properties that make them suitable for feature selection in the context of large amount of data.
 By taking advantage of their intrinsic properties, it is possible to increase the computational efficiency with respect to standard implementation.
 
-% Several strategy have been explored to perform feature selection and, among them, wrapper methods receive a certain interest of the community.
-Wrapper methods can be seen as a search method to determine the best subset of variables for a given learning model.
-As exhaustive searches are out of question in a practical amount of time, numerous search strategies have been designed.
-Some optimal are under particular hypothesis \cite{narendra1977branch} and other suboptimal but easier to set up \cite{whitney1971direct}\cite{somol1999adaptive}.
-Such methods have the advantage to tune the selection to make the most of a particular classifier but require the training of multiple models to test various set of variables which can make them slow if the training is not optimized.
-
-This work proposes to develop a forward feature selection method using GMM in continuation of \cite{fauvel2015fast} and an upgraded floating forward method.
-The basic method selects iteratively the meaningful features.
-At each step, the pool of selected features is used to train a GMM which allows to compute a criterion function then used to rank features at the next iteration.
-The upgraded method introduces possible backward steps after each addition of a feature.
-An efficient implementation of the method is presented in order to handle large amount of data.
-Moreover, the developed algorithm is made available as a remote module of the C++ Orfeo Toolbox \cite{christophe2008orfeo}.
-Finally, tests are conducted to compare different variations of the method and to compare also to other classifiers (GMM, Random Forest, k-nearest-neighbor) available in the Orfeo Toolbox.
+The contribution of the paper is a forward feature selection method using GMM in continuation of~\cite{fauvel2015fast}.
+An efficient implementation of the method, using linear algebra on block matrix, is presented in order to handle large amount of data.
+Furthermore, a floating version of the algorithm is proposed.
+Several criteria are proposed to handle unbalanced training set.
+Finally, the developed algorithm is made available as a remote module of the C++ Orfeo Toolbox~\cite{christophe2008orfeo}.
+
+% Finally, tests are conducted to compare different variations of the method and to compare also to other classifiers (GMM, Random Forest, k-nearest-neighbor) available in the Orfeo Toolbox .
 
 The remaining of this article is organized as follows.
-Section~\ref{sec:gmm-hd} presents Gaussian Mixture Model classifiers and the problems link to high-dimensionality.
-The features selection methods used to make them suitable for high-dimension space is detailed in Section~\ref{sec:selection}.
-The work done to develop an efficient implementation of the proposed selection method is presented in Section~\ref{sec:implementation}.
-And finally, after introducing the two data-sets in Section~\ref{sec:datasets}, the tests conducted to explore variations of the algorithm and to compare it to other standard classifiers are shown and commented in Section~\ref{sec:test}.
+Section~\ref{sec:gmm-hd} presents GMM classifiers and the problems related to high-dimensional feature space.
+The features selection methods are detailed in Section~\ref{sec:selection}.
+The efficient implementation of the proposed selection method is presented in Section~\ref{sec:implementation}.
+Experimental results on two real high dimensional data set are given Section~\ref{sec:test}.
+Conclusion and perspectives conclude the paper in Section~\ref{sec:conclusion}.
 
@@
 
-        Table~\ref{tab:crit} summarized the presented criterion functions and their characteristics.
+        Table~\ref{tab:crit} summarizes the presented criterion functions and their characteristics.
 
@@
         \begin{tikzpicture}
-            \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
+            \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
                 \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
@@
 
-\section{Conclusion}
+\section{Conclusion and perspectives}
+\label{sec:conclusion}
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{biblio}
+\bibliography{IEEEabrv,biblio}
 
@@
 %%% End:
-
-working...

### 2016-10-12 paper.tex
Conclusion

@@
 
-An algorithm for the classification of high dimension data as hyperspectral image has been presented.
-The classifier uses GMM model to select iteratively the most relevant features.
-Several variation of the algorithms has been explored.
-Experimentations show that allowing backward step in the selection to discard already selected features do not give significant advantages.
-
-Additionally, from the comparison between all the measures used to rank the features, it has been shown that the Jeffries-Matusita distance can be a fast and accurate solution in some case but suffer from some limitations which remain to be determine more precisely.
-A possible explanation could be that this distance does not perform well when data distribution is not similar to a Gaussian distribution.
-More investigation is needed to confirm this hypothesis.
-In any case, it is possible to use directly kappa as criterion function even if it implies a loss of time.
-
-Finally, experiments show that the developed GMM classifier performs, in most cases, at least as best as standard classifiers without features selection in particular Random Forest and even outperforms all of them in term of classification time.
-
-The second achievement of this work is the development of an efficient implementation using GMM properties to derive fast update rules of the model.
-The resulting code is available as a remote module of the Orfeo toolbox on Github and makes it possible to process huge quantity of high dimension data.
-
-Finally, an improvement could be made to increase the stability of the selected features.
+An algorithm for the classification of high dimensional earth observation images has been proposed.
+The algorithm is based on Gaussian mixture model and a forward feature selection strategy to reduce the dimension of the data to be processed.
+From experimental results, this strategy has shown to be robust the \emph{curse of dimensionality}.
+As a side effect, the volume of data is reduced and the final classification processing time is reduced.
+
+To cope with the large volume of data during the learning step, updates rules from the forward search have been split into two parts in order to save computation.
+One part is only computed once per iteration, and the other part needs to be computed for each tested features.
+Several criteria have been included, three based on classification accuracy and two based on divergence measures.
+
+Experiments have been conducted on two real high dimensional data set, and the results have been compared to standards classifiers.
+Results show that the proposed approach  performs, in most cases, at least as best as classifiers (Random Forest) and even outperforms all of them in term of classification time.
+
+The resulting code is available as a remote module of the Orfeo toolbox on Github and makes it possible to process large high dimensional images efficiently.
+The \texttt{C++} code is freely available for download: \url{https://github.com/Laadr/otbFFSforGMM}.
+
+Perspectives of this work concern the selection of continuous interval of features rather than a single feature~\cite{serpico2007extraction,1468090}.
+It will be of  highest interest for continuous features, such as temporal feature or spectral feature.
+
+% Finally, an improvement could be made to increase the stability of the selected features.
 For example, with hyperspectral data, a selection of continuous intervals and not band is a possible solution which has already been explored in \cite{serpico2007extraction}.
 
-The python and C++ code are available freely for download: \url{https://github.com/Laadr/FFFS}, \url{https://github.com/Laadr/otbFFSforGMM}.
+% The python and C++ code are available freely for download: \url{https://github.com/Laadr/FFFS}, .
 
@@
 \begin{proof}[Proof of proposition~(\ref{eq:update-log})]
-    {\footnotesize
+    % {\footnotesize
+    % \begin{align*}
+    %     \log &\left(|\boldsymbol{\Sigma}^{(k)}|\right) =
+    %     \log \left(\left|
+    %         \bigg[\begin{array}{cc}
+    %             \boldsymbol{\Sigma}^{(k-1)} & \mathbf{u}      \\
+    %             \mathbf{u}^t          & \sigma_{kk} \\
+    %         \end{array}\bigg]
+    %     \right|\right) \\
+    %     =&\log \left(\left|
+    %         \bigg[\begin{array}{cc}
+    %             \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
+    %             \mathbf{u}^t          & 1 \\
+    %         \end{array}\bigg]
+    %         \bigg[\begin{array}{cc}
+    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
+    %             \mathbf{0}_{k-1,1}^t          & \sigma_{kk} - \mathbf{u}^t(\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}\\
+    %         \end{array}\bigg]
+    %     \right|\right) \\
+    %     =&\log \left(\left|
+    %         \bigg[\begin{array}{cc}
+    %             \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
+    %             \mathbf{u}^t          & 1 \\
+    %         \end{array}\bigg]
+    %     \right|\right)
+    %     +\log \left(\left|
+    %         \bigg[\begin{array}{cc}
+    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
+    %             \mathbf{0}_{k-1,1}^t          & \alpha \\
+    %         \end{array}\bigg]
+    %     \right|\right) \\
+    %     =&\log \left( |\boldsymbol{\Sigma}^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha| \right)\\
+    %     =&\log \left( |\boldsymbol{\Sigma}^{(k-1)}| \right) + \log ( \alpha )
+    % \end{align*}
+    % }
+  From eq.~(\ref{eq:cov}) and standard results for the determinant of block matrix~\cite[Chapter 9]{IMM2012-03274} we have immediatly:
   \begin{align*}
-        \log &\left(|\boldsymbol{\Sigma}^{(k)}|\right) =
-        \log \left(\left|
-            \bigg[\begin{array}{cc}
-                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{u}      \\
-                \mathbf{u}^t          & \sigma_{kk} \\
-            \end{array}\bigg]
-        \right|\right) \\
-        =&\log \left(\left|
-            \bigg[\begin{array}{cc}
-                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
-                \mathbf{u}^t          & 1 \\
-            \end{array}\bigg]
-            \bigg[\begin{array}{cc}
-                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
-                \mathbf{0}_{k-1,1}^t          & \sigma_{kk} - \mathbf{u}^t(\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}\\
-            \end{array}\bigg]
-        \right|\right) \\
-        =&\log \left(\left|
-            \bigg[\begin{array}{cc}
-                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
-                \mathbf{u}^t          & 1 \\
-            \end{array}\bigg]
-        \right|\right)
-        +\log \left(\left|
-            \bigg[\begin{array}{cc}
-                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
-                \mathbf{0}_{k-1,1}^t          & \alpha \\
-            \end{array}\bigg]
-        \right|\right) \\
-        =&\log \left( |\boldsymbol{\Sigma}^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha| \right)\\
-        =&\log \left( |\boldsymbol{\Sigma}^{(k-1)}| \right) + \log ( \alpha )
+    \log \left(|\boldsymbol{\Sigma}^{(k)}|\right)  & = \log\left(|\boldsymbol{\Sigma}^{(k-1)}|\right)\log\left(\sigma_{kk}  - \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}  \mathbf{u}\right)\\
+    &= \log\left(|\boldsymbol{\Sigma}^{(k-1)}|\right)\log\left(\alpha\right)
   \end{align*}
-    }
-    \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.
 

### 2016-10-12 paper.tex
add _c+change for crit evolution+add line in table 7

--- 2016-10-11 paper.tex
@@
         \begin{align}
-            &KL_{ij} = \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j) \nonumber \\
-            & + (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \boldsymbol{\Sigma}_i^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - d + \log \left( \frac{|\boldsymbol{\Sigma}_i|}{|\boldsymbol{\Sigma}_j|} \right) \biggr\},
+            &KL_{cc^\prime} = \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_c^{-1} \boldsymbol{\Sigma}_{c^\prime}) \nonumber \\
+            & + (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t \boldsymbol{\Sigma}_c^{-1} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) - d + \log \left( \frac{|\boldsymbol{\Sigma}_c|}{|\boldsymbol{\Sigma}_{c^\prime}|} \right) \biggr\},
         \end{align}
@@
         The    KL   divergence    is   not    symmetric,   \emph{i.e.}
-        $KL_{ij} \ne KL_{ji}$.  A symmetrized version  is used to
+        $KL_{cc^\prime} \ne KL_{c^\prime c}$.  A symmetrized version  is used to
         compute the criterion function:
         \begin{align}\label{eq:skl}
-          SKL_{ij} &=KL_{ij} + KL_{i,j} \nonumber \\
-            &= \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j + \boldsymbol{\Sigma}_j^{-1} \boldsymbol{\Sigma}_i) \nonumber \\
-            &~~+ (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t (\boldsymbol{\Sigma}_i^{-1} + \boldsymbol{\Sigma}_j^{-1}) (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - 2d \biggr\}.
+          SKL_{cc^\prime} &=KL_{cc^\prime} + KL_{c^\prime c} \nonumber \\
+            &= \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_c^{-1} \boldsymbol{\Sigma}_{c^\prime} + \boldsymbol{\Sigma}_{c^\prime}^{-1} \boldsymbol{\Sigma}_c) \nonumber \\
+            &~~+ (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t (\boldsymbol{\Sigma}_c^{-1} + \boldsymbol{\Sigma}_{c^\prime}^{-1}) (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) - 2d \biggr\}.
         \end{align}
@@
         \begin{equation}
-            C_{SKL} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_i \pi_j SKL_{ij}.
+            C_{SKL} = \sum_{c=1}^{C} \sum_{c^\prime = c + 1}^{C} \pi_c \pi_{c^\prime} SKL_{cc^\prime}.
         \end{equation}
@@
         \begin{align}
-            {B}_{ij} = &\frac{1}{8} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \left( \frac{\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j}{2} \right)^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) \nonumber \\
-            &+ \frac{1}{2} \ln \left( \frac{|\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j|}{\sqrt{|\boldsymbol{\Sigma}_i| |\boldsymbol{\Sigma}_j|}} \right).
+            {B}_{cc^\prime} = &\frac{1}{8} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t \left( \frac{\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}}{2} \right)^{-1} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) \nonumber \\
+            &+ \frac{1}{2} \ln \left( \frac{|\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}|}{\sqrt{|\boldsymbol{\Sigma}_c| |\boldsymbol{\Sigma}_{c^\prime}|}} \right).
         \end{align}
@@
         \begin{equation}\label{eq:jm}
-            {JM}_{ij} = \sqrt{ 2 \{1 - \text{exp}(-B_{ij})\} }.
+            {JM}_{cc^\prime} = \sqrt{ 2 \{1 - \text{exp}(-B_{cc^\prime})\} }.
         \end{equation}
@@
         \begin{equation}
-            {C}_{JM} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_i \pi_j {JM}_{ij}.
+            {C}_{JM} = \sum_{c=1}^{C} \sum_{c^{\prime} = c + 1}^{C} \pi_c \pi_{c^\prime} {JM}_{cc^\prime}.
         \end{equation}
@@
 
-        In the  remaining of the  paper, $\boldsymbol{\Sigma}^{(k-1)}$
+        In the  remaining of the  paper, $\boldsymbol{\Sigma}_c^{(k-1)}$
         denotes the  covariance matrix of the  $(k-1)^{th}$ iteration,
         \emph{i.e.}, the  covariance matrix  of the  selected features
-        and $\boldsymbol{\Sigma}^{(k)}$ denotes a covariance matrix at
+        and $\boldsymbol{\Sigma}_c^{(k)}$ denotes a covariance matrix at
         the $k^{th}$ iteration, \emph{i.e.},  the covariance matrix of
-        an augmented set. Then, since $\boldsymbol{\Sigma}^{(k)}$ is a
+        an augmented set. Then, since $\boldsymbol{\Sigma}_c^{(k)}$ is a
         positive definite symmetric matrix,  the covariance matrix can
@@
             \bigg[\begin{array}{cc}
-            \boldsymbol{\Sigma}^{(k-1)} & \mathbf{u}      \\
-            \mathbf{u}^t          & \sigma_{kk} \\
+            \boldsymbol{\Sigma}^{(k-1)}_c & \mathbf{u}_c      \\
+            \mathbf{u}_c^t          & \sigma^{(k)}_c \\
             \end{array}\bigg],
         \end{equation}
-        where  $\mathbf{u}$  is  the  $k^{th}$ column  of  the  matrix
+        where  $\mathbf{u}_c$  is  the  $k^{th}$ column  of  the  matrix
         without      the      diagonal      element,      \emph{i.e.},
-        $\mathbf{u}_{i}   =    \boldsymbol{\Sigma}^{(k)}_{i,k}$   with
+        $\mathbf{u}_{c}(i)   =    \boldsymbol{\Sigma}^{(k)}_{c}(i,k)$   with
         $i  \in [1,k-1]$.  Using  block matrix  inverse formulae,  the
@@
                 \bigg[\begin{array}{cc}
-                \mathbf{A} & \mathbf{v} \\
-                \mathbf{v}^t  & \frac{1}{\alpha} \\
+                \mathbf{A}_c & \mathbf{v}_c \\
+                \mathbf{v}_c^t  & \frac{1}{\alpha_c} \\
                 \end{array}\bigg]
@@
         \noindent                                                where
-        $\mathbf{A}     =     (\boldsymbol{\Sigma}^{(k-1)})^{-1}     +
-        \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}
-        \mathbf{u}^t              (\boldsymbol{\Sigma}^{(k-1)})^{-1}$,
-        $\mathbf{v}           =           -           \frac{1}{\alpha}
-        (\boldsymbol{\Sigma}^{(k-1)})^{-1}       \mathbf{u}$       and
-        $      \alpha      =      \sigma_{kk}      -      \mathbf{u}^t
-        (\boldsymbol{\Sigma}^{(k-1)})^{-1}  \mathbf{u}$.  This  update
+        $\mathbf{A}_c     =     (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}     +
+        \frac{1}{\alpha_c} (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} \mathbf{u}_c
+        \mathbf{u}_c^t              (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}$,
+        $\mathbf{v}_c           =           -           \frac{1}{\alpha_c}
+        (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}       \mathbf{u}_c$       and
+        $      \alpha_c      =      \sigma^{(k)}_c      -      \mathbf{u}_c^t
+        (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}  \mathbf{u}_c$.  This  update
         formulae      is     used      to      obtain     all      the
-        $(\boldsymbol{\Sigma}^{(k)})^{-1}$  corresponding  to all  the
+        $(\boldsymbol{\Sigma}^{(k)}_c)^{-1}$  corresponding  to all  the
         possible augmented set of a given selection iteration. Similar
@@
             \begin{align}
-              (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once per selection step}}} \nonumber \\
-                                                                                       &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for each augmented set}}}.
+              (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)}_c)^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once per selection step}}} \nonumber \\
+                                                                                       &+ \underbrace{\alpha_c ( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for each augmented set}}}.
             \end{align}
@@
             \begin{equation}
-                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)}_{\substack{\text{computed once}\\ \text{per selection step}}} + \underbrace{\log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
+                \log \left(|\boldsymbol{\Sigma}^{(k)}_c|\right) = \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right)}_{\substack{\text{computed once}\\ \text{per selection step}}} + \underbrace{\log \alpha_c}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
             \end{equation}
@@
             \label{eq:q-update}
-                Q(\mathbf{x}) = &- \underbrace{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
-                &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once per selection step}}} \nonumber \\
-                &- \underbrace{\alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 - \log \alpha}_{\substack{\text{computed for each augmented set}}}.
+                Q_c(\mathbf{x}) = &- \underbrace{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}_c^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)}_c)}_{\substack{\text{computed once per selection step}}} \nonumber \\
+                &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once per selection step}}} \nonumber \\
+                &- \underbrace{\alpha_c \left( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}_c^{(k)}) \right)^2 - \log \alpha_c}_{\substack{\text{computed for each augmented set}}}.
             \end{align}
@@
         \label{eq:skl-update}
-            & SKL_{ij} = \frac{1}{2} \biggl\{ \underbrace{\text{\normalfont Tr} \left((\boldsymbol{\Sigma}_i^{(k)})^{-1} \boldsymbol{\Sigma}_j^{(k)} + (\boldsymbol{\Sigma}_j^{(k)})^{-1} \boldsymbol{\Sigma}_i^{(k)}\right)}_{\substack{\text{computed for each augmented set}}} \nonumber \\
-            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2}_{\substack{\text{computed for each augmented set}}} \nonumber \\
-            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2 - 2k}_{\substack{\text{computed for each augmented set}}} \nonumber \\
-            &+ \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t (\boldsymbol{\Sigma}_i^{(k-1)})^{-1}(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber\\
-          & + \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t(\boldsymbol{\Sigma}_j^{(k-1)})^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \biggr\}.
+            & SKL_{cc^\prime} = \frac{1}{2} \biggl\{ \underbrace{\text{\normalfont Tr} \left((\boldsymbol{\Sigma}_c^{(k)})^{-1} \boldsymbol{\Sigma}_{c^\prime}^{(k)} + (\boldsymbol{\Sigma}_{c^\prime}^{(k)})^{-1} \boldsymbol{\Sigma}_c^{(k)}\right)}_{\substack{\text{computed for each augmented set}}} \nonumber \\
+            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2}_{\substack{\text{computed for each augmented set}}} \nonumber \\
+            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}_{c^\prime}^t & \frac{1}{\alpha_{c^\prime}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 - 2k}_{\substack{\text{computed for each augmented set}}} \nonumber \\
+            &+ \underbrace{(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber\\
+          & + \underbrace{(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t(\boldsymbol{\Sigma}_{c^\prime}^{(k-1)})^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \biggr\}.
         \end{align}
-        with $(\boldsymbol{\Sigma}_i^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
+        with $(\boldsymbol{\Sigma}_c^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
       \end{prop}
@@
         \label{eq:jm-update}
-            \text{B}_{ij} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
-            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
-            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right)}_{\substack{\text{computed for each augmented set}}},
+            \text{B}_{cc^\prime} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
+            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_c^{(k-1)}| |\boldsymbol{\Sigma}_{c^\prime}^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
+            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_c \alpha_{c^\prime}}} \right)}_{\substack{\text{computed for each augmented set}}},
       \end{align}
-      where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha$ and $\mathbf{v}$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}$.
     \end{prop}
+    \noindent where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha_c$ and $\mathbf{v}_c$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}_c$.
 
@@
 
-    To deal with this limitation, the choice has been made to perform an eigendecomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}$:
+    To deal with this limitation, the choice has been made to perform an eigendecomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}_c$:
 
     \begin{eqnarray}
-    \boldsymbol{\Sigma}^{(k)} = \mathbf{Q}^{(k)} \boldsymbol{\Lambda}^{(k)} (\mathbf{Q}^{(k)})^t\label{eq:eigendecomp}
+    \boldsymbol{\Sigma}^{(k)}_c = \mathbf{P}^{(k)}_c \boldsymbol{\Lambda}^{(k)}_c (\mathbf{P}^{(k)}_c)^t\label{eq:eigendecomp}
     \end{eqnarray}
 
-    where  $\boldsymbol{\Lambda}^{(k)}$ and  $\mathbf{Q}^{(k)}$ are  the
+    where  $\boldsymbol{\Lambda}^{(k)}_c$ and  $\mathbf{P}^{(k)}_c$ are  the
     diagonal matrix  of eigenvalues of  the covariance matrix  and the
@@
     \begin{eqnarray}
-      \boldsymbol{\Sigma}^{(k)} = \mathbf{Q}^{(k)} (\tilde{\boldsymbol{\Lambda}}^{(k)})^{-1} (\mathbf{Q}^{(k)})^t\label{eq:eigendecomp:inv}
+      (\boldsymbol{\Sigma}^{(k)}_c)^{-1} = \mathbf{P}^{(k)}_c (\tilde{\boldsymbol{\Lambda}}^{(k)}_c)^{-1} (\mathbf{P}^{(k)}_c)^t\label{eq:eigendecomp:inv}
     \end{eqnarray}
@@
       \label{eq:log:det}
-      \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_i^{(k)}),
+      \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_{c}^{(k)}(i)),
     \end{eqnarray}
-    where the \textasciitilde{} indicated thresholded values.
+    where the \textasciitilde{} indicated thresholded values and $\lambda_{c}(i)$ the ith eigenvalue.
 
@@
     \WHILE{$k \neq \text{maxVarNb}$}
-    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
-    \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
-    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
+    \FORALL{$c \in \{1,...,C\}$}
+    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)}_c = \mathbf{P}_c \boldsymbol{\Lambda}_c \mathbf{P}_c^t$
+    \STATE {\bfseries for all} $\lambda_{c}(i)$ {\bfseries do} $\lambda_{c}(i) = \max (\text{EPS\_FLT},\lambda_{c}(i))$
+    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)}_c)^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)}_c)$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
+    \ENDFOR
     \FORALL{$f_i \in F$}
-    \STATE Compute update constant $\alpha$
-    \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
+    \FORALL{$c \in \{1,...,C\}$}
+    \STATE Compute update constant $\alpha_c$
+    \STATE $\alpha_c = \max (\text{EPS\_FLT},\alpha_c)$
+    \ENDFOR
     \STATE $R_i = J(\{\Omega_k + f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
@@
     \WHILE{$k > 2 \text{ and } \text{flag}=1$}
-    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
-    \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
-    \STATE Compute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
+    \FORALL{$c \in \{1,...,C\}$}
+    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)}_c = \mathbf{P}_c \boldsymbol{\Lambda}_c \mathbf{P}_c^t$
+    \STATE {\bfseries for all} $\lambda_{c}(i)$ {\bfseries do} $\lambda_{c}(i) = \max (\text{EPS\_FLT},\lambda_{c}(i))$
+    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)}_c)^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)}_c)$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
+    \ENDFOR
     \FORALL{$f_i \in \Omega_k$}
-    \STATE Compute update constant $\alpha$
-    \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
+    \FORALL{$c \in \{1,...,C\}$}
+    \STATE Compute update constant $\alpha_c$
+    \STATE $\alpha_c = \max (\text{EPS\_FLT},\alpha_c)$
+    \ENDFOR
     \STATE $R_i = J(\{\Omega_k \setminus f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
@@
         \begin{tikzpicture}
-            \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
-                \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
-                \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
-                \draw[thick,black] (axis cs:47,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:47,\pgfkeysvalueof{/pgfplots/ymax});
+            \begin{axis}[ymin=0.2,ymax=0.9,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
+                \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl_aisa.txt};
+                % \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
+                % \draw[thick,black] (axis cs:30,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:30,\pgfkeysvalueof{/pgfplots/ymax});
             \end{axis};
+            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
+            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
+            %     \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
+            %     \draw[thick,black] (axis cs:47,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:47,\pgfkeysvalueof{/pgfplots/ymax});
+            % \end{axis};
             % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
@@
         \end{tikzpicture}
-        \caption{Criterion evolution in function of the number of selected variables.
+        \caption{Criterion evolution in function of the number of selected variables for first trial with Aisa dataset.
 Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.\label{fig:crit-evol}}
@@
         \caption{Kappa coefficient and processing time for 50,000 samples by class and averaged over 5 trials (standard deviation in parenthesis).
-Processing times are given in second.\label{tab:potsdam-otbsimu-big}}
+Processing times are given in second.
+NB: the test has not been conduct with KNN because of a too long processing time for classification.\label{tab:potsdam-otbsimu-big}}
         \begin{tabular}{lcccccc}\toprule
@@
             GMM ridge &     0.641 (0.015) & 0.611 (0.026) & 0.440 (0.015) & 460 & 2000 & all \\
+            KNN &           /             & /             & /             & / & / & / \\
             Random Forest & {\bfseries 0.851 (0.001)} & {\bfseries 0.715 (0.001)} & {\bfseries 0.573 (0.002)} & 2000 & 2000 & all \\
@@
     \begin{alignat*}{3}
-    (&\mathbf{x}^{(k)})^t && (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}
+    (&\mathbf{x}^{(k)})^t && (\boldsymbol{\Sigma}^{(k)}_c)^{-1} \mathbf{x}^{(k)}
        = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
         \left[\begin{array}{cc}
-            \mathbf{A}   & \mathbf{v} \\
-            \mathbf{v}^t & \frac{1}{\alpha}
+            \mathbf{A}_c   & \mathbf{v}_c \\
+            \mathbf{v}_c^t & \frac{1}{\alpha_c}
         \end{array}\right]
         \left[\begin{array}{c} \mathbf{x}^{(k-1)} \\ x_k \end{array}\right] \\
-     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
+     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A}_c + x_k \mathbf{v}_c^t & (\mathbf{x}^{(k-1)})^t \mathbf{v}_c + \frac{x_k}{\alpha_c} \end{array}\right]
             \left[\begin{array}{c} \mathbf{x}^{(k-1)} \\ x_k \end{array}\right] \\
-     &= &&(\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
-     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
-     & &&+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
-     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \alpha \mathbf{v} \mathbf{v}^t) \mathbf{x}^{(k-1)}\\
-     & &&+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
-     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} \mathbf{v}^t \mathbf{x}^{(k-1)} \\
-     & &&+ 2 \frac{x_k}{\alpha} \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha^2}) \\
-     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} )^2 \\
-     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
+     &= &&(\mathbf{x}^{(k-1)})^t \mathbf{A}_c \mathbf{x}^{(k-1)} + x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v}_c x_k + \frac{(x_k)^2}{\alpha_c} \\
+     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} + \frac{1}{\alpha_c} (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
+     & &&+ 2 x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c} \\
+     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} + \alpha_c \mathbf{v}_c \mathbf{v}_c^t) \mathbf{x}^{(k-1)}\\
+     & &&+ 2 x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c} \\
+     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c \mathbf{v}_c^t \mathbf{x}^{(k-1)} \\
+     & &&+ 2 \frac{x_k}{\alpha_c} \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c^2}) \\
+     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c + \frac{x_k}{\alpha_c} )^2 \\
+     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] \mathbf{x}^{(k)} )^2
     \end{alignat*}
@@
     \begin{align*}
-        \log &\left(|\boldsymbol{\Sigma}^{(k)}|\right) =
+        \log &\left(|\boldsymbol{\Sigma}_c^{(k)}|\right) =
         \log \left(\left|
             \bigg[\begin{array}{cc}
-                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{u}      \\
-                \mathbf{u}^t          & \sigma_{kk} \\
+                \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{u}_c      \\
+                \mathbf{u}_c^t          & \sigma^{(k)}_c \\
             \end{array}\bigg]
@@
             \bigg[\begin{array}{cc}
-                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
-                \mathbf{u}^t          & 1 \\
+                \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
+                \mathbf{u}_c^t          & 1 \\
             \end{array}\bigg]
             \bigg[\begin{array}{cc}
-                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
-                \mathbf{0}_{k-1,1}^t          & \sigma_{kk} - \mathbf{u}^t(\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}\\
+                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
+                \mathbf{0}_{k-1,1}^t          & \sigma^{(k)}_c - \mathbf{u}_c^t(\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c\\
             \end{array}\bigg]
@@
             \bigg[\begin{array}{cc}
-                \boldsymbol{\Sigma}^{(k-1)} & \mathbf{0}_{k-1,1}     \\
-                \mathbf{u}^t          & 1 \\
+                \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
+                \mathbf{u}_c^t          & 1 \\
             \end{array}\bigg]
@@
             \bigg[\begin{array}{cc}
-                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}      \\
-                \mathbf{0}_{k-1,1}^t          & \alpha \\
+                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
+                \mathbf{0}_{k-1,1}^t          & \alpha_c \\
             \end{array}\bigg]
         \right|\right) \\
-        =&\log \left( |\boldsymbol{\Sigma}^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha| \right)\\
-        =&\log \left( |\boldsymbol{\Sigma}^{(k-1)}| \right) + \log ( \alpha )
+        =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha_c| \right)\\
+        =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| \right) + \log ( \alpha_c )
     \end{align*}

### 2016-10-12 paper.tex

@@
 \begin{proof}[Proof of proposition~(\ref{eq:update-log})]
-    {\footnotesize
+
+    % {\footnotesize
+    % \begin{align*}
+    %     \log &\left(|\boldsymbol{\Sigma}_c^{(k)}|\right) =
+    %     \log \left(\left|
+    %         \bigg[\begin{array}{cc}
+    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{u}_c      \\
+    %             \mathbf{u}_c^t          & \sigma^{(k)}_c \\
+    %         \end{array}\bigg]
+    %     \right|\right) \\
+    %     =&\log \left(\left|
+    %         \bigg[\begin{array}{cc}
+    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
+    %             \mathbf{u}_c^t          & 1 \\
+    %         \end{array}\bigg]
+    %         \bigg[\begin{array}{cc}
+    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
+    %             \mathbf{0}_{k-1,1}^t          & \sigma^{(k)}_c - \mathbf{u}_c^t(\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c\\
+    %         \end{array}\bigg]
+    %     \right|\right) \\
+    %     =&\log \left(\left|
+    %         \bigg[\begin{array}{cc}
+    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
+    %             \mathbf{u}_c^t          & 1 \\
+    %         \end{array}\bigg]
+    %     \right|\right)
+    %     +\log \left(\left|
+    %         \bigg[\begin{array}{cc}
+    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
+    %             \mathbf{0}_{k-1,1}^t          & \alpha_c \\
+    %         \end{array}\bigg]
+    %     \right|\right) \\
+    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha_c| \right)\\
+    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| \right) + \log ( \alpha_c )
+    % \end{align*}
+    % }
+    % \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.
+  From eq.~(\ref{eq:cov}) and standard results for the determinant of block matrix~\cite[Chapter 9]{IMM2012-03274} we have immediatly:
   \begin{align*}
-        \log &\left(|\boldsymbol{\Sigma}_c^{(k)}|\right) =
-        \log \left(\left|
-            \bigg[\begin{array}{cc}
-                \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{u}_c      \\
-                \mathbf{u}_c^t          & \sigma^{(k)}_c \\
-            \end{array}\bigg]
-        \right|\right) \\
-        =&\log \left(\left|
-            \bigg[\begin{array}{cc}
-                \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
-                \mathbf{u}_c^t          & 1 \\
-            \end{array}\bigg]
-            \bigg[\begin{array}{cc}
-                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
-                \mathbf{0}_{k-1,1}^t          & \sigma^{(k)}_c - \mathbf{u}_c^t(\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c\\
-            \end{array}\bigg]
-        \right|\right) \\
-        =&\log \left(\left|
-            \bigg[\begin{array}{cc}
-                \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
-                \mathbf{u}_c^t          & 1 \\
-            \end{array}\bigg]
-        \right|\right)
-        +\log \left(\left|
-            \bigg[\begin{array}{cc}
-                \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
-                \mathbf{0}_{k-1,1}^t          & \alpha_c \\
-            \end{array}\bigg]
-        \right|\right) \\
-        =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha_c| \right)\\
-        =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| \right) + \log ( \alpha_c )
+    \log \left(|\boldsymbol{\Sigma}_c^{(k)}|\right)  & = \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\sigma^{(k)}_c  - \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}  \mathbf{u}_c\right)\\
+    &= \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\alpha_c\right)
   \end{align*}
-    }
-    \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.
 

### 2016-10-12 paper.tex
add figure of mean nb of selected var

@@
                 \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl_aisa.txt};
-                % \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
-                % \draw[thick,black] (axis cs:30,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:30,\pgfkeysvalueof{/pgfplots/ymax});
+                \draw[thick,red] (axis cs:18,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:18,\pgfkeysvalueof{/pgfplots/ymax});
+                \draw[thick,black] (axis cs:24,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:24,\pgfkeysvalueof{/pgfplots/ymax});
             \end{axis};
@@
         \end{tikzpicture}
-        \caption{Criterion evolution in function of the number of selected variables for first trial with Aisa dataset.
+        \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Aisa dataset with 500 samples by class.
 Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.\label{fig:crit-evol}}
@@
     Table~\ref{tab:aisa-otbsimu-time}  the   corresponding  processing
-    time.
+    time. Additionally, Figure~\ref{fig:meanNbVar-aisa} summarizes the mean of the number of selected variables for each variation of the GMM classifier with selection.
 
@@
     \end{table}
+
+    \begin{figure}
+      \centering
+      \begin{tikzpicture}
+        \begin{axis}[ylabel=$\bar{n}_s$,ymax=250,
+          symbolic x coords={0,250,500,1000,1050},
+          enlargelimits=0.15,legend pos=outer north east,xtick=data,legend cell align=left,
+          ybar,
+          bar width=5pt,grid,xlabel=Number of training samples per class,small,area legend]
+          \addplot coordinates {(250,15.9) (500,21.85) (1000,26.65)};
+          \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
+          \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
+          \addplot[red,very thick,sharp plot,update limits=false]coordinates {(0,252) (1050,252)};
+          \legend{{\footnotesize SFS kappa},{\footnotesize SFS JM},{\footnotesize SFFS JM}}
+        \end{axis}
+      \end{tikzpicture}
+      \caption{Mean number $\bar{n}_s$ of selected features for the different variation of selection methods for Aisa data set. The red line indicates the original number of spectral features.}
+      \label{fig:meanNbVar-aisa}
+    \end{figure}
 

### 2016-10-13 paper.tex
DOC: tipo

@@
 Satellite missions produce a huge amount of data on a regularly (daily) basis.
-From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites missioned by the German space agency will produce images with 244 spectral bands with a resolution of 30x30m with a frequency of revisit of 4 days\cite{Müller09enmap}.
+From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites managed by the German space agency will produce images with 244 spectral bands with a resolution of 30x30m with a frequency of revisit of 4 days\cite{Müller09enmap}.
 The Hyperspectral Infrared Imager (HyspIRI) of NASA will also deliver images with 212 spectral bands.
@@
 In the spectral or temporal domain, a pixel is represented by a vector
-for   wich  each   component   corresponds   to  a   spectral/temporal
+for   which  each   component   corresponds   to  a   spectral/temporal
 measurement.  The  size of  the vector  is the  number of  spectral or
@@
 community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
-related to the use of  parallel approach (multicore, CPU, clusters) to
+related to the use of  parallel approach (multi-core, CPU, clusters) to
 improve the  processing time, and  to the use of  streaming techniques
@@
 Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
-are  convenientely  proposed  to  the  user  through  several  modules
+are  conveniently  proposed  to  the  user  through  several  modules
 ``\emph{ready to use}''.
@@
 
-There are two ways to reduce dimension~\cite[Chapter 1]{Guyon:2006:FEF:1208773}: features extraction and feature selection.
-Feature extraction means reformulate and summarize the information by creating new features in combining the existing ones, it is sometimes referred to \emph{feature construction}.
-Linear combination of the initial features can be extracted using Principal Component Analysis (PCA)~\cite{jimenez1998supervised} or Independent Component Analysis \cite{villa2011hyperspectral}.
-Supervised extraction method has also been investigated such as Fischer discriminant analysis and decision boundary feature extraction~\cite{book:landgrebe}.
-To the contrary, features selection selects a subset of existing features identified as the most relevant for a given criterion.
-It has the additional advantage to be much more interpretable for the end-user.
+There are two ways to reduce dimension~\cite[Chapter 1]{Guyon:2006:FEF:1208773}:
+features extraction and feature selection. Feature extraction means reformulate
+and summarize the information by creating new features in combining the existing
+ones, it is sometimes referred to \emph{feature construction}. Linear
+combination of the initial features can be extracted using Principal Component
+Analysis (PCA)~\cite{jimenez1998supervised} or Independent Component Analysis
+\cite{villa2011hyperspectral}. Supervised extraction method has also been
+investigated such as Fischer discriminant analysis and decision boundary feature
+extraction~\cite{book:landgrebe}. To the contrary, features selection selects a
+subset of existing features identified as the most relevant for a given
+criterion. It has the additional advantage to be much more easily understandable for the end-user.
 
@@
     where $\pi_c$ is  the prior, \emph{i.e.}, the  proportion of class
-    $c$  and  $f_c$  a  parametric density  function  parametrized  by
+    $c$  and  $f_c$  a  parametric density  function  control by
     $\theta$.
@@
         The    KL   divergence    is   not    symmetric,   \emph{i.e.}
-        $KL_{cc^\prime} \ne KL_{c^\prime c}$.  A symmetrized version  is used to
+        $KL_{cc^\prime} \ne KL_{c^\prime c}$.  A symmetrical version  is used to
         compute the criterion function:
@@
     selection algorithm (SFS or SFFS) with the one given criterion (to
-    choose  between  Jeffries-Matusita  dist.,  Kullback-Lieber  div.,
+    choose  between  Jeffries-Matusita  dist.,  Kullback-Leibler  div.,
     Cohen’s  kappa, overall  accuracy,  mean  F1-score).  The  second,

### 2016-10-13 paper.tex
earth -> Earth

@@
 For instance, the European satellites Sentinel-2 were launched successfully recently and 2 Terabytes of data will be released every days~\cite{drusch2012sentinel}.
-This high volume of earth observation data provides an accurate information of land state and functioning, and help to improve the understanding of the earth~\cite{rs1010001}.
+This high volume of Earth observation data provides an accurate information of land state and functioning, and help to improve the understanding of the Earth~\cite{rs1010001}.
 However, processing such data is more and more challenging because of statistical and computational issues.
@@
 
-An algorithm for the classification of high dimensional earth observation images has been proposed.
+An algorithm for the classification of high dimensional Earth observation images has been proposed.
 The algorithm is based on Gaussian mixture model and a forward feature selection strategy to reduce the dimension of the data to be processed.

### 2016-10-13 paper.tex
repetition

@@
 For instance, the European satellites Sentinel-2 were launched successfully recently and 2 Terabytes of data will be released every days~\cite{drusch2012sentinel}.
-This high volume of Earth observation data provides an accurate information of land state and functioning, and help to improve the understanding of the Earth~\cite{rs1010001}.
+This high volume of Earth observation data provides an accurate information of land state and functioning, and help to improve the understanding of the planet~\cite{rs1010001}.
 However, processing such data is more and more challenging because of statistical and computational issues.

### 2016-10-13 paper.tex
relecture globale

--- 2016-10-13 paper.tex
@@
 % Do not put math or special symbols in the title.
-\title{Operational Feature Selection with Gaussian Mixture Models for remote sensing images}
+\title{Large-scale feature selection with Gaussian mixture models for remote sensing images}
 %
@@
 
-\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available data for a given landscape becomes larger and larger.
+\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available earth observation data for a given landscape becomes larger and larger.
 Satellite missions produce a huge amount of data on a regularly (daily) basis.
-From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites managed by the German space agency will produce images with 244 spectral bands with a resolution of 30x30m with a frequency of revisit of 4 days\cite{Müller09enmap}.
+From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites managed by the German space agency will produce images with 244 spectral bands, a spatial resolution of 30x30m per pixel and with a frequency of revisit of 4 days\cite{Müller09enmap}.
 The Hyperspectral Infrared Imager (HyspIRI) of NASA will also deliver images with 212 spectral bands.
@@
 For instance, the European satellites Sentinel-2 were launched successfully recently and 2 Terabytes of data will be released every days~\cite{drusch2012sentinel}.
-This high volume of Earth observation data provides an accurate information of land state and functioning, and help to improve the understanding of the Earth~\cite{rs1010001}.
+The Landsat open archive (\url{http://landsat.usgs.gov/products_data_at_no_charge.php}) has released thousands of images.
+This high volume of Earth observation data provides an accurate information of land state and functioning, and helps to improve the understanding of the planet~\cite{rs1010001}.
 However, processing such data is more and more challenging because of statistical and computational issues.
@@
 for   which  each   component  corresponds   to  a   spectral/temporal
-measurement.  The  size of  the vector  is the  number of  spectral or
-temporal  measurements.   For  hyperspectral images,  this  number  is
-typically   about   several   hundreds  while   for   the   Sentinel-2
+measurement.   The size  of  the  vector is  therefore  the number  of
+spectral  or temporal  measurements.  For  hyperspectral images,  this
+number is  typically about several  hundreds while for  the Sentinel-2
 multitemporal images, the number of spectro-temporal measurement for a
@@
 classification,  these  problems are  related  to  the \emph{curse  of
-  dimensionality}. This is a major drawback in remote sensing since it
-is  difficult to  collect a  large and  an accurate  ground-truth.  An
-intensive work has  been performed in the remote  sensing community to
-build  accurate classifiers  for  high  dimensional images.   Bayesian
-models~\cite{book:landgrebe}, feature extraction and feature reduction
-techniques~\cite{book:landgrebe,DR:guided:tour},                random
-forest~\cite{1396322},   neural  networks~\cite{5411821}   and  kernel
-methods~\cite{kernel:methods:rs}  have   been  investigated   for  the
+  dimensionality}. This  is a  major drawback  in many  remote sensing
+application since it  is difficult to collect a large  and an accurate
+ground-truth.   An intensive  work has  been performed  in the  remote
+sensing community  to build accurate classifiers  for high dimensional
+images.  Bayesian models~\cite{book:landgrebe}, feature extraction and
+feature   reduction   techniques~\cite{book:landgrebe,DR:guided:tour},
+random   forest~\cite{1396322},  neural   networks~\cite{5411821}  and
+kernel methods~\cite{kernel:methods:rs} have been investigated for the
 classification of such images.
 
-The  size of  the data  is increasing  with respect  to the  number of
-measurement per  pixel. The data  volume of an hyperspectral  image is
-typically    several    Gigabytes     per    acquisition    ($\approx$
-300km$^2$). Multitemporal data are  now available freely from internet
-stream    (see    for    instance     the    Copernicus    data    hub
+The volume of the data is  increasing dramatically with respect to the
+number of measurement  per pixel. The data volume  of an hyperspectral
+image  is  typically several  hundreds  of  Gigabytes per  acquisition
+($\approx$  300km$^2$). Multitemporal  data are  now available  freely
+from  internet  stream  (see  for instance  the  Copernicus  data  hub
 \url{https://cophub.copernicus.eu/}).  This very  large volume of data
@@
 community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
-related to the use of  parallel approach (multi-core, CPU, clusters) to
+related to the use of parallel approach (multi-core, GPU, clusters) to
 improve the  processing time, and  to the use of  streaming techniques
@@
 Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
-are  conveniently  proposed  to  the  user  through  several  modules
-``\emph{ready to use}''.
+are  conveniently proposed  to the  users/developpers through  several
+modules ``\emph{ready to use}''.
 
@@
 space   exhibits  interesting   properties   for  the   classification
-purpose. In  particular, it is possible  to extract a set  of relevant
-features    to   get    a   parsimonious    representation   of    the
-data~\cite{jimenez1998supervised}.    For   instance,  in   land-cover
-classification,  given  a  set   of  spatial,  temporal  and  spectral
-features,  it  is  possible  to  extract  those  which  are  the  most
+purpose.  In  particular,  it  is   possible  to  get  a  parsimonious
+representation  of  the  data  while  maintaining  or  increasing  the
+classification  accuracy~\cite{jimenez1998supervised}.  For  instance,
+in land-cover  classification, given  a set  of spatial,  temporal and
+spectral features, it is possible to  extract those which are the most
 discriminant      for      the     purpose      of      classification
@@
 
-There are two ways to reduce dimension~\cite[Chapter 1]{Guyon:2006:FEF:1208773}:
-features extraction and feature selection. Feature extraction means reformulate
-and summarize the information by creating new features in combining the existing
-ones, it is sometimes referred to \emph{feature construction}. Linear
-combination of the initial features can be extracted using Principal Component
-Analysis (PCA)~\cite{jimenez1998supervised} or Independent Component Analysis
-\cite{villa2011hyperspectral}. Supervised extraction method has also been
-investigated such as Fischer discriminant analysis and decision boundary feature
-extraction~\cite{book:landgrebe}. To the contrary, features selection selects a
-subset of existing features identified as the most relevant for a given
-criterion. It has the additional advantage to be much more easily understandable for the end-user.
+There  are  two  main  strategies  to  reduce  dimension~\cite[Chapter
+1]{Guyon:2006:FEF:1208773}: features extraction and feature selection.
+Feature extraction means reformulate  and summarize the information by
+creating new features in combining  the existing ones, it is sometimes
+referred to  \emph{feature construction}.   Linear combination  of the
+initial features  can be extracted using  Principal Component Analysis
+(PCA)~\cite{jimenez1998supervised}  or Independent  Component Analysis
+\cite{villa2011hyperspectral}.  Supervised extraction  method has also
+been investigated  such as Fischer discriminant  analysis and decision
+boundary feature  extraction~\cite{book:landgrebe}.  To  the contrary,
+features selection  extracts a subset of  existing features identified
+as the  most relevant  by a  given criterion.   It has  the additional
+advantage to be much more easily understandable for the end-user.
 
@@
 They select features independently of the classifier.
-Features are ranked according to some statistical measures (correlation, independence).
-For example, the PCA is a typical unsupervised filter method.
+Features are ranked according to some statistical measures, \emph{e.g.}, correlation or independence.
+For example,  PCA is a typical unsupervised filter method.
 Bruzzone \emph{et al.}\cite{bruzzone1995extension} develop a supervised filter method based on Jeffries-Matusita distance to maximize the separability of class distribution.
-The use of correlation between bands has been explored with hyperspectral data\cite{demir2008phase}.
-In general, these methods have the advantage to be fast and usable with any classifier but they do not take into account the strength and weakness of the chosen classifier and do not optimize directly the classification process.
+Correlation between bands has been explored for feature selection in hyperspectral data~\cite{demir2008phase}.
+In general, these methods are fast and do not depend on any classifier.
+But they do not take into account the properties of the chosen classifier and do not optimize directly the classification accuracy.
 
@@
 The advantage of such methods compare to filter methods is that they are dedicated to a particular model and to a particular learning problem.
-On the other hand, as they require the training of multiple models to test various set of variables, they tend to be more time consuming.
+On the other hand, as they require the training of multiple models to test various set of variables, they are more time consuming.
 
@@
 They do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
-The basic example of such method is the decision tree algorithm in which a feature is selected for the creation of each node.
-Embedded methods also exist for other models, e.g.
-SVM \cite{guyon2002gene,weston2003use,tuia2015multiclass}.
+A popular method embedded method is the \emph{Random Forest}.
+Embedded methods also exist for other models, \emph{e.g.} SVM \cite{guyon2002gene,weston2003use,tuia2015multiclass}.
 
@@
 The training computational load is too important to compensate the reduced prediction computational load.
-Hence, feature selection is not so used in operational situation.
+Hence, feature selection is not widely used in operational situation.
 However, methods based on Gaussian Mixture Models (GMM) have several interesting properties that make them suitable for feature selection in the context of large amount of data.
@@
 Several criteria are proposed to handle unbalanced training set.
-Finally, the developed algorithm is made available as a remote module of the C++ Orfeo Toolbox~\cite{christophe2008orfeo}.
+Finally, the developed algorithm is made available as a remote module of the Orfeo Toolbox~\cite{christophe2008orfeo}.
 
@@
 
-The remaining of this article is organized as follows.
+The remaining of the article is organized as follows.
 Section~\ref{sec:gmm-hd} presents GMM classifiers and the problems related to high-dimensional feature space.
 The features selection methods are detailed in Section~\ref{sec:selection}.
-The efficient implementation of the proposed selection method is presented in Section~\ref{sec:implementation}.
+Then an efficient implementation of the proposed selection method is presented in Section~\ref{sec:implementation}.
 Experimental results on two real high dimensional data set are given Section~\ref{sec:test}.
@@
 set where $\mathbf{x}_i \in \mathbb{R}^d$ is the vector of features of
-the $i^{th}$ sample, $y_i = 1,...,C$ the associated label, C the total
-number of classes,  $n$ the number of samples and  $n_c$ the number of
-samples of class $c$.
+the  $i^{th}$ sample,  $d$  the number  of spectral/temporal  features,
+$y_i = 1,...,C$ the associated label, $C$ the total number of classes,
+$n$ the  number of samples  and $n_c$ the  number of samples  of class
+$c$.
 
@@
     distribution~\cite{Fraley00model-basedclustering}:
-    \begin{equation}
+    \begin{align}
         p(\mathbf{x}) = \sum_{c=1}^{C} \pi_c f_c(\mathbf{x}|\theta)
-    \end{equation}
+    \end{align}
     where $\pi_c$ is  the prior, \emph{i.e.}, the  proportion of class
@@
     $\boldsymbol{\mu}_c$    and    $\boldsymbol{\Sigma}_c$:
-    \begin{equation*}
-        f_c(\mathbf{x}|\boldsymbol{\mu}_c, \boldsymbol{\Sigma}_c) = \frac{1}{(2\pi)^{\frac{d}{2}} |\boldsymbol{\Sigma}_c|^{\frac{1}{2}}} \exp \left( -\frac{1}{2} (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \right).
-    \end{equation*}
+    \begin{multline}
+        f_c(\mathbf{x}|\boldsymbol{\mu}_c, \boldsymbol{\Sigma}_c) = \\ \frac{1}{(2\pi)^{\frac{d}{2}} |\boldsymbol{\Sigma}_c|^{\frac{1}{2}}} \exp \left( -\frac{1}{2} (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \right).
+    \end{multline}
 
@@
         \hat{\boldsymbol{\mu}}_c &= \frac{1}{n_c} \sum_{\{i|y_i = c\}} \mathbf{x}_i ,\\
-        \hat{\boldsymbol{\Sigma}}_c &= \frac{1}{(n_c - 1)} \sum_{\{i|y_i = c\}} (\mathbf{x}_i - \boldsymbol{\mu}_c) (\boldsymbol{}x_i - \boldsymbol{\mu}_c)^t.
+        \hat{\boldsymbol{\Sigma}}_c &= \frac{1}{(n_c - 1)} \sum_{\{i|y_i = c\}} (\mathbf{x}_i - \boldsymbol{\mu}_c) (\mathbf{x}_i - \boldsymbol{\mu}_c)^t.
     \end{align}
@@
     Under the GMM,  and identifying $p(c)$ as $\pi_c$  and $p(\mathbf{x}|c)$ as
-    $f_c(\mathbf{x}|\theta)$ and by taking the log, a decision function is obtained
+    $f_c(\mathbf{x}|\theta)$ and by taking the log, the decision function is obtained
     \begin{eqnarray}\label{eq:decision}
@@
 
-    The computation of eq.~(\ref{eq:decision}) requires the inversion of the
-    covariance matrix and the  computation of the logarithm of the
-    determinant.  The estimation of these terms suffers from the curse
-    of  dimensionality\cite{bouveyron2014model}, because  in practice,
-    the  number of  parameters  $\eta_c$ to  estimate  for each  class
-    increases quadratically with respect to the number of features, as
+    The computation of  eq.~(\ref{eq:decision}) requires the inversion
+    of the covariance  matrix and the computation of  the logarithm of
+    the determinant.  The  estimation of these terms  suffers from the
+    curse of dimensionality\cite{bouveyron2014model}. In practice, the
+    number of parameters $\eta_c$ to estimate for each class increases
+    quadratically  with   respect  to  the  number   of  features,  as
     illustrated in Figure~\ref{fig:nb-param}.  Hence, if the number of
@@
     thus the computation  of its inverse and its  determinant would be
-    unstable.   The worst  situation  is $n_c<\eta_c$  which leads  to
-    a singular covariance matrix.  Unfortunately, this situation happens
+    unstable.  The  worst situation is  $n_c<\eta_c$ which leads  to a
+    singular covariance matrix.  Unfortunately, this situation happens
     regularly in remote sensing.   For instance in hyperspectral image
     classification,  very few  labeled samples  are usually  available
-    because of the difficulty and the high cost to collect ground-truth.
+    because of the difficulty and the cost to collect ground-truth.
 
@@
 Qu'en dis-tu ?
-    Second option is to reduce dimension.
+    Second option is to reduce the dimension.
 Features extraction/selection methods have been developed in order to reduce the dimension with various approaches described in Section~\ref{sec:intro}.
-In this study, this latter option is explored and a feature selection method named sequential forward features selection is presented after a brief overview of the selection algorithms in the next Section.
+In this study, this latter option is explored and a feature selection method named sequential forward features selection is presented after an overview of the selection algorithms.
 
@@
         A measure of correct classification is based on an error matrix $M$, or \emph{confusion matrix}~\cite[Chapter 4]{congalton2008assessing}.
-An element $M_{ij}$ of the matrix represents the number of samples of class $i$ classified as class $j$.
 The confusion matrix allows the computation of several global and per-class indices related to the classification accuracy~\cite{congalton2008assessing}.
-Three global criteria were used :
+Three global criteria were used:
 
@@
 
-        The classification rate is  estimated by a $n_{cv}$- cross-validation ($n_{cv}$-CV)~\cite{opac-b1127878}.
-To compute the $n_{cv}$-CV, a subset is removed  from $\mathcal{S}$ and  the GMM is learned with the remaining training samples.
-A test error is computed with the removed  training  samples  used  as  validation  samples.
-The process is  iterated $n_{cv}$ times and  the estimated classification rate is computed as the mean  test error over the $n_{cv}$ subsets of $\mathcal{S}$.
+        These  indices  are  estimated  from the  training  set  by  a
+        $n_{cv}$- cross-validation ($n_{cv}$-CV)~\cite{opac-b1127878}.
+        To  compute   the  $n_{cv}$-CV,  a  subset   is  removed  from
+        $\mathcal{S}$  and  the  GMM  is learned  with  the  remaining
+        training samples.  A test error  is computed with  the removed
+        training samples  used as validation samples.   The process is
+        iterated $n_{cv}$ times and  the estimated classification rate
+        is computed as  the mean test error over  the $n_{cv}$ subsets
+        of $\mathcal{S}$.
 
@@
         of correct  classification, divergences are  computed directly
-        on the trained model, with no need of validation samples.  Two
-        particular   divergences   are   used  in   this   work:   the
-        \emph{Kullback-Leibler}        divergence       and        the
+        on  the  trained  model,  with  no  need  of  cross-validation
+        estimation.  Two particular divergences are used in this work:
+        the     \emph{Kullback-Leibler}     divergence     and     the
         \emph{Jeffries-Matusita}  distance.   The advantage  of  these
@@
 
-        The extension  to the multiclass  situation is done  by taking
+        The extension  to the multiclass  problem is done  by taking
         the weighted mean  of the KL divergences computed  on all pair
@@
 
-        Table~\ref{tab:crit} summarizes the presented criterion functions and their characteristics.
+        Table~\ref{tab:crit} summarizes the presented criterion functions and their characteristics. In the following $J$ denotes one criterion from Table~\ref{tab:crit}.
 
@@
         The Sequential Forward Selection (SFS) starts with an empty set of selected features.
-At each  step, the feature that most  improves the criterion function $J$ selected in Table~\ref{tab:crit} is added  to the set.
-This feature is definitively added to the pool of selected features and the algorithm stops when a given number of variables \emph{maxVarNb} has been selected.
+At each  step, the feature that most  improves the criterion function $J$  is added  to the set.
+This feature is definitively added to the pool of selected features and the algorithm stops when a given number of variables \emph{maxVarNb} has been reached.
 The Algorithm~\ref{alg:sfs} presents the process in details.
@@
 
-        The SFFS works like the SFS but between each step of the SFS algorithm, a backward selection is operated to identify the less important feature.
+        The SFFS works as the SFS but between each step of the SFS algorithm, a backward selection is operated to identify the less important feature.
 If the criterion value is higher than the best value ever obtained with a set of same size, the identified feature is picked out.
@@
 
-        This SFFS algorithm eventually tests more solutions than the SFS algorithm.
+        This SFFS algorithm eventually evaluates more solutions than the SFS algorithm.
 The results are expected to be better but the trade-off is an increased computational time which is dependent on the complexity of the dataset.
@@
         Based on \cite{fauvel2015fast}, a method to accelerate the $n_{cv}$-fold cross-validation process in the case of criterion functions based on correct classification measures was implemented.
-The idea is to estimate the GMM with the whole training set once and then, instead of training models on $(n_{cv}-1)$ folds, parameters of the complete model are used to derive those of sub-models.
-
-        A sub-GMM model can be estimated at a reduced complexity when one fold is remove from the training set.
-From~\cite{fauvel2015fast}, the following formulae for unbiased estimators can be derived:
-        \begin{prop}[Cross-validation mean update]
+The idea is to estimate the GMM with the whole training set once and then, instead of training models on $(n_{cv}-1)$ folds, parameters of the complete model are used to derive those of sub-models, thus reducing the whole complexity.
+
+        \begin{prop}[Mean update for cross-validation]
             \label{eq:update-cv1}
@@
         \end{prop}
-        \begin{prop}[Cross-validation covariance matrix update]
+        \begin{prop}[Covariance matrix update cross-validation]
             \label{eq:update-cv2}
@@
         logarithm of the determinant of  the covariance matrix and the
-        quadratic  term in  eq.(~\ref{eq:decision}). However,  all these
+        quadratic term in  eq.~(\ref{eq:decision}). However, all these
         terms  have already  been computed  for iteration  $(k-1)$. By
-        using  the  fact that  the  covariance  matrix is  a  positive
-        definite     symmetric    matrix     and    block     matrices
-        formulae~\cite[Chapter   9.2]{webb2003statistical},    it   is
-        possible to factorize these terms at iteration $k$.
+        using the  positive definiteness of the  the covariance matrix
+        and block formulae~\cite[Chapter 9.2]{webb2003statistical}, it
+        is possible to factorize these terms at iteration $k$.
 
@@
         \emph{i.e.}, the  covariance matrix  of the  selected features
-        and $\boldsymbol{\Sigma}_c^{(k)}$ denotes a covariance matrix at
-        the $k^{th}$ iteration, \emph{i.e.},  the covariance matrix of
-        an augmented set. Then, since $\boldsymbol{\Sigma}_c^{(k)}$ is a
-        positive definite symmetric matrix,  the covariance matrix can
-        be written as
+        and $\boldsymbol{\Sigma}_c^{(k)}$ denotes  a covariance matrix
+        at the $k^{th}$ iteration,  \emph{i.e.}, the covariance matrix
+        of  an  augmented  set  by  the  feature  $x_k$.  Then,  since
+        $\boldsymbol{\Sigma}_c^{(k)}$ is a positive definite symmetric
+        matrix, the covariance matrix can be written as
 
@@
         \begin{equation}\label{eq:cov}
-            \boldsymbol{\Sigma}^{(k)} =
+            \boldsymbol{\Sigma}_c^{(k)} =
             \bigg[\begin{array}{cc}
@@
         \end{equation}
-        where  $\mathbf{u}_c$  is  the  $k^{th}$ column  of  the  matrix
-        without      the      diagonal      element,      \emph{i.e.},
+        where  $\sigma^{(k)}_c$  is  the variance  of  $\mathbf{x}_c$,
+        $\mathbf{u}_c$ is  the $k^{th}$  column of the  matrix without
+        the          diagonal           element,          \emph{i.e.},
         $\mathbf{u}_{c}(i) =  \boldsymbol{\Sigma}^{(k)}_{c}(i,k)$ with
         $i  \in [1,k-1]$.   Using block  matrix inverse  formulae, the
-        following proposition holds:
+        inverse of  the covariance  matrix is  given by  the following
+        proposition.
         \begin{prop}[Forward update rule for the inverse of the covariance matrix]
@@
         \begin{equation}\label{eq:cov:inv}
-                (\boldsymbol{\Sigma}^{(k)})^{-1} =
+                (\boldsymbol{\Sigma}_c^{(k)})^{-1} =
                 \bigg[\begin{array}{cc}
@@
 
-    where  $\boldsymbol{\Lambda}^{(k)}_c$ and  $\mathbf{P}^{(k)}_c$ are  the
-    diagonal matrix  of eigenvalues of  the covariance matrix  and the
-    orthonormal matrix of corresponding eigenvectors, respectively. To
-    prevent numerical instability, the  eigenvalues are thresholded to
-    EPS\_FLT. In our  implementation, EPS\_FLT is set  to the floating
-    machine precision.
+    where $\boldsymbol{\Lambda}^{(k)}_c$  and $\mathbf{P}^{(k)}_c$ are
+    the diagonal  matrix of eigenvalues  of the covariance  matrix and
+    the    orthonormal   matrix    of   corresponding    eigenvectors,
+    respectively.  To prevent  numerical instability,  the eigenvalues
+    are thresholded to a given  value EPS\_FLT. In our implementation,
+    EPS\_FLT is set to the floating machine precision.
 
@@
       \label{eq:log:det}
-      \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_{c}^{(k)}(i)),
+      \log \left(|\boldsymbol{\Sigma}_c^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_{c}^{(k)}(i)),
     \end{eqnarray}
@@
 
-    Following OTB framework, two application are available. The first,
-    called  \texttt{otbcli\_TrainGMMSelectionApp}, performs  a feature
-    selection algorithm (SFS or SFFS) with the one given criterion (to
-    choose  between  Jeffries-Matusita  dist.,  Kullback-Leibler  div.,
-    Cohen’s  kappa, overall  accuracy,  mean  F1-score).  The  second,
-    called \texttt{otbcli\_PredictGMMApp}, generates the thematic maps
+    Following  OTB framework,  two  applications  are available.   The
+    first,  called  \texttt{otbcli\_TrainGMMSelectionApp}, performs  a
+    feature selection algorithm  (SFS or SFFS) with  the one criterion
+    given      from~\ref{tab:crit}.       The      second,      called
+    \texttt{otbcli\_PredictGMMApp},   generates   the  thematic   maps
     according to the  learn model. The only  technical limitation that
@@
 Numerical experiments have been  conducted on two different data-sets.
-The first  dataset, called  \emph{Aisa}, is an  airborne hyperspectral
+The  first  one,  called  \emph{Aisa}, is  an  airborne  hyperspectral
 dataset  and  the  second,  called \emph{Potsdam},  is  an  very  high
@@
     \begin{itemize}
-        \item 15 Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2~\cite{otb}.
-        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)~\cite{fauvel2013advances};
-        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)~\cite{dalla2010morphological}.
-        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)~\cite{dalla2010morphological}.
+        \item Fifteen Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2~\cite{otb}.
+        \item Morphological profile build on each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)~\cite{fauvel2013advances};
+        \item Attribute profile build on each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)~\cite{dalla2010morphological}.
+        \item Attribute profile build on each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)~\cite{dalla2010morphological}.
         \item Textural features for  each channel with neighborhood of
@@
     The aim  of the experiments is  to compare the proposed  method to
-    some   standard  classifiers   used   in   operational  land   map
+    standard    classifiers    used    in   operational    land    map
     production~\cite{rs70912356}. A non-optimized  previous version of
@@
     \begin{itemize}
-    \item one with forward selection and JM distance as criterion (GMM SFS JM);
-    \item one with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
-    \item one with floating forward selection and JM distance as criterion (GMM SFFS JM).
+    \item One with forward selection and JM distance as criterion (GMM SFS JM);
+    \item One with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
+    \item One with floating forward selection and JM distance as criterion (GMM SFFS JM).
     \end{itemize}
@@
     set to  30 for the  Aisa dataset and  60 for the  Potsdam dataset.
-    After  the selection  procedure,  the optimal  number of  selected
-    variable is selected  as follow. Rather than  selecting the number
+    After  the selection  procedure, the  optimal number  of extracted
+    variables is selected as follow.  Rather than selecting the number
     of variables corresponding to the  highest value of the criterion,
@@
     to increase significantly.  It is  found by computing the discrete
-    derivative of the criteria. See \ref{fig:crit-evol}.
+    derivative of the criteria.  See Figure~\ref{fig:crit-evol} for an
+    example.
 
@@
       \begin{tikzpicture}
-        \begin{axis}[ylabel=$\bar{n}_s$,ymax=250,
+        \begin{axis}[ylabel=$\bar{n}_s$,ymax=25,
           symbolic x coords={0,250,500,1000,1050},
@@
           \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
-          \addplot[red,very thick,sharp plot,update limits=false]coordinates {(0,252) (1050,252)};
+          %%\addplot[red,very thick,sharp plot,update limits=false]coordinates {(0,252) (1050,252)};
           \legend{{\footnotesize SFS kappa},{\footnotesize SFS JM},{\footnotesize SFFS JM}}
@@
       \end{tikzpicture}
-      \caption{Mean number $\bar{n}_s$ of selected features for the different variation of selection methods for Aisa data set. The red line indicates the original number of spectral features.}
+      \caption{Mean number $\bar{n}_s$ of selected features for the different variation of selection methods for Aisa data set. The original number of features is 252.}
       \label{fig:meanNbVar-aisa}
@@
     tile (5\_11)  and validation  samples were all  the pixel  of tile
-    5\_12  or  3\_10.  Table~\ref{tab:potsdam-otbsimu}  presents  the
-    results in terms of classification accuracy and processing time.
+    5\_12     or    3\_10.      Table~\ref{tab:potsdam-otbsimu}    and
+    Table~\ref{tab:potsdam-otbsimu-big} present  the results  in terms
+    of classification accuracy and processing time.
 
@@
 When using 1000 samples per class, no significant difference of classification rate has been observed on test set.
-But, with 50,000 samples per class, the Random Forest classifier becomes significantly better.
-
-    It can be guessed that the fact that the Gaussian hypothesis is unlikely to be verified explains the difficulty to extract the finer information available in a larger dataset.
+But, with 50,000 samples per class, the Random Forest classifier becomes significantly better in terms of classification accuracy.
+
+    % It can be guessed that the fact that the Gaussian hypothesis is unlikely to be verified explains the difficulty to extract the finer information available in a larger dataset.
 The JM distance has stressed in Table~\ref{tab:crit} has a strong Gaussian hypothesis and the same hypotheses could also explain the drop of the classification rate of GMM classifiers with JM distance as criterion.
 
-    The KNN classifier and the GMM classifier with ridge regularization are again outperformed even if they get more stable results than GMM classifiers with JM as criterion.
-
-    The number of variables identified by the selection process shows that selection with JM distance as criterion identifies less relevant samples than with kappa as criterion.
-Moreover, the GMM classifier with kappa manages to get good performance with only 6.7\ of the variables with 1000 samples and 15\% with 50,000 samples.
-
-    In term of processing time, results are similar than with the Aisa dataset.
-GMM classifiers with selection are very fast for prediction.
-For example, the GMM classifier with kappa as criterion is 63\% faster than the Random Forest classifier for prediction with 1000 samples and 83\% faster with 50,000 samples.
+    The  Postdam classes  are  more difficult  to discriminate,  since
+    Gaussianity assumption  does not  hold. For instance,  a buildings
+    can  be  made of  various  materials,  resulting in  heterogeneous
+    distribution.   Hence,  GMM  with  ridge  regularization  performs
+    baldy. Random  Forest classifier is  more adapted to  this problem
+    and reached the  best classification accuracy. However,  it can be
+    note  that letting  the algorithm  be driven  by a  classification
+    quality criterion such as the Kappa coefficient helps in improving
+    the  classification   accuracy.  The   KNN  classifier   is  again
+    outperformed.
+
+    From        the        tables~\ref{tab:potsdam-otbsimu}        and
+    Table~\ref{tab:potsdam-otbsimu-big},   the  number   of  extracted
+    variables shows that JM criterion identifies less relevant samples
+    than with the kappa criterion. Moreover, the selection method with
+    criterion kappa manages to get good performance with only 6.7\% of
+    the  initial variables  with  1000 samples  and  15\% with  50,000
+    samples.
+
+    In term of processing time, results are similar than with the Aisa
+    dataset.  GMM  classifiers  with   selection  are  very  fast  for
+    prediction.  For  example,  the   GMM  classifier  with  kappa  as
+    criterion for the selection is  63\% faster than the Random Forest
+    classifier for prediction  with 1000 samples and  83\% faster with
+    50,000  samples.  However, the  training  time  is increased  with
+    respect to random forest.
 

### 2016-10-14 paper.tex
change url

--- 2016-10-13 paper.tex
@@
     the           source            code           is           freely
-    available\footnote{\url{https://github.com/Laadr/otbFFSforGMM}}.
+    available\footnote{\url{https://www.orfeo-toolbox.org/external-projects/}}.
 
@@
 The resulting code is available as a remote module of the Orfeo toolbox on Github and makes it possible to process large high dimensional images efficiently.
-The \texttt{C++} code is freely available for download: \url{https://github.com/Laadr/otbFFSforGMM}.
+The \texttt{C++} code is freely available for download: \url{https://www.orfeo-toolbox.org/external-projects/}.
 
@@
 
-% The python and C++ code are available freely for download: \url{https://github.com/Laadr/FFFS}, .
+% The python and C++ code are available freely for download: \url{https://www.orfeo-toolbox.org/external-projects/}, .
 

### 2016-10-14 paper.tex
misc

@@
             {B}_{cc^\prime} = &\frac{1}{8} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t \left( \frac{\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}}{2} \right)^{-1} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) \nonumber \\
-            &+ \frac{1}{2} \ln \left( \frac{|\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}|}{\sqrt{|\boldsymbol{\Sigma}_c| |\boldsymbol{\Sigma}_{c^\prime}|}} \right).
+            &+ \frac{1}{2} \log \left( \frac{|\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}|}{\sqrt{|\boldsymbol{\Sigma}_c| |\boldsymbol{\Sigma}_{c^\prime}|}} \right).
         \end{align}
@@
         \STATE $\Omega=\emptyset$
-        \STATE $F=\text{\{all variables\}}$
-        \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
+        \STATE $F=\text{\{all variables $f_i$\}}$
+        \WHILE{$\text{card}(\Omega) < maxVarNb$}
         \FORALL{$f_i \in F$}
@@
         \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
-        \STATE $F=\text{\{all variables\}}$
+        \STATE $F=\text{\{all variables $f_i$\}}$
         \STATE $k=0$
-        \WHILE{$k \neq \text{maxVarNb}$}
+        \WHILE{$k < \text{maxVarNb}$}
         \FORALL{$f_i \in F$}
@@
             \text{B}_{cc^\prime} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
-            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_c^{(k-1)}| |\boldsymbol{\Sigma}_{c^\prime}^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
-            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_c \alpha_{c^\prime}}} \right)}_{\substack{\text{computed for each augmented set}}},
+            &+ \underbrace{\frac{1}{2} \log \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_c^{(k-1)}| |\boldsymbol{\Sigma}_{c^\prime}^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
+            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 + \frac{1}{2} \log \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_c \alpha_{c^\prime}}} \right)}_{\substack{\text{computed for each augmented set}}},
       \end{align}
@@
     \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
-    \STATE $F=\text{\{all variables\}}$
+    \STATE $F=\text{\{all variables $f_i$\}}$
     \STATE $k=0$
-    \WHILE{$k \neq \text{maxVarNb}$}
+    \WHILE{$k < \text{maxVarNb}$}
     \FORALL{$c \in \{1,...,C\}$}
@@
     Table~\ref{tab:aisa-otbsimu-time}  the   corresponding  processing
-    time. Additionally, Figure~\ref{fig:meanNbVar-aisa} summarizes the mean of the number of selected variables for each variation of the GMM classifier with selection.
+    time.
+Bold values corresponds to best results.
+In Table~\ref{tab:aisa-otbsimu}, when several bold scores appears for the same experiment, it means that the scores has been assessed as equivalent with a Wilcoxon rank-sum test \cite{mann1947test}.
+Additionally, Figure~\ref{fig:meanNbVar-aisa} summarizes the mean of the number of selected variables for each variation of the GMM classifier with selection.
 
@@
 
-            GMM SFS kappa & {\bfseries 0.684 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.711 (0.028)} \\
-            GMM SFS JM &    {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
-            GMM SFFS JM &   {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
+            GMM SFS kappa & {\bfseries 0.678 (0.029)} & {\bfseries 0.687 (0.029)} & {\bfseries 0.699 (0.028)} \\
+            GMM SFS JM &    {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.701 (0.029)} \\
+            GMM SFFS JM &   {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.701 (0.029)} \\
             GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
@@
     \end{table}
+
+    % this one is result without changing the number of selected variables
+    % \begin{table}[!t]
+    %     \centering
+    %     \caption{Average classification accuracy 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
+    %     \begin{tabular}{lccc}\toprule
+    %          & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-4}
+    %         \# samples by class & 250 & 500 & 1000 \\ \midrule
+
+    %         GMM SFS kappa & {\bfseries 0.684 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.711 (0.028)} \\
+    %         GMM SFS JM &    {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
+    %         GMM SFFS JM &   {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
+    %         GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
+    %         KNN &           0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
+    %         Random Forest & 0.645 (0.026) & 0.673 (0.023) & 0.693 (0.023) \\
+    %         \bottomrule
+    %     \end{tabular}
+    % \end{table}
 
@@
     of classification accuracy and processing time.
+Bold values corresponds to best results.
+In Table~\ref{tab:potsdam-otbsimu}, when several bold scores appears for the same experiment, it means that the scores has been assessed as equivalent with a Wilcoxon rank-sum test \cite{mann1947test}.
 
@@
 \begin{proof}[Proof of proposition~(\ref{eq:update-quad})]
-    {\footnotesize
     \begin{alignat*}{3}
-    (&\mathbf{x}^{(k)})^t && (\boldsymbol{\Sigma}^{(k)}_c)^{-1} \mathbf{x}^{(k)}
-       = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
+    (&\mathbf{x}^{(k)})^t && (\boldsymbol{\Sigma}^{(k)}_c)^{-1} \mathbf{x}^{(k)} \\
+     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
         \left[\begin{array}{cc}
@@
         \left[\begin{array}{c} \mathbf{x}^{(k-1)} \\ x_k \end{array}\right] \\
-     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A}_c + x_k \mathbf{v}_c^t & (\mathbf{x}^{(k-1)})^t \mathbf{v}_c + \frac{x_k}{\alpha_c} \end{array}\right]
-            \left[\begin{array}{c} \mathbf{x}^{(k-1)} \\ x_k \end{array}\right] \\
-     &= &&(\mathbf{x}^{(k-1)})^t \mathbf{A}_c \mathbf{x}^{(k-1)} + x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v}_c x_k + \frac{(x_k)^2}{\alpha_c} \\
-     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} + \frac{1}{\alpha_c} (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
+     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
+            \left[\begin{array}{c} \mathbf{A}_c \mathbf{x}^{(k-1)} + x_k \mathbf{v}_c \\ \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{x_k}{\alpha_c} \end{array}\right] \\
+     &= &&(\mathbf{x}^{(k-1)})^t \mathbf{A}_c \mathbf{x}^{(k-1)} + x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} \\
+     & &&+ (\mathbf{x}^{(k-1)})^t \mathbf{v}_c x_k + \frac{(x_k)^2}{\alpha_c} \\
+     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \\
+     & &&+ \frac{1}{\alpha_c} (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
      & &&+ 2 x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c} \\
@@
     \end{alignat*}
-    }
 \end{proof}
@@
 
-    % {\footnotesize
     % \begin{align*}
@@
     % \end{align*}
-    % }
     % \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.

### 2016-10-14 paper.tex
relecture complete

@@
 Several variations of this algorithm are explored by changing the criterion function and also by testing a floating forward variation allowing backward step to discard already selected features.
-
 An important effort is made in exploiting GMM properties to implement a fast algorithm.
@@
 The result is a C++ remote module for the remote sensing processing toolbox Orfeo (OTB) developed by CNES.
-
 Finally, the method is compared to standard classifiers using two different data-sets, one of hyperspectral images with a lot of spectral variables and one with heterogeneous spatial features.
@@
 
-\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available earth observation data for a given landscape becomes larger and larger.
+\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available Earth observation data for a given landscape becomes larger and larger.
 Satellite missions produce a huge amount of data on a regularly (daily) basis.
@@
 statistical complexity  and computational load. In  short, by reducing
-the dimension better classification results is expected with a reduced
+the dimension better, classification results is expected with a reduced
 computational load.
@@
 creating new features in combining  the existing ones, it is sometimes
-referred to  \emph{feature construction}.   Linear combination  of the
+referred to as \emph{feature construction}.   Linear combination  of the
 initial features  can be extracted using  Principal Component Analysis
@@
 
-The third class corresponds to the \emph{embedded method}.
+The third class corresponds to the \emph{embedded methods}.
 They do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
@@
 The remaining of the article is organized as follows.
-Section~\ref{sec:gmm-hd} presents GMM classifiers and the problems related to high-dimensional feature space.
+Section~\ref{sec:gmm-hd} presents GMM classifiers and the problems related to high-dimensional feature spaces.
 The features selection methods are detailed in Section~\ref{sec:selection}.
 Then an efficient implementation of the proposed selection method is presented in Section~\ref{sec:implementation}.
-Experimental results on two real high dimensional data set are given Section~\ref{sec:test}.
+Experimental results on two real high dimensional data-set are given Section~\ref{sec:test}.
 Conclusion and perspectives conclude the paper in Section~\ref{sec:conclusion}.
 
-\section{Gaussian Mixture Models in high dimensional space}
+\section{Gaussian Mixture Models in high dimensional spaces}
 \label{sec:gmm-hd}
@@
 
-    %% Question: tu suggérais de reprendre ici "Ensuite, il y a la réduction de dimension données des refs (voir intro du papier JSTARS) et dans la réduction  de dimension, il  y a  la sélection de features." Le soucis que je vois, c'est que je détaille déjà dans l'intro les différentes méthodes de sélection/extraction et je trouve que ça devient pas très clair si je reprends tout en me basant sur le papier JSTARS.
-J'aurais dit avt "il y a trois types de sélection: filtre, wrapper, embedded blablabla" et ensuite ça me fait reprendre ici une autre distinction "supervisé/non-supervisé" si je me base sur jstars qui est un peu redondante il me semble.
-Qu'en dis-tu ?
     Second option is to reduce the dimension.
 Features extraction/selection methods have been developed in order to reduce the dimension with various approaches described in Section~\ref{sec:intro}.
-In this study, this latter option is explored and a feature selection method named sequential forward features selection is presented after an overview of the selection algorithms.
+In this study, this latter option is explored and a feature selection method named sequential forward features selection is presented.
 
@@
 \begin{enumerate}
-\item A function that  ranks the features according to some good classification criteria,
+\item A function that ranks the features according to some good classification or class separability criteria,
 \item A search strategy to optimize the function.
@@
 
-        The \emph{Jeffries–Matusita distance} is a measure based on the Bhattacharyya distance.
-It saturates if the separability between the two distribution increases~\cite{bruzzone2009novel}.
+        The \emph{Jeffries-Matusita distance} is a measure based on the Bhattacharyya distance.
+It saturates if the separability between the two distributions increases~\cite{bruzzone2009novel}.
 The JM distance is defined as
@@
         \label{sec:selection:method}
-        Two  sequential  search  algorithms were  implemented  in  this
+        Two  sequential  search  algorithms have been  implemented  in  this
         work~\cite{Guyon:2006:FEF:1208773}:  the   sequential  forward
@@
 An efficient implementation of the criterion optimization is proposed.
-It based on  the symmetry property of the covariance  matrix and block
+It is based on  the symmetry property of the covariance  matrix and block
 inverse  formula~\cite{IMM2012-03274}. In  the following  it is  shown
-that the criterion  can be split into  two part: one that  needs to be
+that the criterion  can be split into  two parts: one that  needs to be
 computed for each  tested variable, and one that needs  to be computed
@@
             &+ \underbrace{(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber\\
-          & + \underbrace{(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t(\boldsymbol{\Sigma}_{c^\prime}^{(k-1)})^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \biggr\}.
+          & + \underbrace{(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t(\boldsymbol{\Sigma}_{c^\prime}^{(k-1)})^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \biggr\},
         \end{align}
-        with $(\boldsymbol{\Sigma}_c^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
       \end{prop}
+      \noindent with $(\boldsymbol{\Sigma}_c^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv}.
 
@@
     Algorithm~\ref{alg:sffs-update}    details   when    computational
-    stability is enforced (lines 6, 10, 20 and 24).
+    stability is enforced (lines 7, 13, 25 and 31).
 
@@
     feature selection algorithm  (SFS or SFFS) with  the one criterion
-    given      from~\ref{tab:crit}.       The      second,      called
+    given      from Table~\ref{tab:crit}.       The      second,      called
     \texttt{otbcli\_PredictGMMApp},   generates   the  thematic   maps
@@
     The normalized DSM and the raw  RGBIR image are added to these 191
-    features and  then then  stacked to  create a  new image  with 196
+    features and  then stacked to  create a  new image  with 196
     bands. %  The  training is  made with  tile 5\_11  and test  with two

### 2016-10-14 paper.tex
correct table

@@
 
-            GMM SFS kappa & 257             & 496             & 955             & {\bfseries 7.7} & {\bfseries 8.6} & {\bfseries 8.7} \\
-            GMM SFS JM &    {\bfseries 8.6} & {\bfseries 8.9} & {\bfseries 9.1} & {\bfseries 9.7} & {\bfseries 9.8} & {\bfseries 9.6} \\
-            GMM SFFS JM &   {\bfseries 8.8} & {\bfseries 9.0} & {\bfseries 9.3} & {\bfseries 9.7} & {\bfseries 9.8} & {\bfseries 9.8} \\
+            GMM SFS kappa & 257             & 496             & 955             & {\bfseries 5.2} & {\bfseries 5.2} & {\bfseries 5.5} \\
+            GMM SFS JM &    {\bfseries 8.6} & {\bfseries 8.9} & {\bfseries 9.1} & {\bfseries 5.7} & {\bfseries 5.7} & {\bfseries 5.9} \\
+            GMM SFFS JM &   {\bfseries 8.8} & {\bfseries 9.0} & {\bfseries 9.3} & {\bfseries 5.0} & {\bfseries 5.0} & {\bfseries 5.4} \\
             GMM ridge &     71.7            & 105             & 167             & 530 & 530 & 530 \\

### 2016-10-14 paper.tex
derniere relecture
interrompu au critere OA, Kappa ...

@@
 Additionally to hyperspectral data, the amount of available hypertemporal data increases a lot.
-For instance, the European satellites Sentinel-2 were launched successfully recently and 2 Terabytes of data will be released every days~\cite{drusch2012sentinel}.
+For instance, the European satellites Sentinel-2 were launched recently and 2 Terabytes of data will be released every day~\cite{drusch2012sentinel}.
 The Landsat open archive (\url{http://landsat.usgs.gov/products_data_at_no_charge.php}) has released thousands of images.
-This high volume of Earth observation data provides an accurate information of land state and functioning, and helps to improve the understanding of the planet~\cite{rs1010001}.
+Such high volume of Earth observation data provides an accurate information of land state and functioning, and helps to improve the understanding of the planet~\cite{rs1010001}.
 However, processing such data is more and more challenging because of statistical and computational issues.
@@
 features selection  extracts a subset of  existing features identified
-as the  most relevant  by a  given criterion.   It has  the additional
-advantage to be much more easily understandable for the end-user.
+as  the most  relevant  by a  given criterion.   This  subset has  the
+additional advantage to  be much more understandable  for the end-user
+than those constructed by a (non-)linear combination.
 
@@
 They do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
-A popular method embedded method is the \emph{Random Forest}.
+A popular embedded method is the \emph{Random Forest}.
 Embedded methods also exist for other models, \emph{e.g.} SVM \cite{guyon2002gene,weston2003use,tuia2015multiclass}.
@@
 
-The remaining of the article is organized as follows.
-Section~\ref{sec:gmm-hd} presents GMM classifiers and the problems related to high-dimensional feature spaces.
-The features selection methods are detailed in Section~\ref{sec:selection}.
-Then an efficient implementation of the proposed selection method is presented in Section~\ref{sec:implementation}.
-Experimental results on two real high dimensional data-set are given Section~\ref{sec:test}.
-Conclusion and perspectives conclude the paper in Section~\ref{sec:conclusion}.
+The     remaining    of     the     article     is    organized     as
+follows.   Section~\ref{sec:gmm-hd}  presents   GMM  classifiers   and
+problems  related to  high-dimensional  feature  spaces. The  features
+selection methods are detailed in Section~\ref{sec:selection}. Then an
+efficient         implementation        is         presented        in
+Section~\ref{sec:implementation}.  Experimental  results on  two  real
+high dimensional data-set are given Section~\ref{sec:test}. Conclusion
+and perspectives conclude the paper in Section~\ref{sec:conclusion}.
 
@@
 The confusion matrix allows the computation of several global and per-class indices related to the classification accuracy~\cite{congalton2008assessing}.
-Three global criteria were used:
+Three global criteria were used in this work:
 

### 2016-10-14 paper.tex
last changes ...

@@
         The Sequential Forward Selection (SFS) starts with an empty set of selected features.
-At each  step, the feature that most  improves the criterion function $J$  is added  to the set.
+At each  step, the feature associated to the highest criterion function $J$  is added  to the set.
 This feature is definitively added to the pool of selected features and the algorithm stops when a given number of variables \emph{maxVarNb} has been reached.
@@
 The algorithm stops when a given number of features \emph{maxVarNb} has been selected.
-The Algorithm~\ref{alg:sffs} provides details of the process.
-
-        This SFFS algorithm eventually evaluates more solutions than the SFS algorithm.
+The Algorithm~\ref{alg:sffs} provides details about the process.
+
+        This SFFS algorithm evaluates more solutions than the SFS algorithm.
 The results are expected to be better but the trade-off is an increased computational time which is dependent on the complexity of the dataset.
@@
 
-An efficient implementation of the criterion optimization is proposed.
-It is based on  the symmetry property of the covariance  matrix and block
-inverse  formula~\cite{IMM2012-03274}. In  the following  it is  shown
-that the criterion  can be split into  two parts: one that  needs to be
-computed for each  tested variable, and one that needs  to be computed
-only once per  selection step. For the  cross-validation part, updates
-rules are given to derive sub-models  without the necessity to learn a
-GMM models for each fold.
+An efficient implementation of  the criterion optimization is detailed
+in  the following.   It  is  based on  the  symmetry  property of  the
+covariance matrix  and block inverse  formula~\cite{IMM2012-03274}.  It is shown  that the criterion  can be split  into two
+parts: one that needs to be computed for each tested variable, and one
+that  needs to  be  computed only  once per  selection  step. For  the
+cross-validation part,  updates rules  are given to  derive sub-models
+without the necessity to learn a GMM models for each fold.
 
@@
         \end{equation}
-        where  $\sigma^{(k)}_c$  is  the variance  of  $\mathbf{x}_c$,
+        where  $\sigma^{(k)}_c$  is  the variance  of  $x_c$,
         $\mathbf{u}_c$ is  the $k^{th}$  column of the  matrix without
@@
     \subsection{Numerical issues}
-    For each  iteration $k$, after  the selection of  optimal features
-    w.r.t  the  selected criterion,  the  inverses  of the  covariance
-    matrices and their log-determinant  needs to be computed. However,
-    the lack of training samples or the highly correlated features may
-    induce  a  badly-conditioned  matrix  with  very  small,  or  even
-    negative, eigenvalues.   Such values will degrade  drastically the
-    estimation of the  inverse and of the log-determinant,  and so the
-    numerical stability.
-
-    To deal with this limitation, the choice has been made to perform an eigendecomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}_c$:
-
-    \begin{eqnarray}
-    \boldsymbol{\Sigma}^{(k)}_c = \mathbf{P}^{(k)}_c \boldsymbol{\Lambda}^{(k)}_c (\mathbf{P}^{(k)}_c)^t\label{eq:eigendecomp}
-    \end{eqnarray}
-
-    where $\boldsymbol{\Lambda}^{(k)}_c$  and $\mathbf{P}^{(k)}_c$ are
-    the diagonal  matrix of eigenvalues  of the covariance  matrix and
-    the    orthonormal   matrix    of   corresponding    eigenvectors,
-    respectively.  To prevent  numerical instability,  the eigenvalues
-    are thresholded to a given  value EPS\_FLT. In our implementation,
-    EPS\_FLT is set to the floating machine precision.
-
-    Then, the inverse of the covariance matrix can be computed as
-
-    \begin{eqnarray}
-      (\boldsymbol{\Sigma}^{(k)}_c)^{-1} = \mathbf{P}^{(k)}_c (\tilde{\boldsymbol{\Lambda}}^{(k)}_c)^{-1} (\mathbf{P}^{(k)}_c)^t\label{eq:eigendecomp:inv}
-    \end{eqnarray}
-
-    and the log-determinant as
-
-    \begin{eqnarray}
-      \label{eq:log:det}
-      \log \left(|\boldsymbol{\Sigma}_c^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_{c}^{(k)}(i)),
-    \end{eqnarray}
-    where the \textasciitilde{} indicated thresholded values and $\lambda_{c}(i)$ the ith eigenvalue.
-
-    Same reasoning applied  to the term $\alpha$ in  the update rules:
-    it       is       also        thresholded       to       EPS\_FLT.
-    Algorithm~\ref{alg:sffs-update}    details   when    computational
-    stability is enforced (lines 7, 13, 25 and 31).
 
@@
 
+    For each  iteration $k$, after  the selection of  optimal features
+    w.r.t  the  selected criterion,  the  inverses  of the  covariance
+    matrices and their log-determinant  needs to be computed. However,
+    the lack of training samples or the highly correlated features may
+    induce  a  badly-conditioned  matrix  with  very  small,  or  even
+    negative, eigenvalues.   Such values will degrade  drastically the
+    estimation of the  inverse and of the log-determinant,  and so the
+    numerical stability.
+
+    To deal with this limitation, the choice has been made to perform an eigendecomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}_c$:
+
+    \begin{eqnarray}
+    \boldsymbol{\Sigma}^{(k)}_c = \mathbf{P}^{(k)}_c \boldsymbol{\Lambda}^{(k)}_c (\mathbf{P}^{(k)}_c)^t\label{eq:eigendecomp}
+    \end{eqnarray}
+
+    where $\boldsymbol{\Lambda}^{(k)}_c$  and $\mathbf{P}^{(k)}_c$ are
+    the diagonal  matrix of eigenvalues  of the covariance  matrix and
+    the    orthonormal   matrix    of   corresponding    eigenvectors,
+    respectively.  To prevent  numerical instability,  the eigenvalues
+    are thresholded to a given  value EPS\_FLT. In our implementation,
+    EPS\_FLT is set to the floating machine precision.
+
+    Then, the inverse of the covariance matrix can be computed as
+
+    \begin{eqnarray}
+      (\boldsymbol{\Sigma}^{(k)}_c)^{-1} = \mathbf{P}^{(k)}_c (\tilde{\boldsymbol{\Lambda}}^{(k)}_c)^{-1} (\mathbf{P}^{(k)}_c)^t\label{eq:eigendecomp:inv}
+    \end{eqnarray}
+
+    and the log-determinant as
+
+    \begin{eqnarray}
+      \label{eq:log:det}
+      \log \left(|\boldsymbol{\Sigma}_c^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_{c}^{(k)}(i)),
+    \end{eqnarray}
+    where the \textasciitilde{} indicated thresholded values and $\lambda_{c}(i)$ the ith eigenvalue.
+
+    Same reasoning applied  to the term $\alpha$ in  the update rules:
+    it       is       also        thresholded       to       EPS\_FLT.
+    Algorithm~\ref{alg:sffs-update}    details   when    computational
+    stability is enforced (lines 7, 13, 25 and 31).
+
     \subsection{Implementation}
     \label{sec:otb-module}
+    \begin{figure} 
+        \centering %Pas touche l'indentation !
+        \begin{minted}[fontsize=\tiny,bgcolor=bg]{sh} 
+otbcli_TrainGMMSelectionApp -io.il hyper.tif \
+                            -io.vd reference.shp \
+                            -gmm.varnb 20 -gmm.method forward -gmm.crit jm\
+                            -gmm.best 1 -gmm.seed 0\
+                            -io.out model.txt
+
+otbcli_PredictGMMApp -in hyper.tif \
+                     -model model.txt -modeltype selection\
+                     -out ThematicMap.tif
+        \end{minted}
+        \caption{OTB Module: The feature selection is done on the image \emph{hyper.tif} using the training set from \emph{reference.shp}.
+The feature selection algorithm is the forward search used with the Jeffries-Matusita criterion, 20 features are extracted and the corresponding GMM model is saved in \emph{model.txt}.
+Then the whole image is classified using the model and the results is saved in the geotiff \emph{ThematicMap.tif}.}
+        \label{fig:otb:ffs}
+    \end{figure}
 
@@
     to run the application.
-
-    \begin{figure} 
-        \centering %Pas touche l'indentation !
-        \begin{minted}[fontsize=\tiny,bgcolor=bg]{sh} 
-otbcli_TrainGMMSelectionApp -io.il hyper.tif \
-                            -io.vd reference.shp \
-                            -gmm.varnb 20 -gmm.method forward -gmm.crit jm\
-                            -gmm.best 1 -gmm.seed 0\
-                            -io.out model.txt
-
-otbcli_PredictGMMApp -in hyper.tif \
-                     -model model.txt -modeltype selection\
-                     -out ThematicMap.tif
-        \end{minted}
-        \caption{OTB Module: The feature selection is done on the image \emph{hyper.tif} using the training set from \emph{reference.shp}.
-The feature selection algorithm is the forward search used with the Jeffries-Matusita criterion, 20 features are extracted and the corresponding GMM model is saved in \emph{model.txt}.
-Then the whole image is classified using the model and the results is saved in the geotiff \emph{ThematicMap.tif}.}
-        \label{fig:otb:ffs}
-    \end{figure}
 
@@
      & &&+ (\mathbf{x}^{(k-1)})^t \mathbf{v}_c x_k + \frac{(x_k)^2}{\alpha_c} \\
-     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \\
-     & &&+ \frac{1}{\alpha_c} (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
+     &= &&(\mathbf{x}^{(k-1)})^t \Big((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \\
+     & &&+ \frac{1}{\alpha_c} (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}\Big) \mathbf{x}^{(k-1)}\\
      & &&+ 2 x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c} \\
-     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} + \alpha_c \mathbf{v}_c \mathbf{v}_c^t) \mathbf{x}^{(k-1)}\\
+     &= &&(\mathbf{x}^{(k-1)})^t \Big((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} + \alpha_c \mathbf{v}_c \mathbf{v}_c^t\Big) \mathbf{x}^{(k-1)}\\
      & &&+ 2 x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c} \\
-     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c \mathbf{v}_c^t \mathbf{x}^{(k-1)} \\
-     & &&+ 2 \frac{x_k}{\alpha_c} \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c^2}) \\
-     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c + \frac{x_k}{\alpha_c} )^2 \\
+     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c \Big( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c \mathbf{v}_c^t \mathbf{x}^{(k-1)} \\
+     & &&+ 2 \frac{x_k}{\alpha_c} \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c^2}\Big) \\
+     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c \Big( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c + \frac{x_k}{\alpha_c} \Big)^2 \\
      &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] \mathbf{x}^{(k)} )^2

### 2016-10-14 paper.tex
abstract

@@
 % Do not put math or special symbols in the title.
-\title{Large-scale feature selection with Gaussian mixture models for remote sensing images}
+\title{Large-scale feature selection with Gaussian mixture models for the classification of high dimensional remote sensing images}
 %
@@
 \begin{abstract}
-This article presents a forward feature selection algorithm based on Gaussian mixture model (GMM) classifiers.
-The algorithm selects iteratively features that maximize a criterion function which can be either a classification rate or a measure of divergence.
-Several variations of this algorithm are explored by changing the criterion function and also by testing a floating forward variation allowing backward step to discard already selected features.
-An important effort is made in exploiting GMM properties to implement a fast algorithm.
-In particular, update rules of the GMM model are used to compute the criterion function with various sets of features.
-The result is a C++ remote module for the remote sensing processing toolbox Orfeo (OTB) developed by CNES.
-Finally, the method is compared to standard classifiers using two different data-sets, one of hyperspectral images with a lot of spectral variables and one with heterogeneous spatial features.
-The results validate the fact that the method performs well in terms of processing time and classification accuracy in comparison to the standard classifiers available in OTB.\\
+  A  large  scale  feature  selection wrapper  is  discussed  for  the
+  classification  of high  dimensional  remote  sensing. An  efficient
+  implementation is proposed based on intrinsic properties of Gaussian
+  mixtures models and  block matrix.  The criterion  function is split
+  into two parts  : one that is  updated to test each  feature and one
+  that needs to be updated only once per feature selection. This split
+  saved  a  lot  of  computation  for each  test.   The  algorithm  is
+  implemented in  \texttt{C++} and integrated into  the Orfeo Toolbox.
+  It has been compared to  other classification algorithms on two high
+  dimension remote sensing images. The  results show that the approach
+  provide good classification accuracies with low computation time.
 \end{abstract}

### 2016-12-12 response.tex
add review response

% -*- coding: utf-8 -*-
\documentclass[a4paper,10pt,DIV=16]{scrartcl}  
\usepackage{ucs} 
\usepackage{array}
\usepackage[utf8x]{inputenc}                   
\usepackage[T1]{fontenc}
\usepackage{lmodern}                        
\usepackage[english]{babel} 
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[pdftex]{graphicx}
\usepackage{graphicx,amsfonts,amssymb,amsmath,amsthm}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\hypersetup{colorlinks=true,hypertexnames=false}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage[]{tcolorbox}
\usepackage{paralist}
\usepackage{tikz}
\usepackage{pgfplots}

\subject{Response to the reviewers}
\title{Large-scale feature selection with Gaussian mixture models for the classification of high dimensional remote sensing images}
\date{\today}
\author{Adrien Lagrange \and Mathieu Fauvel \and Manuel Grizonnet}

\newcommand{\rev}[1]{\textcolor{magenta}{#1}}
%%% Definition of the boxes
\newtcolorbox[auto counter,number within=section]{revbox}[1][]{%
colback=red!5!white,colframe=red!50!white,fonttitle=\sffamily\bfseries,
title=Comment~\thetcbcounter: #1}
\newtcolorbox[auto counter,number within=section]{resbox}[1][]{%
colback=blue!5!white,colframe=blue!5!white,title=Response~:,coltitle=black,fonttitle=\sffamily\bfseries}

\begin{document}
\maketitle

We have revised  the paper taking into account the  comments raised by
the reviewers.   We provide an  item-by-item response detailed  in the
following.   We hope  that this  improved  paper is  now suitable  for
publication  in  {\sc  Special Issue on Computational Imaging for Earth Sciences}.   We would  like  to  thank  the
anonymous reviewers for their dedication in reviewing our work and for
their constructive comments.

For clarity,  the modification appear  in \textcolor{magenta}{magenta}
in the revised paper.

\section{Reviewer 1}

\begin{revbox}
In Algorithm 3, a threshold EPS\_FLT on eigenvalues is used to prevent numerical instability. It’s not clear how the authors select the proper value for this threshold. And how does this threshold influence the classification accuracy?

  \begin{resbox}
    In order to influence as less as possible the classification accuracy, the threshold has been set as small as possible. The smallest possible float number is define by the computer architecture (in our case 64bits) and C++ specifications. The C++ documentation states that epsilon is the \emph{difference between 1 and the least value greater than 1 that is representable}(\url{http://www.cplusplus.com/reference/cfloat/}).
    This thresholding can be seen as a ridge regularization when applied to eigenvalues. 
  \end{resbox}

\end{revbox}

\begin{revbox}
  The authors show the experimental results on two different types of hyper-dimensional images, which is very good. But I have some doubts on the construction of the second high dimensional cube where 191 features are constructed based on the low dimensional RGBIR images. Are these features simply used for the purpose of increasing the data dimensionality? Or are there benefits in terms of classification accuracy?
  \begin{resbox}
    The interest of this second data cube is first to test the algorithm in a different context. In particular, this second dataset should be seen as a method to learn the most relevant kind of features and the best parameters to extract those features. In the first dataset, only spectral features has been considered and in the particular context of hyperspectral imaging which imply contiguous features highly correlated.
    (reponse bof à revoir et developper mais c'est un debut)
  \end{resbox}
\end{revbox}

\begin{revbox}
  In Fig. 5, the optimal number of variables are selected if the criterion stops to increase significantly. It is unclear to me how to make this selection. For example in Fig. 5, I would pick the optimal number of variables to be 7 instead of 18 which the authors suggest, since after 7 variables I do not observe a significant increase. Maybe plotting the discrete derivative of criteria together with Fig. 5 can help in the explanation.
  \begin{resbox}
  With the same example, 7 features gives a kappa of $0.620$ and 18 features a kappa of $0.657$.

  \end{resbox}
\end{revbox}

  \begin{figure}[!ht]
      \centering
      \begin{tikzpicture}
          \begin{axis}[ymin=0,ymax=0.3,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
              \addplot[thick,black] table[x expr=\coordindex,y=deriv] {derivate_criterion_1000spl_aisa.txt};
              % \draw[thick,red] (axis cs:18,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:18,\pgfkeysvalueof{/pgfplots/ymax});
              % \draw[thick,black] (axis cs:24,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:24,\pgfkeysvalueof{/pgfplots/ymax});
          \end{axis};
      \end{tikzpicture}
      \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Aisa dataset with 500 samples by class. Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.}
  \end{figure}

\begin{revbox}
  Following the question 3, why choosing the optimal number of variables based on the increase of criterion instead of the highest value?
  \begin{resbox}

  \end{resbox}
\end{revbox}

\begin{revbox}
  In the experiment with Aisa dataset, although several numbers of training samples are tested, the number of training samples seems carefully selected by the authors to show the cases where the GMM classifiers perform better. I guess when a larger number of samples are trained, the random forest will have a higher classification rate. And when a smaller number of samples are used, the KNN will have a shorter training time. The GMM only wins in the classification time regardless of the sample number. It’ll be good to show these cases.
  \begin{resbox}
  \end{resbox}
\end{revbox}

\begin{revbox}
  Correct some typos, for example:
  \begin{enumerate}
    \item In Proposition 2, in the sentence after the covariance matrix equation, I guess it should be ``u\_c'' instead of ``v\_c''.
    \item Grammar mistakes such as ``a buildings''.
  \end{enumerate}
  \begin{resbox}
    % Done. Thank you for the careful reading.
    NB:
    no $v_c$ after prop 2...
    building corrected. Need to correct english at the end
  \end{resbox}
\end{revbox}

\section{Reviewer 2}

\begin{revbox}
  page 4 :  C. Model inference equations seem correct  but what is the
  typical range of p-hat (fixed parameter)?
  \begin{resbox}

  \end{resbox}
\end{revbox}

\begin{revbox}
  Experimental Results : Absence of such visual graphics such as final
  classification map,  original imagery and accuracy  studies based on
  classification  performance   (not  just  compute  time   and  kappa
  coefficients, but  in terms  of miss classification  rate, confusion
  matrix from classifications etc.)

  \begin{resbox}

  \end{resbox}
\end{revbox}

\clearpage

\section{Reviewer 2}
\begin{revbox}
  In section III-A, 5 criterion functions are proposed. With 2 feature selection methods (SFS and SFFS), there should be 10 possible configurations. However, in the experimental parts (page 14, line 21-24), results of only three configurations are shown. The authors said “other configurations have been investigated and performs either equally or lower in terms of classification accuracy.” I would suggest that the authors either show the results of more configurations or not introducing those unrelated criterion functions.
  \begin{resbox}

  \end{resbox}
\end{revbox}

\begin{revbox}
  Only Cohen’s Kappa is utilized to analyze the classification accuracy. The experimental results would be more persuasive if other criterion functions are utilized as well. I am interested to see how different methods’ performance varies under different criterion functions.
  \begin{resbox}

  \end{resbox}
\end{revbox}

\begin{revbox}
  According to Table IV in page 16. GMM SFS JM outperforms GMM SFS kappa a little bit under the kappa criterion. Since GMM SFS kappa updates the features according to the kappa criterion while GMM SFS JM updates the features according to the JM criterion, shouldn’t GMM SFS kappa outperforms GMM SFS JM under the kappa criterion? Or is there some explanation to it?
  \begin{resbox}

  \end{resbox}
\end{revbox}

\begin{revbox}
  In Table VI and Table VII, the \# of selected features are different for different proposed methods. And the classification performance drops when the \# of selected features decreases. According to Algorithm 2 and Algorithm 3, the \# of selected features should be equal to maxVarNb. So why the \# of selected features are different for different proposed methods? The comparison is not fair to me if the \# of selected features are different. The authors should also show the \# of selected features in Table IV or Table V as well.
  \begin{resbox}

  \end{resbox}
\end{revbox}

\begin{revbox}
  A reference or an explanation should be made in Page 1, line 36 regarding the curse of dimensionality
  \begin{resbox}

  \end{resbox}
\end{revbox}

\section{Reviewer 3}

\begin{revbox}
  Although the method seems promising, there are some significant issues the authors should address to. The key novelty and contribution of this technique is unfortunately unclear. The authors dedicate a lot of space in explaining already known mathematical formulations of the Gaussian Mixture Models and the feature selection techniques, in comparison with their proposed scheme.
  \begin{resbox}

  \end{resbox}
\end{revbox}

\begin{revbox}
  Additionally, critical references are missing, such as:
  \begin{enumerate}
  \item K. Karalas, G. Tsagkatakis, M. Zervakis, and P. Tsakalides, "Land Classification Using Remotely Sensed Data: Going Multi-Label," IEEE Transactions on Geoscience and Remote Sensing, 2016, doi: 10.1109/TGRS.2016.2520203
  \item K. Karalas, G. Tsagkatakis, M. Zervakis, and P. Tsakalides. "Deep learning for multi-label land cover classification", in SPIE Remote Sensing, Toulouse
  \end{enumerate}

  \begin{resbox}

  \end{resbox}
\end{revbox}

\begin{revbox}
  Another drawback is that the English language is poorly written, since it contains several critical syntax and typographical errors. As a result, some parts of the text are unclarified and misunderstood.
  \begin{resbox}

  \end{resbox}
\end{revbox}

\section{Auxillary data}
\label{sec:auxillary}

\begin{table}[h]
  \footnotesize
  \centering
  \caption{University}
  \begin{tabular}{lrrrrrrrrrrrrrrr}
    Method & M0 & M1 & M2 & M3 & M4 & M5 & M6 & NM0 & NM1 & NM2 & NM3 & NM3 & KDA & RF & SVM \\
    OA & 82.0 & 83.9 & 69.6 & 68.0 & 73.3 & 65.2 & 68.7 & 78.7 & 83.8 & 67.7 & 65.8 & 65.8 & 83.5 & 71.9 & 84.5 \\
    Kappa & 0.768 & 0.793 & 0.617 & 0.603 & 0.661 & 0.567 & 0.61 & 0.73 & 0.792 & 0.599 & 0.578 & 0.578 & 0.786 & 0.646 & 0.799 \\
    CPT & 18 & 18 & 18 & 19 & 19 & 18 & 19 & 17 & 18 & 18 & 19 & 19 & 98 & 3 & 10 \\
    PA 0 & 82.0 & 81.6 & 74.0 & 67.5 & 73.5 & 64.0 & 68.7 & 83.4 & 87.5 & 71.8 & 66.4 & 66.4 & 73.0 & 73.1 & 80.0 \\
    PA 1 & 79.5 & 82.5 & 65.1 & 59.3 & 69.4 & 64.3 & 61.5 & 72.9 & 80.1 & 60.7 & 56.9 & 56.9 & 84.1 & 65.1 & 84.6 \\
    PA 2 & 72.1 & 75.9 & 58.4 & 57.4 & 75.8 & 63.3 & 64.9 & 72.7 & 75.6 & 54.5 & 53.5 & 53.5 & 82.4 & 68.2 & 79.6 \\
    PA 3 & 96.3 & 96.7 & 81.7 & 93.6 & 91.9 & 68.2 & 91.3 & 96.4 & 97.7 & 94.0 & 93.3 & 93.3 & 92.1 & 93.6 & 92.5 \\
    PA 4 & 99.4 & 99.5 & 88.1 & 95.6 & 98.2 & 80.4 & 95.7 & 99.6 & 99.7 & 91.8 & 96.5 & 96.5 & 98.6 & 99.0 & 99.3 \\
    PA 5 & 79.6 & 82.9 & 64.8 & 68.0 & 65.7 & 56.9 & 66.4 & 75.3 & 85.3 & 67.9 & 64.8 & 64.8 & 83.3 & 68.6 & 83.6 \\
    PA 6 & 86.9 & 91.3 & 71.8 & 89.5 & 83.2 & 71.1 & 88.0 & 86.4 & 84.0 & 62.9 & 86.9 & 86.9 & 93.4 & 87.2 & 90.1 \\
    PA 7 & 79.1 & 78.6 & 72.9 & 72.8 & 67.4 & 66.2 & 67.8 & 78.5 & 77.3 & 68.0 & 70.0 & 70.0 & 79.8 & 71.1 & 78.5 \\
    PA 8 & 99.8 & 99.7 & 99.2 & 98.9 & 99.0 & 99.5 & 98.8 & 99.5 & 99.6 & 97.3 & 96.6 & 96.6 & 99.7 & 99.9 & 99.9 \\
    UA 0 & 94.7 & 96.0 & 81.5 & 95.1 & 94.4 & 86.8 & 93.1 & 94.8 & 94.5 & 89.6 & 95.0 & 95.0 & 96.0 & 93.7 & 94.5 \\
    UA 1 & 93.4 & 94.5 & 88.9 & 88.2 & 88.3 & 88.7 & 88.2 & 91.4 & 95.1 & 88.8 & 86.8 & 86.8 & 94.8 & 88.3 & 94.2 \\
    UA 2 & 60.8 & 62.4 & 41.2 & 45.5 & 46.8 & 35.3 & 43.6 & 60.0 & 62.4 & 41.9 & 43.7 & 43.7 & 60.3 & 49.2 & 65.6 \\
    UA 3 & 60.2 & 65.7 & 61.0 & 51.5 & 59.7 & 65.5 & 53.1 & 60.3 & 63.9 & 49.5 & 48.7 & 48.7 & 77.1 & 57.6 & 77.6 \\
    UA 4 & 97.9 & 97.9 & 99.9 & 75.8 & 81.8 & 98.1 & 74.1 & 92.8 & 96.7 & 68.9 & 56.3 & 56.3 & 98.9 & 92.6 & 96.9 \\
    UA 5 & 67.1 & 69.4 & 41.4 & 39.4 & 45.1 & 41.7 & 40.3 & 53.3 & 66.0 & 39.9 & 37.4 & 37.4 & 66.4 & 43.1 & 66.2 \\
    UA 6 & 65.9 & 64.7 & 44.3 & 48.7 & 58.3 & 24.9 & 50.2 & 71.9 & 81.2 & 55.6 & 53.5 & 53.5 & 55.4 & 52.8 & 59.9 \\
    UA 7 & 76.8 & 79.3 & 68.9 & 73.1 & 78.8 & 65.8 & 75.5 & 80.2 & 80.2 & 71.8 & 73.0 & 73.0 & 75.9 & 75.2 & 80.4 \\
    UA 8 & 99.9 & 99.9 & 90.6 & 100.0 & 100.0 & 44.6 & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 & 99.8 & 99.8 & 99.9 \\
  \end{tabular}
\end{table}

\begin{table}[h]
  \footnotesize
  \centering
  \caption{KSC}
  \begin{tabular}{lrrrrrrrrrrrrrrr}
    Method & M0 & M1 & M2 & M3 & M4 & M5 & M6 & NM0 & NM1 & NM2 & NM3 & NM3 & KDA & RF & SVM \\
    OA & 92.9 & 93.1 & 86.0 & 85.9 & 88.4 & 83.9 & 86.1 & 92.1 & 92.9 & 85.5 & 83.6 & 83.6 & 93.2 & 86.9 & 93.5 \\
    Kappa & 0.92 & 0.922 & 0.844 & 0.842 & 0.87 & 0.82 & 0.845 & 0.911 & 0.921 & 0.838 & 0.817 & 0.817 & 0.924 & 0.853 & 0.928 \\
    CPT & 31 & 33 & 31 & 33 & 34 & 32 & 34 & 31 & 33 & 31 & 33 & 33 & 253 & 3 & 28 \\
    PA 0 & 90.7 & 90.8 & 87.0 & 85.9 & 88.2 & 86.8 & 85.6 & 89.9 & 90.3 & 85.3 & 84.2 & 84.2 & 87.0 & 88.1 & 90.0 \\
    PA 1 & 88.9 & 89.6 & 83.7 & 80.7 & 85.5 & 83.3 & 80.8 & 90.6 & 90.3 & 82.6 & 80.2 & 80.2 & 87.7 & 83.5 & 91.0 \\
    PA 2 & 90.1 & 91.7 & 86.1 & 89.9 & 86.6 & 86.6 & 89.9 & 85.0 & 89.2 & 86.3 & 87.0 & 87.0 & 88.2 & 89.6 & 93.0 \\
    PA 3 & 77.2 & 78.3 & 61.3 & 67.7 & 70.3 & 64.1 & 70.9 & 74.3 & 77.9 & 62.6 & 61.9 & 61.9 & 79.8 & 64.1 & 78.5 \\
    PA 4 & 81.5 & 83.4 & 63.1 & 61.8 & 71.7 & 66.7 & 64.1 & 79.1 & 82.2 & 62.2 & 59.2 & 59.2 & 82.8 & 65.0 & 80.4 \\
    PA 5 & 80.7 & 80.4 & 55.3 & 51.1 & 58.0 & 36.2 & 48.0 & 84.2 & 83.4 & 53.0 & 45.5 & 45.5 & 82.5 & 56.3 & 79.6 \\
    PA 6 & 95.5 & 95.6 & 89.5 & 93.9 & 92.0 & 91.4 & 94.6 & 92.4 & 94.0 & 88.1 & 90.4 & 90.4 & 96.8 & 93.1 & 94.8 \\
    PA 7 & 91.9 & 92.3 & 82.0 & 81.7 & 88.0 & 84.8 & 82.8 & 93.2 & 93.2 & 82.0 & 74.1 & 74.1 & 93.0 & 81.0 & 93.8 \\
    PA 8 & 96.4 & 96.6 & 92.6 & 91.8 & 94.3 & 93.9 & 92.1 & 94.9 & 96.1 & 90.1 & 86.7 & 86.7 & 97.5 & 89.4 & 95.3 \\
    PA 9 & 95.1 & 94.9 & 81.2 & 88.5 & 90.5 & 73.5 & 88.8 & 95.4 & 95.5 & 94.5 & 93.0 & 93.0 & 96.8 & 86.4 & 96.9 \\
    PA 10 & 97.7 & 98.1 & 90.2 & 90.4 & 94.7 & 85.0 & 90.9 & 97.7 & 97.9 & 89.6 & 90.2 & 90.2 & 97.4 & 95.8 & 97.9 \\
    PA 11 & 93.0 & 92.5 & 88.0 & 81.0 & 82.1 & 78.5 & 80.8 & 90.0 & 92.4 & 80.5 & 77.7 & 77.7 & 96.3 & 83.2 & 95.1 \\
    PA 12 & 98.9 & 98.9 & 97.8 & 98.2 & 98.4 & 97.9 & 98.3 & 98.6 & 98.6 & 97.1 & 97.6 & 97.6 & 99.9 & 99.4 & 100.0 \\
    UA 0 & 97.5 & 97.6 & 96.4 & 96.8 & 96.3 & 96.9 & 96.9 & 97.5 & 97.4 & 96.7 & 96.8 & 96.8 & 97.8 & 95.7 & 97.0 \\
    UA 1 & 93.4 & 93.3 & 82.4 & 81.0 & 84.8 & 78.9 & 81.2 & 93.1 & 93.6 & 86.2 & 83.3 & 83.3 & 92.9 & 79.0 & 93.7 \\
    UA 2 & 92.1 & 92.2 & 75.8 & 81.4 & 92.7 & 58.5 & 86.2 & 95.4 & 94.0 & 88.7 & 84.2 & 84.2 & 91.3 & 87.1 & 87.6 \\
    UA 3 & 78.9 & 81.2 & 59.9 & 62.9 & 65.8 & 56.3 & 63.2 & 78.0 & 81.2 & 60.9 & 60.4 & 60.4 & 71.4 & 63.7 & 79.0 \\
    UA 4 & 72.9 & 73.4 & 53.8 & 60.2 & 57.0 & 47.0 & 60.3 & 69.3 & 74.6 & 52.7 & 53.5 & 53.5 & 73.9 & 62.1 & 74.6 \\
    UA 5 & 69.9 & 71.2 & 45.9 & 49.9 & 58.1 & 55.7 & 54.6 & 65.2 & 67.3 & 49.5 & 47.4 & 47.4 & 67.0 & 55.2 & 69.8 \\
    UA 6 & 71.3 & 71.6 & 62.2 & 55.9 & 65.8 & 27.8 & 56.2 & 77.8 & 76.6 & 63.6 & 58.0 & 58.0 & 71.7 & 59.6 & 72.0 \\
    UA 7 & 88.0 & 87.5 & 81.1 & 80.2 & 81.6 & 78.1 & 79.8 & 84.6 & 88.6 & 83.2 & 79.0 & 79.0 & 92.5 & 74.1 & 91.1 \\
    UA 8 & 94.5 & 94.6 & 89.5 & 89.4 & 91.9 & 88.2 & 90.3 & 95.6 & 95.8 & 92.1 & 88.6 & 88.6 & 95.1 & 86.1 & 95.0 \\
    UA 9 & 95.6 & 95.7 & 98.1 & 76.5 & 84.1 & 99.8 & 75.3 & 92.2 & 93.2 & 65.2 & 62.0 & 62.0 & 99.0 & 90.3 & 99.5 \\
    UA 10 & 98.5 & 98.5 & 99.2 & 96.4 & 92.8 & 99.8 & 92.4 & 97.6 & 98.4 & 96.6 & 93.6 & 93.6 & 98.5 & 97.3 & 97.4 \\
    UA 11 & 94.1 & 94.4 & 79.9 & 88.2 & 93.6 & 92.7 & 88.5 & 94.4 & 93.7 & 91.4 & 88.7 & 88.7 & 97.6 & 91.5 & 98.0 \\
    UA 12 & 99.8 & 99.9 & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 & 99.9 & 100.0 & 100.0 & 100.0 & 99.7 & 98.7 & 99.7 \\
  \end{tabular}
\end{table}

\begin{table}[h]
  \footnotesize
  \centering
  \caption{Heves}
  \begin{tabular}{lrrrrrrrrrrrrrrr}
    Method & M0 & M1 & M2 & M3 & M4 & M5 & M6 & NM0 & NM1 & NM2 & NM3 & NM3 & KDA & RF & SVM \\
    OA & 71.6 & 72.2 & 64.5 & 65.1 & 65.1 & 63.9 & 64.1 & 69.5 & 72.9 & 63.1 & 64.2 & 64.2 & 71.5 & 64.4 & 70.8 \\
    Kappa & 0.664 & 0.671 & 0.588 & 0.594 & 0.595 & 0.582 & 0.583 & 0.64 & 0.677 & 0.573 & 0.585 & 0.585 & 0.666 & 0.585 & 0.658 \\
    CPT & 148 & 151 & 148 & 152 & 152 & 148 & 152 & 148 & 151 & 148 & 152 & 152 & 695 & 18 & 171 \\
    PA 0 & 82.1 & 81.9 & 75.5 & 77.4 & 82.0 & 77.0 & 77.8 & 84.8 & 83.8 & 76.9 & 76.8 & 76.8 & 81.0 & 77.1 & 82.5 \\
    PA 1 & 51.7 & 51.2 & 36.9 & 37.1 & 46.0 & 39.2 & 38.2 & 53.5 & 53.5 & 39.2 & 38.9 & 38.9 & 50.3 & 41.8 & 55.0 \\
    PA 2 & 73.2 & 73.3 & 62.0 & 62.7 & 66.5 & 61.2 & 61.6 & 71.0 & 72.4 & 60.6 & 61.3 & 61.3 & 73.6 & 59.6 & 71.4 \\
    PA 3 & 67.2 & 68.1 & 51.2 & 53.5 & 61.2 & 51.1 & 54.2 & 68.1 & 71.5 & 56.2 & 56.9 & 56.9 & 65.0 & 55.1 & 67.6 \\
    PA 4 & 94.8 & 94.8 & 93.0 & 92.0 & 91.6 & 92.5 & 90.5 & 88.6 & 92.8 & 89.0 & 88.4 & 88.4 & 96.1 & 88.0 & 94.6 \\
    PA 5 & 51.2 & 43.8 & 47.2 & 44.1 & 61.4 & 49.5 & 47.5 & 55.1 & 44.3 & 46.1 & 42.7 & 42.7 & 65.8 & 51.5 & 65.5 \\
    PA 6 & 51.5 & 52.5 & 51.7 & 51.3 & 50.9 & 50.2 & 51.0 & 46.3 & 51.2 & 49.9 & 50.2 & 50.2 & 58.2 & 48.6 & 54.3 \\
    PA 7 & 72.9 & 72.7 & 74.2 & 74.3 & 75.8 & 74.0 & 71.6 & 74.7 & 73.8 & 78.2 & 75.7 & 75.7 & 79.7 & 70.0 & 79.2 \\
    PA 8 & 61.0 & 60.0 & 54.3 & 55.3 & 61.6 & 54.4 & 56.8 & 64.6 & 61.8 & 55.0 & 55.0 & 55.0 & 62.1 & 51.3 & 58.9 \\
    PA 9 & 90.8 & 91.0 & 89.1 & 86.8 & 86.9 & 88.5 & 85.3 & 89.8 & 90.9 & 85.9 & 84.8 & 84.8 & 91.3 & 87.2 & 91.5 \\
    PA 10 & 64.6 & 64.8 & 60.6 & 55.4 & 61.3 & 62.4 & 57.2 & 58.3 & 57.1 & 55.7 & 51.3 & 51.3 & 65.2 & 45.9 & 60.9 \\
    PA 11 & 72.4 & 77.4 & 66.4 & 67.4 & 53.5 & 63.5 & 63.4 & 59.6 & 75.2 & 63.3 & 65.5 & 65.5 & 74.2 & 61.6 & 74.4 \\
    PA 12 & 42.6 & 42.5 & 29.9 & 39.6 & 43.3 & 30.2 & 40.3 & 54.7 & 53.6 & 47.1 & 50.9 & 50.9 & 49.9 & 31.7 & 53.0 \\
    PA 13 & 90.3 & 90.4 & 87.6 & 87.6 & 88.0 & 87.7 & 86.9 & 89.8 & 90.4 & 87.4 & 86.6 & 86.6 & 90.3 & 85.3 & 90.5 \\
    PA 14 & 71.6 & 71.8 & 61.9 & 63.5 & 64.9 & 61.6 & 62.8 & 72.8 & 74.6 & 60.4 & 63.3 & 63.3 & 66.9 & 65.5 & 65.9 \\
    PA 15 & 92.4 & 92.4 & 90.3 & 88.4 & 86.6 & 89.9 & 86.8 & 87.6 & 89.8 & 84.5 & 83.4 & 83.4 & 90.5 & 84.1 & 89.9 \\
    UA 0 & 65.7 & 65.5 & 70.0 & 67.6 & 65.0 & 66.2 & 67.1 & 63.4 & 64.0 & 69.9 & 68.8 & 68.8 & 69.0 & 56.9 & 66.4 \\
    UA 1 & 44.0 & 44.1 & 35.2 & 34.0 & 37.7 & 33.7 & 32.8 & 39.6 & 41.6 & 35.9 & 33.4 & 33.4 & 42.7 & 28.8 & 41.5 \\
    UA 2 & 65.1 & 65.9 & 65.1 & 63.8 & 64.1 & 65.4 & 64.4 & 65.7 & 66.9 & 65.8 & 65.2 & 65.2 & 71.7 & 59.4 & 67.2 \\
    UA 3 & 31.7 & 31.2 & 35.9 & 28.2 & 31.0 & 36.1 & 29.0 & 23.9 & 24.9 & 23.0 & 21.0 & 21.0 & 30.9 & 22.9 & 22.2 \\
    UA 4 & 76.3 & 78.1 & 71.9 & 72.0 & 73.0 & 74.4 & 73.9 & 85.4 & 83.6 & 79.2 & 78.6 & 78.6 & 83.7 & 68.6 & 82.0 \\
    UA 5 & 40.8 & 43.0 & 31.9 & 34.2 & 29.2 & 30.1 & 31.9 & 32.1 & 42.7 & 29.6 & 32.9 & 32.9 & 49.1 & 33.0 & 42.4 \\
    UA 6 & 73.2 & 74.1 & 58.4 & 57.2 & 64.6 & 58.9 & 56.5 & 76.4 & 74.6 & 59.3 & 58.5 & 58.5 & 70.9 & 55.6 & 74.9 \\
    UA 7 & 50.0 & 52.2 & 25.8 & 25.2 & 32.3 & 27.1 & 26.8 & 51.4 & 54.2 & 24.7 & 26.5 & 26.5 & 43.5 & 25.0 & 46.7 \\
    UA 8 & 16.3 & 16.4 & 13.6 & 12.7 & 12.9 & 14.4 & 12.7 & 12.7 & 14.9 & 12.7 & 12.1 & 12.1 & 14.8 & 11.3 & 15.6 \\
    UA 9 & 97.9 & 98.0 & 86.1 & 94.3 & 98.4 & 89.7 & 95.6 & 98.3 & 98.1 & 98.0 & 98.8 & 98.8 & 98.4 & 97.3 & 97.6 \\
    UA 10 & 12.4 & 12.3 & 11.2 & 12.3 & 11.7 & 10.5 & 11.5 & 14.9 & 15.4 & 11.7 & 13.5 & 13.5 & 11.7 & 13.5 & 13.6 \\
    UA 11 & 86.8 & 85.2 & 85.3 & 85.5 & 89.1 & 85.7 & 86.0 & 88.1 & 85.8 & 85.7 & 85.9 & 85.9 & 91.4 & 87.1 & 91.5 \\
    UA 12 & 27.1 & 25.2 & 14.2 & 9.4 & 16.1 & 14.1 & 8.7 & 16.3 & 19.9 & 7.6 & 6.9 & 6.9 & 16.9 & 12.1 & 20.0 \\
    UA 13 & 74.2 & 74.4 & 83.8 & 81.1 & 73.5 & 84.3 & 85.0 & 73.3 & 72.4 & 88.2 & 88.8 & 88.8 & 73.5 & 59.6 & 72.6 \\
    UA 14 & 96.8 & 96.7 & 97.6 & 97.9 & 98.1 & 97.5 & 97.9 & 97.2 & 96.5 & 98.0 & 97.9 & 97.9 & 97.6 & 97.6 & 97.2 \\
    UA 15 & 58.4 & 60.1 & 32.3 & 51.2 & 65.8 & 34.8 & 51.4 & 80.8 & 75.5 & 64.1 & 69.8 & 69.8 & 58.6 & 56.2 & 56.5 \\
  \end{tabular}
\end{table}

\end{document}

### 2016-12-12 response.tex
add resp to 1.3 1.4

@@
   \begin{resbox}
+  The value of the derivative of criterion is compared to a user-defined threshold.
+The first time the value reaches below the threshold correspond to the number of selected variables.
+The derivative for the example presented on Fig.
+5 of the paper is presented in Figure~\ref{fig:deriv}.
   With the same example, 7 features gives a kappa of $0.620$ and 18 features a kappa of $0.657$.
@@
   \begin{tikzpicture}
-          \begin{axis}[ymin=0,ymax=0.3,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
+    \begin{axis}[ymin=0,ymax=0.25,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Derivative of\\Cohen's kappa},ylabel style={align=center},yticklabel style={/pgf/number format/fixed,/pgf/number format/precision=3}]
       \addplot[thick,black] table[x expr=\coordindex,y=deriv] {derivate_criterion_1000spl_aisa.txt};
-              % \draw[thick,red] (axis cs:18,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:18,\pgfkeysvalueof{/pgfplots/ymax});
-              % \draw[thick,black] (axis cs:24,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:24,\pgfkeysvalueof{/pgfplots/ymax});
+      \draw[thick,red] (axis cs:18,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:18,\pgfkeysvalueof{/pgfplots/ymax});
+      \draw[thick,black] (axis cs:24,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:24,\pgfkeysvalueof{/pgfplots/ymax});
     \end{axis};
   \end{tikzpicture}
-      \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Aisa dataset with 500 samples by class.
-Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.}
+  \caption{Derivative of criterion (kappa) in function of the number of selected variables for first trial with Aisa dataset with 500 samples by class.
+Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.\label{fig:deriv}}
 \end{figure}
@@
   \begin{resbox}
-
+    We also considered this possibility and it is a choice available in the code of the OTB module we provide.
+But taking the maximum appears not to be the optimal choice in the sense that the algorithm selects sometimes more variables than necessary.
+Sometimes, at a given point, the best new selected variable is one which contains no information (for example one whose contribution is thresholded by EPS\_FLT) because all others decrease the criterion value.
+If it happens several times, it can lead to a final drop of the criterion.
+The algorithm then tries to counterbalance the drop and regains the original maximum but can be slightly higher.
+Figure~\ref{fig:crit-evol} illustrated this case.
+    The problem is that we introduce a few irrelevant features and we want to avoid that if possible.
+This case appears more often with the second dataset but we choose to use the same method for both dataset.
   \end{resbox}
 \end{revbox}
+
+\begin{figure}[!t]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[ymin=0.2,ymax=0.9,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
+      \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
+    \end{axis};
+  \end{tikzpicture}
+  \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Potsdam dataset with 1000 samples by class.\label{fig:crit-evol}}
+\end{figure}
 
@@
   \begin{resbox}
+    It is true that we can try with fewer samples. NB: TODO
+    But in the case of a larger training set is not possible to test with the chosen validation method.
+We use an equal number of samples in each class and 50\% of the samples are used for test.
+Thus, as the less represented class contains 2107 samples, we are limited to 1053 samples by class with this dataset.
+Nevertheless, the second dataset illustrates the case and we also add the following comment in the paper to make it clear that RF could perform better with more samples:
+    \emph{Random Forest has similar performance in term of classification rate with 1000 samples \rev{ and one could expect to get a better classification rate with RF if more samples were available}.}
   \end{resbox}

### 2016-12-12 paper.tex
add review response
add resp to 1.3 1.4

@@
-%\documentclass[journal,peerreview,onecolumn]{IEEEtran}
+% \documentclass[journal,peerreview,onecolumn]{IEEEtran}
 \documentclass[journal]{IEEEtran}
@@
     The difference with the second best classifier, Random Forest, appears to be significant when using 250 and 500 samples.
-Random Forest has similar performance in term of classification rate with 1000 samples.
+Random Forest has similar performance in term of classification rate with 1000 samples and one could expect to get a better classification rate with RF if more samples were available.
 The GMM classifier with ridge regularization and the KNN classifier are both outperformed.
@@
     The  Postdam classes  are  more difficult  to discriminate,  since
-    Gaussianity assumption  does not  hold. For instance,  a buildings
+    Gaussianity assumption  does not  hold. For instance,  a building
     can  be  made of  various  materials,  resulting in  heterogeneous

### 2016-12-12 response.tex
delete section of former doc...

@@
   \begin{resbox}
-    % Done. Thank you for the careful reading.
+    % Thank you for the careful reading.
     NB:
@@
 \end{revbox}
-
-\section{Reviewer 2}
-
-\begin{revbox}
-  page 4 :  C. Model inference equations seem correct  but what is the
-  typical range of p-hat (fixed parameter)?
-  \begin{resbox}
-
-  \end{resbox}
-\end{revbox}
-
-\begin{revbox}
-  Experimental Results : Absence of such visual graphics such as final
-  classification map,  original imagery and accuracy  studies based on
-  classification  performance   (not  just  compute  time   and  kappa
-  coefficients, but  in terms  of miss classification  rate, confusion
-  matrix from classifications etc.)
-
-  \begin{resbox}
-
-  \end{resbox}
-\end{revbox}
-
-\clearpage
 

### 2016-12-15 response.tex
add data file to plot derivate of criterion
add res simu wth 50 spls

@@
   \begin{resbox}
-    It is true that we can try with fewer samples. NB: TODO
+    It is true that we can try with fewer samples. We tried with 50 samples by class and it leads to the same behavior than with 250 samples (see Table~\ref{tab:aisa-otbsimu}).
+
     But in the case of a larger training set is not possible to test with the chosen validation method.
@@
 \end{revbox}
+
+\begin{table}[!t]
+    \centering
+    \caption{Average classification accuracy 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
+    \begin{tabular}{lcccc}\toprule
+         & \multicolumn{4}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-5}
+        \# samples by class & 50 & 250 & 500 & 1000 \\ \midrule
+
+        GMM SFS kappa &  0.612 (0.028) & {\bfseries 0.678 (0.029)} & {\bfseries 0.687 (0.029)} & {\bfseries 0.699 (0.028)} \\
+        GMM SFS JM &     0.622 (0.026) & {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.701 (0.029)} \\
+        GMM SFFS JM &    0.622 (0.026) & {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.701 (0.029)} \\
+        GMM ridge &      0.597 (0.035) & 0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
+        KNN &            0.481 (0.054) & 0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
+        Random Forest &  0.555 (0.031) & 0.645 (0.026) & 0.673 (0.023) & {\bfseries 0.693 (0.023)} \\
+        \bottomrule
+    \end{tabular}
+\end{table}
 
@@
     NB:
+
     no $v_c$ after prop 2...
+
     building corrected. Need to correct english at the end

### 2016-12-16 response.tex
update

@@
     The interest of this second data cube is first to test the algorithm in a different context.
-In particular, this second dataset should be seen as a method to learn the most relevant kind of features and the best parameters to extract those features.
+In particular, the algorithm applied to this second dataset should be seen as a method to learn the most relevant kind of features and the best parameters to extract those features.
+
     In the first dataset, only spectral features has been considered and in the particular context of hyperspectral imaging which imply contiguous features highly correlated.
+
     (reponse bof à revoir et developper mais c'est un debut)
@@
   \begin{resbox}
-
+    see Table~\ref{tab:aisa-otbsimu-all}
   \end{resbox}
 \end{revbox}
+
+    \begin{table}[!t]
+        \centering
+        \caption{Average classification accuracy 20 trials with all configurations of criterion and method (standard deviation in parenthesis).\label{tab:aisa-otbsimu-all}}
+        \begin{tabular}{lccc}\toprule
+             & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-4}
+            \# samples by class & 250 & 500 & 1000 \\ \midrule
+
+            GMM SFS JM &    0.685 (0.030) & 0.689 (0.030) & 0.701 (0.029) \\
+            GMM SFS divKL &    &&\\
+            GMM SFS kappa & 0.678 (0.029) & 0.687 (0.029) & 0.699 (0.028) \\
+            GMM SFS OA &    &&\\
+            GMM SFS F1mean &    &&\\
+            GMM SFFS JM &   0.685 (0.030) & 0.689 (0.030) & 0.701 (0.029) \\
+            GMM SFFS divKL &   &&\\
+            GMM SFFS kappa &   &&\\
+            GMM SFFS OA &   &&\\
+            GMM SFFS F1mean &   &&\\
+            \bottomrule
+        \end{tabular}
+    \end{table}
 
@@
   \begin{resbox}
-
+    see Table~\ref{tab:aisa-otbsimu-othereval}
   \end{resbox}
 \end{revbox}
+
+    \begin{table}[!t]
+        \centering
+        \caption{Average classification accuracy 20 trials evaluated with overall accuracy and mean of F1-score (standard deviation in parenthesis).\label{tab:aisa-otbsimu-othereval}}
+        \begin{tabular}{lcccccc}\toprule
+             & \multicolumn{3}{c}{\bfseries Overall Accuracy} & \multicolumn{3}{c}{\bfseries F1-mean} \\ \cmidrule{2-7}
+            \# samples by class & 250 & 500 & 1000  & 250 & 500 & 1000 \\ \midrule
+
+            GMM SFS kappa & & & & & & \\
+            GMM SFS JM &    & & & & & \\
+            GMM SFFS JM &   & & & & & \\
+            GMM ridge &     & & & & & \\
+            KNN &           & & & & & \\
+            Random Forest & & & & & & \\
+            \bottomrule
+        \end{tabular}
+    \end{table}
 
@@
   \begin{resbox}
-
+    We totally agree that some mathematical formulations we used are already known.
+Nevertheless, we think that these formulations in this particular context has not been exploited so far.
+The results of our experiment shows that the proposed algorithm has the advantage to be very efficient in term of processing time with a fairly reasonable cost in term of classification accuracy.
+It has to be noticed that compared to similar methods using Gaussian Mixture model, no additional hypothesis are made on the covariance matrix which is tractable only because of the efficient implementation.
   \end{resbox}
@@
   \begin{resbox}
+    After reading with interest this two references, we fail to see the link between these publications and our submission except for the type of data used to test algorithms.
 
+    First reference reviews ways to assign several labels to one sample and so performs spectral unmixing.
+Second reference focus on the same topic but develop a deep learning approach.
+Although very interesting, we do not address the subject of multi-labeling in our work.
+
+    According to these observations, we choose to respectfully disagree with the reviewer and take the decision not to add a digression to cite these references.
   \end{resbox}
@@
 
+  TODO
+
   \end{resbox}
@@
 
-\begin{table}[h]
-  \footnotesize
-  \centering
-  \caption{University}
-  \begin{tabular}{lrrrrrrrrrrrrrrr}
-    Method & M0 & M1 & M2 & M3 & M4 & M5 & M6 & NM0 & NM1 & NM2 & NM3 & NM3 & KDA & RF & SVM \\
-    OA & 82.0 & 83.9 & 69.6 & 68.0 & 73.3 & 65.2 & 68.7 & 78.7 & 83.8 & 67.7 & 65.8 & 65.8 & 83.5 & 71.9 & 84.5 \\
-    Kappa & 0.768 & 0.793 & 0.617 & 0.603 & 0.661 & 0.567 & 0.61 & 0.73 & 0.792 & 0.599 & 0.578 & 0.578 & 0.786 & 0.646 & 0.799 \\
-    CPT & 18 & 18 & 18 & 19 & 19 & 18 & 19 & 17 & 18 & 18 & 19 & 19 & 98 & 3 & 10 \\
-    PA 0 & 82.0 & 81.6 & 74.0 & 67.5 & 73.5 & 64.0 & 68.7 & 83.4 & 87.5 & 71.8 & 66.4 & 66.4 & 73.0 & 73.1 & 80.0 \\
-    PA 1 & 79.5 & 82.5 & 65.1 & 59.3 & 69.4 & 64.3 & 61.5 & 72.9 & 80.1 & 60.7 & 56.9 & 56.9 & 84.1 & 65.1 & 84.6 \\
-    PA 2 & 72.1 & 75.9 & 58.4 & 57.4 & 75.8 & 63.3 & 64.9 & 72.7 & 75.6 & 54.5 & 53.5 & 53.5 & 82.4 & 68.2 & 79.6 \\
-    PA 3 & 96.3 & 96.7 & 81.7 & 93.6 & 91.9 & 68.2 & 91.3 & 96.4 & 97.7 & 94.0 & 93.3 & 93.3 & 92.1 & 93.6 & 92.5 \\
-    PA 4 & 99.4 & 99.5 & 88.1 & 95.6 & 98.2 & 80.4 & 95.7 & 99.6 & 99.7 & 91.8 & 96.5 & 96.5 & 98.6 & 99.0 & 99.3 \\
-    PA 5 & 79.6 & 82.9 & 64.8 & 68.0 & 65.7 & 56.9 & 66.4 & 75.3 & 85.3 & 67.9 & 64.8 & 64.8 & 83.3 & 68.6 & 83.6 \\
-    PA 6 & 86.9 & 91.3 & 71.8 & 89.5 & 83.2 & 71.1 & 88.0 & 86.4 & 84.0 & 62.9 & 86.9 & 86.9 & 93.4 & 87.2 & 90.1 \\
-    PA 7 & 79.1 & 78.6 & 72.9 & 72.8 & 67.4 & 66.2 & 67.8 & 78.5 & 77.3 & 68.0 & 70.0 & 70.0 & 79.8 & 71.1 & 78.5 \\
-    PA 8 & 99.8 & 99.7 & 99.2 & 98.9 & 99.0 & 99.5 & 98.8 & 99.5 & 99.6 & 97.3 & 96.6 & 96.6 & 99.7 & 99.9 & 99.9 \\
-    UA 0 & 94.7 & 96.0 & 81.5 & 95.1 & 94.4 & 86.8 & 93.1 & 94.8 & 94.5 & 89.6 & 95.0 & 95.0 & 96.0 & 93.7 & 94.5 \\
-    UA 1 & 93.4 & 94.5 & 88.9 & 88.2 & 88.3 & 88.7 & 88.2 & 91.4 & 95.1 & 88.8 & 86.8 & 86.8 & 94.8 & 88.3 & 94.2 \\
-    UA 2 & 60.8 & 62.4 & 41.2 & 45.5 & 46.8 & 35.3 & 43.6 & 60.0 & 62.4 & 41.9 & 43.7 & 43.7 & 60.3 & 49.2 & 65.6 \\
-    UA 3 & 60.2 & 65.7 & 61.0 & 51.5 & 59.7 & 65.5 & 53.1 & 60.3 & 63.9 & 49.5 & 48.7 & 48.7 & 77.1 & 57.6 & 77.6 \\
-    UA 4 & 97.9 & 97.9 & 99.9 & 75.8 & 81.8 & 98.1 & 74.1 & 92.8 & 96.7 & 68.9 & 56.3 & 56.3 & 98.9 & 92.6 & 96.9 \\
-    UA 5 & 67.1 & 69.4 & 41.4 & 39.4 & 45.1 & 41.7 & 40.3 & 53.3 & 66.0 & 39.9 & 37.4 & 37.4 & 66.4 & 43.1 & 66.2 \\
-    UA 6 & 65.9 & 64.7 & 44.3 & 48.7 & 58.3 & 24.9 & 50.2 & 71.9 & 81.2 & 55.6 & 53.5 & 53.5 & 55.4 & 52.8 & 59.9 \\
-    UA 7 & 76.8 & 79.3 & 68.9 & 73.1 & 78.8 & 65.8 & 75.5 & 80.2 & 80.2 & 71.8 & 73.0 & 73.0 & 75.9 & 75.2 & 80.4 \\
-    UA 8 & 99.9 & 99.9 & 90.6 & 100.0 & 100.0 & 44.6 & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 & 99.8 & 99.8 & 99.9 \\
-  \end{tabular}
-\end{table}
-
-\begin{table}[h]
-  \footnotesize
-  \centering
-  \caption{KSC}
-  \begin{tabular}{lrrrrrrrrrrrrrrr}
-    Method & M0 & M1 & M2 & M3 & M4 & M5 & M6 & NM0 & NM1 & NM2 & NM3 & NM3 & KDA & RF & SVM \\
-    OA & 92.9 & 93.1 & 86.0 & 85.9 & 88.4 & 83.9 & 86.1 & 92.1 & 92.9 & 85.5 & 83.6 & 83.6 & 93.2 & 86.9 & 93.5 \\
-    Kappa & 0.92 & 0.922 & 0.844 & 0.842 & 0.87 & 0.82 & 0.845 & 0.911 & 0.921 & 0.838 & 0.817 & 0.817 & 0.924 & 0.853 & 0.928 \\
-    CPT & 31 & 33 & 31 & 33 & 34 & 32 & 34 & 31 & 33 & 31 & 33 & 33 & 253 & 3 & 28 \\
-    PA 0 & 90.7 & 90.8 & 87.0 & 85.9 & 88.2 & 86.8 & 85.6 & 89.9 & 90.3 & 85.3 & 84.2 & 84.2 & 87.0 & 88.1 & 90.0 \\
-    PA 1 & 88.9 & 89.6 & 83.7 & 80.7 & 85.5 & 83.3 & 80.8 & 90.6 & 90.3 & 82.6 & 80.2 & 80.2 & 87.7 & 83.5 & 91.0 \\
-    PA 2 & 90.1 & 91.7 & 86.1 & 89.9 & 86.6 & 86.6 & 89.9 & 85.0 & 89.2 & 86.3 & 87.0 & 87.0 & 88.2 & 89.6 & 93.0 \\
-    PA 3 & 77.2 & 78.3 & 61.3 & 67.7 & 70.3 & 64.1 & 70.9 & 74.3 & 77.9 & 62.6 & 61.9 & 61.9 & 79.8 & 64.1 & 78.5 \\
-    PA 4 & 81.5 & 83.4 & 63.1 & 61.8 & 71.7 & 66.7 & 64.1 & 79.1 & 82.2 & 62.2 & 59.2 & 59.2 & 82.8 & 65.0 & 80.4 \\
-    PA 5 & 80.7 & 80.4 & 55.3 & 51.1 & 58.0 & 36.2 & 48.0 & 84.2 & 83.4 & 53.0 & 45.5 & 45.5 & 82.5 & 56.3 & 79.6 \\
-    PA 6 & 95.5 & 95.6 & 89.5 & 93.9 & 92.0 & 91.4 & 94.6 & 92.4 & 94.0 & 88.1 & 90.4 & 90.4 & 96.8 & 93.1 & 94.8 \\
-    PA 7 & 91.9 & 92.3 & 82.0 & 81.7 & 88.0 & 84.8 & 82.8 & 93.2 & 93.2 & 82.0 & 74.1 & 74.1 & 93.0 & 81.0 & 93.8 \\
-    PA 8 & 96.4 & 96.6 & 92.6 & 91.8 & 94.3 & 93.9 & 92.1 & 94.9 & 96.1 & 90.1 & 86.7 & 86.7 & 97.5 & 89.4 & 95.3 \\
-    PA 9 & 95.1 & 94.9 & 81.2 & 88.5 & 90.5 & 73.5 & 88.8 & 95.4 & 95.5 & 94.5 & 93.0 & 93.0 & 96.8 & 86.4 & 96.9 \\
-    PA 10 & 97.7 & 98.1 & 90.2 & 90.4 & 94.7 & 85.0 & 90.9 & 97.7 & 97.9 & 89.6 & 90.2 & 90.2 & 97.4 & 95.8 & 97.9 \\
-    PA 11 & 93.0 & 92.5 & 88.0 & 81.0 & 82.1 & 78.5 & 80.8 & 90.0 & 92.4 & 80.5 & 77.7 & 77.7 & 96.3 & 83.2 & 95.1 \\
-    PA 12 & 98.9 & 98.9 & 97.8 & 98.2 & 98.4 & 97.9 & 98.3 & 98.6 & 98.6 & 97.1 & 97.6 & 97.6 & 99.9 & 99.4 & 100.0 \\
-    UA 0 & 97.5 & 97.6 & 96.4 & 96.8 & 96.3 & 96.9 & 96.9 & 97.5 & 97.4 & 96.7 & 96.8 & 96.8 & 97.8 & 95.7 & 97.0 \\
-    UA 1 & 93.4 & 93.3 & 82.4 & 81.0 & 84.8 & 78.9 & 81.2 & 93.1 & 93.6 & 86.2 & 83.3 & 83.3 & 92.9 & 79.0 & 93.7 \\
-    UA 2 & 92.1 & 92.2 & 75.8 & 81.4 & 92.7 & 58.5 & 86.2 & 95.4 & 94.0 & 88.7 & 84.2 & 84.2 & 91.3 & 87.1 & 87.6 \\
-    UA 3 & 78.9 & 81.2 & 59.9 & 62.9 & 65.8 & 56.3 & 63.2 & 78.0 & 81.2 & 60.9 & 60.4 & 60.4 & 71.4 & 63.7 & 79.0 \\
-    UA 4 & 72.9 & 73.4 & 53.8 & 60.2 & 57.0 & 47.0 & 60.3 & 69.3 & 74.6 & 52.7 & 53.5 & 53.5 & 73.9 & 62.1 & 74.6 \\
-    UA 5 & 69.9 & 71.2 & 45.9 & 49.9 & 58.1 & 55.7 & 54.6 & 65.2 & 67.3 & 49.5 & 47.4 & 47.4 & 67.0 & 55.2 & 69.8 \\
-    UA 6 & 71.3 & 71.6 & 62.2 & 55.9 & 65.8 & 27.8 & 56.2 & 77.8 & 76.6 & 63.6 & 58.0 & 58.0 & 71.7 & 59.6 & 72.0 \\
-    UA 7 & 88.0 & 87.5 & 81.1 & 80.2 & 81.6 & 78.1 & 79.8 & 84.6 & 88.6 & 83.2 & 79.0 & 79.0 & 92.5 & 74.1 & 91.1 \\
-    UA 8 & 94.5 & 94.6 & 89.5 & 89.4 & 91.9 & 88.2 & 90.3 & 95.6 & 95.8 & 92.1 & 88.6 & 88.6 & 95.1 & 86.1 & 95.0 \\
-    UA 9 & 95.6 & 95.7 & 98.1 & 76.5 & 84.1 & 99.8 & 75.3 & 92.2 & 93.2 & 65.2 & 62.0 & 62.0 & 99.0 & 90.3 & 99.5 \\
-    UA 10 & 98.5 & 98.5 & 99.2 & 96.4 & 92.8 & 99.8 & 92.4 & 97.6 & 98.4 & 96.6 & 93.6 & 93.6 & 98.5 & 97.3 & 97.4 \\
-    UA 11 & 94.1 & 94.4 & 79.9 & 88.2 & 93.6 & 92.7 & 88.5 & 94.4 & 93.7 & 91.4 & 88.7 & 88.7 & 97.6 & 91.5 & 98.0 \\
-    UA 12 & 99.8 & 99.9 & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 & 99.9 & 100.0 & 100.0 & 100.0 & 99.7 & 98.7 & 99.7 \\
-  \end{tabular}
-\end{table}
-
-\begin{table}[h]
-  \footnotesize
-  \centering
-  \caption{Heves}
-  \begin{tabular}{lrrrrrrrrrrrrrrr}
-    Method & M0 & M1 & M2 & M3 & M4 & M5 & M6 & NM0 & NM1 & NM2 & NM3 & NM3 & KDA & RF & SVM \\
-    OA & 71.6 & 72.2 & 64.5 & 65.1 & 65.1 & 63.9 & 64.1 & 69.5 & 72.9 & 63.1 & 64.2 & 64.2 & 71.5 & 64.4 & 70.8 \\
-    Kappa & 0.664 & 0.671 & 0.588 & 0.594 & 0.595 & 0.582 & 0.583 & 0.64 & 0.677 & 0.573 & 0.585 & 0.585 & 0.666 & 0.585 & 0.658 \\
-    CPT & 148 & 151 & 148 & 152 & 152 & 148 & 152 & 148 & 151 & 148 & 152 & 152 & 695 & 18 & 171 \\
-    PA 0 & 82.1 & 81.9 & 75.5 & 77.4 & 82.0 & 77.0 & 77.8 & 84.8 & 83.8 & 76.9 & 76.8 & 76.8 & 81.0 & 77.1 & 82.5 \\
-    PA 1 & 51.7 & 51.2 & 36.9 & 37.1 & 46.0 & 39.2 & 38.2 & 53.5 & 53.5 & 39.2 & 38.9 & 38.9 & 50.3 & 41.8 & 55.0 \\
-    PA 2 & 73.2 & 73.3 & 62.0 & 62.7 & 66.5 & 61.2 & 61.6 & 71.0 & 72.4 & 60.6 & 61.3 & 61.3 & 73.6 & 59.6 & 71.4 \\
-    PA 3 & 67.2 & 68.1 & 51.2 & 53.5 & 61.2 & 51.1 & 54.2 & 68.1 & 71.5 & 56.2 & 56.9 & 56.9 & 65.0 & 55.1 & 67.6 \\
-    PA 4 & 94.8 & 94.8 & 93.0 & 92.0 & 91.6 & 92.5 & 90.5 & 88.6 & 92.8 & 89.0 & 88.4 & 88.4 & 96.1 & 88.0 & 94.6 \\
-    PA 5 & 51.2 & 43.8 & 47.2 & 44.1 & 61.4 & 49.5 & 47.5 & 55.1 & 44.3 & 46.1 & 42.7 & 42.7 & 65.8 & 51.5 & 65.5 \\
-    PA 6 & 51.5 & 52.5 & 51.7 & 51.3 & 50.9 & 50.2 & 51.0 & 46.3 & 51.2 & 49.9 & 50.2 & 50.2 & 58.2 & 48.6 & 54.3 \\
-    PA 7 & 72.9 & 72.7 & 74.2 & 74.3 & 75.8 & 74.0 & 71.6 & 74.7 & 73.8 & 78.2 & 75.7 & 75.7 & 79.7 & 70.0 & 79.2 \\
-    PA 8 & 61.0 & 60.0 & 54.3 & 55.3 & 61.6 & 54.4 & 56.8 & 64.6 & 61.8 & 55.0 & 55.0 & 55.0 & 62.1 & 51.3 & 58.9 \\
-    PA 9 & 90.8 & 91.0 & 89.1 & 86.8 & 86.9 & 88.5 & 85.3 & 89.8 & 90.9 & 85.9 & 84.8 & 84.8 & 91.3 & 87.2 & 91.5 \\
-    PA 10 & 64.6 & 64.8 & 60.6 & 55.4 & 61.3 & 62.4 & 57.2 & 58.3 & 57.1 & 55.7 & 51.3 & 51.3 & 65.2 & 45.9 & 60.9 \\
-    PA 11 & 72.4 & 77.4 & 66.4 & 67.4 & 53.5 & 63.5 & 63.4 & 59.6 & 75.2 & 63.3 & 65.5 & 65.5 & 74.2 & 61.6 & 74.4 \\
-    PA 12 & 42.6 & 42.5 & 29.9 & 39.6 & 43.3 & 30.2 & 40.3 & 54.7 & 53.6 & 47.1 & 50.9 & 50.9 & 49.9 & 31.7 & 53.0 \\
-    PA 13 & 90.3 & 90.4 & 87.6 & 87.6 & 88.0 & 87.7 & 86.9 & 89.8 & 90.4 & 87.4 & 86.6 & 86.6 & 90.3 & 85.3 & 90.5 \\
-    PA 14 & 71.6 & 71.8 & 61.9 & 63.5 & 64.9 & 61.6 & 62.8 & 72.8 & 74.6 & 60.4 & 63.3 & 63.3 & 66.9 & 65.5 & 65.9 \\
-    PA 15 & 92.4 & 92.4 & 90.3 & 88.4 & 86.6 & 89.9 & 86.8 & 87.6 & 89.8 & 84.5 & 83.4 & 83.4 & 90.5 & 84.1 & 89.9 \\
-    UA 0 & 65.7 & 65.5 & 70.0 & 67.6 & 65.0 & 66.2 & 67.1 & 63.4 & 64.0 & 69.9 & 68.8 & 68.8 & 69.0 & 56.9 & 66.4 \\
-    UA 1 & 44.0 & 44.1 & 35.2 & 34.0 & 37.7 & 33.7 & 32.8 & 39.6 & 41.6 & 35.9 & 33.4 & 33.4 & 42.7 & 28.8 & 41.5 \\
-    UA 2 & 65.1 & 65.9 & 65.1 & 63.8 & 64.1 & 65.4 & 64.4 & 65.7 & 66.9 & 65.8 & 65.2 & 65.2 & 71.7 & 59.4 & 67.2 \\
-    UA 3 & 31.7 & 31.2 & 35.9 & 28.2 & 31.0 & 36.1 & 29.0 & 23.9 & 24.9 & 23.0 & 21.0 & 21.0 & 30.9 & 22.9 & 22.2 \\
-    UA 4 & 76.3 & 78.1 & 71.9 & 72.0 & 73.0 & 74.4 & 73.9 & 85.4 & 83.6 & 79.2 & 78.6 & 78.6 & 83.7 & 68.6 & 82.0 \\
-    UA 5 & 40.8 & 43.0 & 31.9 & 34.2 & 29.2 & 30.1 & 31.9 & 32.1 & 42.7 & 29.6 & 32.9 & 32.9 & 49.1 & 33.0 & 42.4 \\
-    UA 6 & 73.2 & 74.1 & 58.4 & 57.2 & 64.6 & 58.9 & 56.5 & 76.4 & 74.6 & 59.3 & 58.5 & 58.5 & 70.9 & 55.6 & 74.9 \\
-    UA 7 & 50.0 & 52.2 & 25.8 & 25.2 & 32.3 & 27.1 & 26.8 & 51.4 & 54.2 & 24.7 & 26.5 & 26.5 & 43.5 & 25.0 & 46.7 \\
-    UA 8 & 16.3 & 16.4 & 13.6 & 12.7 & 12.9 & 14.4 & 12.7 & 12.7 & 14.9 & 12.7 & 12.1 & 12.1 & 14.8 & 11.3 & 15.6 \\
-    UA 9 & 97.9 & 98.0 & 86.1 & 94.3 & 98.4 & 89.7 & 95.6 & 98.3 & 98.1 & 98.0 & 98.8 & 98.8 & 98.4 & 97.3 & 97.6 \\
-    UA 10 & 12.4 & 12.3 & 11.2 & 12.3 & 11.7 & 10.5 & 11.5 & 14.9 & 15.4 & 11.7 & 13.5 & 13.5 & 11.7 & 13.5 & 13.6 \\
-    UA 11 & 86.8 & 85.2 & 85.3 & 85.5 & 89.1 & 85.7 & 86.0 & 88.1 & 85.8 & 85.7 & 85.9 & 85.9 & 91.4 & 87.1 & 91.5 \\
-    UA 12 & 27.1 & 25.2 & 14.2 & 9.4 & 16.1 & 14.1 & 8.7 & 16.3 & 19.9 & 7.6 & 6.9 & 6.9 & 16.9 & 12.1 & 20.0 \\
-    UA 13 & 74.2 & 74.4 & 83.8 & 81.1 & 73.5 & 84.3 & 85.0 & 73.3 & 72.4 & 88.2 & 88.8 & 88.8 & 73.5 & 59.6 & 72.6 \\
-    UA 14 & 96.8 & 96.7 & 97.6 & 97.9 & 98.1 & 97.5 & 97.9 & 97.2 & 96.5 & 98.0 & 97.9 & 97.9 & 97.6 & 97.6 & 97.2 \\
-    UA 15 & 58.4 & 60.1 & 32.3 & 51.2 & 65.8 & 34.8 & 51.4 & 80.8 & 75.5 & 64.1 & 69.8 & 69.8 & 58.6 & 56.2 & 56.5 \\
-  \end{tabular}
-\end{table}
-
 \end{document}

### 2017-01-04 response.tex
update

@@
 
-            GMM SFS JM &    0.685 (0.030) & 0.689 (0.030) & 0.701 (0.029) \\
-            GMM SFS divKL &    &&\\
-            GMM SFS kappa & 0.678 (0.029) & 0.687 (0.029) & 0.699 (0.028) \\
-            GMM SFS OA &    &&\\
-            GMM SFS F1mean &    &&\\
-            GMM SFFS JM &   0.685 (0.030) & 0.689 (0.030) & 0.701 (0.029) \\
-            GMM SFFS divKL &   &&\\
-            GMM SFFS kappa &   &&\\
-            GMM SFFS OA &   &&\\
-            GMM SFFS F1mean &   &&\\
+            GMM SFS JM &      0.680 (0.028) & 0.700 (0.028) & 0.710 (0.030) \\
+            GMM SFS divKL &   0.680 (0.027) & 0.699 (0.028) & 0.708 (0.031) \\
+            GMM SFS kappa &   0.684 (0.025) & 0.698 (0.027) & 0.711 (0.028) \\
+            GMM SFS OA &      0.684 (0.025) & 0.698 (0.027) & 0.710 (0.028) \\
+            GMM SFS F1mean &  0.683 (0.027) & 0.699 (0.027) & 0.711 (0.028) \\
+            GMM SFFS JM &     0.680 (0.028) & 0.700 (0.028) & 0.710 (0.030) \\
+            GMM SFFS divKL &  0.668 (0.039) & 0.689 (0.029) & 0.669 (0.091) \\
+            GMM SFFS kappa &  0.648 (0.087) & 0.627 (0.100) & 0.615 (0.105) \\
+            GMM SFFS OA &     0.647 (0.087) & 0.604 (0.104) & 0.613 (0.100) \\
+            GMM SFFS F1mean & 0.650 (0.112) & 0.638 (0.105) & 0.625 (0.107) \\
             \bottomrule
@@
 
-            GMM SFS kappa & & & & & & \\
-            GMM SFS JM &    & & & & & \\
-            GMM SFFS JM &   & & & & & \\
-            GMM ridge &     & & & & & \\
-            KNN &           & & & & & \\
-            Random Forest & & & & & & \\
+            GMM SFS JM &      0.730 (0.026) & 0.748 (0.026) & 0.757 (0.027) & 0.591 (0.016) & 0.611 (0.016) & 0.622 (0.018) \\
+            GMM SFS divKL &   0.730 (0.024) & 0.747 (0.026) & 0.754 (0.029) & 0.587 (0.017) & 0.607 (0.017) & 0.616 (0.019) \\
+            GMM SFS kappa &   0.734 (0.024) & 0.746 (0.025) & 0.757 (0.026) & 0.593 (0.014) & 0.608 (0.015) & 0.621 (0.018) \\
+            GMM SFS OA &      0.734 (0.024) & 0.746 (0.025) & 0.757 (0.026) & 0.593 (0.014) & 0.608 (0.016) & 0.621 (0.018) \\
+            GMM SFS F1mean &  0.733 (0.025) & 0.746 (0.025) & 0.757 (0.026) & 0.591 (0.015) & 0.610 (0.017) & 0.622 (0.018) \\
+            GMM SFFS JM &     0.730 (0.026) & 0.748 (0.026) & 0.757 (0.027) & 0.591 (0.016) & 0.611 (0.016) & 0.622 (0.018) \\
+            GMM SFFS divKL &  0.720 (0.035) & 0.738 (0.027) & 0.718 (0.087) & 0.572 (0.036) & 0.594 (0.028) & 0.579 (0.065) \\
+            GMM SFFS kappa &  0.702 (0.079) & 0.681 (0.097) & 0.672 (0.094) & 0.559 (0.082) & 0.538 (0.087) & 0.511 (0.116) \\
+            GMM SFFS OA &     0.700 (0.080) & 0.661 (0.099) & 0.671 (0.089) & 0.558 (0.083) & 0.509 (0.100) & 0.511 (0.110) \\
+            GMM SFFS F1mean & 0.701 (0.109) & 0.691 (0.101) & 0.680 (0.103) & 0.555 (0.092) & 0.543 (0.095) & 0.524 (0.107) \\
+            GMM ridge &       0.665 (0.038) & 0.672 (0.034) & 0.692 (0.033) & 0.580 (0.017) & 0.593 (0.016) & 0.609 (0.018) \\
+            KNN &             0.611 (0.033) & 0.622 (0.030) & 0.632 (0.027) & 0.485 (0.019) & 0.502 (0.018) & 0.515 (0.019) \\
+            Random Forest &   0.699 (0.023) & 0.723 (0.020) & 0.741 (0.021) & 0.554 (0.018) & 0.578 (0.017) & 0.598 (0.019) \\
             \bottomrule
@@
   \begin{resbox}
-
+    First of all, the score difference between GMM SFS JM and GMM SFS kappa in Table IV is statistically significant.
+But, for the reason given by the reviewer, we expected that GMM SFS kappa performs better before carrying the experiment.
+Our guess is that the Gaussian hypothesis is sufficiently accurate on this data and thus JM metric has good generalization performances in this case.
   \end{resbox}
@@
   \begin{resbox}
-
+    As explained in Section VI.A, maxVarNb is the same for all methods for a given dataset.
+The number of variables used for classification is selected after the use of the selection algorithm by studying the behavior of the selection criterion in function of the number of selected variables which is a collateral output of the selection.
+We studied to select this number of variables.
+First we tried to use the number which maximize the criterion but due to some limitation (cf.
+Q3 of reviewer 1), we then considered to put a threshold on the derivative of the criterion.
+We believe that it is not unfair to use a different number of variables for each method because it also reflects the ability of a criterion to identify relevant features.
+And taking more features when the criterion shows on training data that it does not give more information should not increase the score whatever.
+    TODO:put nb of var in table 4-5
   \end{resbox}
@@
   \begin{resbox}
-
+    We added a reference as suggested.
   \end{resbox}
@@
 
-\section{Auxillary data}
+\section{Auxiliary data}
 \label{sec:auxillary}

### 2017-01-10 response.tex
update

--- 2016-12-16 response.tex
@@
 \begin{resbox}
-    In order to influence as less as possible the classification accuracy, the threshold has been set as small as possible.
-The smallest possible float number is define by the computer architecture (in our case 64bits) and C++ specifications.
-The C++ documentation states that epsilon is the \emph{difference between 1 and the least value greater than 1 that is representable}(\url{http://www.cplusplus.com/reference/cfloat/}).
-    This thresholding can be seen as a ridge regularization when applied to eigenvalues. 
+  We acknowledge for the missing information.
+Numerical instability comes from the computation of the inverse and of the log-determinant of the covariance matrix.
+When the matrix is badly conditionned, e.g.
+when it is estimated with less samples than the number of parameters to estimate, some eigenvalues might be negative or null.
+Therefore the covariance matrix is non-invertible and its determinant is zero.
+To prevent such situations, the eigenvalues are shrunk.
+
+  In practice, it consists in thresholding eigenvalues to the smallest significant positive value: $\lambda_c = \max\{EPS\_FLT,\lambda_c\}$, where $EPS\_FLT$ is defined by the computer architecture and C++ specifications.
+This thresholding can be seen as a ridge regularization with a very small value.
+
+  In the context of feature selection, it is not expected to increase the classification accuracy. It rather ensures the computation to be tractable.
+
+  The paper has been changed to make the sentence more clear: \emph{To
+    prevent  numerical  instability,   \rev{non  strictly  positive
+      eigenvalues are thresholded to a fixed value EPS\_FLT}.}
   \end{resbox}
@@
   \begin{resbox}
-    The interest of this second data cube is first to test the algorithm in a different context.
-In particular, the algorithm applied to this second dataset should be seen as a method to learn the most relevant kind of features and the best parameters to extract those features.
+    The extraction of spatial features from very high spatial resolution images is a standard practice in remote sensing.
+There are a vast literature, let us cite for instance the \emph{morphological profile} and its extension, all the possible spectral features such as \emph{NDVI} and recently all the convolutional filters.
+There is in general a significant increase of the classification accuracy.
 
-    In the first dataset, only spectral features has been considered and in the particular context of hyperspectral imaging which imply contiguous features highly correlated.
+    However, the reviewer is right, there is also an increase of the data dimensionality which can result in many statistical issues.
+Usually, the choice of features is performed by experts.
+In this work, to illustrate the potential of the feature selection, a large number of features is extracted and the optimal selection is done by the algorithm, as in ref [26], thus alleviate the \emph{curse of dimensionality}.
 
-    (reponse bof à revoir et developper mais c'est un debut)
+    To make it clear, the following part of the paper has been modified:
+
+    \emph{\rev{Conventionally,  the following  features are  extracted using  the
+      RGBIR  images in  order to  increase the  classification accuracy,
+      similarly to [26]:}\\
+    $\vdots$\\    
+    The normalized DSM and the raw  RGBIR image are added to these 191
+    features and  then stacked to create  a new image with  196 bands.
+    \rev{The  resulting  data cube  is  therefore  high-dimensional.}}
+
   \end{resbox}
@@
   \begin{resbox}
-  The value of the derivative of criterion is compared to a user-defined threshold.
-The first time the value reaches below the threshold correspond to the number of selected variables.
+    We would like to thank the reviewer for this very important comment.
+The sentence was indeed not clear.
+In order to check if the criterion stops to increase significantly, we compute the discrete derivative between two iterations and normalize it by its maximum value.
+The number of selected features corresponds to the last iteration before the value drops below $10^{-3}$ for all datasets.
+
     The derivative for the example presented on Fig.
 5 of the paper is presented in Figure~\ref{fig:deriv}.
-  With the same example, 7 features gives a kappa of $0.620$ and 18 features a kappa of $0.657$.
+For this example, 7 features gives a kappa of $0.620$ and 18 features a kappa of $0.657$.
+This figure was not included in the paper in order to keep the paper not too long.
+If the reviewer asks, it is of course possible to include it in the paper.
+
+    The paper has been updated as follow:
+
+    \rev{It  is found  by  computing the  discrete  derivative of  the
+      criteria  between  two  iterations  and normalizing  it  by  its
+      maximum value.   The number of selected  features corresponds to
+      the last  iteration before the  value drops below  $10^{-3}$ for
+      all datasets}.
 
@@
     We also considered this possibility and it is a choice available in the code of the OTB module we provide.
-But taking the maximum appears not to be the optimal choice in the sense that the algorithm selects sometimes more variables than necessary.
-Sometimes, at a given point, the best new selected variable is one which contains no information (for example one whose contribution is thresholded by EPS\_FLT) because all others decrease the criterion value.
-If it happens several times, it can lead to a final drop of the criterion.
-The algorithm then tries to counterbalance the drop and regains the original maximum but can be slightly higher.
-Figure~\ref{fig:crit-evol} illustrated this case.
-    The problem is that we introduce a few irrelevant features and we want to avoid that if possible.
-This case appears more often with the second dataset but we choose to use the same method for both dataset.
+However, selecting the maximum value of the criteria without considering the slope of the curve might be hazardous.
+At some point, adding more feature does not improve significantly the discrimination of the classes, until a too large number of variables is reached and statistical issues arise.
+It is therefore safer to stop when the increase of criteria stops to be significant.
 \end{resbox}
@@
 
-\begin{figure}[!t]
-  \centering
-  \begin{tikzpicture}
-    \begin{axis}[ymin=0.2,ymax=0.9,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
-      \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
-    \end{axis};
-  \end{tikzpicture}
-  \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Potsdam dataset with 1000 samples by class.\label{fig:crit-evol}}
-\end{figure}
+% \begin{figure}[!t]
+%   \centering
+%   \begin{tikzpicture}
+%     \begin{axis}[ymin=0.2,ymax=0.9,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
+%       \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
+%     \end{axis};
+%   \end{tikzpicture}
+%   \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Potsdam dataset with 1000 samples by class.\label{fig:crit-evol}}
+% \end{figure}
 
@@
   \begin{resbox}
-    It is true that we can try with fewer samples. We tried with 50 samples by class and it leads to the same behavior than with 250 samples (see Table~\ref{tab:aisa-otbsimu}).
+    The reviewer is right regarding the number of training samples.
+GMM (not restricted to the feature selection case) does not benefit as much as other machine learning algorithm of the increase of the number of training samples.
+For the feature selection case, it is discussed in ref [27] for instance, with KNN, RF and SVM.
+Please note it is also discussed in the paper for the Potsdam dataset.
 
-    But in the case of a larger training set is not possible to test with the chosen validation method.
+    We tried with 50 samples by class and it leads to the same behavior than with 250 samples (see Table~\ref{tab:aisa-otbsimu}).
+We therefore choose to not include this additional results in the final version of the paper.
+
+    For larger training set,  it is not possible to test with the chosen validation method for the AISA dataset.
 We use an equal number of samples in each class and 50\% of the samples are used for test.
@@
   \begin{resbox}
-    % Thank you for the careful reading.
-    NB:
+    The changes have been done. Thank you for the careful reading.
 
-    no $v_c$ after prop 2...
-
-    building corrected. Need to correct english at the end
   \end{resbox}
@@
   \begin{resbox}
-    see Table~\ref{tab:aisa-otbsimu-all}
+    We agree with the reviewer comment.
+In order to keep the paper short, we provide the additional results as a supplementary material.
+The reviewer can find the results in Table~\ref{tab:aisa-otbsimu-all} for AISA dataset and Table~\ref{tab:potsdam-otbsimu-all} for Potsdam.
+
+    The paper has been modified:
+
+    \emph{For the sake of clarity, only the three aforementioned configurations are discussed here \rev{and the results for all other configurations is available in the supplementary material}.}
   \end{resbox}
@@
 
-            GMM SFS JM &    0.685 (0.030) & 0.689 (0.030) & 0.701 (0.029) \\
+            GMM SFS JM &      0.685 (0.030) & 0.689 (0.030) & 0.693 (0.032) \\
+            GMM SFS divKL &   0.680 (0.027) & 0.699 (0.028) & 0.708 (0.031) \\
+            GMM SFS kappa &   0.682 (0.025) & 0.698 (0.027) & 0.710 (0.029) \\
+            GMM SFS OA &      0.682 (0.025) & 0.698 (0.027) & 0.710 (0.028) \\
+            GMM SFS F1mean &  0.682 (0.028) & 0.699 (0.027) & 0.710 (0.028) \\
+            GMM SFFS JM &     0.685 (0.030) & 0.689 (0.030) & 0.693 (0.032) \\
+            GMM SFFS divKL &  0.547 (0.158) & 0.615 (0.113) & 0.636 (0.120) \\
+            GMM SFFS kappa &  0.519 (0.126) & 0.537 (0.102) & 0.545 (0.141) \\
+            GMM SFFS OA &     0.531 (0.126) & 0.537 (0.093) & 0.560 (0.127) \\
+            GMM SFFS F1mean & 0.447 (0.158) & 0.579 (0.097) & 0.529 (0.129) \\
+            \bottomrule
+        \end{tabular}
+    \end{table}
+
+    \begin{table}[!t]
+        \centering
+        \caption{Average classification accuracy for 1,000 samples by class averaged over 5 trials  with all configurations of criterion and method (standard deviation in parenthesis).\label{tab:potsdam-otbsimu-all}}
+        \begin{tabular}{lccc}\toprule
+            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} \\ \cmidrule{2-4}
+            GMM SFS JM &      &&\\
             GMM SFS divKL &   &&\\
-            GMM SFS kappa & 0.678 (0.029) & 0.687 (0.029) & 0.699 (0.028) \\
+            GMM SFS kappa &   &&\\
             GMM SFS OA &      &&\\
             GMM SFS F1mean &  &&\\
-            GMM SFFS JM &   0.685 (0.030) & 0.689 (0.030) & 0.701 (0.029) \\
+            GMM SFFS JM &     &&\\
             GMM SFFS divKL &  &&\\
@@
   \begin{resbox}
-    see Table~\ref{tab:aisa-otbsimu-othereval}
+    The reviewer is right by suggesting to look at different criteria.
+Again, in order to keep the length of the paper reasonable, only the Kappa coefficient is reported in the paper.
+This choice has been motivated by previous publications that recommend this criterion, see for instance ref [34] page 105.
+
+    The other criterion are now provided in the supplementary material, and are given here Table~\ref{tab:aisa-otbsimu-othereval} and  Table~\ref{tab:potsdam-otbsimu-othereval}.
+
+    The paper has been modified:
+
+    \emph{The    classification   rate    is    presented   using    Cohen's kappa \rev{but scores computed with overall accuracy and mean of f1-score are available in supplementary material}.}
   \end{resbox}
@@
 
-            GMM SFS kappa & & & & & & \\
-            GMM SFS JM &    & & & & & \\
-            GMM SFFS JM &   & & & & & \\
-            GMM ridge &     & & & & & \\
-            KNN &           & & & & & \\
-            Random Forest & & & & & & \\
+            GMM SFS JM &      0.735 (0.027) & 0.738 (0.027) & 0.742 (0.029) & 0.591 (0.018) & 0.596 (0.018) & 0.600 (0.020) \\
+            GMM SFS divKL &   0.730 (0.024) & 0.747 (0.026) & 0.754 (0.029) & 0.587 (0.017) & 0.607 (0.017) & 0.616 (0.019) \\
+            GMM SFS kappa &   0.733 (0.023) & 0.746 (0.025) & 0.756 (0.027) & 0.591 (0.014) & 0.607 (0.015) & 0.620 (0.018) \\
+            GMM SFS OA &      0.732 (0.023) & 0.746 (0.025) & 0.756 (0.026) & 0.591 (0.014) & 0.607 (0.015) & 0.619 (0.018) \\
+            GMM SFS F1mean &  0.732 (0.026) & 0.747 (0.025) & 0.757 (0.026) & 0.590 (0.018) & 0.609 (0.016) & 0.620 (0.018) \\
+            GMM SFFS JM &     0.735 (0.027) & 0.738 (0.027) & 0.742 (0.029) & 0.591 (0.018) & 0.596 (0.018) & 0.600 (0.020) \\
+            GMM SFFS divKL &  0.611 (0.149) & 0.672 (0.105) & 0.690 (0.110) & 0.441 (0.152) & 0.515 (0.114) & 0.541 (0.116) \\
+            GMM SFFS kappa &  0.583 (0.119) & 0.599 (0.099) & 0.605 (0.141) & 0.427 (0.117) & 0.439 (0.085) & 0.442 (0.119) \\
+            GMM SFFS OA &     0.596 (0.115) & 0.600 (0.090) & 0.620 (0.127) & 0.437 (0.124) & 0.436 (0.079) & 0.455 (0.109) \\
+            GMM SFFS F1mean & 0.515 (0.154) & 0.640 (0.093) & 0.590 (0.127) & 0.352 (0.128) & 0.476 (0.085) & 0.431 (0.114) \\
+            GMM ridge &       0.665 (0.038) & 0.672 (0.034) & 0.692 (0.033) & 0.580 (0.017) & 0.593 (0.016) & 0.609 (0.018) \\
+            KNN &             0.611 (0.033) & 0.622 (0.030) & 0.632 (0.027) & 0.485 (0.019) & 0.502 (0.018) & 0.515 (0.019) \\
+            Random Forest &   0.699 (0.023) & 0.723 (0.020) & 0.741 (0.021) & 0.554 (0.018) & 0.578 (0.017) & 0.598 (0.019) \\
+            \bottomrule
+        \end{tabular}
+    \end{table}
+
+    \begin{table}[!t]
+        \centering
+        \caption{Average classification accuracy for 1,000 samples by class averaged over 5 trials  evaluated with overall accuracy and mean of F1-score (standard deviation in parenthesis).\label{tab:potsdam-otbsimu-othereval}}
+        \begin{tabular}{lcccccc}\toprule
+             & \multicolumn{2}{c}{\bfseries 5\_11 (train)} & \multicolumn{2}{c}{\bfseries 5\_12 (test)} & \multicolumn{2}{c}{\bfseries 3\_10 (test)} \\ \cmidrule{2-7}
+             & {\bfseries OA} & {\bfseries F1-mean} & {\bfseries OA} & {\bfseries F1-mean} & {\bfseries OA} & {\bfseries F1-mean} \\ \cmidrule{2-7}
+
+            GMM SFS JM &      &&&&&\\
+            GMM SFS divKL &   &&&&&\\
+            GMM SFS kappa &   &&&&&\\
+            GMM SFS OA &      &&&&&\\
+            GMM SFS F1mean &  &&&&&\\
+            GMM SFFS JM &     &&&&&\\
+            GMM SFFS divKL &  &&&&&\\
+            GMM SFFS kappa &  &&&&&\\
+            GMM SFFS OA &     &&&&&\\
+            GMM SFFS F1mean & &&&&&\\
             \bottomrule
@@
   \begin{resbox}
+    We would like to thank the reviewer for this important comment.
+First, please note that the score difference between GMM SFS JM and GMM SFS kappa in Table IV is not statistically significant according to the Wilcoxon rank-sum test.
+Second, the Kappa is estimated by cross-validation during the feature selection, which is a (little) biased estimators and therefore, it is possible to observe significant differences between the estimated kappa and the true Kappa (usually, a lower value), see ref [36] Chapter 7.
 
+    Our guess is that if the Gaussian hypothesis is a good approximation for the data, the JM metric performs well  with  performances similar to the ones obtained with the Kappa.
+It is the case for the AISA data set, while for the Potsdam data, the Kappa coefficient outperformed the JM metric.
+Our recommendation would be to start with JM since it is fast, and then try Kappa if necessary.
   \end{resbox}
@@
   \begin{resbox}
+    We apologize  for the misunderstanding,  this part was  unclear in
+    the primary  version of  the paper. The  \emph{maxVarNb} parameter
+    corresponds  to  the number  of  bands  searched by  the  iterative
+    algorithm. To select  the optimal number of  selected features, the
+    criterion  evolution is  analyzed, as  explain in  the end  of the
+    Section VI.A. The optimal number  of retained features is found by
+    computing  the discrete  derivative  of the  criterion between  two
+    iterations and normalizing it by  its maximum value. The number of
+    selected  features corresponds  to the  last iteration  before the
+    value drops below $10^{-3}$ for  all datasets. See response to the
+    Reviewer 1 (1.3 and 1.4).
 
+    Therefore, in practice, the optimal number can be different from one method to another.
+We believe that it is fair to use a different number of variables for each method because it also reflects the ability of a criterion to identify relevant features.
+
+    The Table V has been modified to include the \# of selected features.
   \end{resbox}
@@
 
-  TODO
+    We proofread the article to correct as many errors as possible. Thank you for the careful reading.
 
@@
 
-\section{Auxillary data}
-\label{sec:auxillary}
+\end{document}
 
-\end{document}
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2017-01-10 paper.tex
add data file to plot derivate of criterion
update

@@
 % \documentclass[journal,peerreview,onecolumn]{IEEEtran}
-\documentclass[journal]{IEEEtran}
+\documentclass[journal,10pt]{IEEEtran}
 \usepackage{etex}
@@
 % \hyphenation{op-tical net-works semi-conduc-tor}
+\newcommand{\rev}[1]{\textcolor{magenta}{\emph{#1}}}
 
@@
 classification,  these  problems are  related  to  the \emph{curse  of
-  dimensionality}. This  is a  major drawback  in many  remote sensing
+  dimensionality}~\cite{hughes}. This  is a  major drawback  in many  remote sensing
 application since it  is difficult to collect a large  and an accurate
@@
 when data  does not fit in  memory.  One popular open  source software
-solution  is  the Orfeo  Toolbox,  developed  by the  French  National
+solution  is  the Orfeo  Toolbox,  developed  by the  French  Space
 Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
@@
         \end{prop}
-        \noindent where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exponents on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\nu}_c$ denotes the set of samples used to compute them.
+        \noindent where $n_c$ is the number of samples of class $c$, $\nu_c$ is the number of samples of class $c$ removed from the initial set, exponents on $\boldsymbol{\Sigma}_c$ and $\boldsymbol{\mu}_c$ denotes the set of samples used to compute them.
 
@@
     the    orthonormal   matrix    of   corresponding    eigenvectors,
-    respectively.  To prevent  numerical instability,  the eigenvalues
-    are thresholded to a given  value EPS\_FLT. In our implementation,
-    EPS\_FLT is set to the floating machine precision.
+    respectively.  To prevent numerical instability, \rev{non strictly
+      positive eigenvalues are thresholded to a fixed value EPS\_FLT}.
+    In our  implementation, EPS\_FLT  is set  to the  floating machine
+    precision.
 
@@
 
-    In order to increase the dimensionality of the data, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
+    In order to increase the dimensionality of the data,
+
+    \rev{Conventionally,  the following  features are  extracted using  the
+    RGBIR  images in  order to  increase the  classification accuracy,
+    similarly to \cite{tuia2015multiclass}:}
     \begin{itemize}
@@
     The normalized DSM and the raw  RGBIR image are added to these 191
-    features and  then stacked to  create a  new image  with 196
-    bands. %  The  training is  made with  tile 5\_11  and test  with two
-    % tiles: 5\_12 which very similar with  tile 5\_11 and 3\_10 which a
-    % further area a bit less urban and more industrial.
+    features and  then stacked to create  a new image with  196 bands.
+    \rev{The  resulting  data cube  is  therefore  high-dimensional.}
 
@@
     accuracy~\cite{al:report}. For the sake of clarity, only the three
-    aforementioned configurations are discussed here.
+    aforementioned configurations are discussed here \rev{and the results for all other configurations is available in the supplementary material}.
 
@@
     the number of  retained variables is set when  the criterion stops
-    to increase significantly.  It is  found by computing the discrete
-    derivative of the criteria.  See Figure~\ref{fig:crit-evol} for an
-    example.
+    to  increase significantly.   \rev{It  is found  by computing  the
+      discrete derivative  of the criteria between  two iterations and
+      normalizing it  by its  maximum value.   The number  of selected
+      features  corresponds to  the  last iteration  before the  value
+      drops    below    $10^{-3}$     for    all    datasets}.     See
+    Figure~\ref{fig:crit-evol} for an example.
 
@@
     The    classification   rate    is    presented   using    Cohen's
-    kappa. Processing  time has been  evaluated on a  desktop computer
+    kappa \rev{but scores computed with overall accuracy and mean of f1-score are available in supplementary material}. Processing  time has been  evaluated on a  desktop computer
     with  8Gb of  RAM  and  Intel(R) Core(TM)  i5-3570  CPU @  3.40GHz
@@
 
-    \begin{table}[!t]
+    \begin{table*}[!t]
         \centering
         \caption{Mean processing time for training and classification for results in Table \ref{tab:aisa-otbsimu}.\label{tab:aisa-otbsimu-time}}
-        \begin{tabularx}{\columnwidth}{l*{6}{>{\centering\arraybackslash}X}}
+        \begin{tabularx}{0.8\textwidth}{l*{9}{>{\centering\arraybackslash}X}}
             \toprule
-             & \multicolumn{3}{c}{\bfseries Training time (s)} & \multicolumn{3}{c}{\bfseries Classification time (s)} \\ \cmidrule{2-7}
-            \# samples by class & 250 & 500 & 1000 & 250 & 500 & 1000 \\ \midrule
-
-            GMM SFS kappa & 257             & 496             & 955             & {\bfseries 5.2} & {\bfseries 5.2} & {\bfseries 5.5} \\
-            GMM SFS JM &    {\bfseries 8.6} & {\bfseries 8.9} & {\bfseries 9.1} & {\bfseries 5.7} & {\bfseries 5.7} & {\bfseries 5.9} \\
-            GMM SFFS JM &   {\bfseries 8.8} & {\bfseries 9.0} & {\bfseries 9.3} & {\bfseries 5.0} & {\bfseries 5.0} & {\bfseries 5.4} \\
-            GMM ridge &     71.7            & 105             & 167             & 530 & 530 & 530 \\
-            KNN &           {\bfseries 8.9} & 19.6            & 59.7            & 387 & 639 & 887 \\
-            Random Forest & 24.5            & 49.3            & 105             & 33.0 & 41.7 & 45.9 \\
+             & \multicolumn{3}{c}{\bfseries Training time (s)} & \multicolumn{3}{c}{\bfseries Classification time (s)} & \multicolumn{3}{c}{\bfseries \# of selected features} \\ \cmidrule{2-10}
+            \# samples by class & 250 & 500 & 1000 & 250 & 500 & 1000 & 250 & 500 & 1000 \\ \midrule
+
+            GMM SFS kappa & 257             & 496             & 955             & {\bfseries 5.2} & {\bfseries 5.2} & {\bfseries 5.5} & 11.95 & 12    & 12.05\\
+            GMM SFS JM &    {\bfseries 8.6} & {\bfseries 8.9} & {\bfseries 9.1} & {\bfseries 5.7} & {\bfseries 5.7} & {\bfseries 5.9} & 11.95 & 12    & 12.05\\
+            GMM SFFS JM &   {\bfseries 8.8} & {\bfseries 9.0} & {\bfseries 9.3} & {\bfseries 5.0} & {\bfseries 5.0} & {\bfseries 5.4} & 21.45 & 24.35 & 27.05\\
+            GMM ridge &     71.7            & 105             & 167             & 530 & 530 & 530    & & & \\
+            KNN &           {\bfseries 8.9} & 19.6            & 59.7            & 387 & 639 & 887    & & & \\
+            Random Forest & 24.5            & 49.3            & 105             & 33.0 & 41.7 & 45.9 & & & \\
             \bottomrule
         \end{tabularx}
-    \end{table}
+    \end{table*}
 
@@
     The difference with the second best classifier, Random Forest, appears to be significant when using 250 and 500 samples.
-Random Forest has similar performance in term of classification rate with 1000 samples and one could expect to get a better classification rate with RF if more samples were available.
+Random Forest has similar performance in term of classification rate with 1000 samples \rev{and one could expect to get a better classification rate with RF if more samples were available}.
 The GMM classifier with ridge regularization and the KNN classifier are both outperformed.

### 2017-01-11 response.tex
update

@@
     The reviewer is right regarding the number of training samples.
-GMM (not restricted to the feature selection case) does not benefit as much as other machine learning algorithm of the increase of the number of training samples.
+GMM (not restricted to the feature selection case) does not benefit as much as other machine learning algorithms of the increase of the number of training samples.
 For the feature selection case, it is discussed in ref [27] for instance, with KNN, RF and SVM.
@@
 First, please note that the score difference between GMM SFS JM and GMM SFS kappa in Table IV is not statistically significant according to the Wilcoxon rank-sum test.
-Second, the Kappa is estimated by cross-validation during the feature selection, which is a (little) biased estimators and therefore, it is possible to observe significant differences between the estimated kappa and the true Kappa (usually, a lower value), see ref [36] Chapter 7.
+Second, the Kappa is estimated by cross-validation during the feature selection, which is a  biased estimators and therefore, it is possible to observe significant differences between the estimated kappa and the true Kappa (usually, a lower value), see ref [36] Chapter 7.
 
-    Our guess is that if the Gaussian hypothesis is a good approximation for the data, the JM metric performs well  with  performances similar to the ones obtained with the Kappa.
+    Our guess is that if the Gaussian hypothesis is a good approximation for the used data, the JM metric performs well  with  performances similar to the ones obtained with the Kappa.
 It is the case for the AISA data set, while for the Potsdam data, the Kappa coefficient outperformed the JM metric.
@@
   \begin{resbox}
-    We added a reference as suggested.
+    We added the following reference: \emph{D. Donoho, ``High-dimensional data analysis: the curses and blessing of dimensionality,'' in AMS Mathematical challenges of the 21st century, 2000}.
   \end{resbox}
@@
   \begin{resbox}
-    We totally agree that some mathematical formulations we used are already known.
+    We totally agree that some mathematical formulations are already known.
 Nevertheless, we think that these formulations in this particular context has not been exploited so far.

### 2017-01-12 response.tex
Réponse au reviewer III
- Modif des réponses au Rev III
- Modif de l'avant derniere paragaph sur les contributions

@@
   \begin{resbox}
-    We totally agree that some mathematical formulations are already known.
-Nevertheless, we think that these formulations in this particular context has not been exploited so far.
+    We totally agree that some mathematical formulations are conventional.
+We believe that it is necessary to provide to readers sufficient information to explain clearly the contribution of the proposed paper.
+The contributions are based on specific derivation of the GMM and the FE formulation.
+Without these primary recall, we are afraid that readers can be lost.
+Concerning the contribution of the paper, they are stated in the antepenultimate paragraph of the introduction.
+They concern the expression of efficient updates rules using block matrix processing.
+Furthermore, an open source C++ library is made available for the community.
 The results of our experiment shows that the proposed algorithm has the advantage to be very efficient in term of processing time with a fairly reasonable cost in term of classification accuracy.
 It has to be noticed that compared to similar methods using Gaussian Mixture model, no additional hypothesis are made on the covariance matrix which is tractable only because of the efficient implementation.
+
+    To make the contribution more clear, the antepenultimate paragraph has been modified to
+
+\rev{The contribution of the paper is a extension of the forward feature selection method proposed in~\cite{fauvel2015fast}.
+An smart implementation of the feature selection updates rules are presented in order to perform efficiently on large amount of data.
+They are based on linear algebra on block matrix applied to the covariance matrix of the conditional class density function.
+Furthermore, a floating version of the algorithm is proposed and evaluated.
+Several \emph{goodness of fit} criteria are proposed to handle unbalanced training set, extending the conventional \emph{overall accuracy} measure.
+Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~[13].}
   \end{resbox}
@@
   \begin{resbox}
-    After reading with interest this two references, we fail to see the link between these publications and our submission except for the type of data used to test algorithms.
-
+    We thank the reviewer for these two interesting references.
+However, after reading them with interest, we think the issues address in these papers are somewhat unrelated to the feature selection proposed in this work.
 First reference reviews ways to assign several labels to one sample and so performs spectral unmixing.
@@
 Although very interesting, we do not address the subject of multi-labeling in our work.
-
-    According to these observations, we choose to respectfully disagree with the reviewer and take the decision not to add a digression to cite these references.
+According to these observations and to keep the paper short, we choose to not include these references in the revised paper.
   \end{resbox}
@@
   \begin{resbox}
-
-    We proofread the article to correct as many errors as possible. Thank you for the careful reading.
+    We proofread the article to correct as many errors as possible. If the reviewer still see weird sentences or typo, we will happy pleased to modify or correct them.
 

### 2017-01-12 paper.tex
update
Réponse au reviewer III
- Modif des réponses au Rev III
- Modif de l'avant derniere paragaph sur les contributions

@@
 classification,  these  problems are  related  to  the \emph{curse  of
-  dimensionality}~\cite{hughes}. This  is a  major drawback  in many  remote sensing
-application since it  is difficult to collect a large  and an accurate
-ground-truth.   An intensive  work has  been performed  in the  remote
-sensing community  to build accurate classifiers  for high dimensional
-images.  Bayesian models~\cite{book:landgrebe}, feature extraction and
-feature   reduction   techniques~\cite{book:landgrebe,DR:guided:tour},
-random   forest~\cite{1396322},  neural   networks~\cite{5411821}  and
-kernel methods~\cite{kernel:methods:rs} have been investigated for the
+  dimensionality}~\cite{donoho}.  This is  a  major  drawback in  many
+remote sensing  application since it  is difficult to collect  a large
+and an accurate ground-truth.  An intensive work has been performed in
+the remote  sensing community to  build accurate classifiers  for high
+dimensional  images.   Bayesian models~\cite{book:landgrebe},  feature
+extraction              and             feature              reduction
+techniques~\cite{book:landgrebe,DR:guided:tour},                random
+forest~\cite{1396322},   neural  networks~\cite{5411821}   and  kernel
+methods~\cite{kernel:methods:rs}  have   been  investigated   for  the
 classification of such images.
@@
 
-The contribution of the paper is a forward feature selection method using GMM in continuation of~\cite{fauvel2015fast}.
-An efficient implementation of the method, using linear algebra on block matrix, is presented in order to handle large amount of data.
-Furthermore, a floating version of the algorithm is proposed.
-Several criteria are proposed to handle unbalanced training set.
-Finally, the developed algorithm is made available as a remote module of the Orfeo Toolbox~\cite{christophe2008orfeo}.
+\rev{The contribution of the paper is a extension of the forward feature selection method proposed in~\cite{fauvel2015fast}.
+An smart implementation of the feature selection updates rules are presented in order to perform efficiently on large amount of data.
+They are based on linear algebra on block matrix applied to the covariance matrix of the conditional class density function.
+Furthermore, a floating version of the algorithm is proposed and evaluated.
+Several \emph{goodness of fit} criteria are proposed to handle unbalanced training set, extending the conventional \emph{overall accuracy} measure.
+Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~\cite{christophe2008orfeo}.}
 

### 2017-01-12 response.tex
improve english

@@
 
-            GMM SFS JM &      0.685 (0.030) & 0.689 (0.030) & 0.693 (0.032) \\
-            GMM SFS divKL &   0.680 (0.027) & 0.699 (0.028) & 0.708 (0.031) \\
-            GMM SFS kappa &   0.682 (0.025) & 0.698 (0.027) & 0.710 (0.029) \\
-            GMM SFS OA &      0.682 (0.025) & 0.698 (0.027) & 0.710 (0.028) \\
-            GMM SFS F1mean &  0.682 (0.028) & 0.699 (0.027) & 0.710 (0.028) \\
-            GMM SFFS JM &     0.685 (0.030) & 0.689 (0.030) & 0.693 (0.032) \\
+            GMM SFS JM &      {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.693 (0.032)} \\
+            GMM SFS divKL &   {\bfseries 0.680 (0.027)} & {\bfseries 0.699 (0.028)} & {\bfseries 0.708 (0.031)} \\
+            GMM SFS kappa &   {\bfseries 0.682 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.710 (0.029)} \\
+            GMM SFS OA &      {\bfseries 0.682 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.710 (0.028)} \\
+            GMM SFS F1mean &  {\bfseries 0.682 (0.028)} & {\bfseries 0.699 (0.027)} & {\bfseries 0.710 (0.028)} \\
+            GMM SFFS JM &     {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.693 (0.032)} \\
             GMM SFFS divKL &  0.547 (0.158) & 0.615 (0.113) & 0.636 (0.120) \\
@@
             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} \\ \cmidrule{2-4}
-            GMM SFS JM &      &&\\
-            GMM SFS divKL &   &&\\
-            GMM SFS kappa &   &&\\
-            GMM SFS OA &      &&\\
-            GMM SFS F1mean &  &&\\
-            GMM SFFS JM &     &&\\
-            GMM SFFS divKL &  &&\\
-            GMM SFFS kappa &  &&\\
-            GMM SFFS OA &     &&\\
-            GMM SFFS F1mean & &&\\
+            GMM SFS JM &      0.624 (0.028) & & \\
+            GMM SFS divKL &   0.552 (0.092) & & \\
+            GMM SFS kappa &   {\bfseries 0.694 (0.002)} & & \\
+            GMM SFS OA &      {\bfseries 0.696 (0.003)} & & \\
+            GMM SFS F1mean &  {\bfseries 0.689 (0.002)} & & \\
+            GMM SFFS JM &     0.624 (0.028) & & \\
+            GMM SFFS divKL &  0.161 (0.066) & & \\
+            GMM SFFS kappa &  0.474 (0.103) & & \\
+            GMM SFFS OA &     0.407 (0.114) & & \\
+            GMM SFFS F1mean & 0.319 (0.052) & & \\
             \bottomrule
@@
 
-            GMM SFS JM &      0.735 (0.027) & 0.738 (0.027) & 0.742 (0.029) & 0.591 (0.018) & 0.596 (0.018) & 0.600 (0.020) \\
-            GMM SFS divKL &   0.730 (0.024) & 0.747 (0.026) & 0.754 (0.029) & 0.587 (0.017) & 0.607 (0.017) & 0.616 (0.019) \\
-            GMM SFS kappa &   0.733 (0.023) & 0.746 (0.025) & 0.756 (0.027) & 0.591 (0.014) & 0.607 (0.015) & 0.620 (0.018) \\
-            GMM SFS OA &      0.732 (0.023) & 0.746 (0.025) & 0.756 (0.026) & 0.591 (0.014) & 0.607 (0.015) & 0.619 (0.018) \\
-            GMM SFS F1mean &  0.732 (0.026) & 0.747 (0.025) & 0.757 (0.026) & 0.590 (0.018) & 0.609 (0.016) & 0.620 (0.018) \\
-            GMM SFFS JM &     0.735 (0.027) & 0.738 (0.027) & 0.742 (0.029) & 0.591 (0.018) & 0.596 (0.018) & 0.600 (0.020) \\
+            GMM SFS JM &      {\bfseries 0.735 (0.027)} & {\bfseries 0.738 (0.027)} & {\bfseries 0.742 (0.029)} & {\bfseries 0.591 (0.018)} & {\bfseries 0.596 (0.018)} & {\bfseries 0.600 (0.020)} \\
+            GMM SFS divKL &   {\bfseries 0.730 (0.024)} & {\bfseries 0.747 (0.026)} & {\bfseries 0.754 (0.029)} & {\bfseries 0.587 (0.017)} & {\bfseries 0.607 (0.017)} & {\bfseries 0.616 (0.019)} \\
+            GMM SFS kappa &   {\bfseries 0.733 (0.023)} & {\bfseries 0.746 (0.025)} & {\bfseries 0.756 (0.027)} & {\bfseries 0.591 (0.014)} & {\bfseries 0.607 (0.015)} & {\bfseries 0.620 (0.018)} \\
+            GMM SFS OA &      {\bfseries 0.732 (0.023)} & {\bfseries 0.746 (0.025)} & {\bfseries 0.756 (0.026)} & {\bfseries 0.591 (0.014)} & {\bfseries 0.607 (0.015)} & {\bfseries 0.619 (0.018)} \\
+            GMM SFS F1mean &  {\bfseries 0.732 (0.026)} & {\bfseries 0.747 (0.025)} & {\bfseries 0.757 (0.026)} & {\bfseries 0.590 (0.018)} & {\bfseries 0.609 (0.016)} & {\bfseries 0.620 (0.018)} \\
+            GMM SFFS JM &     {\bfseries 0.735 (0.027)} & {\bfseries 0.738 (0.027)} & {\bfseries 0.742 (0.029)} & {\bfseries 0.591 (0.018)} & {\bfseries 0.596 (0.018)} & {\bfseries 0.600 (0.020)} \\
             GMM SFFS divKL &  0.611 (0.149) & 0.672 (0.105) & 0.690 (0.110) & 0.441 (0.152) & 0.515 (0.114) & 0.541 (0.116) \\
@@
             GMM SFFS F1mean & 0.515 (0.154) & 0.640 (0.093) & 0.590 (0.127) & 0.352 (0.128) & 0.476 (0.085) & 0.431 (0.114) \\
-            GMM ridge &       0.665 (0.038) & 0.672 (0.034) & 0.692 (0.033) & 0.580 (0.017) & 0.593 (0.016) & 0.609 (0.018) \\
+            GMM ridge &       0.665 (0.038) & 0.672 (0.034) & 0.692 (0.033) & {\bfseries 0.580 (0.017)} & {\bfseries 0.593 (0.016)} & {\bfseries 0.609 (0.018)} \\
             KNN &             0.611 (0.033) & 0.622 (0.030) & 0.632 (0.027) & 0.485 (0.019) & 0.502 (0.018) & 0.515 (0.019) \\
-            Random Forest &   0.699 (0.023) & 0.723 (0.020) & 0.741 (0.021) & 0.554 (0.018) & 0.578 (0.017) & 0.598 (0.019) \\
+            Random Forest &   0.699 (0.023) & 0.723 (0.020) & {\bfseries 0.741 (0.021)} & 0.554 (0.018) & 0.578 (0.017) & {\bfseries 0.598 (0.019)} \\
             \bottomrule
@@
 
-            GMM SFS JM &      &&&&&\\
-            GMM SFS divKL &   &&&&&\\
-            GMM SFS kappa &   &&&&&\\
-            GMM SFS OA &      &&&&&\\
-            GMM SFS F1mean &  &&&&&\\
-            GMM SFFS JM &     &&&&&\\
-            GMM SFFS divKL &  &&&&&\\
-            GMM SFFS kappa &  &&&&&\\
-            GMM SFFS OA &     &&&&&\\
-            GMM SFFS F1mean & &&&&&\\
+            GMM SFS JM &      0.725 (0.025) & 0.528 (0.022) &  &&&\\
+            GMM SFS divKL &   0.677 (0.065) & 0.503 (0.044) &  &&&\\
+            GMM SFS kappa &   0.771 (0.002) & 0.642 (0.005) &  &&&\\
+            GMM SFS OA &      0.773 (0.002) & 0.642 (0.006) &  &&&\\
+            GMM SFS F1mean &  0.767 (0.002) & 0.646 (0.004) &  &&&\\
+            GMM SFFS JM &     0.725 (0.025) & 0.528 (0.022) &  &&&\\
+            GMM SFFS divKL &  0.408 (0.037) & 0.237 (0.051) &  &&&\\
+            GMM SFFS kappa &  0.601 (0.087) & 0.431 (0.065) &  &&&\\
+            GMM SFFS OA &     0.544 (0.097) & 0.392 (0.063) &  &&&\\
+            GMM SFFS F1mean & 0.473 (0.043) & 0.364 (0.027) &  &&&\\
             \bottomrule
@@
 
-\rev{The contribution of the paper is a extension of the forward feature selection method proposed in~\cite{fauvel2015fast}.
+\rev{The contribution of the paper is a extension of the forward feature selection method proposed in~[27].
 An smart implementation of the feature selection updates rules are presented in order to perform efficiently on large amount of data.

### 2017-01-12 paper.tex
improve english

@@
   dimension remote sensing images. The  results show that the approach
-  provide good classification accuracies with low computation time.
+  provides good classification accuracies with low computation time.
 \end{abstract}
@@
 For instance, the European satellites Sentinel-2 were launched recently and 2 Terabytes of data will be released every day~\cite{drusch2012sentinel}.
-The Landsat open archive (\url{http://landsat.usgs.gov/products_data_at_no_charge.php}) has released thousands of images.
-Such high volume of Earth observation data provides an accurate information of land state and functioning, and helps to improve the understanding of the planet~\cite{rs1010001}.
+The Landsat open archive (\url{http://landsat.usgs.gov/products_data_at_no_charge.php}) has also released thousands of images.
+Such high volume of Earth observation data provides accurate information of land state and functions, and helps to improve the understanding of the planet~\cite{rs1010001}.
 However, processing such data is more and more challenging because of statistical and computational issues.
@@
 given  year  is approximately  one  thousand.   When working  in  high
-dimensional  space,  statistical  methods  made for  low  or  moderate
-dimensional  space do  not  adapt  well.  For  instance,  the rate  of
+dimensional  spaces,  statistical  methods  made for  low  or  moderate
+dimensional  spaces do  not  adapt  well.  For  instance,  the rate  of
 convergence of the statistical estimation decreases when the dimension
-grows while conjointly the number of parameters to estimate increases,
+grows while jointly the number of parameters to estimate increases,
 making    the    estimation    of   the    model    parameters    very
@@
   dimensionality}~\cite{donoho}.  This is  a  major  drawback in  many
-remote sensing  application since it  is difficult to collect  a large
-and an accurate ground-truth.  An intensive work has been performed in
+remote sensing  applications since it  is difficult to collect  a large
+and accurate ground-truth.  An intensive work has been performed in
 the remote  sensing community to  build accurate classifiers  for high
@@
 community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
-related to the use of parallel approach (multi-core, GPU, clusters) to
+related to the use of parallel approaches (multi-core, GPU, clusters) to
 improve the  processing time, and  to the use of  streaming techniques
@@
 Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
-are  conveniently proposed  to the  users/developpers through  several
+are  conveniently proposed  to  users/developpers through  several
 modules ``\emph{ready to use}''.
@@
   dimensionality}~\cite{bouveyron2014model}:  high   dimensional  data
-space   exhibits  interesting   properties   for  the   classification
+spaces   exhibit  interesting   properties   for    classification
 purpose.  In  particular,  it  is   possible  to  get  a  parsimonious
@@
 statistical complexity  and computational load. In  short, by reducing
-the dimension better, classification results is expected with a reduced
+the dimension, better classification results are expected with a reduced
 computational load.

### 2017-01-12 paper.tex
improve english

@@
 \IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available Earth observation data for a given landscape becomes larger and larger.
-Satellite missions produce a huge amount of data on a regularly (daily) basis.
+Satellite missions produce a huge amount of data on a regular (daily) basis.
 From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites managed by the German space agency will produce images with 244 spectral bands, a spatial resolution of 30x30m per pixel and with a frequency of revisit of 4 days\cite{Müller09enmap}.
@@
 Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
-are  conveniently proposed  to  users/developpers through  several
-modules ``\emph{ready to use}''.
+are  conveniently proposed  to  users/developpers through  several ``\emph{ready to use}'' modules.
 
@@
 There  are  two  main  strategies  to  reduce  dimension~\cite[Chapter
-1]{Guyon:2006:FEF:1208773}: features extraction and feature selection.
+1]{Guyon:2006:FEF:1208773}: feature extraction and feature selection.
 Feature extraction means reformulate  and summarize the information by
@@
 boundary feature  extraction~\cite{book:landgrebe}.  To  the contrary,
-features selection  extracts a subset of  existing features identified
+feature selection  extracts a subset of  existing features identified
 as  the most  relevant  by a  given criterion.   This  subset has  the
@@
 
-Features selection/extraction algorithms can be divided into three classes.
+Feature selection/extraction algorithms can be divided into three classes.
 The first class is called \emph{filter methods}.
@@
 Since exhaustive searches are too expensive in terms of processing time, several sub-optimal search strategies have been designed, mainly iterative forward or backward search~\cite{whitney1971direct,marill1963effectiveness} or a combination of both~\cite{somol1999adaptive}.
-The advantage of such methods compare to filter methods is that they are dedicated to a particular model and to a particular learning problem.
+The advantage of such methods compared to filter methods is that they are dedicated to a particular model and to a particular learning problem.
 On the other hand, as they require the training of multiple models to test various set of variables, they are more time consuming.
@@
 The third class corresponds to the \emph{embedded methods}.
-They do not separate the features selection process from the learning algorithm and allow interaction between the two processes.
+They do not separate the feature selection process from the learning algorithm and allow interactions between the two processes.
 A popular embedded method is the \emph{Random Forest}.
@@
 The training computational load is too important to compensate the reduced prediction computational load.
-Hence, feature selection is not widely used in operational situation.
+Hence, feature selection is not widely used in operational situations.
 However, methods based on Gaussian Mixture Models (GMM) have several interesting properties that make them suitable for feature selection in the context of large amount of data.
@@
 
-\rev{The contribution of the paper is a extension of the forward feature selection method proposed in~\cite{fauvel2015fast}.
-An smart implementation of the feature selection updates rules are presented in order to perform efficiently on large amount of data.
-They are based on linear algebra on block matrix applied to the covariance matrix of the conditional class density function.
+\rev{The contribution of this paper is a extension of the forward feature selection method proposed in~\cite{fauvel2015fast}.
+A smart implementation of the feature selection update rules are presented in order to perform efficiently on large amount of data.
+The rules use on linear algebra on block matrices applied to the covariance matrix of the conditional class density function.
 Furthermore, a floating version of the algorithm is proposed and evaluated.
-Several \emph{goodness of fit} criteria are proposed to handle unbalanced training set, extending the conventional \emph{overall accuracy} measure.
+Several \emph{correctness of fit} criteria are proposed to handle unbalanced training sets, extending the conventional \emph{overall accuracy} measure.
 Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~\cite{christophe2008orfeo}.}
@@
 follows.   Section~\ref{sec:gmm-hd}  presents   GMM  classifiers   and
-problems  related to  high-dimensional  feature  spaces. The  features
-selection methods are detailed in Section~\ref{sec:selection}. Then an
+problems  related to  high-dimensional  feature  spaces. The  feature
+selection methods are detailed in Section~\ref{sec:selection}. Then, an
 efficient         implementation        is         presented        in
 Section~\ref{sec:implementation}.  Experimental  results on  two  real
-high dimensional data-set are given Section~\ref{sec:test}. Conclusion
+high dimensional data-sets are given Section~\ref{sec:test}. Conclusion
 and perspectives conclude the paper in Section~\ref{sec:conclusion}.
@@
     (convex     combination)     of      several     class     conditioned
-    distribution~\cite{Fraley00model-basedclustering}:
+    distributions~\cite{Fraley00model-basedclustering}:
     \begin{align}
@@
     where $\pi_c$ is  the prior, \emph{i.e.}, the  proportion of class
-    $c$  and  $f_c$  a  parametric density  function  control by
+    $c$  and  $f_c$  a  parametric density  function  controlled by
     $\theta$.
 
-    Among the possible parametric model,  the Gaussian one is the most
+    Among the possible parametric models,  the Gaussian one is the most
     used~\cite{bouveyron2014model}.   It assumes  that each  $f_c$ is,
@@
     There are two major solutions to this problem.
-First option is to stabilize the inversion of the covariance matrices.
+The first option is to stabilize the inversion of the covariance matrices.
 Some methods investigate the use of constraints on the direct problem.
@@
 Celeux and Govaert \cite{celeux1995gaussian} suggest to use equality constraints between coefficients of the covariance matrix in a parsimonious cluster-based GMM framework.
-Others papers propose to work on the inverse problem.
+Other papers propose to work on the inverse problem.
 A classical method is to use a regularization method as the well-known ridge regularization \cite{hoerl1970ridge}.
@@
 
-    Second option is to reduce the dimension.
-Features extraction/selection methods have been developed in order to reduce the dimension with various approaches described in Section~\ref{sec:intro}.
+    The second option is to reduce the dimension.
+Feature extraction/selection methods have been developed in order to reduce the dimension with various approaches described in Section~\ref{sec:intro}.
 In this study, this latter option is explored and a feature selection method named sequential forward features selection is presented.
@@
 \begin{enumerate}
-\item A function that ranks the features according to some good classification or class separability criteria,
+\item A function that ranks the features according to some good classification or class separability criterion,
 \item A search strategy to optimize the function.
@@
 
-Section~\ref{sec:criterion} describes the different criterion used in this work and two search strategies are discussed in~\ref{sec:selection:method}.
+Section~\ref{sec:criterion} describes the various criteria used in this work and two search strategies are discussed in~\ref{sec:selection:method}.
 
@@
     The criterion evaluates how a given model built with a subset of features performs for the classification task.
-It can be an estimation of the correct classification or a measure of separability/similarity between class distribution.
+It can be an estimation of the correct classification or a measure of separability/similarity between class distributions.
 The former are in general more demanding in terms of processing time than the later.
@@
         \end{itemize}
-        High  value  of  theses  indices  correspond  to  an  accurate
+        High  values  of  theses  indices  correspond  to  an  accurate
         classification.
@@
         \subsubsection{Similarity between distributions}
-        The similarity  between two distributions is  quantified using
-        divergence  measure~\cite{opac-b1097517}. Contrary  to measure
+        The similarity  between two distributions can be  quantified using
+        divergence  measures~\cite{opac-b1097517}. Contrary  to measures
         of correct  classification, divergences are  computed directly
@@
         case of Gaussian models. The  simplification allows to get rid
-        of  any integration  calculus which  is a  major problem  when
+        of  any integration  calculations which  is a  major problem  when
         dealing with high-dimensional data.
@@
 criterion  for  all   the  remaining  variables  (see   lines  5-7  in
-Algorithm~\ref{alg:sfs}). Calculus are based on linear algebra, and the
+Algorithm~\ref{alg:sfs}). Calculations are based on linear algebra, and the
 numerical  complexity is  on average  $O(d^3)$.  Furthermore,  for the
@@
 
-    To deal with this limitation, the choice has been made to perform an eigendecomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}_c$:
+    To deal with this limitation, the choice has been made to perform an eigenvalues decomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}_c$:
 

### 2017-01-12 response.tex
improve english

@@
     To make the contribution more clear, the antepenultimate paragraph has been modified to
-
-\rev{The contribution of the paper is a extension of the forward feature selection method proposed in~[27].
-An smart implementation of the feature selection updates rules are presented in order to perform efficiently on large amount of data.
-They are based on linear algebra on block matrix applied to the covariance matrix of the conditional class density function.
+\rev{The contribution of this paper is a extension of the forward feature selection method proposed in~[27].
+A smart implementation of the feature selection update rules are presented in order to perform efficiently on large amount of data.
+The rules use on linear algebra on block matrices applied to the covariance matrix of the conditional class density function.
 Furthermore, a floating version of the algorithm is proposed and evaluated.
-Several \emph{goodness of fit} criteria are proposed to handle unbalanced training set, extending the conventional \emph{overall accuracy} measure.
-Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~[13].}
+Several \emph{correctness of fit} criteria are proposed to handle unbalanced training sets, extending the conventional \emph{overall accuracy} measure.
+Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~[13].
   \end{resbox}

### 2017-01-12 response.tex
typo

@@
     To make the contribution more clear, the antepenultimate paragraph has been modified to
-\rev{The contribution of this paper is a extension of the forward feature selection method proposed in~[27].
+\emph{\rev{The contribution of this paper is a extension of the forward feature selection method proposed in~[27].
 A smart implementation of the feature selection update rules are presented in order to perform efficiently on large amount of data.
@@
 Several \emph{correctness of fit} criteria are proposed to handle unbalanced training sets, extending the conventional \emph{overall accuracy} measure.
-Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~[13].
+Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~[13].}}
   \end{resbox}

### 2017-01-13 response.tex
update table + add supplementary material

@@
             & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} \\ \cmidrule{2-4}
-            GMM SFS JM &      0.624 (0.028) & & \\
-            GMM SFS divKL &   0.552 (0.092) & & \\
-            GMM SFS kappa &   {\bfseries 0.694 (0.002)} & & \\
-            GMM SFS OA &      {\bfseries 0.696 (0.003)} & & \\
-            GMM SFS F1mean &  {\bfseries 0.689 (0.002)} & & \\
-            GMM SFFS JM &     0.624 (0.028) & & \\
-            GMM SFFS divKL &  0.161 (0.066) & & \\
-            GMM SFFS kappa &  0.474 (0.103) & & \\
-            GMM SFFS OA &     0.407 (0.114) & & \\
-            GMM SFFS F1mean & 0.319 (0.052) & & \\
+            GMM SFS JM &      0.624 (0.028) & 0.631 (0.034) & 0.461 (0.027) \\
+            GMM SFS divKL &   0.552 (0.092) & 0.557 (0.086) & 0.388 (0.084) \\
+            GMM SFS kappa &   {\bfseries 0.694 (0.002)} & {\bfseries 0.669 (0.005)} & {\bfseries 0.533 (0.008)} \\
+            GMM SFS OA &      {\bfseries 0.696 (0.003)} & {\bfseries 0.670 (0.005)} & {\bfseries 0.535 (0.011)} \\
+            GMM SFS F1mean &  {\bfseries 0.689 (0.002)} & {\bfseries 0.668 (0.006)} & {\bfseries 0.529 (0.005)} \\
+            GMM SFFS JM &     0.624 (0.028) & 0.631 (0.034) & 0.461 (0.027) \\
+            GMM SFFS divKL &  0.161 (0.066) & 0.142 (0.065) & 0.148 (0.058) \\
+            GMM SFFS kappa &  0.474 (0.103) & 0.499 (0.144) & 0.330 (0.099) \\
+            GMM SFFS OA &     0.407 (0.114) & 0.424 (0.148) & 0.335 (0.099) \\
+            GMM SFFS F1mean & 0.319 (0.052) & 0.316 (0.067) & 0.316 (0.034) \\
             \bottomrule
@@
 
-            GMM SFS JM &      0.725 (0.025) & 0.528 (0.022) &  &&&\\
-            GMM SFS divKL &   0.677 (0.065) & 0.503 (0.044) &  &&&\\
-            GMM SFS kappa &   0.771 (0.002) & 0.642 (0.005) &  &&&\\
-            GMM SFS OA &      0.773 (0.002) & 0.642 (0.006) &  &&&\\
-            GMM SFS F1mean &  0.767 (0.002) & 0.646 (0.004) &  &&&\\
-            GMM SFFS JM &     0.725 (0.025) & 0.528 (0.022) &  &&&\\
-            GMM SFFS divKL &  0.408 (0.037) & 0.237 (0.051) &  &&&\\
-            GMM SFFS kappa &  0.601 (0.087) & 0.431 (0.065) &  &&&\\
-            GMM SFFS OA &     0.544 (0.097) & 0.392 (0.063) &  &&&\\
-            GMM SFFS F1mean & 0.473 (0.043) & 0.364 (0.027) &  &&&\\
+            GMM SFS JM &      0.725 (0.025) & 0.528 (0.022) & 0.750 (0.029) & 0.514 (0.023) & 0.585 (0.026) & 0.467 (0.020) \\
+            GMM SFS divKL &   0.677 (0.065) & 0.503 (0.044) & 0.702 (0.063) & 0.488 (0.036) & 0.521 (0.071) & 0.414 (0.072) \\
+            GMM SFS kappa &   {\bfseries 0.771 (0.002)} & {\bfseries 0.642 (0.005)} & {\bfseries 0.769 (0.004)} & {\bfseries 0.602 (0.004)} & {\bfseries 0.637 (0.007)} & {\bfseries 0.562 (0.010)} \\
+            GMM SFS OA &      {\bfseries 0.773 (0.002)} & {\bfseries 0.642 (0.006)} & {\bfseries 0.771 (0.004)} & {\bfseries 0.602 (0.004)} & {\bfseries 0.639 (0.009)} & {\bfseries 0.568 (0.016)} \\
+            GMM SFS F1mean &  {\bfseries 0.767 (0.002)} & {\bfseries 0.646 (0.004)} & {\bfseries 0.768 (0.005)} & {\bfseries 0.605 (0.003)} & {\bfseries 0.634 (0.005)} & {\bfseries 0.569 (0.008)} \\
+            GMM SFFS JM &     0.725 (0.025) & 0.528 (0.022) & 0.750 (0.029) & 0.514 (0.023) & 0.585 (0.026) & 0.467 (0.020) \\
+            GMM SFFS divKL &  0.408 (0.037) & 0.237 (0.051) & 0.466 (0.020) & 0.241 (0.045) & 0.292 (0.072) & 0.195 (0.048) \\
+            GMM SFFS kappa &  0.601 (0.087) & 0.431 (0.065) & 0.641 (0.125) & 0.439 (0.069) & 0.467 (0.081) & 0.373 (0.071) \\
+            GMM SFFS OA &     0.544 (0.097) & 0.392 (0.063) & 0.579 (0.127) & 0.396 (0.064) & 0.473 (0.087) & 0.369 (0.066) \\
+            GMM SFFS F1mean & 0.473 (0.043) & 0.364 (0.027) & 0.491 (0.061) & 0.361 (0.025) & 0.462 (0.028) & 0.355 (0.023) \\
             \bottomrule

### 2017-01-13 paper.tex
Mathieu's corrections

@@
 
-\author{Adrien~Lagrange,~Mathieu~Fauvel~and~Manuel~Grizonnet% <-this % stops a space
+\author{Adrien~Lagrange ̃\IEEEmembership{Student, IEEE},~Mathieu~Fauvel~\IEEEmembership{Senior, IEEE}~and~Manuel~Grizonnet% <-this % stops a space
 \thanks{A. Lagrange and M. Fauvel are with the Universit\'{e} de Toulouse,
@@
 solution  is  the Orfeo  Toolbox,  developed  by the  French  Space
-Agency~\cite{christophe2008orfeo}.   Streaming and  parallel computing
+Agency (CNES)~\cite{christophe2008orfeo}.   Streaming and  parallel computing
 are  conveniently proposed  to  users/developpers through  several ``\emph{ready to use}'' modules.
@@
         \end{tikzpicture}
-        \caption{Number of parameters $\eta_c$ by class in function of dimension $d$: $\eta_c=d(d+3)/2+1$.\label{fig:nb-param}}
+        \caption{Number of parameters $\eta_c$ per class in function of dimension $d$: $\eta_c=d(d+3)/2+1$.\label{fig:nb-param}}
     \end{figure}
@@
 
-        The    KL   divergence    is   not    symmetric,   \emph{i.e.}
+        The    KL   divergence    is   not    symmetric,   \emph{i.e.},
         $KL_{cc^\prime} \ne KL_{c^\prime c}$.  A symmetrical version  is used to
@@
         The \emph{Jeffries-Matusita distance} is a measure based on the Bhattacharyya distance.
-It saturates if the separability between the two distributions increases~\cite{bruzzone2009novel}.
+It saturates when the separability between the two distributions increases~\cite{bruzzone2009novel}.
 The JM distance is defined as
@@
         at the $k^{th}$ iteration,  \emph{i.e.}, the covariance matrix
-        of  an  augmented  set  by  the  feature  $x_k$.  Then,  since
+        augmented  by  the  feature  $x_k$.  Then,  since
         $\boldsymbol{\Sigma}_c^{(k)}$ is a positive definite symmetric
@@
             GMM SFFS JM &   {\bfseries 8.8} & {\bfseries 9.0} & {\bfseries 9.3} & {\bfseries 5.0} & {\bfseries 5.0} & {\bfseries 5.4} & 21.45 & 24.35 & 27.05\\
-            GMM ridge &     71.7            & 105             & 167             & 530 & 530 & 530    & & & \\
-            KNN &           {\bfseries 8.9} & 19.6            & 59.7            & 387 & 639 & 887    & & & \\
-            Random Forest & 24.5            & 49.3            & 105             & 33.0 & 41.7 & 45.9 & & & \\
+            GMM ridge &     71.7            & 105             & 167             & 530 & 530 & 530    & all & all & all \\
+            KNN &           {\bfseries 8.9} & 19.6            & 59.7            & 387 & 639 & 887    & all & all & all \\
+            Random Forest & 24.5            & 49.3            & 105             & 33.0 & 41.7 & 45.9 & all & all & all \\
             \bottomrule

### 2017-01-16 paper.tex
minor updates in paper

@@
   It has been compared to  other classification algorithms on two high
-  dimension remote sensing images. The  results show that the approach
+  dimension remote sensing images. Results show that the approach
   provides good classification accuracies with low computation time.
@@
 
-\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the quantity of available Earth observation data for a given landscape becomes larger and larger.
-Satellite missions produce a huge amount of data on a regular (daily) basis.
-From 2018, the EnMAP (Environmental Mapping and Analysis Program) satellites managed by the German space agency will produce images with 244 spectral bands, a spatial resolution of 30x30m per pixel and with a frequency of revisit of 4 days\cite{Müller09enmap}.
-The Hyperspectral Infrared Imager (HyspIRI) of NASA will also deliver images with 212 spectral bands.
-Additionally to hyperspectral data, the amount of available hypertemporal data increases a lot.
-For instance, the European satellites Sentinel-2 were launched recently and 2 Terabytes of data will be released every day~\cite{drusch2012sentinel}.
-The Landsat open archive (\url{http://landsat.usgs.gov/products_data_at_no_charge.php}) has also released thousands of images.
-Such high volume of Earth observation data provides accurate information of land state and functions, and helps to improve the understanding of the planet~\cite{rs1010001}.
-However, processing such data is more and more challenging because of statistical and computational issues.
+\IEEEPARstart{W}{ith} the increasing number of remote sensing missions, the
+quantity of available Earth observation data for a given landscape becomes
+larger and larger. Satellite missions produce a huge amount of data on a regular
+(daily) basis. From 2018, the EnMAP (Environmental Mapping and Analysis Program)
+satellites managed by the German space agency will produce images with 244
+spectral bands, a spatial resolution of 30x30m per pixel and with a frequency of
+revisit of 4 days\cite{Müller09enmap}. The Hyperspectral Infrared Imager
+(HyspIRI) of NASA will also deliver images with 212 spectral bands. Additionally
+to hyperspectral data, the amount of available hypertemporal data increases a
+lot. For instance, the European satellites Sentinel-2 were launched recently and
+2 Terabytes of data will be released every day~\cite{drusch2012sentinel}. The
+Landsat open archive
+(\url{http://landsat.usgs.gov/products_data_at_no_charge.php}) has also released
+thousands of images. Such high volume of Earth observation data provides
+accurate information of land state and functions, and helps to improve the
+understanding of the planet~\cite{rs1010001}. However, processing such data is
+more and more challenging because of statistical and computational issues.
 
@@
 Agency (CNES)~\cite{christophe2008orfeo}.   Streaming and  parallel computing
-are  conveniently proposed  to  users/developpers through  several ``\emph{ready to use}'' modules.
+are  conveniently proposed  to  users/developers through  several ``\emph{ready to use}'' modules.
 
@@
 
-The resulting code is available as a remote module of the Orfeo toolbox on Github and makes it possible to process large high dimensional images efficiently.
+The resulting code is available as a remote module of the Orfeo ToolBox on GitHub and makes it possible to process large high dimensional images efficiently.
 The \texttt{C++} code is freely available for download: \url{https://www.orfeo-toolbox.org/external-projects/}.
@@
     % \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.
-  From eq.~(\ref{eq:cov}) and standard results for the determinant of block matrix~\cite[Chapter 9]{IMM2012-03274} we have immediatly:
+  From eq.~(\ref{eq:cov}) and standard results for the determinant of block matrix~\cite[Chapter 9]{IMM2012-03274} we have immediately:
   \begin{align*}

### 2017-01-16 paper.tex
del -

@@
 Section~\ref{sec:implementation}.  Experimental  results on  two  real
-high dimensional data-sets are given Section~\ref{sec:test}. Conclusion
+high dimensional datasets are given Section~\ref{sec:test}. Conclusion
 and perspectives conclude the paper in Section~\ref{sec:conclusion}.
@@
 
-\section{Data-sets}
+\section{Datasets}
 \label{sec:datasets}
-Numerical experiments have been  conducted on two different data-sets.
+Numerical experiments have been  conducted on two different datasets.
 The  first  one,  called  \emph{Aisa}, is  an  airborne  hyperspectral

### 2017-02-06 paper.tex
update

@@
 \usepackage[caption=false,font=footnotesize]{subfig}
-\usepackage{fixltx2e}
+% \usepackage{fixltx2e} % it trigers a warning. It may be useless ?
 \usepackage{url}
@@
 % \hyphenation{op-tical net-works semi-conduc-tor}
-\newcommand{\rev}[1]{\textcolor{magenta}{\emph{#1}}}
+% \newcommand{\rev}[1]{\textcolor{black}{\emph{#1}}}
+\newcommand{\rev}[1]{\textcolor{black}{#1}}
 
@@
 % \clearpage
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
 \begin{abstract}
@@
 % Note that keywords are not normally used for peerreview papers.
-% \begin{IEEEkeywords}
-% remote sensing, hyperspectral imaging, feature selection, gaussian mixture model, fast computing.
-% \end{IEEEkeywords}
+\begin{IEEEkeywords}
+remote sensing, hyperspectral imaging, feature selection, gaussian mixture model, fast computing.
+\end{IEEEkeywords}
 
@@
 
-% Among the various dimension reduction techniques, feature selection demonstrates several appealing for practical situation.
-The objective of feature selection is to find a subset of the original features with a minimum loss according the objective.
-Thus,
-
-% retrieve the most compact representation of the data with minimum loss of information \cite{Guyon:2006:FEF:1208773}.
-By deleting the redundant information, feature selection allows to reduce dimension and thus avoid statistical issue, increase algorithm speed and limit storage requirements.
-Moreover, it adds several beneficial side effects.
-It improves data understanding by identify where is the most relevant information and it is then possible to save resources when organizing a new data acquisition.
-
 There  are  two  main  strategies  to  reduce  dimension~\cite[Chapter
@@
 
-% The method proposed in this paper is actually a wrapper method which has been optimized in order to accelerate model learning and testing which is the weak point of this family.
-
 The third class corresponds to the \emph{embedded methods}.
@@
 Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~\cite{christophe2008orfeo}.}
-
-% Finally, tests are conducted to compare different variations of the method and to compare also to other classifiers (GMM, Random Forest, k-nearest-neighbor) available in the Orfeo Toolbox .
 
@@
     \begin{align}
-        p(\mathbf{x}) = \sum_{c=1}^{C} \pi_c f_c(\mathbf{x}|\theta)
+        p(\mathbf{x}) = \sum_{c=1}^{C} \pi_c f_c(\mathbf{x}|\theta),
     \end{align}
@@
         \begin{tikzpicture}
-          \begin{axis}[ymode=log,xmin=0,xmax=200,ymin=0,width=0.8\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = $d$,ylabel = $\eta_c$,small]
+          \begin{axis}[ymode=log,xmin=0,xmax=200,width=0.8\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = $d$,ylabel = $\eta_c$,small]
             \addplot+[very thick,mark=none,smooth,domain=0:200,samples=200] (\x,{\x*(\x+3)/2+1});
@@
 
-    % Criterion functions aim to evaluate either a rate of correct classification or the separability/similarity of class distributions.
-These functions are used to estimate which sets of variables are the best to represent data, to assure the separability of the classes and to perform classification.
-The choice of a criterion function is the choice of a way to compare sets of variables.
-
         \subsubsection{Measures of correct classification}
@@
         of $\mathcal{S}$.
-
-        \hspace{0pt} \\
 
@@
         \end{equation}
-
-        \vspace{10 mm}
-
-        % According to \cite{bruzzone2009novel}, it is interesting to notice that the KL divergence increases quadratically with respect to the distance between the mean vectors of the class distributions whereas the measures of correct classification asymptotically tends to one when distributions are perfectly separable.
-On the contrary, the JM distance tends to saturate as the measures of correct classification.
 
@@
         matrix, the covariance matrix can be written as
-
-        % Then, the inverse of  the covariance matrix
-        % $(\boldsymbol{\Sigma}^{(k)})^{-1}$,   the   quadratical   term
-        % $(\mathbf{x}^{(k)})^t         (\boldsymbol{\Sigma}^{(k)})^{-1}
-        % \mathbf{x}^{(k)}$        and        the        log-determinant
-        % $\log   |\boldsymbol{\Sigma}^{(k)}|$  can   be  expressed   in
-        % function of terms of the $(k-1)^{th}$ iteration.
 
@@
 
-        % In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$.
-The update rule becomes (proof in Appendix~\ref{app:proof-update})
-        % \begin{prop}[Backward update rule for inverse of covariance matrix]
-        % \label{eq:update-inv-back}
-        %     \begin{equation*}
-        %         (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \underbrace{\mathbf{A}}_{\substack{\text{computed once}\\ \text{per selection step}}} - \underbrace{\alpha \mathbf{v} \mathbf{v}^t}_{\substack{\text{computed for} \\ \text{each augmented set}}}
-        %     \end{equation*}
-        % \end{prop}
-
         Using eq.~(\ref{eq:cov}) and~(\ref{eq:cov:inv}), it is possible to
@@
         \label{eq:jm-update}
-            \text{B}_{cc^\prime} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
+            &\text{B}_{cc^\prime} = \underbrace{\frac{1}{4} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
             &+ \underbrace{\frac{1}{2} \log \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_c^{(k-1)}| |\boldsymbol{\Sigma}_{c^\prime}^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
@@
         \begin{tabular}{c}
-            \includegraphics[width=0.7\columnwidth]{Fig/aisa.png} \\
+            \includegraphics[width=0.6\columnwidth]{Fig/aisa.png} \\
             {\bfseries{(a)}} \\
-            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt.png} \\
+            \includegraphics[width=0.6\columnwidth]{Fig/aisa_gt.png} \\
             {\bfseries{(b)}} \\
@@
             \end{axis};
-            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
-            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
-            %     \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
-            %     \draw[thick,black] (axis cs:47,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:47,\pgfkeysvalueof{/pgfplots/ymax});
-            % \end{axis};
-            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
-            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_50000spl.txt};
-            %     \draw[thick,red] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
-            %     \draw[thick,black] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
-            % \end{axis};
         \end{tikzpicture}
@@
     \end{figure}
-
-    % Then, a small processing is made to select the number of variables used for prediction.
-The evolution of the criterion function is retrieved from the model file and the discrete derivative of the criterion \emph{DJ} is computed and normalized by its maximum.
-The subset of variables used for classification $\Omega^{*}$ is the first subset for which this value is lower than a given threshold \emph{th}, equal  to 0.1\% in this experiment,
-    % \begin{equation*}
-    %     \Omega^{*} = \Omega_{\min \{k|DJ_k<th\}}
-    % \end{equation*}
-    % with $DJ_k = \frac{J(\Omega_k) - J(\Omega_{k-1})}{\max_k (J(\Omega_k) - J(\Omega_{k-1}))}$.
 
@@
 In Table~\ref{tab:aisa-otbsimu}, when several bold scores appears for the same experiment, it means that the scores has been assessed as equivalent with a Wilcoxon rank-sum test \cite{mann1947test}.
-Additionally, Figure~\ref{fig:meanNbVar-aisa} summarizes the mean of the number of selected variables for each variation of the GMM classifier with selection.
+Additionally, \ref{tab:aisa-otbsimu-time} summarizes the mean of the number of selected variables for each variation of the GMM classifier with selection.
 
@@
         \begin{tabular}{c}
-            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt_train.png} \\
+            \includegraphics[width=0.6\columnwidth]{Fig/aisa_gt_train.png} \\
             {\bfseries{(a)}} \\
-            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt_test.png} \\
+            \includegraphics[width=0.6\columnwidth]{Fig/aisa_gt_test.png} \\
             {\bfseries{(b)}} \\
@@
 
-    % this one is result without changing the number of selected variables
-    % \begin{table}[!t]
-    %     \centering
-    %     \caption{Average classification accuracy 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
-    %     \begin{tabular}{lccc}\toprule
-    %          & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-4}
-    %         \# samples by class & 250 & 500 & 1000 \\ \midrule
-
-    %         GMM SFS kappa & {\bfseries 0.684 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.711 (0.028)} \\
-    %         GMM SFS JM &    {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
-    %         GMM SFFS JM &   {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
-    %         GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
-    %         KNN &           0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
-    %         Random Forest & 0.645 (0.026) & 0.673 (0.023) & 0.693 (0.023) \\
-    %         \bottomrule
-    %     \end{tabular}
-    % \end{table}
-
-    \begin{figure}
-      \centering
-      \begin{tikzpicture}
-        \begin{axis}[ylabel=$\bar{n}_s$,ymax=25,
-          symbolic x coords={0,250,500,1000,1050},
-          enlargelimits=0.15,legend pos=outer north east,xtick=data,legend cell align=left,
-          ybar,
-          bar width=5pt,grid,xlabel=Number of training samples per class,small,area legend]
-          \addplot coordinates {(250,15.9) (500,21.85) (1000,26.65)};
-          \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
-          \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
-          %%\addplot[red,very thick,sharp plot,update limits=false]coordinates {(0,252) (1050,252)};
-          \legend{{\footnotesize SFS kappa},{\footnotesize SFS JM},{\footnotesize SFFS JM}}
-        \end{axis}
-      \end{tikzpicture}
-      \caption{Mean number $\bar{n}_s$ of selected features for the different variation of selection methods for Aisa data set. The original number of features is 252.}
-      \label{fig:meanNbVar-aisa}
-    \end{figure}
-
     \begin{table*}[!t]
@@
 
+    % \begin{figure}
+    %   \centering
+    %   \begin{tikzpicture}
+    %     \begin{axis}[ylabel=$\bar{n}_s$,ymax=25,
+    %       symbolic x coords={0,250,500,1000,1050},
+    %       enlargelimits=0.15,legend pos=outer north east,xtick=data,legend cell align=left,
+    %       ybar,
+    %       bar width=5pt,grid,xlabel=Number of training samples per class,small,area legend]
+    %       \addplot coordinates {(250,15.9) (500,21.85) (1000,26.65)};
+    %       \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
+    %       \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
+    %       \legend{{\footnotesize SFS kappa},{\footnotesize SFS JM},{\footnotesize SFFS JM}}
+    %     \end{axis}
+    %   \end{tikzpicture}
+    %   \caption{Mean number $\bar{n}_s$ of selected features for the different variation of selection methods for Aisa data set. The original number of features is 252.}
+    %   \label{fig:meanNbVar-aisa}
+    % \end{figure}
+
     The results show that, on this  dataset, GMM classifiers with feature selection get the best classification rate.
@@
 Finally, for the Random Forest classifier, the trees tends to be deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.
-    % The computation of the kappa seems  to suffer from  a lack of  parallelization   which could be  improved. ->  plus tard dans la conclusion.
 
@@
 
-    % It can be guessed that the fact that the Gaussian hypothesis is unlikely to be verified explains the difficulty to extract the finer information available in a larger dataset.
-The JM distance has stressed in Table~\ref{tab:crit} has a strong Gaussian hypothesis and the same hypotheses could also explain the drop of the classification rate of GMM classifiers with JM distance as criterion.
-
     The  Postdam classes  are  more difficult  to discriminate,  since
@@
 
-% Finally, an improvement could be made to increase the stability of the selected features.
-For example, with hyperspectral data, a selection of continuous intervals and not band is a possible solution which has already been explored in \cite{serpico2007extraction}.
-
-% The python and C++ code are available freely for download: \url{https://www.orfeo-toolbox.org/external-projects/}, .
-
 \appendices
@@
 \label{app:proof-update}
-    % \subsubsection{Proposition~(\ref{eq:update-inv-back})}
-    %     \begin{proof}
-    %         \begin{align*}
-    %             \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
-    %             &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
-    %             &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
-    %             &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
-    %         \end{align*}
-    %     \end{proof}
 
@@
 
-    % \begin{align*}
-    %     \log &\left(|\boldsymbol{\Sigma}_c^{(k)}|\right) =
-    %     \log \left(\left|
-    %         \bigg[\begin{array}{cc}
-    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{u}_c      \\
-    %             \mathbf{u}_c^t          & \sigma^{(k)}_c \\
-    %         \end{array}\bigg]
-    %     \right|\right) \\
-    %     =&\log \left(\left|
-    %         \bigg[\begin{array}{cc}
-    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
-    %             \mathbf{u}_c^t          & 1 \\
-    %         \end{array}\bigg]
-    %         \bigg[\begin{array}{cc}
-    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
-    %             \mathbf{0}_{k-1,1}^t          & \sigma^{(k)}_c - \mathbf{u}_c^t(\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c\\
-    %         \end{array}\bigg]
-    %     \right|\right) \\
-    %     =&\log \left(\left|
-    %         \bigg[\begin{array}{cc}
-    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
-    %             \mathbf{u}_c^t          & 1 \\
-    %         \end{array}\bigg]
-    %     \right|\right)
-    %     +\log \left(\left|
-    %         \bigg[\begin{array}{cc}
-    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
-    %             \mathbf{0}_{k-1,1}^t          & \alpha_c \\
-    %         \end{array}\bigg]
-    %     \right|\right) \\
-    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha_c| \right)\\
-    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| \right) + \log ( \alpha_c )
-    % \end{align*}
-    % \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.
   From eq.~(\ref{eq:cov}) and standard results for the determinant of block matrix~\cite[Chapter 9]{IMM2012-03274} we have immediately:
@@
 
-% \begin{IEEEbiography}{Michael Shell}
-% Biography text here.
-% \end{IEEEbiography}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Fig/lagrange.jpg}}]{Adrien Lagrange} received an Engineering degree in Robotics and Embedded Systems from ENSTA ParisTech, Paris (Palaiseau), France, and the M.Sc.
+degree in Machine Learning from the Paris Saclay University, both in 2016.
+
+He is currently a Ph.D.
+student at the National Polytechnic Institute of Toulouse, within the Signal and Communications Group of the IRIT Laboratory.
+He is working on the subject of multi-resolution learning for hierarchical analysis of hyperspectral and hypertemporal images under the supervision of Nicolas Dobigeon and Mathieu Fauvel.
+His research interests are remote sensing, statistical modeling, image processing.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fauvel.pdf}}]{Mathieu Fauvel}
+ graduated in  electrical engineering  from the Grenoble  Institut of
+ Technology (Grenoble  INP), Grenoble, France, in  2004.  He  received
+ the M.S and  Ph.D.  degrees in image and signal  processing from  the
+ Grenoble INP  in 2004  and 2007,  respectively.  In  2007, he  was a
+ teaching assistant  in Grenoble  INP. From  2008 to  2010, he  was a
+ postdoctoral research associate with the MISTIS Team of the  National
+ Institute  for Research  in  Computer Science  and Control  (INRIA).
+ Since 2011,  Dr.  Fauvel  has been an  Associate Professor  with  the
+ National Polytechnic  Institute of  Toulouse (ENSAT -  University of
+ Toulouse) within  the DYNAFOR lab  (University of Toulouse  -  INRA).
+ His  research interests  are  remote sensing,  data fusion,  pattern
+ recognition, multicomponent signal and image processing. From  2013 to 2016,
+ he was the president of the French Chapter of the GRSS. He is  now in charge of the European GRSS Chapters activities.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Fig/manuel}}]{Manuel Grizonnet} received the Mathematical modeling, Vision, Graphics and
+  Simulation Engineer degree from the \'{E}cole Nationale Sup\'{e}rieure d'Informatique
+  et de Math\'{e}matiques Appliqu\'{e}es de Grenoble, France, in 2007. From 2007 to 2009
+  he worked for the BRGM (French geological survey) in Niamey, Niger where he was a
+  Systems and Geological Information System (GIS) engineer. He is currently with
+  the Centre National d'\'{E}tudes Spatiales (French Space Agency), Toulouse,
+  France, where he is developing image processing algorithms and software for
+  the exploitation of Earth observation images.
+\end{IEEEbiography}
 
@@
 \end{document}
-
-%\begin{figure}[!t]
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively.
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:
